### 2021-09-23 paper/paper.tex
project-brief: add initial background

\documentclass{report}

% Page Size
\usepackage[letterpaper, portrait, margin=1in, left=1.5in]{geometry}

% Spacing
\usepackage{setspace}
\setlength{\parskip}{\baselineskip}
\doublespacing{}

% Bibliography
\usepackage[american]{babel}
\usepackage[style=apa, citestyle=apa, backend=biber]{biblatex}
\DeclareLanguageMapping{american}{american-apa}
\addbibresource{references.bib}
\usepackage[babel,threshold=2]{csquotes}

% Graphics
\usepackage{graphicx}
\graphicspath{ {./images/} }

% Table of Content
\setcounter{tocdepth}{2}

% Code Formatting
\usepackage{minted}
\usemintedstyle{pastie}

\def\code#1{\texttt{#1}}

% Positioning
\usepackage{float}

% Colors
\usepackage[dvipsnames]{xcolor}

% Links
\usepackage{hyperref}
\hypersetup{
	pdftitle={Development of a touch typing trainer with an emphasis on finger and wrist placement},
	pdfstartview={FitH},
	colorlinks=true,
	linkcolor=black,
	citecolor=black,
	filecolor=black,
	urlcolor=MidnightBlue
}

\begin{document}

\pagenumbering{gobble}
\begin{titlepage}
	\centering
	\includegraphics[width=0.25\textwidth]{upc.png}
	\includegraphics[width=0.25\textwidth]{dcs.png}
	\par\vspace{1cm}
	{\scshape\LARGE Research Brief \par}
	\par\vspace{1cm}
	\begin{tabular}{ | l | l | }
		\hline
		Name               & Oscar Vian L. Valles                  \\
		\hline
		Student ID number  & 2018--06491                           \\
		\hline
		Project Title      & Development of a touch typing trainer \\ & with an emphasis on finger and wrist positions\\
		\hline
		Project No.        & 9                                     \\
		\hline
		Name of Supervisor & Dhong Fhel K. Gom-os                  \\
		\hline
	\end{tabular}

\end{titlepage}
\newpage

\tableofcontents
\newpage

\pagenumbering{arabic}

\setcounter{chapter}{1}
\section{Background of the Study}
There are a lot of educational typing tests available that help people learn
touch typing, including Monkeytype, TypeRacer, or Keybr. These typing tests list
out words that are then typed out. The inputted keys are then compared to check
if the user has typed the expected letter. At the end of the test, the time
taken is calculated and certain metrics is given. These metrics include words
per minute (WPM) and accuracy \parencite{bartnik2021}.

However, this method of examination leaves out a crucial part of typing —
ergonomics. Ergonomic typing prevents a lot health issues in the future like
repetitive strain injury or carpal tunnel. One important factor that affects
ergonomics is the typing procedure and posture. This means proper placement of
the wrist, hands, and hitting the keys using the right finger that is assigned
to the key.

Correct finger placement is usually taught at the beginning using a diagram,
with each key being associated with a specific finger. For instance, the letter
Q in a QWERTY layout should be hit using the fifth digit of the left hand, and
this is shown by coloring the fifth digit and the key Q with the same color or
by placing the letters directly on the fingers \parencite{dobson2009touch}.

Incorrect finger placement may cause these hand and wrist positions: ulnar
deviation, forearm pronation, and wrist extension \parencite{serina1999}. These
three are hand and wrist positions that are common in all activities, however,
prolonged periods in these positions may cause cause injuries such as Carpal
tunnel syndrome (CTS) \parencite{toosi2015}

In addition, this type of typing is frequently taught in the beginner level.
However, finger position is not \parencite{donica2018}

There is currently no way to automatically detect which finger is used during
typing. No hardware or software solutions currently exist in the market. This
research aims to solve this by combine finger tracking, with educational typing
tests to determine which finger pressed the key and if that finger is the
correct one to use.

\section{Significance of the Research}
This research is beneficial for all users of physical keyboards. These include a
vast majority of the population as there are a lot of professions that heavily
rely on keyboards. Examples include developers, physicians, educators,
accountants.

This research has a direct impact on people that has onset RSI or other
hand/wrist injuries [ statistics ] that are caused by poor typing habits. By
correcting these poor habits, pain from these injuries will be lessened, and
even be prevented from occurring in the first place.

\newpage
\nocite{*} \printbibliography[heading=bibintoc,title={References}]{}

\end{document}

### 2021-09-23 paper/paper.tex
project-brief: finish background and significance

@@
 
-In addition, this type of typing is frequently taught in the beginner level.
-However, finger position is not \parencite{donica2018}
+In addition, this type of typing is frequently taught in the beginner level
+\parencite{donica2018}. This means that there is a need to weed out bad habits
+that may develop, like using the index finger for pressing the spacebar or
+backspace. However, it is impractical for an educator to check each individual
+student if they are not performing these movements as these may only show for a
+specific which may not be common.
 
-There is currently no way to automatically detect which finger is used during
-typing. No hardware or software solutions currently exist in the market. This
-research aims to solve this by combine finger tracking, with educational typing
-tests to determine which finger pressed the key and if that finger is the
-correct one to use.
+Thus, there is a need for automatically detecting which finger is used during
+typing, and for the position of the wrist in relation to the arm. One way to do
+this is through finger and hand tracking. One solution for tracking is by using
+image processing and machine learning. An example of this is MediaPipe by
+\cite{mediapipe}.
+
+MediaPipe allows for various applications for machine learning in the field of
+image processing. This includes, hand tracking, pose estimation, object
+detection, and others. Another example of a library that allows for hand and
+finger tracking is OpenCV by \cite{opencv}. This is a tool that simplifies for
+computer vision, and image processing. Machine learning can also be used with
+OpenCV.
+
+This research aims to improve touch typing training by combining hand and finger
+tracking, with educational typing tests to determine if the finger used to type
+is correct and if the overall posture of the hand ergonomic and healthy.
 
@@
 rely on keyboards. Examples include developers, physicians, educators,
-accountants.
+accountants. By having better ergonomics while typing, wrist injuries can be
+prevented, and typing speed may be increased
+
+This research also helps educators, especially early educators teaching beginner
+typers. By automatically checking for ergonomics and posture, the burden of
+checking each individual students is lessened, and directed interventions for
+bad habits can be easily created as students with these bad habits are easily
+identified
 
 This research has a direct impact on people that has onset RSI or other
-hand/wrist injuries [ statistics ] that are caused by poor typing habits. By
-correcting these poor habits, pain from these injuries will be lessened, and
-even be prevented from occurring in the first place.
+hand/wrist injuries that are caused by poor typing habits. By correcting these
+poor habits, pain from these injuries will be lessened, and even be prevented
+from occurring in the first place. A specific example of this is by reducing
+ulnar deviation which affects the nerve that is indicative of CTS
+\parencite{toosi2015}.
 

### 2021-09-24 paper/paper.tex
project-brief: add final project brief

@@
 if the user has typed the expected letter. At the end of the test, the time
-taken is calculated and certain metrics is given. These metrics include words
+taken is calculated and certain metrics are given. These metrics include words
 per minute (WPM) and accuracy \parencite{bartnik2021}.
@@
 However, this method of examination leaves out a crucial part of typing —
-ergonomics. Ergonomic typing prevents a lot health issues in the future like
+ergonomics. Ergonomic typing prevents a lot of health issues in the future like
 repetitive strain injury or carpal tunnel. One important factor that affects
-ergonomics is the typing procedure and posture. This means proper placement of
+ergonomics is the typing procedure and posture. This means the proper placement of
 the wrist, hands, and hitting the keys using the right finger that is assigned
@@
 three are hand and wrist positions that are common in all activities, however,
-prolonged periods in these positions may cause cause injuries such as Carpal
+prolonged periods in these positions may cause injuries such as Carpal
 tunnel syndrome (CTS) \parencite{toosi2015}
@@
 that may develop, like using the index finger for pressing the spacebar or
-backspace. However, it is impractical for an educator to check each individual
-student if they are not performing these movements as these may only show for a
-specific which may not be common.
+backspace. However, it is impractical for an educator to check each student if
+they are not performing these movements as these may only show for a small
+period which may not be caught in time.
 
@@
 this is through finger and hand tracking. One solution for tracking is by using
-image processing and machine learning. An example of this is MediaPipe by
-\cite{mediapipe}.
+image processing and machine learning. An example of this is MediaPipe
+by~\cite{mediapipe}.
 
@@
 detection, and others. Another example of a library that allows for hand and
-finger tracking is OpenCV by \cite{opencv}. This is a tool that simplifies for
-computer vision, and image processing. Machine learning can also be used with
+finger tracking is OpenCV by~\cite{opencv}. This is a tool that simplifies
+computer vision and image processing. Machine learning can also be used with
 OpenCV.
 
-This research aims to improve touch typing training by combining hand and finger
-tracking, with educational typing tests to determine if the finger used to type
-is correct and if the overall posture of the hand ergonomic and healthy.
+\section{Research Objectives}
+
+\subsection{General Objectives}
+To create a touch typing trainer that corrects poor finger placement and hand
+position to develop better typing habits and healthier typing ergonomics.
+
+\subsection{Specific Objectives}
+\begin{itemize}
+	\item To develop a program that tracks fingers and hand positions while typing
+	\item To create a subroutine that ascertains which finger was used to type a key
+	\item To detect if incorrect fingers were used to press a key or if the
+	      position of the hand in relation to the wrist is problematic
+	\item To assess the accuracy and performance of the developed program's ability
+	      to perform the previously stated objectives
+	\item To develop a user-friendly interface for users to train touch typing
+	      using the developed program
+	\item To generate key statistics of a user's touch typing performance
+\end{itemize}
+
+\section{Scope and Limitation}
+This research will focus on typing on a 60\% keyboard. This type of keyboard
+only has the alphanumeric part of the keyboard. This limits the number of keys
+to be checked and the expected movement of the hand. Furthermore, The keycaps
+will also be of a light color, while the surface that the keyboard rests upon
+will be of a dark color.
+
+In addition, the keyboard layout will be ANSI. This layout is described by the
+American National Standards Institute \parencite{ansi}. This is the most
+common layout in the United States. However, it is also used in numerous
+English-speaking countries such as the Philippines, Malaysia, and India.
+
+The program will expect the that user has all ten digits and has no hand,
+finger, or wrist deformities. In addition, only the placement of the hands,
+fingers, and wrists will be taken into account when determining if the
+ergonomics of the user while typing is healthy. The program will not check
+seating position, angle of elbows, and other metrics for an ergonomic typing
+posture while typing.
+
+Capturing of the video to be analyzed by the program would be limited to a
+single 1080p webcam that is capturing in 60 frames per second. The camera will
+be pointed downwards facing the keyboard and the hand. This means that the
+vertical angle of the wrist may not be accurate.
 
@@
 This research also helps educators, especially early educators teaching beginner
-typers. By automatically checking for ergonomics and posture, the burden of
-checking each individual students is lessened, and directed interventions for
-bad habits can be easily created as students with these bad habits are easily
-identified
+typers. By automatically checking for ergonomics, posture, and correct
+technique, the burden of checking each student is lessened, and directed
+interventions for bad habits can be easily created as students with these bad
+habits are easily identified
 
-This research has a direct impact on people that has onset RSI or other
-hand/wrist injuries that are caused by poor typing habits. By correcting these
-poor habits, pain from these injuries will be lessened, and even be prevented
-from occurring in the first place. A specific example of this is by reducing
-ulnar deviation which affects the nerve that is indicative of CTS
-\parencite{toosi2015}.
+This research has a direct impact on people that has hand or wrist injuries that
+are caused by poor typing habits. By correcting these poor habits, pain from
+these injuries will be lessened, and even be prevented from occurring in the
+first place. A specific example of this is by reducing ulnar deviation which
+affects the nerve that is indicative of CTS \parencite{toosi2015}.
 

### 2021-12-01 paper/paper.tex
rrl: finish initial rrl

@@
 
+% Acronyms
+\usepackage{acro}
+
+\DeclareAcronym{iso}{
+	short=ISO,
+	long=International Organization for Standardization
+}
+
+\DeclareAcronym{ansi}{
+	short=ANSI,
+	long=American National Standards Institute
+}
+
+\DeclareAcronym{jis}{
+	short=JIS,
+	long=Japanese International Standards
+}
+
+\DeclareAcronym{wpm}{
+	short=WPM,
+	long=Words per Minute
+}
+
+\DeclareAcronym{neck}{
+	short=WRNULD,
+	long=Work-related neck and upper limb disorders
+}
+
+\DeclareAcronym{cv}{
+	short=CV,
+	long=Computer Vision
+}
+
+% Chapter Formatting
+\usepackage{titlesec}
+\titleformat{\chapter}[display]
+{\centering\LARGE\normalfont\bfseries}
+{Chapter \thechapter}
+{0em}
+{\vspace{-1.5ex}}
+
 % Colors
@@
 	\centering
-	\includegraphics[width=0.25\textwidth]{upc.png}
-	\includegraphics[width=0.25\textwidth]{dcs.png}
+
+	\hspace{0pt}
+	\vfill
+
+	\includegraphics[width=0.2\textwidth]{upc.png}
+	\includegraphics[width=0.2\textwidth]{dcs.png}
 	\par\vspace{1cm}
-	{\scshape\LARGE Research Brief \par}
-	\par\vspace{1cm}
-	\begin{tabular}{ | l | l | }
-		\hline
-		Name               & Oscar Vian L. Valles                  \\
-		\hline
-		Student ID number  & 2018--06491                           \\
-		\hline
-		Project Title      & Development of a touch typing trainer \\ & with an emphasis on finger and wrist positions\\
-		\hline
-		Project No.        & 9                                     \\
-		\hline
-		Name of Supervisor & Dhong Fhel K. Gom-os                  \\
-		\hline
-	\end{tabular}
 
+	\textsc{Development of a touch typing trainer\\with an emphasis on finger and wrist positions}
+	\par\vspace{0.5cm}
+
+	\hrulefill
+	\par\vspace{0.25cm}
+	A Special Project Proposal Presented to the\\
+	Faculty of the Department of Computer Science,\\
+	University of the Philippines Cebu
+
+	\par\vspace{0.25cm}
+	In Partial Fulfillment\\
+	Of the Requirements for the Degree\\
+	Bachelor of Science in Computer Science\\
+	\par\vspace{0.25cm}
+	\hrulefill
+	\par\vspace{0.5cm}
+
+	Oscar Vian L. Valles\\
+	BS Computer Science
+	\par\vspace{0.5cm}
+
+	Dhong Fhel K. Gom-os\\
+	Adviser
+	\par\vspace{0.5cm}
+
+	October 2021
+	\vfill
+	\hspace{0pt}
 \end{titlepage}
+\newpage
+
+\printacronyms
 \newpage
@@
 
-\setcounter{chapter}{1}
+\chapter{Introduction}
+
+%FIXME - Lots of passive voice
 \section{Background of the Study}
 There are a lot of educational typing tests available that help people learn
-touch typing, including Monkeytype, TypeRacer, or Keybr. These typing tests list
+touch typing, including Monkeytype, TypeRacer, and Keybr. These typing tests list
 out words that are then typed out. The inputted keys are then compared to check
 if the user has typed the expected letter. At the end of the test, the time
-taken is calculated and certain metrics are given. These metrics include words
+taken is calculated, and certain metrics are given. These metrics include words
 per minute (WPM) and accuracy \parencite{bartnik2021}.
@@
 \subsection{General Objectives}
-To create a touch typing trainer that corrects poor finger placement and hand
-position to develop better typing habits and healthier typing ergonomics.
+To create a touch typing trainer that detects poor finger placement and hand
+position to help develop better typing habits and healthier typing ergonomics.
 
@@
 	      using the developed program
-	\item To generate key statistics of a user's touch typing performance
+	\item To generate key statistics of a user's touch typing performance:
+	      \begin{itemize}
+		      \item Words per Minute
+		      \item Accuracy
+		      \item Finger Placement Accuracy
+	      \end{itemize}
 \end{itemize}
@@
 \section{Scope and Limitation}
-This research will focus on typing on a 60\% keyboard. This type of keyboard
-only has the alphanumeric part of the keyboard. This limits the number of keys
-to be checked and the expected movement of the hand. Furthermore, The keycaps
-will also be of a light color, while the surface that the keyboard rests upon
-will be of a dark color.
+This research will focus on typing on a 60\% keyboard. Figure~\ref{fig:60}
+illustrates this type of keyboard. This type of keyboard only has the
+alphanumeric part of the keyboard. This limits the number of keys to be checked
+and the expected movement of the hand. Furthermore, the keycaps will also be of
+a light color, while the surface that the keyboard rests upon will be of a dark
+color.
 
-In addition, the keyboard layout will be ANSI. This layout is described by the
+In addition, the keyboard layout will be \ac{ansi}. This layout is described by the
 American National Standards Institute \parencite{ansi}. This is the most
@@
 
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{60.jpg}
+	\caption{A 60\% keyboard in \ac{ansi} layout. Reprinted from \fullcite{60}}
+	\label{fig:60}
+	\centering
+\end{figure}
+
 \section{Significance of the Research}
@@
 This research also helps educators, especially early educators teaching beginner
-typers. By automatically checking for ergonomics, posture, and correct
+typists. By automatically checking for ergonomics, posture, and correct
 technique, the burden of checking each student is lessened, and directed
@@
 
+\chapter{Preliminary Review of Related Literature}
+
+\section{Keyboard Typing}
+
+Keyboard typing is the process of using a keyboard to input characters in a
+system. In the context of this paper, keyboard typing will refer to the act of
+using a physical keyboard to input characters in a computer system.
+
+% Insert more about keyboard typing here %
+
+\subsection{Keyboard Layouts and Form Factors}
+
+One key characteristic of a keyboard is its physical attributes. Keyboards come
+in a lot of layouts and form factors. Keyboard layouts are the shapes, size, and
+positions of a key on a keyboard while the form factor of a keyboard refers to
+its shape and dimensions. The form factor also refers to the number of keys included
+in the keyboard \parencite{parkkinen2018}. By combining different layouts and
+form factors, different permutations of a keyboard can be created.
+
+Different keyboard layouts and form factors also produce different effects for
+the user. This is due to how vastly different some keyboard layouts and form
+factors are from one another. Some layouts focus on ergonomics, while others
+focus on typing speed. Some form factors were designed for aesthetics, while
+others focus on comfort and health. As such, different layouts may affect typing
+performance, ergonomics, and long-term health effects \parencite{ciobanu2015}.
+
+\subsubsection{ANSI and ISO Layout}
+
+There are two common keyboard layouts around the world --- \ac{iso} and \ac{ansi}.
+
+\citeauthor{ansi} is the standard that first defined the \ac{ansi} layout.
+Figure~\ref{fig:ansi} illustrates what the \ac{ansi} layout looks like. This layout
+is also used by countries other than America. Examples of countries that use
+this layout as its standard is the Philippines, China, and Korea
+\parencite{apple-layout}. However, these countries also opt to modify the layout
+by adding extra layers to accommodate other character sets.
+
+\citeauthor{iso} is the standard series that defines a framework that is used to
+create other layouts. Layouts created from this standard are colloquially called
+\ac{iso} Layouts. Countries around the world use this framework to create layouts
+that fit the characters in their language. Examples of countries that use this
+framework to create their own layout are France, Greece, Canada, and Sweden
+\parencite{apple-layout}.
+
+Both of these layouts usually utilize the same key ordering. This ordering is
+commonly called QWERTY, based on the first five characters of the first row
+of this specific layout.
+
+There are other layouts available, however, they are not as common as the two
+previously mentioned layouts. Examples of this include \ac{jis}. Other esoteric
+layouts, like Tsangan or split-backspace, also exist. These layouts modify the
+\ac{iso} and \ac{ansi} standards by adding or removing certain keys to fit the
+character set of a language, or for additional keys. Other layouts are also
+exactly the same as \ac{ansi} or \ac{iso}, however, these layouts change the
+arrangement of the alphabet within the keyboard.
+
+Despite the ubiquity of these common layouts, studies have shown that these
+layouts are not ergonomic. The main issue with these layouts is the random
+configurations of the letters. The randomness of the layout necessitates
+memorization of the layout which reduces the ease of learning, reduces performance
+in typing by reducing speed, and increases of typing errors
+\parencite{ciobanu2015}.
+
+\subsubsection{Keyboard Form Factors}
+
+There is only one common keyboard form factor used worldwide: the full-size
+keyboard. This keyboard contains all the keys specified in the keyboard
+layout. This includes the alphanumeric keys, the function keys, the navigation
+cluster, and the numpad.
+
+Other common keyboard form factors are based on the full-size keyboard. The name
+of these layouts, 60\%, 75\%, and 80\% reference the remaining number of keys
+after cutting a portion off from the full-size keyboard. The 60\% keyboards only
+contain the alphanumeric cluster while the 80\% and 75\% layouts retain the
+navigation cluster and the function keys \parencite{parkkinen2018}. The main
+draw for using keyboards with reduced sizes is for aesthetics, space
+constraints, and ergonomics.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{ansi.png}
+	\caption{\ac{ansi} Keyboard layout with form factors. Reprinted from \fullcite{figure-ansi}}
+	\label{fig:ansi}
+	\centering
+\end{figure}
+
+\subsubsection{Ergonomic Keyboards Layouts and Form Factors}
+
+There have been other keyboard layouts and form factors created to mitigate
+common issues associated with QWERTY layouts. These include Colemak, Dvorak, and
+Alphabetical layouts. However, studies have shown that the layout itself does
+not matter as beginners do not necessarily see the keyboard as a structured set,
+but rather as a random collection of characters, even if it is alphabetized
+\parencite{norman1982},
+
+A different form factor has a great effect on ergonomics. One such example of a
+form factor is an ergonomic keyboard developed by Microsoft called Microsoft
+Natural MultiMedia Keyboard. \citeauthor{ripat2010} used this keyboard in
+determining that ergonomic keyboards can help in reducing symptoms of \ac{neck}.
+Figure~\ref{fig:mn} shows the layout of the Microsoft Natural MultiMedia
+Keyboard.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{mn.png}
+	\caption{Microsoft Natural MultiMedia Keyboard. Reprinted from \fullcite{mn}}
+	\label{fig:mn}
+	\centering
+\end{figure}
+
+There are other form factors other than the full-size keyboard and variations
+thereof that focus on ergonomics. One such example is a split keyboard layout
+where the keyboard is split in half, one for the left hand and one for the
+right. One such benefit, according to \citeauthor{ergodox}, is a more relaxed
+position due to typing at shoulder width.
+
+\subsection{Keyboard Typing Metrics}
+
+There are numerous metrics used to quantify keyboard typing performance. Two
+common metrics used in the majority of typing tests include Accuracy and Speed.
+
+\subsubsection{Standardized Keyboard Typing Assessments}
+
+To be able to measure these metrics, a keyboard typing assessment needs to be
+done. However, there are no standardized keyboard typing assessments
+\parencite{donica2018}. As such, teaching methods and assessments, like
+Keyboarding without Tears, Monkeytype, and Keybr, may produce different metrics
+for the same typist due to their difference in conducting the assessment.
+
+\subsubsection{Speed}
+Speed, also called as entry rate by \citeauthor{arif2009}, measures the number
+of characters entered in a specific time frame. The most common metric that
+measures speed is \ac{wpm}. \ac{wpm} as defined by \citeauthor{arif2009} is:
+
+\begin{equation}
+	WPM = \frac{|T| - 1}{S} \cdot 60 \cdot \frac{1}{5}
+\end{equation}
+
+where, $|T|$ is the length of the text, $S$ is the time in seconds spent writing
+the text. This time starts directly after the first character has been pressed,
+and ends when the last letter has been entered. As such, $1$ is subtracted from
+$|T|$, as the time spent to find and press the first character cannot be
+accurately determined. However, some typing assessments do not subtract $1$ from
+$|T|$. $60$ refers to the number of seconds in a minute and $\frac{1}{5}$
+normalizes the metric for the average length of words.
+
+Other metrics also measure speed but they aren't as commonly used as \ac{wpm}.
+These include Characters per Minute, Gestures per Second, Adjusted Words per
+Minute, and Keystrokes per Second
+
+\subsubsection{Accuracy}
+Accuracy measures the number of correctly pressed characters in an input string.
+Accuracy, as defined by \citeauthor{bartnik2021}, is:
+
+\begin{equation}
+	ACC = \frac{|C|}{|T|} \cdot 100\%
+\end{equation}
+
+where $|C|$ is the number of correct characters and $|T|$ is the length of the
+text.
+
+The inverse of accuracy is error rate, where the number of incorrectly pressed
+characters is measured instead. \citeauthor{arif2009} describe 5 common error
+rate metrics: Error Rate, Minimum String Distance Error Rate, Keystroke per
+Character, Erroneous Keystroke Error Rate, and Total Error Rate.
+
+\subsubsection{Limitations of the Metrics}
+These metrics are all based on the inputted characters by the user. These
+metrics do not take into account other aspects of keyboard typing such as
+posture, hand and wrist positions, and finger placement. Consequently, these
+metrics do not give a full picture of the performance of the person typing and
+they only provide a cursory view of how a person types.
+
+\subsection{Keyboard Typing Methodology}
+Keyboard typing can be accomplished in numerous ways. The main difference
+between the different methodologies is the number of fingers used when typing
+and how the typist navigates the keyboard to find the keys. The methodology
+ranges from Hunt and Peck to Touch Typing, with variations of the two in
+between.
+
+Hunt and Peck uses one finger on one hand to press a key. This method is aided
+by using vision to locate the specific key to press \parencite{hoot1986}. On the
+other hand, Touch typing uses standard QWERTY mapping to type without using
+visual cues. \parencite{dobson2009touch} This mapping involves assigning certain
+fingers to certain keys. Figure~\ref{fig:touch-type} is the standard QWERTY
+mapping used for an \ac{ansi} layout. Kinesthesis and proprioception are used in
+locating the keys \parencite{logan2016}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{touch-type.png}
+	\caption{Standard QWERTY mapping for \ac{ansi}. Reprinted from \fullcite{logan2016}}
+	\label{fig:touch-type}
+	\centering
+\end{figure}
+
+\section{Keyboard Typing in Education}
+
+Today, students are expected to type essays, articles, and other submissions
+using word processors \parencite{poole2016}. Testing is also commonly done using
+computerized assessments which require the need for keyboards
+\parencite{moodle}. As such, there is a need for students to be well versed in
+keyboard typing and for keyboard typing to be part of the curriculum.
+
+Keyboard typing has been a part of this curriculum for a long time, with studies
+about effective methods to teach keyboard typing reaching as far back as 1986
+\parencite{hoot1986}. Studies have continued to this day to continue to optimize
+and improve methods of teaching keyboard typing to students.
+
+These studies start teaching kids in the kindergarten level and the studies try
+to optimize the teaching methods to improve the speed and accuracy of typing of
+the learners. By starting to teach touch typing to students early, these
+students will develop the potential for higher-level keyboard typing
+\parencite{donica2018}.
+
+\subsection{Expectations of Keyboard Proficiency}
+
+In the United States, keyboard typing is an expected learning outcome for third
+grade in the Common Core State Standards \parencite{ccs}. At this grade level,
+only basic keyboard typing skills are required. By fourth grade, students are
+expected to have enough proficiency to type one page in one sitting. This is
+increased to two pages by fifth grade.
+
+In the Philippine context, the \citeauthor{deped} expects learners with a mental
+age of 4--6.9 years old to use correct posture and locate characters, learners
+with a mental age of 7--11.9 are introduced to home row finger placement, and
+learners with a mental age of 12 and above are expected to ``use proper typing
+technique with efficiency and accuracy without looking at the keyboard''
+\parencite{deped}.
+
+\subsection{Current Teaching Methods}
+
+Current teaching methods involve replicating a given text. Learners then copy
+the text into a given text field that records the typed characters. Correct and
+incorrect characters are then identified, and suitable errors are presented.
+Afterward, metrics, such as \ac{wpm}, and accuracy are given
+\parencite{bartnik2021, typeracer}.
+
+Through this process, the learner goes through the three stages of Motor
+Learning Theory. The student undergoes the cognitive stage where they try to
+understand and create strategies to accomplish the given task. Then the
+associative stage follows where the strategies and skills learned from the
+previous stage are refined. At this stage, the learners are expected to rely less
+on visuals to locate the keys and more on kinesthesis. By the final stage, the
+autonomous stage, the learner does not rely on visuals at all and focuses on
+using kinesthetic feedback to find the keys. By this point, the learner has
+progressed from using Hunt and Peck, to becoming proficient in touch typing.
+\parencite{donica2018}
+
+\subsubsection{Keyboarding without Tears}
+
+Keyboarding without Tears is a web-based application and curriculum that teaches
+students touch typing. However, one key differentiator of this curriculum is the
+usage of a row-based standard mapping, rather than a column-based standard
+mapping that is common in other teaching guides. Figure~\ref{fig:kwt} shows the
+standard mapping used in this curriculum.
+
+This curriculum is self-directed and learners can learn at their own pace. At
+its core, the curriculum is designed to be 36-week long with 5-10 minutes of
+lessons per day. The lessons in the curriculum follow the three stages of
+Motor Learning Theory \parencite{kwt}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{kwt.png}
+	\caption{Row based standard mapping. Reprinted from \fullcite{kwt}}
+	\label{fig:kwt}
+	\centering
+\end{figure}
+
+\subsubsection{Monkeytype, Typeracer}
+
+These two keyboard typing tests are similar. They follow a common experience
+where users type a predetermined phrase, quote, or random words, and metrics are
+given after the test. Afterward, the learners may try the test again, or choose
+another set of words to type. These typing tests do not have a structured
+curriculum for learning how to touch type. It is left to the learner to practice
+and learn on their own \parencite{bartnik2021, typeracer}.
+
+\subsubsection{Keybr}
+
+Keybr is similar to Monkeytype and Typerace, in that they also have the users
+type a predetermined phrase, quote, or random words. However, this application
+has more guidance compared to the two. Keybr uses statistics to create typing
+lessons that are appropriate to the current typing proficiency of the learner.
+The words selected are random at first, and the skill level of the learner is
+determined by the performance of the user with these words and characters. The
+information gathered is then used to generate new words for the next iteration.
+As an example, if a learner has difficulty in typing the letter q, the next
+iterations will have a lot of words that contain the letter q.
+
+Statistics from their website show that this learning method is successful, with
+some learners improving their typing speed by 20--40 \ac{wpm} \parencite{keybr}.
+
+\section{Keyboard Typing in Health}
+
+There have been a lot of studies that show the effect of keyboard typing, and
+its associated movements (or lack thereof), has an effect on the human body.
+These studies have shown that keyboard typing has an effect on our neck, shoulder,
+upper limb, wrist, arms, and fingers \parencite{szeto2005, baker2007digit}
+
+\subsection{Health Issues arising from Keyboard Typing}
+
+\ac{neck} are a common issue that is associated with an elongated length of time
+maintaining a static posture. When using the computer, the posture commonly
+adapted by users has the neck and shoulder regions in a static hold for a long
+time. This results in forward neck flexion and increased muscle tension
+\parencite{szeto2005}.
+
+In addition, it has been shown that 22\% of computer users sustain
+musculoskeletal disorders of the upper extremity. This includes the neck,
+shoulder, hands, and wrists. \parencite{gerr2002}
+
+Carpal tunnel syndrome is also a common issue in the general population. This is
+caused by the chronic compression of the median nerve. There is a common belief
+that typing is one main cause for the disorder \parencite{carpal-myth}. There
+are no definite conclusions if this myth is true, however, a study by
+\citeauthor{toosi2015} found that typing causes ulnar deviation, especially if
+done without proper form. This ulnar deviation contributes to the swelling of
+the median nerve during and after typing. However, the authors noted that it is
+unclear if this swelling leads to long-term nerve injury.
+
+\subsection{Finger and Wrist Kinematics}
+The way people move their hands, wrists, and fingers differ between each person.
+This can be attributed to the different typing styles each person has. One key
+difference between people is the angle of the 5th digit.
+
+However, there are some common movements and positions regardless of typing
+style: flexion, or the curving of the fingers, across the fingers, is decreasing
+across the hand, with the 2nd digit having the least flexion. This may be due to
+the instinct to reduce pronation of the hand, which in turn increases the
+distance of the 2nd digit to the keyboard. In addition, some people isolate or
+extend one of their thumbs, usually the one not used for pressing a key. This is
+also true for some people that do not use their 5th digit during typing
+\parencite{baker2007}.
+
+The movement and angle of the wrists also depend on the typing style of the
+typists. Some people do not reposition their hands, while others do. This
+difference comes from the way these people reach for certain far-away keys. Some
+stretch their fingers to reach far-away keys, while others move their entire
+hand to reach these keys.
+
+For those that reach their keys by stretching their fingers, there is an
+increased probability that the wrists and fingers adapt non-neutral postures.
+These include wrist extension, ulnar deviation, and pronation, which may cause
+musculoskeletal disorders of the upper extremity (\cite{marklin1999} as cited in
+\cite{baker2007})
+
+\section{Finger and Hand Tracking}
+Finger and Hand tracking is a method of tracking fingers and hands in 3D space
+using motion capture systems or computer vision. This technique allows computers
+to perform actions and analyses on the motions and positions of these body
+parts.
+
+\subsection{Types of Tracking}
+
+\subsubsection{Hardware Aided Solutions}
+
+Motion Capture Systems allow for capturing detailed skeletal motion in humans.
+These systems usually capture full-body motion, focusing on large parts of the
+human body, such as the torso, limbs, and head.
+
+However, motion capture systems have difficulty in tracking more articulated
+body parts --- with the fingers being one of them. The industry standard for
+capturing finger movements is through the use of an optical marker-based motion
+capture system. This is due to its ability to capture natural motion accurately.
+
+This method uses cameras to triangulate the 3D location of markers attached to
+the limbs of a person. For finger tracking, 13--20 markers are placed on the
+fingers, and cameras are brought closer to track the small movements of the
+finger \parencite{wheatland2015}.
+
+But this method is cost-prohibitive, and cannot handle occlusions well.
+\citeauthor{alexanderson2016} present a method for an optical marker-based
+motion capture system that can predictably recover from self-occlusion and has a
+better performance compared to previously used algorithms, however, the issue of
+cost and self-occlusion still persists.
+
+Bend-sensor gloves are also an option for finger tracking. These gloves have
+sensors within them that track joint angles in the hand and fingers. One key
+differentiator of this solution compared to the others is the removal of
+self-occlusion in the data. As such, this is commonly used in sign language, and
+gesture recognition due to its accuracy.
+
+However, these gloves need a lot of time to calibrate as cross-coupling of the
+sensors proves a problem. Cross-coupling is prevalent because the movement of
+one finger also moves other parts of the hand. These movements may cause a
+sensor aimed to track a specific movement of a different part of the hand to
+inadvertently detect a movement when there should be none
+\parencite{wheatland2015}.
+
+\subsubsection{Computer Vision}
+
+At its core, Computer Vision aims to perform tasks that the human visual system
+can do \parencite{cern}. This includes object classification, tracking, and
+gesture recognition, and face recognition. At the present, most computer vision
+systems utilize deep learning algorithms, and convolutional networks to gather
+information from an image, or a set of images. One such example of a
+convolutional network used in computer vision is Inception by
+\citeauthor{szegedy2015} which proposes a convolutional neural network
+architecture for object classification and detection.
+
+\subsection{Available CV Solutions for Tracking}
+
+\subsubsection{OpenCV}
+
+OpenCV is an open-source computer vision and machine learning software library
+that houses ${\approx2500}$ optimized algorithms. This library is widely used by
+companies, researchers, and open source communities that utilize computer vision
+and machine learning in their projects. Examples of companies that use OpenCV
+include Google, Sony, and Honda.
+
+The library has C++, Python, Java, and Matlab interfaces. The library also supports
+Windows, Linux, Android, and macOS, allowing for great developer experience, and
+wide deployment capabilities \parencite{opencv}.
+
+\subsubsection{MediaPipe}
+
+MediaPipe is an open-source computer vision framework that allows developers to
+create a perception pipeline. This perception pipeline is a directed graph of
+calculators. Data passes through the graph as packets and a group of packets
+constitute a data stream. As the data passes through the pipeline, the
+calculators, produce the desired output.
+
+This framework allows for performant object detection, hand and finger tracking,
+human pose detection. The framework also allows for combining multiple features,
+by adding them to the graph as calculators. MediaPipe has C++, Python, JS, and
+Coral interfaces. It also supports Android and iOS devices
+\parencite{mediapipe}.
+
+\subsubsection{MATLAB}
+MATLAB is a programming platform for the analysis and designing of systems.
+MATLAB is commonly used by engineers and scientists for computational
+mathematics \parencite{what-matlab}.
+
+A toolbox offered by MATLAB is the Computer Vision Toolbox that contains
+algorithms, and functions for use in the development of computer vision, 3D
+vision, and video processing systems. By using the available algorithms in the
+toolbox, such as YOLOv2, and ACF, hand detection and gesture recognition is made
+possible in the platform \parencite{matlab}.
+
+\subsection{Applications}
+
+There have been multiple applications and products that utilize hand and finger
+tracking as their main component.
+
+\citeauthor{dorf2001} presents a use case for finger tracking in augmented
+environments. In the paper, interaction in a virtual environment through the use
+of gestures. The tracking system uses an optical marker-based motion capture
+system where the user wears a glove with retroreflective markers.
+
+\citeauthor{chiang2014} used a Kinect, a 3D sensing device by Microsoft that
+uses depth data, to track fingers to play virtual instruments. Virtual Pianos
+and Guitars were created and played with reliable and stable tracking.
+
+\citeauthor{yousaf2014} created a virtual keyboard that operates using finger
+tracking. The tracking uses the movement of the finger joints as the basis for
+selecting which key to press. A camera captures the movement, and the resulting
+video stream is used for hand region detection and finger joint localization.
+Using probabilistic regional density-based kernel tracking, finger joint
+trajectories are gathered. Feature vectors are then interpreted from the
+trajectories. These feature vectors are used in logic-based techniques and
+Dynamic Bayesian Network for classification, detection, and recognition of
+keystrokes.
+
+\section{Summary of the Research Gap}
+While there are a lot of applications and curriculum aimed at teaching touch
+typing, there is no automated system available that detects if a person uses the
+correct finger to press a key.
+
+By having this system, educators can accurately determine if and when a student
+is having a hard time typing and if these students will need an intervention to
+correct mistakes.
+
+This is also important because certain movements and hand positions will cause
+nerve and muscular disorders that will impact the user. By correcting these
+problematic movements and hand positions, these disorders can be prevented.
+
 \newpage
-\nocite{*} \printbibliography[heading=bibintoc,title={References}]{}
+\printbibliography[heading=bibintoc,title={References}]{}
 

### 2022-01-11 paper/paper.tex
methodology: add initial content

@@
 }
+
+% Misc
+\usepackage{gensymb}
+
+%Flowchart
+\usepackage{tikz}
+\usetikzlibrary{shapes,positioning,calc}
 
@@
 Capturing of the video to be analyzed by the program would be limited to a
-single 1080p webcam that is capturing in 60 frames per second. The camera will
+single 480p webcam that is capturing in 24 frames per second. The camera will
 be pointed downwards facing the keyboard and the hand. This means that the
@@
 \subsubsection{MediaPipe}
+\label{section:rrl-mediapipe}
 
@@
 
-\citeauthor{dorf2001} presents a use case for finger tracking in augmented
+\cite{dorf2001} presents a use case for finger tracking in augmented
 environments. In the paper, interaction in a virtual environment through the use
@@
 
-\citeauthor{chiang2014} used a Kinect, a 3D sensing device by Microsoft that
+\cite{chiang2014} used a Kinect, a 3D sensing device by Microsoft that
 uses depth data, to track fingers to play virtual instruments. Virtual Pianos
@@
 
-\citeauthor{yousaf2014} created a virtual keyboard that operates using finger
+\cite{yousaf2014} created a virtual keyboard that operates using finger
 tracking. The tracking uses the movement of the finger joints as the basis for
@@
 
+\chapter{Methodology}
+
+\section{Setup}
+\label{section:metho-setup}
+
+The experimental setup and configuration will be composed of three elements: the
+camera, the keyboard, and the environment.
+
+\subsection{Camera}
+The camera setup will use a single monocular camera positioned in a top-down
+view. The camera should capture the entirety of the keyboard and the movement of
+the ten fingers. Figure~\ref{fig:metho-keeb} shows what the camera will capture
+once placed in its correct position.
+
+The specific camera to be used will be a Logitech C270.
+Figure~\ref{fig:metho-cam} is a picture of this camera. This is a 0.9 mega pixel
+webcam with a CMOS sensor. This camera is capable of capturing color videos in
+720p/30fps and it has a diagonal field of view of 55\degree. This camera has a
+universal mounting clip that will reduce the difficulty in mounting the camera
+\parencite{logitech}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.4\textwidth]{webcam.png}
+	\caption{Logitech C270. Reprinted from \fullcite{logitech}}
+	\label{fig:metho-cam}
+	\centering
+\end{figure}
+
+\subsection{Keyboard}
+\label{section:metho-keeb}
+The keyboard will be a 60\% keyboard as shown in Figure~\ref{fig:metho-keeb}.
+This will limit the necessary mapping for the algorithm to the alphanumeric
+portion of the keyboard. This keyboard choice will also lessen the area that the
+camera will need to capture, as this keyboard type is considerably smaller
+compared to a full-size keyboard. The keyboard will have key caps and a case in
+a light color that will contrast the dark surface that it will be placed on.
+This is to improve initial keyboard detection. In addition, the keyboard layout
+will also be ANSI, due to the availability and widespread adoption of the layout
+in the Philippines.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{actual-keeb.png}
+	\caption{Keyboard, camera angle, and placement to be used for testing.}
+	\label{fig:metho-keeb}
+	\centering
+\end{figure}
+
+\subsection{Environment}
+The environment the setup would be placed in would be a well lit environment
+with a light source beside the camera. This light source will evenly light the
+keyboard and the fingers used for typing. The light source to be used will be a
+common LED desk lamp rated at 9 Watts with 700 lumens. This light will be white
+with a color temperature of 6500k.
+
+In addition, the surface where the keyboard is to be placed on will be solid
+green without any variation of color. This is to improve initial keyboard
+detection.
+
+\section{Algorithm}
+\label{section:metho-algo}
+
+\subsection{Computer Vision based Keyboard Detection, and Mapping}
+\label{section:metho-algo-keyboard}
+
+A computer vision algorithm will be created as a starting point for detecting,
+and mapping the keyboard within a video. The flowchart of the algorithm is shown
+in Figure~\ref{fig:metho-key-algo}.
+
+\begin{figure}[H]
+	\centering
+	\begin{tikzpicture}[font=\small,thick]
+		\node[draw,
+			rounded rectangle,
+			minimum width=2.5cm,
+			minimum height=1cm] (start) {START};
+
+		\node[draw,
+			trapezium,
+			trapezium left angle = 65,
+			trapezium right angle = 115,
+			trapezium stretches,
+			below=of start,
+			minimum width=3.5cm,
+			minimum height=1cm
+		] (input) {Input Frame};
+
+		\node[draw,
+			below=of input,
+			minimum width=3.5cm,
+			minimum height=1cm] (bg-rem) {Background Removal};
+
+		\node[draw,
+			below=of bg-rem,
+			minimum width=3.5cm,
+			minimum height=1cm] (gray) {Convert to Grayscale};
+
+		\node[draw,
+			below=of gray,
+			minimum width=3.5cm,
+			minimum height=1cm] (sobel) {Edge Detection using Sobel Filter};
+
+		\node[draw,
+			below=of sobel,
+			minimum width=3.5cm,
+			minimum height=1cm] (contours) {Find Contours (Simple)};
+
+		\node[draw,
+			below=of contours,
+			minimum width=3.5cm,
+			minimum height=1cm] (map) {Overlay Virtual Keyboard Map using Contour Points};
+
+		\node[draw,
+			trapezium,
+			trapezium left angle = 65,
+			trapezium right angle = 115,
+			trapezium stretches,
+			below=of map,
+			minimum width=3.5cm,
+			minimum height=1cm
+		] (output) {Output Frame};
+
+		\node[draw,
+			below=of output,
+			rounded rectangle,
+			minimum width=2.5cm,
+			minimum height=1cm] (end) {END};
+
+		\draw[-latex] (start) edge (input)
+		(input) edge (bg-rem)
+		(bg-rem) edge (gray)
+		(gray) edge (sobel)
+		(sobel) edge (contours)
+		(contours) edge (map)
+		(map) edge (output)
+		(output) edge (end);
+
+	\end{tikzpicture}
+	\caption{Flowchart of Keyboard Detection and Mapping Algorithm}
+	\label{fig:metho-key-algo}
+\end{figure}
+
+\subsubsection{Background Removal}
+The background of the frame will be removed by using colors as the main metric
+for determining the foreground and the background. This is made easier by the
+setup as the setup will have a green background, with a light colored keyboard
+as a foreground.
+
+\subsubsection{Convert to Grayscale}
+The frame will then be converted to grayscale. This is done because future steps
+of the algorithm does not require color values and this improves the performance
+of these algorithms.
+
+\subsubsection{Edge Detection using Sobel Filter}
+A Sobel Filter is a edge detection filter that uses a 3$times$3 kernel that is
+convolved twice. Once horizontally, and another vertically to produce a
+grayscale image of the outlines within the frame.
+
+\subsubsection{Find Contours (Simple)}
+Contours are curves joining all continuous points that have the same color or
+intensity \parencite{opencv-contours}. In this algorithm, an OpenCV function is
+to be used to find the contours of the outlines of the object found using the
+previous step. This OpenCV function implements the algorithm of \cite{contours}.
+The expected result is four contour points which are positioned on the four
+edges of the keyboard.
+
+\subsubsection{Overlay Virtual Keyboard Map using Contour Points}
+The virtual keyboard map is a rectangular image that has individual region of
+interest (ROI) for each key in a 60\% \ac{ansi} keyboard. Each key has a
+corresponding color assigned to it and this color fills the region the key is
+assigned to.
+
+The virtual map is then stretched over the object, with the four contour points
+as the four edges of the virtual keyboard map. The image generated is then
+returned as the final output of the graph.
+
+\subsection{Computer Vision based Finger Detection, and Tracking}
+\label{section:metho-algo-finger}
+
+A computer vision algorithm will be chosen as a starting point for detecting,
+and tracking fingers within a video.
+
+The first component this algorithm will have is the ability to detect fingers
+within a video. This would involve capturing exactly where the finger is in the
+current frame. This data would then be used for the next component of the
+algorithm: tracking. This component would track the movement of the finger as it
+performs a typing motion. This algorithm should have the ability to separate the
+different fingers and track the movement of the ten fingers individually.
+
+The specific algorithm used for finger detection and tracking will be the
+solution offered by MediaPipe. More details about the specific algoritihm can be
+found in Section~\ref{section:rrl-mediapipe}.
+
+\subsection{Integration for finger-key identification and mapping}
+
+The two previously chosen algorithms will be combined to accomplish finger-key
+identification and mapping. Finger-key identification refers to identifying
+which finger is used to press which key. As an example, the key Q was pressed
+with the pinky finger of the left hand.
+
+The integration of the algorithm is a two step process. The first step is to get
+the mapping of the keyboard using the algorithm shown in Section~\ref{section:metho-algo-keyboard}.
+
+The second step is a continous loop where the mapping of the keyboard is used in
+conjunction with the finger tracking algorithm chosen in
+Section~\ref{section:metho-algo-finger}. The flowchart of the algorithm is shown
+in Figure~\ref{metho-algo-integration}
+
+\begin{figure}[H]
+	\centering
+	\begin{tikzpicture}[font=\small,thick]
+		\node[draw,
+			rounded rectangle,
+			minimum width=2.5cm,
+			minimum height=1cm] (start) {START};
+
+		\node[draw,
+			trapezium,
+			trapezium left angle = 65,
+			trapezium right angle = 115,
+			trapezium stretches,
+			below left=of start,
+			minimum width=3.5cm,
+			minimum height=1cm
+		] (map) {Virtual Keyboard Map};
+
+		\node[draw,
+			trapezium,
+			trapezium left angle = 65,
+			trapezium right angle = 115,
+			trapezium stretches,
+			below right=of start,
+			minimum width=3.5cm,
+			minimum height=1cm
+		] (track) {Finger Tracking Frame};
+
+		\coordinate (CENTER) at ($(map)!0.5!(track)$);
+
+		\node[draw,
+			diamond,
+			aspect=2,
+			below = 3cm of CENTER,
+			minimum width=2.5cm,
+			inner sep=0] (if-packet) {Keypress Packet Received};
+
+		\node[draw,
+			below left=of if-packet,
+			minimum width=3.5cm,
+			minimum height=1cm] (find-roi) {Find ROI of the pressed key in Virtual Keyboard Map};
+
+		\node[draw,
+			below=of find-roi,
+			minimum width=3.5cm,
+			minimum height=1cm] (find-finger) {Find which fingertip is on that ROI};
+
+		\node[draw,
+			trapezium,
+			trapezium left angle = 65,
+			trapezium right angle = 115,
+			trapezium stretches,
+			below=of find-finger,
+			minimum width=3.5cm,
+			minimum height=1cm
+		] (return) {Return which fingertip};
+
+		\node[draw,
+			trapezium,
+			trapezium left angle = 65,
+			trapezium right angle = 115,
+			trapezium stretches,
+			below = 11cm of CENTER,
+			minimum width=3.5cm,
+			minimum height=1cm
+		] (output) {Output Frame};
+
+		\node[draw,
+			below=of output,
+			rounded rectangle,
+			minimum width=2.5cm,
+			minimum height=1cm] (end) {END};
+
+		\draw[-latex] (start) edge (map)
+		(start) edge (track)
+		(track) edge (if-packet)
+		(map) edge (if-packet)
+		(find-roi) edge (find-finger)
+		(find-finger) edge (return)
+		(return) edge (output)
+		(output) edge (end);
+
+		\draw[-latex] (if-packet) -| (find-roi)
+		node[pos=0.25,fill=white,inner sep=0]{true};
+
+		\draw[-latex] (if-packet) edge node[pos=0.4,fill=white,inner sep=2pt]{No}(output);
+
+	\end{tikzpicture}
+	\caption{Flowchart of the overall flow}
+	\label{fig:metho-algo-integration}
+\end{figure}
+
+\subsubsection{Find ROI of the pressend key in Virtual Keyboard Map}
+Since there is a known color-key mapping based on Virtual Keyboard Map, the
+algorithm will find the correct color to search for based on the received keypress
+packet. This color is isolated from the Virtual Keyboard Map, and the region of
+interest and its coordinates are obtained.
+
+\subsubsection{Find which fingertip is on that ROI}
+The Finger Tracking Frame contains the pixel positions of each landmark of the
+hand. For this step, the algorithm will find the landmark which is positioned
+within the region of interest from the previous step.
+
+\subsubsection{Return which fingertip}
+The finger containing the landmark will be returned. This finger will then be
+compared with the expected finger to be used to press the key to determine if it
+is correct or not.
+
+\section{Trainer}
+
+\subsection{Create typing test sequences}
+Typing test sequences are strings that will be used in testing the user in their
+ability to type. The test sequences that will be used in the trainer will come
+from text found in the public domain obtained from Project Gutenberg and the
+Internet Archive. Sentences from these text will be isolated and used as the
+test sequences. There will be a total of 20 test sequences that will be
+gathered. An example test sequence is ``What of it, if some old hunks of a
+sea-captain orders me to get a broom and sweep down the decks?'' from Moby Dick
+by \cite{moby-dick}. The 19 other test sequences can be found in the appendix.
+
+\subsection{UX/UI design and Front-end development for the touch typing trainer}
+The touch typing trainer will have its UX/UI design created in Figma. This
+design will be based on other type tests such as Monkeytype
+\parencite{bartnik2021}, and Keybr \parencite{keybr}, with additional components
+for showing real-time finger-key identification and mapping.
+Figure~\ref{fig:ui-test} and \ref{fig:ui-stat} illustrates the two main pages of
+the design.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{ui-test.png}
+	\caption{Test screen of the front-end design}
+	\label{fig:ui-test}
+	\centering
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{ui-stats.png}
+	\caption{Statistics screen of the front-end design}
+	\label{fig:ui-stat}
+	\centering
+\end{figure}
+
+The front-end will be web based, and the framework the front-end will be
+developed in will be Svelte, due to its community resources, performance, and
+the researcher's familiarity with the framework.
+
+\subsection{Back-end development of a touch typing trainer using the program created
+	in \ref{section:metho-algo} and the experimental setup identified in
+	\ref{section:metho-setup}}
+\subsection{Testing of finger-key identification and mapping accuracy}
+\section{Metric}
+%% 4. Finger Placement Keyboard Typing Metric Development
+%%     1. Determine if the metric is for individual key presses or for the whole typing sequence or both.
+%%     2. Evaluate which variables need to be taken into account for the metric depending on (5.a):
+%%         1. Actual Finger Used
+%%         2. Correct Finger to be Used
+%%         3. Specific key pressed
+%%         4. Number of keys pressed with the correct fingers
+%%         5. Number of keys pressed with the wrong fingers
+
 \newpage

### 2022-01-12 paper/paper.tex
methodology: expound hand tracking

@@
 A computer vision algorithm will be chosen as a starting point for detecting,
-and tracking fingers within a video.
-
-The first component this algorithm will have is the ability to detect fingers
-within a video. This would involve capturing exactly where the finger is in the
-current frame. This data would then be used for the next component of the
-algorithm: tracking. This component would track the movement of the finger as it
-performs a typing motion. This algorithm should have the ability to separate the
-different fingers and track the movement of the ten fingers individually.
-
-The specific algorithm used for finger detection and tracking will be the
-solution offered by MediaPipe. More details about the specific algoritihm can be
-found in Section~\ref{section:rrl-mediapipe}.
+and tracking fingers within a video. The specific algorithm used for finger
+detection and tracking will be the solution offered by MediaPipe dubbed
+MediaPipe Hands. This algorithm is composed of multiple ML models working in
+conjunction to be able to detect the different parts of the hands and track them
+accurately.
+
+\subsubsection{Palm Detection Model}
+The first model, the Palm Detection Model detects the initial hand locations
+using a single shot detector model based on the paper by \cite{ssd}. This
+model achieves an average precision of 95.7\% in palm detection.
+
+MediaPipe Hands detects the palms first, instead of the whole hands with one
+model because the hands lack high contrast patterns. This prevents the model to
+detect the hand accurately. In addition, detecting a palm is simpler compared to
+detecting hands with articulated fingers since estimating a bounding box around
+a rigid objects, i.e.\ a palm, is much simpler. Furthermore, a palm can be
+modeled using only square anchors reducing the number of anchors by a factor of
+3--5.
+
+\subsubsection{Hand Landmark Model}
+After the palms have been detected and an appropriate anchor has been
+established, the Hand Landmark Model pinpoints 21 3D hand-knuckle coordinates
+inside the detected hand. This is done using direct coordinate prediction.
+
+This model was trained using 30,000 manually annotated, real-world images with
+21 3D hand-knuckle coordinates. Using this information, the model can also
+accurately add landmarks to partially visible hands and hands with
+self-occlusion. This is also made possible by the model's consistent internal hand
+pose representation.
 
@@
 
-\subsection{Create typing test sequences}
+\subsection{Typing Test Sequences}
 Typing test sequences are strings that will be used in testing the user in their
@@
 from text found in the public domain obtained from Project Gutenberg and the
-Internet Archive. Sentences from these text will be isolated and used as the
-test sequences. There will be a total of 20 test sequences that will be
-gathered. An example test sequence is ``What of it, if some old hunks of a
-sea-captain orders me to get a broom and sweep down the decks?'' from Moby Dick
-by \cite{moby-dick}. The 19 other test sequences can be found in the appendix.
+Internet Archive. Sentences will be isolated from these text and used as test
+sequences if they fit a criteria.
+
+The criteria for choosing test sequences is as follows: (1) $\approx80\%$ of the
+characters in the keyboard is present in a test sequence. (2) The number of
+words in a test sequence do not exceed 25. (3) Numbers and punctuations should
+be present in at least $\approx30\%$ of the total test sequences.
+
+There will be a total of 10 test sequences that will be gathered. An example
+test sequence is ``What of it, if some old hunks of a sea-captain orders me to
+get a broom and sweep down the decks?'' from Moby Dick by \cite{moby-dick}. The
+9 other test sequences can be found in the appendix.
 

### 2022-01-13 paper/paper.tex
methodology: add new ui and more content

@@
 \usepackage{gensymb}
+\usepackage{amsmath}
 
@@
 	\label{fig:metho-keeb}
-	\centering
 \end{figure}
@@
 and mapping the keyboard within a video. The flowchart of the algorithm is shown
-in Figure~\ref{fig:metho-key-algo}.
+in Figure~\ref{fig:metho-algo-key-flow}.
 
@@
 			minimum width=3.5cm,
-			minimum height=1cm] (bg-rem) {Background Removal};
-
-		\node[draw,
-			below=of bg-rem,
-			minimum width=3.5cm,
 			minimum height=1cm] (gray) {Convert to Grayscale};
@@
 		\draw[-latex] (start) edge (input)
-		(input) edge (bg-rem)
-		(bg-rem) edge (gray)
+		(input) edge (gray)
 		(gray) edge (sobel)
@@
 	\caption{Flowchart of Keyboard Detection and Mapping Algorithm}
-	\label{fig:metho-key-algo}
-\end{figure}
-
-\subsubsection{Background Removal}
-The background of the frame will be removed by using colors as the main metric
-for determining the foreground and the background. This is made easier by the
-setup as the setup will have a green background, with a light colored keyboard
-as a foreground.
+	\label{fig:metho-algo-key-flow}
+\end{figure}
 
 \subsubsection{Convert to Grayscale}
-The frame will then be converted to grayscale. This is done because future steps
-of the algorithm does not require color values and this improves the performance
-of these algorithms.
+The frame will then be converted to a 256 level grayscale image. This step will
+be performed because future steps of the algorithm will not require color values
+to work. In addition, the performance of these steps will be improved.
 
 \subsubsection{Edge Detection using Sobel Filter}
-A Sobel Filter is a edge detection filter that uses a 3$times$3 kernel that is
+A Sobel Filter is an edge detection filter that uses a 3$\times$3 kernel that is
 convolved twice. Once horizontally, and another vertically to produce a
-grayscale image of the outlines within the frame.
+grayscale image of the outlines within the frame. The kernels used by the Sobel
+Filter \parencite{sobel2014} is shown in Figure~\ref{fig:metho-algo-key-sobel}.
+This filter will be used in the algorithm to obtain the four edges of the
+keyboard.
+
+\begin{figure}[H]
+	\centering
+  $\begin{bmatrix}
+  +1 & 0 & -1\\
+  +2 & 0 & -2\\
+  +1 & 0 & -1
+  \end{bmatrix}$
+	and
+  $\begin{bmatrix}
+  +1 & +2 & +1\\
+  0 & 0 & 0\\
+  -1 & -2 & -1
+  \end{bmatrix}$
+	\caption{Sobel Operator Kernels. Reproduced from \fullcite{sobel2014}}
+	\label{fig:metho-algo-key-sobel}
+\end{figure}
 
 \subsubsection{Find Contours (Simple)}
+\label{section:metho-algo-key-contours}
 Contours are curves joining all continuous points that have the same color or
@@
 \subsubsection{Overlay Virtual Keyboard Map using Contour Points}
-The virtual keyboard map is a rectangular image that has individual region of
-interest (ROI) for each key in a 60\% \ac{ansi} keyboard. Each key has a
-corresponding color assigned to it and this color fills the region the key is
-assigned to.
-
-The virtual map is then stretched over the object, with the four contour points
-as the four edges of the virtual keyboard map. The image generated is then
-returned as the final output of the graph.
+The virtual keyboard map is a rectangular image that contains individual region
+of interest (ROI) for each key in a 60\% \ac{ansi} keyboard. Each key has a
+corresponding color assigned to it and this color fills the region where this
+key is located at. Figure~\ref{fig:metho-algo-key-map} is the virtual keyboard
+map.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.6\textwidth]{image-map.png}
+	\caption{Initial Image Map}
+	\label{fig:metho-algo-key-map}
+	\centering
+\end{figure}
+
+The virtual map will then be stretched over the object, with the four contour
+points as the four edges of the virtual keyboard map. The image generated will
+then be returned as the final output of the graph.
+Figure~\ref{fig:metho-algo-key-overlay} illustrates this image, and
+Figure~\ref{fig:metho-algo-key-half} shows how accurate the image map is when
+overlaid over the image.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.6\textwidth]{image-map-final.png}
+	\caption{Expected output of the algorithm}
+	\label{fig:metho-algo-key-overlay}
+	\centering
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.6\textwidth]{image-map-half.png}
+	\caption{Overlaid image map shown at 60\% opacity}
+	\label{fig:metho-algo-key-half}
+	\centering
+\end{figure}
 
@@
 detection and tracking will be the solution offered by MediaPipe dubbed
-MediaPipe Hands. This algorithm is composed of multiple ML models working in
+MediaPipe Hands. This algorithm is composed of two ML models working in
 conjunction to be able to detect the different parts of the hands and track them
@@
 The first model, the Palm Detection Model detects the initial hand locations
-using a single shot detector model based on the paper by \cite{ssd}. This
-model achieves an average precision of 95.7\% in palm detection.
+using a single shot detector model based on the paper by \cite{ssd}. This model
+achieves an average precision of 95.7\% in palm detection
+\parencite{mediapipe-hands}.
 
@@
 modeled using only square anchors reducing the number of anchors by a factor of
-3--5.
+3--5 \parencite{mediapipe-hands}.
 
@@
 accurately add landmarks to partially visible hands and hands with
-self-occlusion. This is also made possible by the model's consistent internal hand
-pose representation.
+self-occlusion. This is also made possible by the model's consistent internal
+hand pose representation \parencite{mediapipe-hands}.
 
@@
 The integration of the algorithm is a two step process. The first step is to get
-the mapping of the keyboard using the algorithm shown in Section~\ref{section:metho-algo-keyboard}.
-
-The second step is a continous loop where the mapping of the keyboard is used in
-conjunction with the finger tracking algorithm chosen in
+the mapping of the keyboard using the algorithm shown in
+Section~\ref{section:metho-algo-keyboard}.
+
+The second step is a continuous loop where the mapping of the keyboard is used
+in conjunction with the finger tracking algorithm chosen in
 Section~\ref{section:metho-algo-finger}. The flowchart of the algorithm is shown
-in Figure~\ref{metho-algo-integration}
+in Figure~\ref{fig:metho-algo-integration}
 
@@
 \subsubsection{Find ROI of the pressend key in Virtual Keyboard Map}
-Since there is a known color-key mapping based on Virtual Keyboard Map, the
-algorithm will find the correct color to search for based on the received keypress
-packet. This color is isolated from the Virtual Keyboard Map, and the region of
-interest and its coordinates are obtained.
+Since there is a known color-key mapping based on the Virtual Keyboard Map, the
+algorithm will find the correct color to search for based on the received
+keypress packet. This color is isolated from the Virtual Keyboard Map, and the
+region of interest and its coordinates are obtained. The edges of this ROI will
+then be obtained by finding its contours. The specific implementation if this is
+explained in Section~\ref{section:metho-algo-key-contours}.
+
+The coordinate system will have its origin at the top left, starting at 0, 0.
+Going to the right increases the value of the x-axis, and going to the bottom
+increases the value of the y-axis. This will result in a coordinate system that
+only has positive values with each value corresponding to a pixel.
+
+The expected output of this is four sets of pairs. Each pair corresponds to an
+edge of the region of interest.
 
 \subsubsection{Find which fingertip is on that ROI}
-The Finger Tracking Frame contains the pixel positions of each landmark of the
-hand. For this step, the algorithm will find the landmark which is positioned
-within the region of interest from the previous step.
+The Finger Tracking Frame will contain the pixel positions of each landmark of
+the hand. For this step, the algorithm will find the landmark which is
+positioned within the region of interest obtained from the previous step. This
+will be done using a series of checks:
 
 \subsubsection{Return which fingertip}
-The finger containing the landmark will be returned. This finger will then be
-compared with the expected finger to be used to press the key to determine if it
-is correct or not.
+The landmark will then be used to determine which specific finger corresponds to
+the keypress. Figure~\ref{fig:metho-algo-integration-landmarks} shows each
+possible landmark that may be returned from the previous step. This step will
+then simply return the name of the landmark, up until the first underscore. As
+an example, if the landmark found is \emph{MIDDLE\_FINGER\_DIP}, this step will
+return \emph{MIDDLE} denoting that the middle finger is the finger that
+corresponds with the keypress.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.6\textwidth]{hand-landmarks.png}
+	\caption{Hand landmarks that may be returned by the algorithm. Reproduced from \fullcite{mediapipe-hands}}.
+	\label{fig:metho-algo-integration-landmarks}
+	\centering
+\end{figure}
 
@@
 Figure~\ref{fig:ui-test} and \ref{fig:ui-stat} illustrates the two main pages of
-the design.
+the design. Additional screens, such as historical statistics, in-platform
+tutorial, and a written guide are shown in the appendix.
 

### 2022-01-26 paper/paper.tex
methodology: expound on algorithm, add backend and testing

@@
 \usetikzlibrary{shapes,positioning,calc}
+
+% Adjust Figure Spacing
+\setlength\intextsep{1cm}
 
@@
 The experimental setup and configuration will be composed of three elements: the
-camera, the keyboard, and the environment.
+camera, the keyboard, and the environment. Figure~\ref{fig:metho-setup-cam}
+shows the sketch of the complete experimental setup.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.4\textwidth]{webcam.png}
+	\caption{Sketch of the experimental setup}
+	\label{fig:metho-setup-cam}
+	\centering
+\end{figure}
 
@@
 The camera setup will use a single monocular camera positioned in a top-down
-view. The camera should capture the entirety of the keyboard and the movement of
-the ten fingers. Figure~\ref{fig:metho-keeb} shows what the camera will capture
-once placed in its correct position.
-
-The specific camera to be used will be a Logitech C270.
-Figure~\ref{fig:metho-cam} is a picture of this camera. This is a 0.9 mega pixel
-webcam with a CMOS sensor. This camera is capable of capturing color videos in
-720p/30fps and it has a diagonal field of view of 55\degree. This camera has a
-universal mounting clip that will reduce the difficulty in mounting the camera
-\parencite{logitech}.
+view. The camera will capture the entirety of the keyboard and the movement of
+the ten fingers. To do so, it will be mounted on top of the monitor and the
+camera will point down towards the table. Figure~\ref{fig:metho-setup-keeb}
+illustrates what the camera will capture once placed in its correct position.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{actual-keeb.png}
+	\caption{Keyboard, camera angle, and placement to be used for testing.}
+	\label{fig:metho-setup-keeb}
+\end{figure}
+
+The specific camera to be used will be a Logitech C920.
+Figure~\ref{fig:metho-setup-cam} is a picture of this camera. It is a 3 mega pixel
+webcam that is capable of capturing color video in 1080p/30fps and 720p/30fps
+with a diagonal field of view of 78\degree. This camera has a universal mounting
+clip that will allow the camera to be correctly positioned within the
+experimental setup \parencite{logitech}.
 
@@
 	\includegraphics[width=0.4\textwidth]{webcam.png}
-	\caption{Logitech C270. Reprinted from \fullcite{logitech}}
-	\label{fig:metho-cam}
+	\caption{Logitech C920. Reprinted from \fullcite{logitech}}
+	\label{fig:metho-setup-cam}
 	\centering
@@
 \label{section:metho-keeb}
-The keyboard will be a 60\% keyboard as shown in Figure~\ref{fig:metho-keeb}.
-This will limit the necessary mapping for the algorithm to the alphanumeric
-portion of the keyboard. This keyboard choice will also lessen the area that the
-camera will need to capture, as this keyboard type is considerably smaller
-compared to a full-size keyboard. The keyboard will have key caps and a case in
-a light color that will contrast the dark surface that it will be placed on.
-This is to improve initial keyboard detection. In addition, the keyboard layout
-will also be ANSI, due to the availability and widespread adoption of the layout
-in the Philippines.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{actual-keeb.png}
-	\caption{Keyboard, camera angle, and placement to be used for testing.}
-	\label{fig:metho-keeb}
-\end{figure}
+
+The keyboard will be a 60\% keyboard as shown in
+Figure~\ref{fig:metho-setup-keeb}. This will limit the necessary mapping for the
+algorithm to the alphanumeric portion of the keyboard. This keyboard choice will
+also lessen the area that the camera will need to capture, as this keyboard type
+is considerably smaller compared to a full-size keyboard. The keyboard will have
+key caps and a case in a light color that will contrast the dark surface that it
+will be placed on. This is to improve initial keyboard detection. In addition,
+the keyboard layout will also be ANSI, due to the availability and widespread
+adoption of the layout in the Philippines.
 
@@
 \subsubsection{Convert to Grayscale}
-The frame will then be converted to a 256 level grayscale image. This step will
-be performed because future steps of the algorithm will not require color values
-to work. In addition, the performance of these steps will be improved.
+The frame will be converted to a 256 level grayscale image. This step will be
+performed because future steps of the algorithm will not require color values to
+work. In addition, the performance of the following steps will be also improved
+as the number of dimensions to be analyzed is reduced.
 
@@
 Filter \parencite{sobel2014} is shown in Figure~\ref{fig:metho-algo-key-sobel}.
-This filter will be used in the algorithm to obtain the four edges of the
-keyboard.
 
@@
 
+The expected result of this step is a black and white image that highlights the
+edges of the keyboard in white.
+
 \subsubsection{Find Contours (Simple)}
@@
 Contours are curves joining all continuous points that have the same color or
-intensity \parencite{opencv-contours}. In this algorithm, an OpenCV function is
-to be used to find the contours of the outlines of the object found using the
-previous step. This OpenCV function implements the algorithm of \cite{contours}.
-The expected result is four contour points which are positioned on the four
-edges of the keyboard.
+intensity \parencite{opencv-contours}. In this algorithm the OpenCV function
+\texttt{findContours} will be used to find the contours of the outlines of the
+object found using the previous step. This function accepts a contour
+approximation method as one of its arguments. Two methods are provided by
+OpenCV, \texttt{CHAIN\_APPROX\_NONE} and \texttt{CHAIN\_APPORX\_SIMPLE}. The
+former returns all contours in the shape, while the latter removes redundant
+points and returns the least amount of points that describes the shape. The
+algorithm will use the latter, as only the extreme edges of the keyboard needs
+to be detected. This OpenCV function implements the algorithm of
+\cite{contours} in their paper ``\citefield{contours}{title}''.
+
+The expected result of this algorithm is four contour points which are
+positioned on the four edges of the keyboard.
 
@@
 	\centering
-	\includegraphics[width=0.6\textwidth]{image-map-final.png}
+	\includegraphics[width=1\textwidth]{image-map-final.png}
 	\caption{Expected output of the algorithm}
@@
 
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.6\textwidth]{image-map-half.png}
+\hfill
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1\textwidth]{image-map-half.png}
 	\caption{Overlaid image map shown at 60\% opacity}
@@
 \label{section:metho-algo-finger}
-
 A computer vision algorithm will be chosen as a starting point for detecting,
@@
 MediaPipe Hands detects the palms first, instead of the whole hands with one
-model because the hands lack high contrast patterns. This prevents the model to
-detect the hand accurately. In addition, detecting a palm is simpler compared to
-detecting hands with articulated fingers since estimating a bounding box around
-a rigid objects, i.e.\ a palm, is much simpler. Furthermore, a palm can be
-modeled using only square anchors reducing the number of anchors by a factor of
-3--5 \parencite{mediapipe-hands}.
+model because the hands lack high contrast patterns. This reduces the model's
+ability to detect the hand with accuracy. In addition, detecting a palm is
+simpler compared to detecting hands with articulated fingers since estimating a
+bounding box around a rigid objects, i.e.\ a palm, is much simpler. Furthermore,
+a palm can be modeled using only square anchors reducing the number of anchors
+by a factor of 3--5 \parencite{mediapipe-hands}.
 
@@
 region of interest and its coordinates are obtained. The edges of this ROI will
-then be obtained by finding its contours. The specific implementation if this is
+then be obtained by finding its contours. The specific implementation of this is
 explained in Section~\ref{section:metho-algo-key-contours}.
@@
 positioned within the region of interest obtained from the previous step. This
-will be done using a series of checks:
+will be done using a series of checks.
+
+Each landmark's coordinates will be compared to the coordinates of the edges of
+the ROI. A landmark will be determined as positioned within the ROI if all of the
+following conditions are true:
+\begin{enumerate}
+	\item In the X axis, the landmark's coordinates is greater than one or both of
+		the two coordinates found of the left side of the ROI
+	\item In the X axis, the landmark's coordinates is less than one or both of
+		the two coordinates found of the right side of the ROI
+	\item In the Y axis, the landmark's coordinates is greater than one or both of
+		the two coordinates found of the top side of the ROI
+	\item In the Y axis, the landmark's coordinates is less than one or both of
+		the two coordinates found of the bottom side of the ROI
+\end{enumerate}
+
+These conditions maximize the total area of the ROI and is not strict about
+exact accuracy. In essence, these conditions will create a perfectly rectangular
+box that contains the quadrilateral formed by the four points from the previous
+step.
 
@@
 possible landmark that may be returned from the previous step. This step will
-then simply return the name of the landmark, up until the first underscore. As
-an example, if the landmark found is \emph{MIDDLE\_FINGER\_DIP}, this step will
-return \emph{MIDDLE} denoting that the middle finger is the finger that
-corresponds with the keypress.
+then return the name of the landmark, up until the first underscore. As an
+example, if the landmark found is \texttt{MIDDLE\_FINGER\_TIP}, this step will
+return \texttt{MIDDLE} denoting that the middle finger is the finger that
+corresponds with the keypress. In addition, the specific hand will also be
+returned as one of two strings, \texttt{Left} and \texttt{Right}, since
+this information is also bundled together with the landmarks.
 
@@
 \end{figure}
+
+\subsection{Implementation}
+\label{section:metho-algo-implementation}
+The previously discussed algorithms will be implemented using Python. The
+Keyboard Detection, and Mapping Algorithm in
+Section~\ref{section:metho-algo-keyboard} will be implemented as a separate
+submodule in Python using OpenCV.\@The Finger Detection, and Tracking Solution
+in Section~\ref{section:metho-algo-finger} will be consumed using the prebuilt
+Python package offered by the MediaPipe team. Finally, the integration of the
+two will be done as another separate submodule in Python.
 
@@
 \subsection{UX/UI design and Front-end development for the touch typing trainer}
-The touch typing trainer will have its UX/UI design created in Figma. This
-design will be based on other type tests such as Monkeytype
+\subsubsection{Design}
+This design will be based on other type tests such as Monkeytype
 \parencite{bartnik2021}, and Keybr \parencite{keybr}, with additional components
@@
 
+Figure~\ref{fig:ui-test} has four main components. From the top, the first
+component is the navigation bar. This directs the user to the different pages
+within the website. The second component in the upper left displays the number
+of remaining words to be typed. This is 16 in the example. Below that is the
+test sequence. This shows the user what characters to type, and highlights
+mistakes in two distinct colors. Red if the character pressed was incorrect and
+purple if the character typed was correct, but the finger used to type it is
+wrong. The last main component in this screen is the keyboard that illustrates
+the correct finger positioning, with each finger corresponding to a different
+color.
+
 \begin{figure}[H]
@@
 
-The front-end will be web based, and the framework the front-end will be
-developed in will be Svelte, due to its community resources, performance, and
-the researcher's familiarity with the framework.
+Figure~\ref{fig:ui-stat} has two main components. The first component at the top
+shows the three major statistics for the typing test: Words per minute,
+Accuracy, and Finger placement accuracy. Finger placement accuracy will be
+explained in Section~\ref{label:metric}. The component below the statistics is a
+graph plotting the major statistics of the user in time. This includes when the
+user made the mistake and when the user slowed down, or increased their speed.
+
+\subsubsection{Implementation}
+The design will be created in Figma and the front-end will be web based.The
+framework the front-end will be developed in will be Svelte, due to its
+community resources, performance, and the researcher's familiarity with the
+framework. The front-end will serve as a display for the information and the
+interface which the user interacts and inputs information into the trainer. No
+calculation will be done within the front-end.
 
@@
 	\ref{section:metho-setup}}
+
+The back-end will handle all of the computation for the trainer---mainly
+finger-key identification. The backend will also handle authentication and data
+storage.
+
+The back-end will expose a REST API. This is a type of application programming
+interface (API) that conforms to the REST architectural style. One key
+characteristic of the this style is its statelessness and cacheable data
+\parencite{rest}. This will be implemented using Django, a Python REST API
+framework.
+
+\subsubsection{Finger-key Identification}
+The submodule that integrates all of the sub algorithms will be called within
+the framework. The framework will then respond to requests for finger-key
+identification using this submodule. This will allow the back-end to expose its
+finger-key identification abilities and allow the front-end to use this data for
+the trainer.
+
+\subsubsection{Authentication}
+The core authentication system will rely on JSON Web Tokens (JWTs) . According
+to \cite{jwt}, ``JSON Web Token (JWT) is a compact, URL-safe means of
+representing claims to be transferred between two parties.'' This allows for a
+user to claim that they are that specific user securely as the claim is
+cryptographically signed. For the backend of the trainer, the JWT will be signed
+using a secret and it will be encoded in HMAC SHA256.
+
+\subsubsection{Data Storage}
+The trainer will store all of the data in a relational database using
+PostgreSQL.\@This specific technology was chosen due to its maturity, community,
+and industry recommendations and the researchers' familiarity with the
+technology.
+
 \subsection{Testing of finger-key identification and mapping accuracy}
+
+There will be 15 videos of the researcher performing the typing test sequences.
+The researcher will then manually perform finger-key identification for all key
+presses present in the video. The finger-key identification submodule will then
+perform the same process for the same videos. The accuracy of the submodule will
+then be obtained by identifying which of its identification is incorrect, based
+on the manual identification.
+
 \section{Metric}
+\label{section:metric}
+
+There will be two total metrics obtained pertaining to finger and key mapping.
+The first will be per test sequence, and the second will be per key.
+
 %% 4. Finger Placement Keyboard Typing Metric Development

### 2022-01-28 paper/paper.tex
methodology: add metrics and backend

@@
 \hypersetup{
-	pdftitle={Development of a touch typing trainer with an emphasis on finger and wrist placement},
+	pdftitle={Development of a touch typing trainer with an emphasis on finger-key identification},
 	pdfstartview={FitH},
@@
 
-	\textsc{Development of a touch typing trainer\\with an emphasis on finger and wrist positions}
+	\textsc{Development of a touch typing trainer\\with an emphasis on finger-key identification }
 	\par\vspace{0.5cm}
@@
 
-	October 2021
+	February 2022
 	\vfill
@@
 The experimental setup and configuration will be composed of three elements: the
-camera, the keyboard, and the environment. Figure~\ref{fig:metho-setup-cam}
+camera, the keyboard, and the environment. Figure~\ref{fig:metho-setup}
 shows the sketch of the complete experimental setup.
@@
 	\centering
-	\includegraphics[width=0.4\textwidth]{webcam.png}
+	\includegraphics[width=0.4\textwidth]{setup.png}
 	\caption{Sketch of the experimental setup}
-	\label{fig:metho-setup-cam}
+	\label{fig:metho-setup}
 	\centering
@@
 perform the same process for the same videos. The accuracy of the submodule will
-then be obtained by identifying which of its identification is incorrect, based
-on the manual identification.
+then be obtained by identifying if its identification is as expected, based on
+the manual identification.
 
@@
 The first will be per test sequence, and the second will be per key.
+
+\subsection{Per Test Sequence}
+The metric to be calculated per test sequence is Finger Placement Accuracy (FP
+ACC). This metric will compute the percentage of accurate key presses pressed
+with the correct finger over the length of the typing test sequence. Inputting
+the wrong character, even if pressed with the correct finger will reduce FP ACC
+since FP ACC measures \emph{accurate key presses}, and incorrect characters will
+be considered as inaccurate key presses.
+
+The equation for calculating FP ACC is as follows:
+
+\begin{equation}
+	FP ACC = \frac{|F|}{|T|} \cdot 100\%
+\end{equation}
+
+Where $|F|$ refers to the number of accurate key presses pressed with the
+correct finger and $|T|$ refers to the length of the text.
+
+\subsection{Per Key}
+This metric will compute the Historical Finger Placement Accuracy (HFP ACC) for
+a certain key. This will include the times the user pressed the wrong key when
+asked for a certain character, the same as in FP ACC. This will allow the user
+to easily verify which keys need attention and retraining if there is a high
+frequency of error.
+
+The equation for calculating a key's HFP ACC is as follows:
+
+\begin{equation}
+	HFP ACC = \frac{|F_{char}|}{|C_{char}|} \cdot 100\%
+\end{equation}
+
+Where $|F_{char}|$ refers to the number of accurate key presses pressed with the
+correct finger for a certain character, identified as $char$. $|C_{char}|$
+refers to the number of times the character has appeared in all test sequences
+for the user.
 

### 2022-02-07 paper/paper.tex
proposal: restructure paper and proofread content

--- 2022-01-26 paper/paper.tex
@@
 \documentclass{report}
+%%% LT-SKIP-BEGIN
 
@@
 	long=Computer Vision
+}
+
+\DeclareAcronym{roi}{
+	short=ROI,
+	long=Region of Interest
+}
+
+\DeclareAcronym{cts}{
+	short=CTS,
+	long=Carpal Tunnel Syndrome
+}
+
+\DeclareAcronym{jwt}{
+	short=JWT,
+	long=JSON Web Tokens
+}
+
+\DeclareAcronym{fpacc}{
+	short=FP ACC,
+	long=Finger Placement Accuracy
+}
+
+\DeclareAcronym{hfpacc}{
+	short=HFP ACC,
+	long=Historical Finger Placement Accuracy
 }
@@
 
+%Tables
+\usepackage{booktabs}
+\usepackage{supertabular}
+
 % Adjust Figure Spacing
 \setlength\intextsep{1cm}
+
+%%% LT-SKIP-END
 
@@
 
-	\textsc{Development of a touch typing trainer\\with an emphasis on finger and wrist positions}
+	\textsc{Development of a touch typing trainer\\with an emphasis on finger-key identification }
 	\par\vspace{0.5cm}
 
-	\hrulefill
+	\hrulefill{}
 	\par\vspace{0.25cm}
@@
 	\par\vspace{0.25cm}
-	\hrulefill
+	\hrulefill{}
 	\par\vspace{0.5cm}
@@
 \end{titlepage}
+
 \newpage
 
-\printacronyms
+\printacronyms{}
+\newpage
+
+\listoffigures
 \newpage
@@
 
-%FIXME - Lots of passive voice
 \section{Background of the Study}
 There are a lot of educational typing tests available that help people learn
-touch typing, including Monkeytype, TypeRacer, and Keybr. These typing tests list
-out words that are then typed out. The inputted keys are then compared to check
-if the user has typed the expected letter. At the end of the test, the time
-taken is calculated, and certain metrics are given. These metrics include words
-per minute (WPM) and accuracy \parencite{bartnik2021}.
+touch typing, including Monkeytype, TypeRacer, and Keybr. These typing tests
+list out words that are then typed out. The entered keys are then compared to
+check if the user has typed the expected letter. At the end of the test, the
+time taken is calculated, and certain metrics are given. These metrics include
+\ac{wpm} and Accuracy \parencite{bartnik2021}.
 
@@
 with each key being associated with a specific finger. For instance, the letter
-Q in a QWERTY layout should be hit using the fifth digit of the left hand, and
-this is shown by coloring the fifth digit and the key Q with the same color or
-by placing the letters directly on the fingers \parencite{dobson2009touch}.
+Q in a QWERTY layout should be hit using the fifth digit, or the little finger,
+of the left hand, and this is shown by coloring the fifth digit and the key Q
+with the same color or by placing the letters directly on the fingers
+\parencite{dobson2009touch}.
 
@@
 three are hand and wrist positions that are common in all activities, however,
-prolonged periods in these positions may cause injuries such as Carpal
-tunnel syndrome (CTS) \parencite{toosi2015}
+prolonged periods in these positions may cause injuries such as \ac{cts}
+\parencite{toosi2015}
 
@@
 Thus, there is a need for automatically detecting which finger is used during
-typing, and for the position of the wrist in relation to the arm. One way to do
-this is through finger and hand tracking. One solution for tracking is by using
-image processing and machine learning. An example of this is MediaPipe
-by~\cite{mediapipe}.
+typing. One way to do this is through finger and hand tracking. One solution for
+tracking is by using image processing and machine learning. An example of this
+is MediaPipe by~\cite{mediapipe}.
 
@@
 computer vision and image processing. Machine learning can also be used with
-OpenCV.
+OpenCV.\@
 
@@
 \subsection{Specific Objectives}
-\begin{itemize}
-	\item To develop a program that tracks fingers and hand positions while typing
-	\item To create a subroutine that ascertains which finger was used to type a key
-	\item To detect if incorrect fingers were used to press a key or if the
-	      position of the hand in relation to the wrist is problematic
-	\item To assess the accuracy and performance of the developed program's ability
-	      to perform the previously stated objectives
-	\item To develop a user-friendly interface for users to train touch typing
-	      using the developed program
-	\item To generate key statistics of a user's touch typing performance:
-	      \begin{itemize}
-		      \item Words per Minute
-		      \item Accuracy
-		      \item Finger Placement Accuracy
-	      \end{itemize}
-\end{itemize}
+\begin{enumerate}
+	\item Develop an algorithm or technique using computer vision that is capable
+	      of identifying which keys are being pressed by which fingers during keyboard
+	      typing --- i.e.\ finger-key identification.
+	\item Identify the setup and configuration of a single optical camera that can
+	      be used to capture images during keyboard typing for real-time finger-key
+	      identification using the algorithm identified in 1
+	\item Design and develop a touch typing trainer using the algorithm in 1 and
+	      camera setup and configuration in 2
+	\item Determine the accuracy of the touch typing trainer in finger-key
+	      identification
+	\item Develop a new keyboard typing metric that uses both the key and
+	      finger used to press as parameters
+\end{enumerate}
 
@@
 
-In addition, the keyboard layout will be \ac{ansi}. This layout is described by the
-American National Standards Institute \parencite{ansi}. This is the most
+In addition, the keyboard layout will be \ac{ansi}. This layout is described by
+the American National Standards Institute \parencite{ansi}. This is the most
 common layout in the United States. However, it is also used in numerous
-English-speaking countries such as the Philippines, Malaysia, and India.
+English-speaking countries such as the Philippines, Malaysia, and India
+\parencite{apple-layout}.
 
@@
 Capturing of the video to be analyzed by the program would be limited to a
-single 480p webcam that is capturing in 24 frames per second. The camera will
-be pointed downwards facing the keyboard and the hand. This means that the
-vertical angle of the wrist may not be accurate.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{60.jpg}
+single 720p webcam that is capturing in 30 frames per second. The camera will be
+pointed downwards facing the keyboard and the hand.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.7\textwidth]{60.jpg}
 	\caption{A 60\% keyboard in \ac{ansi} layout. Reprinted from \fullcite{60}}
@@
 vast majority of the population as there are a lot of professions that heavily
-rely on keyboards. Examples include developers, physicians, educators,
+rely on keyboards. Examples include developers, physicians, educators, and
 accountants. By having better ergonomics while typing, wrist injuries can be
@@
 first place. A specific example of this is by reducing ulnar deviation which
-affects the nerve that is indicative of CTS \parencite{toosi2015}.
-
-\chapter{Preliminary Review of Related Literature}
+affects the nerve that is indicative of \ac{cts} \parencite{toosi2015}.
+
+\chapter{Advanced Review of Related Literature}
 
@@
 
-where, $|T|$ is the length of the text, $S$ is the time in seconds spent writing
+Where, $|T|$ is the length of the text, $S$ is the time in seconds spent writing
 the text. This time starts directly after the first character has been pressed,
@@
 
-Other metrics also measure speed but they aren't as commonly used as \ac{wpm}.
+Other metrics also measure speed, but they aren't as commonly used as \ac{wpm}.
 These include Characters per Minute, Gestures per Second, Adjusted Words per
@@
 
-where $|C|$ is the number of correct characters and $|T|$ is the length of the
+Where $|C|$ is the number of correct characters and $|T|$ is the length of the
 text.
@@
 posture, hand and wrist positions, and finger placement. Consequently, these
-metrics do not give a full picture of the performance of the person typing and
+metrics do not give a full picture of the performance of the person typing, and
 they only provide a cursory view of how a person types.
@@
 This curriculum is self-directed and learners can learn at their own pace. At
-its core, the curriculum is designed to be 36-week long with 5-10 minutes of
+its core, the curriculum is designed to be 36-week long with 5--10 minutes of
 lessons per day. The lessons in the curriculum follow the three stages of
@@
 This can be attributed to the different typing styles each person has. One key
-difference between people is the angle of the 5th digit.
+difference between people is the angle of the fifth digit.
 
@@
 extend one of their thumbs, usually the one not used for pressing a key. This is
-also true for some people that do not use their 5th digit during typing
+also true for some people that do not use their fifth digit during typing
 \parencite{baker2007}.
@@
 musculoskeletal disorders of the upper extremity (\cite{marklin1999} as cited in
-\cite{baker2007})
+~\cite{baker2007})
 
@@
 using motion capture systems or computer vision. This technique allows computers
-to perform actions and analyses on the motions and positions of these body
+to perform actions and analyzes on the motions and positions of these body
 parts.
@@
 calculators. Data passes through the graph as packets and a group of packets
-constitute a data stream. As the data passes through the pipeline, the
+constitutes a data stream. As the data passes through the pipeline, the
 calculators, produce the desired output.
@@
 is considerably smaller compared to a full-size keyboard. The keyboard will have
-key caps and a case in a light color that will contrast the dark surface that it
+keycaps and a case in a light color that will contrast the dark surface that it
 will be placed on. This is to improve initial keyboard detection. In addition,
@@
 \subsection{Environment}
-The environment the setup would be placed in would be a well lit environment
-with a light source beside the camera. This light source will evenly light the
-keyboard and the fingers used for typing. The light source to be used will be a
-common LED desk lamp rated at 9 Watts with 700 lumens. This light will be white
-with a color temperature of 6500k.
+The setup would be placed in would be a well lit environment with a light source
+beside the camera. This light source will evenly light the keyboard and the
+fingers used for typing. The light source to be used will be a common LED desk
+lamp rated at 9 Watts with 700 lumens. This light will be white with a color
+temperature of 6500k.
 
 In addition, the surface where the keyboard is to be placed on will be solid
-green without any variation of color. This is to improve initial keyboard
+black without any variation of color. This is to improve initial keyboard
 detection.
@@
 			rounded rectangle,
-			minimum width=2.5cm,
-			minimum height=1cm] (start) {START};
+			minimum width = 2.5cm,
+			minimum height = 1cm,
+			inner sep = 3mm
+		] (start) {START};
 
@@
 			minimum width=3.5cm,
-			minimum height=1cm
+			minimum height=1cm,
+			inner sep = 3mm
 		] (input) {Input Frame};
@@
 			minimum width=3.5cm,
-			minimum height=1cm] (gray) {Convert to Grayscale};
+			minimum height=1cm,
+			inner sep = 3mm
+		] (gray) {Convert to Grayscale};
 
@@
 			minimum width=3.5cm,
-			minimum height=1cm] (sobel) {Edge Detection using Sobel Filter};
+			minimum height=1cm,
+			inner sep = 3mm
+		] (sobel) {Edge Detection using Sobel Filter};
 
@@
 			minimum width=3.5cm,
-			minimum height=1cm] (contours) {Find Contours (Simple)};
+			minimum height=1cm,
+			inner sep = 3mm
+		] (contours) {Find Contours (Simple)};
 
@@
 			minimum width=3.5cm,
-			minimum height=1cm] (map) {Overlay Virtual Keyboard Map using Contour Points};
+			minimum height=1cm,
+			inner sep = 3mm
+		] (map) {Overlay Virtual Keyboard Map using Contour Points};
 
@@
 			minimum width=3.5cm,
-			minimum height=1cm
+			minimum height=1cm,
+			inner sep = 3mm
 		] (output) {Output Frame};
@@
 			minimum width=2.5cm,
-			minimum height=1cm] (end) {END};
+			minimum height=1cm,
+			inner sep = 3mm
+		] (end) {END};
 
@@
 	$\begin{bmatrix}
-  +1 & 0 & -1\\
-  +2 & 0 & -2\\
+			+1 & 0 & -1 \\
+			+2 & 0 & -2 \\
 			+1 & 0 & -1
@@
 	$\begin{bmatrix}
-  +1 & +2 & +1\\
-  0 & 0 & 0\\
+			+1 & +2 & +1 \\
+			0  & 0  & 0  \\
 			-1 & -2 & -1
@@
 Contours are curves joining all continuous points that have the same color or
-intensity \parencite{opencv-contours}. In this algorithm the OpenCV function
-\texttt{findContours} will be used to find the contours of the outlines of the
-object found using the previous step. This function accepts a contour
-approximation method as one of its arguments. Two methods are provided by
-OpenCV, \texttt{CHAIN\_APPROX\_NONE} and \texttt{CHAIN\_APPORX\_SIMPLE}. The
-former returns all contours in the shape, while the latter removes redundant
-points and returns the least amount of points that describes the shape. The
-algorithm will use the latter, as only the extreme edges of the keyboard needs
-to be detected. This OpenCV function implements the algorithm of
-\cite{contours} in their paper ``\citefield{contours}{title}''.
+intensity \parencite{opencv-contours}. The OpenCV function \texttt{findContours}
+will be used to find the contours of the outlines of the object found using the
+previous step. This function accepts a contour approximation method as one of
+its arguments. Two methods are provided by OpenCV:\@\texttt{CHAIN\_APPROX\_NONE}
+and \texttt{CHAIN\_APPROX\_SIMPLE}. The former returns all contours in the
+shape, while the latter removes redundant points and returns the least amount of
+points that describes the shape. The algorithm will use the latter, as only the
+extreme edges of the keyboard needs to be detected. This OpenCV function
+implements the algorithm of~\cite{contours} in their paper
+\citetitle{contours}.
 
@@
 \subsubsection{Overlay Virtual Keyboard Map using Contour Points}
-The virtual keyboard map is a rectangular image that contains individual region
-of interest (ROI) for each key in a 60\% \ac{ansi} keyboard. Each key has a
-corresponding color assigned to it and this color fills the region where this
-key is located at. Figure~\ref{fig:metho-algo-key-map} is the virtual keyboard
-map.
+The virtual keyboard map is a rectangular image that contains individual
+\ac{roi} for each key in a 60\% \ac{ansi} keyboard. Each key has a corresponding
+color assigned to it and this color fills the region where this key is located
+at. Figure~\ref{fig:metho-algo-key-map} is the virtual keyboard map. Appendix
+Chapter~\ref{appendix:key-map} is a table that shows the hex color code and key
+mapping present in the image map.
 
@@
 The first model, the Palm Detection Model detects the initial hand locations
-using a single shot detector model based on the paper by \cite{ssd}. This model
+using a single shot detector model based on the paper by~\cite{ssd}. This model
 achieves an average precision of 95.7\% in palm detection
@@
 
-MediaPipe Hands detects the palms first, instead of the whole hands with one
-model because the hands lack high contrast patterns. This reduces the model's
-ability to detect the hand with accuracy. In addition, detecting a palm is
-simpler compared to detecting hands with articulated fingers since estimating a
-bounding box around a rigid objects, i.e.\ a palm, is much simpler. Furthermore,
-a palm can be modeled using only square anchors reducing the number of anchors
-by a factor of 3--5 \parencite{mediapipe-hands}.
+MediaPipe Hands detects the palms first, instead of whole hands with one model
+because hands lack high contrast patterns. This reduces the model's ability to
+detect hands with accuracy. In addition, detecting a palm is simpler compared
+to detecting hands with articulated fingers since estimating a bounding box
+around a rigid object, i.e.\ a palm, is much simpler. Furthermore, a palm can be
+modeled using only square anchors reducing the number of anchors by a factor of
+3--5 \parencite{mediapipe-hands}.
 
@@
 
-The integration of the algorithm is a two step process. The first step is to get
+The integration of the algorithm is a two-step process. The first step is to get
 the mapping of the keyboard using the algorithm shown in
@@
 
-The second step is a continuous loop where the mapping of the keyboard is used
-in conjunction with the finger tracking algorithm chosen in
-Section~\ref{section:metho-algo-finger}. The flowchart of the algorithm is shown
-in Figure~\ref{fig:metho-algo-integration}
-
-\begin{figure}[H]
-	\centering
+The second step runs whenever a key press has been detected. This step uses the
+mapping of the keyboard in conjunction with the finger tracking algorithm
+selected in Section~\ref{section:metho-algo-finger}. The flowchart of the
+algorithm is shown in Figure~\ref{fig:metho-algo-integration}
+
+\begin{figure}[H]
+	\centering
+	\tikzset{
+		rectangle connector/.style={
+				-latex,
+				to path={(\tikztostart) -- ++(6cm,0pt) \tikztonodes |- (\tikztotarget) },
+				pos=0.5
+			},
+	}
 	\begin{tikzpicture}[font=\small,thick]
@@
 			rounded rectangle,
-			minimum width=2.5cm,
-			minimum height=1cm] (start) {START};
+			minimum width = 2.5cm,
+			minimum height = 1cm,
+			inner sep = 3mm
+		] (start) {START};
+
+		\node[draw,
+			diamond,
+			aspect = 2,
+			below = of start,
+			minimum width = 4cm,
+			inner sep = 3mm] (if-packet) {Keypress Detected};
 
@@
 			trapezium stretches,
-			below left=of start,
-			minimum width=3.5cm,
-			minimum height=1cm
-		] (map) {Virtual Keyboard Map};
+			below = 2cm of if-packet,
+			minimum width = 3.5cm,
+			minimum height = 1cm,
+			align = center,
+			inner sep = 3mm
+		] (map) {Virtual Keyboard Map \\ Finger Tracking Data};
+
+		\node[draw,
+			below = of map,
+			minimum width = 3.5cm,
+			minimum height = 1cm,
+			inner sep = 3mm
+		] (find-roi) {Find \ac{roi} of the pressed key in Virtual Keyboard Map};
+
+		\node[draw,
+			below=of find-roi,
+			minimum width = 3.5cm,
+			minimum height = 1cm,
+			inner sep = 3mm
+		] (find-finger) {Find which fingertip is on that \ac{roi}};
 
@@
 			trapezium stretches,
-			below right=of start,
-			minimum width=3.5cm,
-			minimum height=1cm
-		] (track) {Finger Tracking Frame};
-
-		\coordinate (CENTER) at ($(map)!0.5!(track)$);
+			below = of find-finger,
+			minimum width = 3.5cm,
+			minimum height = 1cm,
+			inner sep = 3mm
+		] (return) {Return which fingertip};
 
 		\node[draw,
-			diamond,
-			aspect=2,
-			below = 3cm of CENTER,
-			minimum width=2.5cm,
-			inner sep=0] (if-packet) {Keypress Packet Received};
-
-		\node[draw,
-			below left=of if-packet,
-			minimum width=3.5cm,
-			minimum height=1cm] (find-roi) {Find ROI of the pressed key in Virtual Keyboard Map};
-
-		\node[draw,
-			below=of find-roi,
-			minimum width=3.5cm,
-			minimum height=1cm] (find-finger) {Find which fingertip is on that ROI};
-
-		\node[draw,
-			trapezium,
-			trapezium left angle = 65,
-			trapezium right angle = 115,
-			trapezium stretches,
-			below=of find-finger,
-			minimum width=3.5cm,
-			minimum height=1cm
-		] (return) {Return which fingertip};
-
-		\node[draw,
-			trapezium,
-			trapezium left angle = 65,
-			trapezium right angle = 115,
-			trapezium stretches,
-			below = 11cm of CENTER,
-			minimum width=3.5cm,
-			minimum height=1cm
-		] (output) {Output Frame};
-
-		\node[draw,
-			below=of output,
+			below = of return,
 			rounded rectangle,
-			minimum width=2.5cm,
-			minimum height=1cm] (end) {END};
-
-		\draw[-latex] (start) edge (map)
-		(start) edge (track)
-		(track) edge (if-packet)
-		(map) edge (if-packet)
+			minimum width = 2.5cm,
+			minimum height = 1cm,
+			inner sep = 3mm
+		] (end) {END};
+
+		\draw[-latex] (start) edge (if-packet)
+		(map) edge (find-roi)
 		(find-roi) edge (find-finger)
 		(find-finger) edge (return)
-		(return) edge (output)
-		(output) edge (end);
-
-		\draw[-latex] (if-packet) -| (find-roi)
-		node[pos=0.25,fill=white,inner sep=0]{true};
-
-		\draw[-latex] (if-packet) edge node[pos=0.4,fill=white,inner sep=2pt]{No}(output);
+		(return) edge (end);
+
+		\draw[-latex] (if-packet) edge node[pos=0.5, fill=white, inner sep=2] {true} (map);
+		\draw[rectangle connector] (if-packet) to node[fill=white, inner sep=2] {false} (end);
 
@@
 
-\subsubsection{Find ROI of the pressend key in Virtual Keyboard Map}
+\subsubsection{Find \ac{roi} of the pressed key in Virtual Keyboard Map}
 Since there is a known color-key mapping based on the Virtual Keyboard Map, the
-algorithm will find the correct color to search for based on the received
-keypress packet. This color is isolated from the Virtual Keyboard Map, and the
-region of interest and its coordinates are obtained. The edges of this ROI will
-then be obtained by finding its contours. The specific implementation of this is
-explained in Section~\ref{section:metho-algo-key-contours}.
+algorithm will find the correct color to search for based on the received key
+press packet. This color is isolated from the Virtual Keyboard Map, and the
+\ac{roi} and its coordinates are obtained. The edges of this singular \ac{roi}
+will then be obtained by finding its contours. The specific implementation of
+this is explained in Section~\ref{section:metho-algo-key-contours}.
 
@@
 The expected output of this is four sets of pairs. Each pair corresponds to an
-edge of the region of interest.
-
-\subsubsection{Find which fingertip is on that ROI}
-The Finger Tracking Frame will contain the pixel positions of each landmark of
+edge of the single \ac{roi}.
+
+\subsubsection{Find which fingertip is on that \ac{roi}}
+The Finger Tracking Data will contain the pixel positions of each landmark of
 the hand. For this step, the algorithm will find the landmark which is
-positioned within the region of interest obtained from the previous step. This
-will be done using a series of checks.
+positioned within the single \ac{roi} obtained from the previous step.
+This will be done using a series of checks.
 
 Each landmark's coordinates will be compared to the coordinates of the edges of
-the ROI. A landmark will be determined as positioned within the ROI if all of the
+the \ac{roi}. A landmark will be determined as positioned within the \ac{roi} if all the
 following conditions are true:
@@
 	\item In the X axis, the landmark's coordinates is greater than one or both of
-		the two coordinates found of the left side of the ROI
+	      the two coordinates found of the left side of the \ac{roi}
 	\item In the X axis, the landmark's coordinates is less than one or both of
-		the two coordinates found of the right side of the ROI
+	      the two coordinates found of the right side of the \ac{roi}
 	\item In the Y axis, the landmark's coordinates is greater than one or both of
-		the two coordinates found of the top side of the ROI
+	      the two coordinates found of the top side of the \ac{roi}
 	\item In the Y axis, the landmark's coordinates is less than one or both of
-		the two coordinates found of the bottom side of the ROI
+	      the two coordinates found of the bottom side of the \ac{roi}
 \end{enumerate}
 
-These conditions maximize the total area of the ROI and is not strict about
+These conditions maximize the total area of the \ac{roi} and is not strict about
 exact accuracy. In essence, these conditions will create a perfectly rectangular
@@
 The landmark will then be used to determine which specific finger corresponds to
-the keypress. Figure~\ref{fig:metho-algo-integration-landmarks} shows each
+the key press. Figure~\ref{fig:metho-algo-integration-landmarks} shows each
 possible landmark that may be returned from the previous step. This step will
@@
 return \texttt{MIDDLE} denoting that the middle finger is the finger that
-corresponds with the keypress. In addition, the specific hand will also be
+corresponds with the key press. In addition, the specific hand will also be
 returned as one of two strings, \texttt{Left} and \texttt{Right}, since
@@
 Internet Archive. Sentences will be isolated from these text and used as test
-sequences if they fit a criteria.
+sequences if they fit the criteria.
 
@@
 test sequence is ``What of it, if some old hunks of a sea-captain orders me to
-get a broom and sweep down the decks?'' from Moby Dick by \cite{moby-dick}. The
-9 other test sequences can be found in the appendix.
+get a broom and sweep down the decks?'' from Moby Dick by~\cite{moby-dick}. The
+9 other test sequences can be found in Appendix
+Chapter~\ref{appendix:test-sequences}.
 
@@
 for showing real-time finger-key identification and mapping.
-Figure~\ref{fig:ui-test} and \ref{fig:ui-stat} illustrates the two main pages of
+Figure~\ref{fig:ui-test} and~\ref{fig:ui-stat} illustrates the two main pages of
 the design. Additional screens, such as historical statistics, in-platform
@@
 shows the three major statistics for the typing test: Words per minute,
-Accuracy, and Finger placement accuracy. Finger placement accuracy will be
-explained in Section~\ref{label:metric}. The component below the statistics is a
+Accuracy, and Finger Placement Accuracy. Finger Placement Accuracy will be
+explained in Section~\ref{section:metric}. The component below the statistics is a
 graph plotting the major statistics of the user in time. This includes when the
@@
 \subsubsection{Implementation}
-The design will be created in Figma and the front-end will be web based.The
+The design will be created in Figma and the front-end will be web based. The
 framework the front-end will be developed in will be Svelte, due to its
@@
 \subsection{Back-end development of a touch typing trainer using the program created
-	in \ref{section:metho-algo} and the experimental setup identified in
-	\ref{section:metho-setup}}
-
-The back-end will handle all of the computation for the trainer---mainly
+	in~\ref{section:metho-algo} and the experimental setup identified in~\ref{section:metho-setup}}
+
+The back-end will handle all the computation for the trainer---mainly
 finger-key identification. The backend will also handle authentication and data
@@
 
-The back-end will expose a REST API. This is a type of application programming
+The back-end will expose a REST API.\@This is a type of application programming
 interface (API) that conforms to the REST architectural style. One key
-characteristic of the this style is its statelessness and cacheable data
+characteristic of this style is its statelessness and cacheable data
 \parencite{rest}. This will be implemented using Django, a Python REST API
@@
 \subsubsection{Finger-key Identification}
-The submodule that integrates all of the sub algorithms will be called within
-the framework. The framework will then respond to requests for finger-key
+The submodule that integrates all the algorithms will be called within the
+framework. The framework will then respond to requests for finger-key
 identification using this submodule. This will allow the back-end to expose its
@@
 \subsubsection{Authentication}
-The core authentication system will rely on JSON Web Tokens (JWTs) . According
-to \cite{jwt}, ``JSON Web Token (JWT) is a compact, URL-safe means of
-representing claims to be transferred between two parties.'' This allows for a
-user to claim that they are that specific user securely as the claim is
-cryptographically signed. For the backend of the trainer, the JWT will be signed
-using a secret and it will be encoded in HMAC SHA256.
+The core authentication system will rely on \ac{jwt}. According to~\cite{jwt},
+``JSON Web Token (JWT) is a compact, URL-safe means of representing claims to be
+transferred between two parties.'' This allows for a user to claim that they are
+that specific user securely as the claim is cryptographically signed. For the
+backend of the trainer, the \ac{jwt} will be signed using a secret, and it will
+be encoded in HMAC SHA256.
 
 \subsubsection{Data Storage}
-The trainer will store all of the data in a relational database using
+The trainer will store all the data in a relational database using
 PostgreSQL.\@This specific technology was chosen due to its maturity, community,
@@
 
-There will be 15 videos of the researcher performing the typing test sequences.
+There will be 3 groups of 10 videos to be captured by the researcher, for a
+total of 30 videos. Each group corresponds to three predetermined speeds of
+typing: slow (15wpm), average (35wpm), and fast (80wpm). These speeds are based
+on test data from~\cite{keybr}. The researcher will type the 10 typing test
+sequences at these predetermined speeds. During typing, errors will not be
+consciously taken into account, and errors will happen naturally as part of the
+typing process.
+
 The researcher will then manually perform finger-key identification for all key
-presses present in the video. The finger-key identification submodule will then
-perform the same process for the same videos. The accuracy of the submodule will
-then be obtained by identifying which of its identification is incorrect, based
-on the manual identification.
+presses present in the video. The annotated data will then be divided into
+training and test data at a ratio of 60:40. The finger-key identification
+submodule will then perform the finger-key identification for the same videos.
+The accuracy of the submodule will then be obtained by verifying if its
+identification is as expected based on the manual annotation.
 
@@
 
-%% 4. Finger Placement Keyboard Typing Metric Development
-%%     1. Determine if the metric is for individual key presses or for the whole typing sequence or both.
-%%     2. Evaluate which variables need to be taken into account for the metric depending on (5.a):
-%%         1. Actual Finger Used
-%%         2. Correct Finger to be Used
-%%         3. Specific key pressed
-%%         4. Number of keys pressed with the correct fingers
-%%         5. Number of keys pressed with the wrong fingers
+\subsection{Per Test Sequence}
+The metric to be calculated per test sequence is \ac{fpacc}. This metric will
+compute the percentage of accurate keys pressed with the correct finger over the
+length of the typing test sequence. Inputting the wrong character, even if
+pressed with the correct finger will reduce \ac{fpacc} since \ac{fpacc} measures
+\emph{accurate key presses}, and incorrect characters will be considered as
+inaccurate key presses.
+
+The equation for calculating \ac{fpacc} is as follows:
+
+\begin{equation}
+	FP ACC = \frac{|F|}{|T|} \cdot 100\%
+\end{equation}
+
+Where $|F|$ refers to the number of accurate keys pressed with the
+correct finger and $|T|$ refers to the length of the text.
+
+\subsection{Per Key}
+This metric will compute the \ac{hfpacc} for a certain key. This metric will
+take the times the user pressed the wrong key when asked for a certain key into
+account, the same as in \ac{fpacc}.\@This will allow the user to easily verify
+which keys need attention and retraining if there is a high frequency of error.
+
+The equation for calculating a key's \ac{hfpacc} is as follows:
+
+\begin{equation}
+	HFP ACC = \frac{|F_{char}|}{|C_{char}|} \cdot 100\%
+\end{equation}
+
+Where $|F_{char}|$ refers to the number of accurate keys pressed with the
+correct finger for a certain character, identified as $char$. $|C_{char}|$
+refers to the number of times the character has appeared in all test sequences
+for the user.
+
+\newpage
+\appendix
+
+\twocolumn
+\chapter{Key-Color Values Map}
+\label{appendix:key-map}
+
+\tablehead{\toprule Hex Color Code&\multicolumn{1}{l}{Corresponding Key} \\ \midrule}
+\tabletail{%
+	\midrule \multicolumn{2}{r}{{Continued on next column}} \\ \midrule}
+\tablelasttail{%
+	\\\midrule
+	\multicolumn{2}{r}{{Concluded}} \\ \bottomrule}
+\centering
+\begin{supertabular}{ll}
+	\#FF0000 & $\sim$            \\
+	\#FF9900 & 1                 \\
+	\#FFE600 & 2                 \\
+	\#BDFF00 & 3                 \\
+	\#8FFF00 & 4                 \\
+	\#33FF00 & 5                 \\
+	\#00FF66 & 6                 \\
+	\#00FFC2 & 7                 \\
+	\#00E0FF & 8                 \\
+	\#00A3FF & 9                 \\
+	\#0066FF & 0                 \\
+	\#000AFF & -                 \\
+	\#5200FF & +                 \\
+	\#BD00FF & Backspace         \\
+
+	\#E00000 & Tab               \\
+	\#D68000 & Q								 \\
+	\#E1CB00 & W								 \\
+	\#A0D900 & E								 \\
+	\#7EE000 & R								 \\
+	\#2BD900 & T								 \\
+	\#00D856 & Y								 \\
+	\#00C395 & U								 \\
+	\#00B7D0 & I								 \\
+	\#0088D4 & O								 \\
+	\#004FC5 & P								 \\
+	\#0008D3 & [								 \\
+			\#3F00C5 & ]								 \\
+	\#8600B5 & $\vert$					 \\
+
+	\#C30000 & CapsLock          \\
+	\#A76400 & A								 \\
+	\#B2A100 & S								 \\
+	\#7CA800 & D								 \\
+	\#65B400 & F								 \\
+	\#1F9D00 & G								 \\
+	\#008B38 & H								 \\
+	\#018B6A & J								 \\
+	\#008294 & K								 \\
+	\#005E93 & L								 \\
+	\#003482 & :								 \\
+	\#0007AA & "								 \\
+	\#23006D & Enter						 \\
+
+	\#960000 & LeftShift         \\
+	\#683E00 & Z								 \\
+	\#827500 & X								 \\
+	\#4D6800 & C								 \\
+	\#3F7100 & V								 \\
+	\#125800 & B								 \\
+	\#00421A & N								 \\
+	\#005440 & M								 \\
+	\#004B55 & ,								 \\
+	\#00466D & .								 \\
+	\#002050 & /								 \\
+	\#000580 & RightShift				 \\
+
+	\#550000 & LeftControl       \\
+	\#2D0000 & LeftSuper         \\
+	\#3A2300 & LeftAlt           \\
+	\#3E3700 & Spacebar					 \\
+	\#003238 & RightAlt					 \\
+	\#001C2C & RightMeta 				 \\
+	\#000349 & RightSuper				 \\
+	\#18004D & RightControl 		 \\
+\end{supertabular}
+
+\onecolumn
+\chapter{Typing Test Sequences}
+\label{appendix:test-sequences}
+
+\begin{enumerate}
+	\item ``What of it, if some old hunks of a sea-captain orders me to get a broom and sweep down the decks?'' \parencite{moby-dick}
+	\item ``I sat down on an old wooden settle, carved all over like a bench on the Battery.'' \parencite{moby-dick}
+	\item ``I lay there dismally calculating that sixteen entire hours must elapse before I could hope for a resurrection.'' \parencite{moby-dick}
+	\item ``How slowly the time passes here, encompassed as I am by frost and snow!'' \parencite{frankenstein}
+	\item ``I listened to my father in silence and remained for some time incapable of offering any reply.'' \parencite{frankenstein}
+	\item ``Think you’re escaping and run into yourself. Longest way round is the shortest way home.'' \parencite{ulysses}
+	\item ``History, Stephen said, is a nightmare from which I am trying to awake'' \parencite{ulysses}
+	\item ``A man of genius makes no mistakes. His errors are volitional and are the portals of discovery.'' \parencite{ulysses}
+	\item ``Never trust to general impressions, my boy, but concentrate yourself upon details'' \parencite{sherlock}
+	\item ``I have no data yet. It is a capital mistake to theorise before one has data.'' \parencite{sherlock}
+\end{enumerate}
+
+\chapter{Front-End Screens}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{frontend-statistics.png}
+	\caption{Page showing historical statistics of the user}
+	\centering
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{frontend-guide-1.png}
+	\caption{Introduction to the platform during first open}
+	\centering
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{frontend-guide-2.png}
+	\caption{Step 1 of the tour of the platform}
+	\centering
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{frontend-guide-3.png}
+	\caption{Step 2 of the tour of the platform}
+	\centering
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{frontend-guide-4.png}
+	\caption{Step 3 of the tour of the platform}
+	\centering
+\end{figure}
 

### 2022-03-04 paper/paper.tex
manuscript: proofread and correct paper

@@
 \hypersetup{
-	pdftitle={Development of a touch typing trainer with an emphasis on finger-key identification},
+	pdftitle={Development of a finger-key identification submodule for a touch typing trainer},
 	pdfstartview={FitH},
@@
 
-%Flowchart
+% Flowchart
 \usepackage{tikz}
@@
 
-%Tables
+% Tables
 \usepackage{booktabs}
@@
 \setlength\intextsep{1cm}
+
+% Add Gantt Chart
+\usepackage{pdfpages}
+\usepackage{pdflscape}
 
@@
 
-	\textsc{Development of a touch typing trainer\\with an emphasis on finger-key identification }
+	\textsc{Development of a finger-key identification submodule\\for a touch typing trainer}
 	\par\vspace{0.5cm}
@@
 OpenCV.\@
+
+\section{Research Questions}
+\begin{enumerate}
+	\item What algorithm or technique using computer vision is capable of
+	      identifying which keys are being pressed by which fingers during keyboard
+	      typing?
+	\item What setup and configuration of a single optical camera can be used to
+	      capture images during keyboard typing for real-time finger-key identification
+	      using the algorithm identified in 1?
+	\item How to design and develop a touch typing trainer using the algorithm in 1
+	      and camera setup and configuration in 2?
+	\item How accurate is the touch typing trainer in identifying which finger is
+	      used to press which key?
+	\item What new keyboard typing metric can be developed in order to consider both
+	      the key and finger used to press as parameters?
+\end{enumerate}
 
@@
 
+The setup that will be captured will be simplified and fixed for the development
+and testing of the finger-key identification submodule. This includes the
+keyboard, surface it is placed on, and lighting of the entire area. Furthermore,
+the setup will not be tested by other respondents, but by the researchers only
+--- as the focus is on the development and testing of the finger-key
+identification submodule and not on the touch typing trainer and its effects on
+typing.
+
 The program will expect the that user has all ten digits and has no hand,
@@
 rely on keyboards. Examples include developers, physicians, educators, and
-accountants. By having better ergonomics while typing, wrist injuries can be
-prevented, and typing speed may be increased
+accountants. By having better ergonomics while typing --- by touch typing
+correctly --- wrist injuries can be prevented, and typing speed may be increased
 
@@
 this information is also bundled together with the landmarks.
+
+As such, the expected output of the submodule is the hand and the finger used to
+press the key. This allows for a more flexible utilization of the data as it
+carries greater context, compared to just returning if the finger used was
+correct or not.
 
@@
 
+\chapter{Gantt Chart}
+\begin{landscape}
+	\includepdf[pages=-,landscape,scale=0.8,pagecommand={}]{images/gantt-chart.pdf}
+\end{landscape}
+
 \newpage

### 2022-03-04 application/algorithms/keyboard-detection.py
algorithm(keyboard): add edge detection and contour finding code

@@ -1,0 +1,146 @@
+from cv2 import cv2
+import numpy as np
+
+SHOW_DEBUG_IMAGES: bool = True
+USE_DEBUG_IMAGE: bool = True
+
+
+def getKeyboardImage() -> np.ndarray | None:
+    cam = cv2.VideoCapture(0)
+
+    while True:
+        ret, frame = cam.read()
+
+        if not ret:
+            return None
+
+        cv2.imshow("keyboard", frame)
+        key = cv2.waitKey(1)
+
+        if key % 256 == 155:
+            cv2.destroyWindow("keyboard")
+            return frame
+
+
+def applySobelFilter(frame: np.ndarray) -> np.ndarray:
+    grayscaleFrame = cv2.cvtColor(frame, cv2.COLOR_BGR2GRAY)
+
+    blurredFrame = cv2.GaussianBlur(grayscaleFrame, (3, 3), 0)
+
+    sobelX = cv2.Sobel(
+        src=blurredFrame,
+        ddepth=cv2.CV_64F,
+        dx=1,
+        dy=0,
+        ksize=3,
+        borderType=cv2.BORDER_DEFAULT,
+    )
+
+    sobelY = cv2.Sobel(
+        src=blurredFrame,
+        ddepth=cv2.CV_64F,
+        dx=0,
+        dy=1,
+        ksize=3,
+        borderType=cv2.BORDER_DEFAULT,
+    )
+
+    absSobelX = cv2.convertScaleAbs(sobelX)
+    absSobelY = cv2.convertScaleAbs(sobelY)
+
+    finalSobel = cv2.addWeighted(absSobelX, 0.5, absSobelY, 0.5, 0)
+
+    # Apply Adaptive Thresholding
+    _, threshedFrame = cv2.threshold(
+        finalSobel, 0, 255, cv2.THRESH_BINARY + cv2.THRESH_OTSU
+    )
+
+    if SHOW_DEBUG_IMAGES:
+        cv2.imshow("sobel", finalSobel)
+        cv2.imshow("threshed-sobel", threshedFrame)
+        cv2.waitKey(0)
+        cv2.destroyWindow("sobel")
+
+    return threshedFrame
+
+
+# def applyThreshold(frame: np.ndarray) -> np.ndarray:
+#     grayscaleFrame = cv2.cvtColor(frame, cv2.COLOR_BGR2GRAY)
[…]

### 2022-03-04 application/algorithms/keyboard-detection.py
algorithm(keyboard): sort contour points and overlay image map correctly

@@ -1,12 +1,12 @@
-from cv2 import cv2
+from cv2 import cv2 as cv
 import numpy as np
 
 SHOW_DEBUG_IMAGES: bool = True
-USE_DEBUG_IMAGE: bool = True
+USE_DEBUG_IMAGE: bool = False
 
 
 def getKeyboardImage() -> np.ndarray | None:
-    cam = cv2.VideoCapture(0)
+    cam = cv.VideoCapture(0)
 
     while True:
         ret, frame = cam.read()
@@ -14,121 +14,150 @@ def getKeyboardImage() -> np.ndarray | None:
         if not ret:
             return None
 
-        cv2.imshow("keyboard", frame)
-        key = cv2.waitKey(1)
+        cv.imshow("keyboard", frame)
+        key = cv.waitKey(1)
 
         if key % 256 == 155:
-            cv2.destroyWindow("keyboard")
+            cv.destroyWindow("keyboard")
             return frame
 
 
 def applySobelFilter(frame: np.ndarray) -> np.ndarray:
-    grayscaleFrame = cv2.cvtColor(frame, cv2.COLOR_BGR2GRAY)
+    grayscaleFrame = cv.cvtColor(frame, cv.COLOR_BGR2GRAY)
 
-    blurredFrame = cv2.GaussianBlur(grayscaleFrame, (3, 3), 0)
+    bilateralFilteredFrame = cv.bilateralFilter(grayscaleFrame, 11, 17, 17)
 
-    sobelX = cv2.Sobel(
-        src=blurredFrame,
-        ddepth=cv2.CV_64F,
+    sobelX = cv.Sobel(
+        src=bilateralFilteredFrame,
+        ddepth=cv.CV_64F,
         dx=1,
         dy=0,
         ksize=3,
-        borderType=cv2.BORDER_DEFAULT,
+        borderType=cv.BORDER_DEFAULT,
     )
 
-    sobelY = cv2.Sobel(
-        src=blurredFrame,
-        ddepth=cv2.CV_64F,
+    sobelY = cv.Sobel(
+        src=bilateralFilteredFrame,
+        ddepth=cv.CV_64F,
         dx=0,
         dy=1,
         ksize=3,
-        borderType=cv2.BORDER_DEFAULT,
+        borderType=cv.BORDER_DEFAULT,
     )
 
-    absSobelX = cv2.convertScaleAbs(sobelX)
[…]

### 2022-03-04 application/algorithms/keyboard-detection.py
algorithm(keyboard): cleanup and rename function names

@@ -1,8 +1,8 @@
 from cv2 import cv2 as cv
 import numpy as np
 
-SHOW_DEBUG_IMAGES: bool = True
-USE_DEBUG_IMAGE: bool = False
+DEBUG_SHOW_IMAGES: bool = True
+DEBUG_USE_WEBCAM: bool = True
 
 
 def getKeyboardImage() -> np.ndarray | None:
@@ -22,9 +22,10 @@ def getKeyboardImage() -> np.ndarray | None:
             return frame
 
 
-def applySobelFilter(frame: np.ndarray) -> np.ndarray:
+def getEdgeFrame(frame: np.ndarray) -> np.ndarray:
     grayscaleFrame = cv.cvtColor(frame, cv.COLOR_BGR2GRAY)
 
+    # Denoise while maintaining edges
     bilateralFilteredFrame = cv.bilateralFilter(grayscaleFrame, 11, 17, 17)
 
     sobelX = cv.Sobel(
@@ -54,12 +55,9 @@ def applySobelFilter(frame: np.ndarray) -> np.ndarray:
         finalSobel, 0, 255, cv.THRESH_BINARY + cv.THRESH_OTSU
     )
 
-    if SHOW_DEBUG_IMAGES:
-        cv.imshow("sobel", finalSobel)
-        cv.imshow("threshed-sobel", threshedFrame)
+    if DEBUG_SHOW_IMAGES:
+        cv.imshow("debug", threshedFrame)
         cv.waitKey(0)
-        cv.destroyWindow("sobel")
-        cv.destroyWindow("threshed-sobel")
 
     return threshedFrame
 
@@ -67,46 +65,30 @@ def applySobelFilter(frame: np.ndarray) -> np.ndarray:
 def findContours(frame: np.ndarray) -> tuple[np.ndarray]:
     contours, _ = cv.findContours(frame, cv.RETR_EXTERNAL, cv.CHAIN_APPROX_SIMPLE)
 
-    if SHOW_DEBUG_IMAGES:
+    if DEBUG_SHOW_IMAGES:
         contourFrame = cv.cvtColor(frame, cv.COLOR_GRAY2BGR)
-        cv.drawContours(contourFrame, contours, -1, (0, 255, 0), 3)
-        cv.imshow("contours", contourFrame)
+        cv.drawContours(contourFrame, contours, -1, (0, 255, 0), 2)
+        cv.imshow("debug", contourFrame)
         cv.waitKey(0)
-        cv.destroyWindow("contours")
 
[…]

### 2022-05-17 application/algorithms/finger_detection.py
feat(test): add initial testing script

@@ -4,13 +4,11 @@ from cv2 import cv2 as cv
 import mediapipe as mp
 import numpy as np
 
-from common import DEBUG
+from common import DEBUG, convertResultToStr
 
 # Heavily lifted from https://google.github.io/mediapipe/getting_started/python.html.
 # Cited as (Lugaresi et al., n.d.) in the paper
-def classifyPressedFinger(
-    image: np.ndarray, roiPoints: np.ndarray
-) -> tuple[str, str] | None:
+def classifyPressedFinger(image: np.ndarray, roiPoints: np.ndarray) -> list[str]:
 
     mp_drawing = mp.solutions.drawing_utils
     mp_drawing_styles = mp.solutions.drawing_styles
@@ -110,7 +108,7 @@ def classifyPressedFinger(
 
         # Get all fingertip landmarks
         tipLandmarkMatches = [
-            identifier
+            convertResultToStr(identifier)
             for identifier in possibleMatches
             if identifier[1].split("_")[-1] == "TIP"
         ]
@@ -120,12 +118,8 @@ def classifyPressedFinger(
             print(tipLandmarkMatches)
             cv.waitKey(0)
 
-        if len(tipLandmarkMatches) > 0:
-            return tipLandmarkMatches[0]
+        return tipLandmarkMatches
 
         # Don't consider fallback. Later if naay instances lang
         # if len(possibleMatches) > 0:
         #     return possibleMatches[0]
-
-        # No fingertip is inside the roi
-        return None

### 2022-05-17 application/algorithms/integration.py

@@ -2,89 +2,26 @@
 
 from cv2 import cv2 as cv
 import numpy as np
-from copy import deepcopy
 
 # import time
 
 from common import (
-    findContours,
-    approximateLargestContour,
-    sortContourPoints,
     convertResultToStr,
-    DEBUG_displayContours,
     DEBUG,
 )
 
-from keyboard_map import keyDict
-from keyboard_detection import getImageMap
+from keyboard_detection import getContourPointsFromImage
 from finger_detection import classifyPressedFinger
-
-
-def filterVirtualMap(virtualMap: np.ndarray, key: list[int]) -> np.ndarray:
-    # Flip key, as they are stored in RGB, but the image is in BGR
-    key = [key[-1], key[1], key[0]]
-
-    filteredVirtualMap = deepcopy(virtualMap)
-
-    # Create mask where all other colors are blacked out .all is used here as
-    # np.where(virtualMap != key, ...) will consider other pixel values where
-    # only 1 channel is the same. i.e. pixel [100, 0, 0] will not be blacked out
-    # if the key is [100, 243, 0] as the blue channel of the pixel matches the
-    # key
-    mask = np.invert((filteredVirtualMap == key).all(axis=2))
-
-    filteredVirtualMap[mask] = [0, 0, 0]
-
-    # virtualMap = np.where(virtualMap != key, [0, 0, 0], virtualMap).astype(np.uint8)
-
-    if DEBUG:
-        cv.imshow("debug", filteredVirtualMap)
-
-        cv.waitKey(0)
-
-    return filteredVirtualMap
-
-
-def getKeyContourPoints(virtualMap: np.ndarray, key: list[int]) -> np.ndarray:
-    filteredMap = filterVirtualMap(virtualMap, key)
-
-    if DEBUG:
-        cv.imshow("debug", filteredMap)
-
-        cv.waitKey(0)
-
-    grayMap = cv.cvtColor(filteredMap, cv.COLOR_BGR2GRAY)
-
-    contours = findContours(grayMap)
-    largestContour = approximateLargestContour(contours)
-
-    DEBUG_displayContours(grayMap, largestContour)
-
-    largestContour = np.reshape(largestContour, (4, 2))
-
-    return sortContourPoints(largestContour)
-
-
-def getAllContourPoints(virtualMap: np.ndarray) -> dict[str, np.ndarray]:
-    return {
[…]

### 2022-05-17 application/algorithms/keyboard_detection.py

@@ -1,6 +1,9 @@
 from cv2 import cv2 as cv
 import numpy as np
 
+from copy import deepcopy
+
+from keyboard_map import keyDict
 from common import (
     findContours,
     approximateLargestContour,
@@ -128,6 +131,66 @@ def getImageMap(image: np.ndarray) -> np.ndarray:
         cv.destroyAllWindows()
 
     return imageMap
+
+
+def filterVirtualMap(virtualMap: np.ndarray, key: list[int]) -> np.ndarray:
+    # Flip key, as they are stored in RGB, but the image is in BGR
+    key = [key[-1], key[1], key[0]]
+
+    filteredVirtualMap = deepcopy(virtualMap)
+
+    # Create mask where all other colors are blacked out .all is used here as
+    # np.where(virtualMap != key, ...) will consider other pixel values where
+    # only 1 channel is the same. i.e. pixel [100, 0, 0] will not be blacked out
+    # if the key is [100, 243, 0] as the blue channel of the pixel matches the
+    # key
+    mask = np.invert((filteredVirtualMap == key).all(axis=2))
+
+    filteredVirtualMap[mask] = [0, 0, 0]
+
+    # virtualMap = np.where(virtualMap != key, [0, 0, 0], virtualMap).astype(np.uint8)
+
+    if DEBUG:
+        cv.imshow("debug", filteredVirtualMap)
+
+        cv.waitKey(0)
+
+    return filteredVirtualMap
+
+
+def getKeyContourPoints(virtualMap: np.ndarray, key: list[int]) -> np.ndarray:
+    filteredMap = filterVirtualMap(virtualMap, key)
+
+    if DEBUG:
+        cv.imshow("debug", filteredMap)
+
+        cv.waitKey(0)
+
+    grayMap = cv.cvtColor(filteredMap, cv.COLOR_BGR2GRAY)
+
+    contours = findContours(grayMap)
+    largestContour = approximateLargestContour(contours)
+
+    DEBUG_displayContours(grayMap, largestContour)
+
+    largestContour = np.reshape(largestContour, (4, 2))
+
+    return sortContourPoints(largestContour)
+
+
+def getAllContourPoints(virtualMap: np.ndarray) -> dict[str, np.ndarray]:
+    return {
+        key: getKeyContourPoints(virtualMap, color) for key, color in keyDict.items()
+    }
+
+
[…]

### 2022-05-17 application/algorithms/test-scripts/confirm-label.py

@@ -20,13 +20,13 @@ keypressDict = {
 
 
 def main():
-    (_, _, filenames) = next(walk("../dataset/test"))
+    (_, _, filenames) = next(walk("../dataset/training"))
     files = sorted(
         [file for file in filenames if not "mp4" in file and not "labeled" in file]
     )
 
     for index, file in enumerate(files):
-        fileLocation = f"../dataset/test/{file}"
+        fileLocation = f"../dataset/training/{file}"
         cap = cv.VideoCapture(fileLocation + ".mp4")
         labeledFile = f"{fileLocation}-labeled"
 
@@ -34,12 +34,6 @@ def main():
 
         with open(labeledFile) as fp:
             fileContent = [line.rstrip() for line in fp]
-
-        cap.set(cv.CAP_PROP_POS_FRAMES, 10)
-        ret, frame = cap.read()
-        frame = cv.rotate(frame, cv.ROTATE_180)
-
-        cv.imshow("setup", frame)
 
         for keypressIndex, keypress in enumerate(fileContent):
             data = keypress.split(":")
@@ -53,7 +47,7 @@ def main():
             cv.putText(
                 frame,
                 keypress,
-                (10, 25),
+                (200, 150),
                 cv.FONT_HERSHEY_SIMPLEX,
                 0.5,
                 (0, 255, 255),

### 2022-05-17 application/algorithms/test.py

@@ -1,0 +1,129 @@
+#!/usr/bin/env python3
+
+from os import walk, mkdir
+from cv2 import cv2 as cv
+import time
+import csv
+
+from finger_detection import classifyPressedFinger
+from keyboard_detection import getContourPointsFromImage
+
+
+def main():
+    (_, _, filenames) = next(walk("./dataset/training"))
+
+    # Create output folder
+    folder = f"./results/training-{time.time()}/"
+    mkdir(folder)
+
+    files = sorted(
+        [file for file in filenames if not "mp4" in file and not "labeled" in file]
+    )
+
+    # Create log for entire run
+    logFileContent = []
+
+    for file in files:
+        # Output Files
+        logFileContent.append(f"INFO: Start {file}")
+        resultFileContent = [
+            [
+                "is_correct",
+                "fingertips_detected",
+                "keyboard_detection_time",
+                "finger_classification_time",
+                "total_time",
+            ]
+        ]
+
+        fileLocation = f"./dataset/training/{file}"
+        labeledFile = f"{fileLocation}-labeled"
+
+        cap = cv.VideoCapture(fileLocation + ".mp4")
+        cap.set(cv.CAP_PROP_POS_FRAMES, 10)
+        ret, frame = cap.read()
+
+        keyboardDetectionTimeStart = time.time()
+
+        allContourPoints = {}
+        try:
+            allContourPoints = getContourPointsFromImage(frame)
+        except Exception as e:
+            logFileContent.append(f"ERROR: Failed to get contour points for {file}")
+            logFileContent.append(f"ERROR: {e}")
+            continue
+
+        keyboardDetectionTimeEnd = time.time() - keyboardDetectionTimeStart
+
+        if not ret:
+            logFileContent.append(f"ERROR: Failed to get frame 10 for {file}")
+            continue
+
+        with open(labeledFile) as fp:
+            fileContent = [line.rstrip() for line in fp]
+
+        for keypress in fileContent:
+            data = keypress.split(":")
+            cap.set(cv.CAP_PROP_POS_FRAMES, int(data[0]))
+            ret, frame = cap.read()
+
+            if not ret:
+                logFileContent.append(
+                    f"ERROR: Failed to get frame for {keypress} of {file}"
+                )
+                continue
+
+            contourPoints = allContourPoints[fileContent[1]]
+
+            fingerClassificationTimeStart = time.time()
+            result = []
+            try:
+                result = classifyPressedFinger(frame, contourPoints)
+            except Exception as e:
+                logFileContent.append(
[…]

### 2022-05-17 application/algorithms/test-scripts/clean-data.py
data: clean key error

@@ -3,12 +3,12 @@ from os import walk
 
 
 def main():
-    (_, _, filenames) = next(walk("../dataset/test"))
+    (_, _, filenames) = next(walk("../dataset/training"))
     files = [file for file in filenames if not "mp4" in file]
 
     for file in files:
-        with open(f"../dataset/test/{file}", "r+") as f:
-            content = f.read().replace(" ", "")
+        with open(f"../dataset/training/{file}", "r+") as f:
+            content = f.read().replace(" ", "").replace("//", "/")
             f.seek(0)
             f.write(content)
             f.truncate()

### 2022-05-17 application/algorithms/test.py
test: first test results

@@ -13,7 +13,8 @@ def main():
     (_, _, filenames) = next(walk("./dataset/training"))
 
     # Create output folder
-    folder = f"./results/training-{time.time()}/"
+    uuid = f"training-{time.time()}"
+    folder = f"./results/{uuid}"
     mkdir(folder)
 
     files = sorted(
@@ -22,42 +23,56 @@ def main():
 
     # Create log for entire run
     logFileContent = []
+    resultFileContent = [
+        [
+            "file_name",
+            "frame",
+            "key",
+            "finger",
+            "is_correct",
+            "fingertips_detected",
+            "keyboard_detection_time",
+            "finger_classification_time",
+        ]
+    ]
 
     for file in files:
         # Output Files
         logFileContent.append(f"INFO: Start {file}")
-        resultFileContent = [
-            [
-                "is_correct",
-                "fingertips_detected",
-                "keyboard_detection_time",
-                "finger_classification_time",
-                "total_time",
-            ]
-        ]
 
         fileLocation = f"./dataset/training/{file}"
         labeledFile = f"{fileLocation}-labeled"
 
         cap = cv.VideoCapture(fileLocation + ".mp4")
-        cap.set(cv.CAP_PROP_POS_FRAMES, 10)
-        ret, frame = cap.read()
 
         keyboardDetectionTimeStart = time.time()
 
         allContourPoints = {}
-        try:
-            allContourPoints = getContourPointsFromImage(frame)
-        except Exception as e:
+
+        for i in [10, 15, 20]:
+            cap.set(cv.CAP_PROP_POS_FRAMES, i)
+            ret, frame = cap.read()
+            if not ret:
+                logFileContent.append(f"ERROR: Failed to get frame {i} for {file}")
+                break
+
+            try:
+                keyboardDetectionTimeStart = time.time()
+                allContourPoints = getContourPointsFromImage(frame)
+            except Exception as e:
+                logFileContent.append(
+                    f"ERROR: Failed to get contour points for {file} with frame {i}"
+                )
+                logFileContent.append(f"ERROR: {e}")
+                cv.imwrite(f"{folder}/{file}:{i}.png", frame)
+            else:
+                break
+
+        if len(allContourPoints) == 0:
             logFileContent.append(f"ERROR: Failed to get contour points for {file}")
-            logFileContent.append(f"ERROR: {e}")
             continue
 
         keyboardDetectionTimeEnd = time.time() - keyboardDetectionTimeStart
-
[…]

### 2022-05-17 application/algorithms/common.py
train: scale contour, fix image map

@@ -3,7 +3,7 @@ import numpy as np
 from cv2 import cv2 as cv
 
 
-DEBUG = False
+DEBUG = True
 
 
 def findContours(frame: np.ndarray) -> tuple[np.ndarray]:
@@ -48,6 +48,32 @@ def sortContourPoints(contour: np.ndarray) -> np.ndarray:
 
     # arrange clockwise
     return np.array([left[0], right[0], right[1], left[1]])
+
+
+def scaleContours(contours: np.ndarray) -> np.ndarray:
+    scaleFactor = 10
+
+    upperLeft = [contours[0][0] - scaleFactor, contours[0][1] - scaleFactor]
+    upperRight = [contours[1][0] + scaleFactor, contours[1][1] - scaleFactor]
+    lowerRight = [contours[2][0] + scaleFactor, contours[2][1] + scaleFactor]
+    lowerLeft = [contours[3][0] - scaleFactor, contours[3][1] + scaleFactor]
+
+    return np.array([upperLeft, upperRight, lowerRight, lowerLeft])
+
+
+# def scaleContours(contours: np.ndarray) -> np.ndarray:
+#     M = cv.moments(contours)
+#     cx = int(M["m10"] / M["m00"])
+#     cy = int(M["m01"] / M["m00"])
+
+#     cnt_norm = contours - [cx, cy]
+#     cnt_scaled = cnt_norm * 1.5
+#     cnt_scaled = cnt_scaled + [cx, cy]
+#     cnt_scaled = cnt_scaled.astype(np.int32)
+
+#     print(cnt_scaled.shape)
+
+#     return cnt_scaled
 
 
 def DEBUG_displayContours(frame: np.ndarray, contour: np.ndarray):

### 2022-05-17 application/algorithms/integration.py

@@ -5,10 +5,7 @@ import numpy as np
 
 # import time
 
-from common import (
-    convertResultToStr,
-    DEBUG,
-)
+from common import DEBUG
 
 from keyboard_detection import getContourPointsFromImage
 from finger_detection import classifyPressedFinger
@@ -35,7 +32,7 @@ def main(keyboardImage: np.ndarray, handImage: np.ndarray):
     result = classifyPressedFinger(handImage, contourPoints)
 
     if result is not None:
-        print(convertResultToStr(result))
+        print(result)
 
     # total = time.time() - start
     # classifying = total - elapsed

### 2022-05-17 application/algorithms/keyboard_detection.py

@@ -8,6 +8,7 @@ from common import (
     findContours,
     approximateLargestContour,
     sortContourPoints,
+    scaleContours,
     DEBUG_displayContours,
     DEBUG,
 )
@@ -171,11 +172,13 @@ def getKeyContourPoints(virtualMap: np.ndarray, key: list[int]) -> np.ndarray:
     contours = findContours(grayMap)
     largestContour = approximateLargestContour(contours)
 
-    DEBUG_displayContours(grayMap, largestContour)
-
     largestContour = np.reshape(largestContour, (4, 2))
+    sortedContour = sortContourPoints(largestContour)
+    scaledContour = scaleContours(sortedContour)
 
-    return sortContourPoints(largestContour)
+    DEBUG_displayContours(grayMap, np.reshape(scaledContour, (4, 1, 2)))
+
+    return scaledContour
 
 
 def getAllContourPoints(virtualMap: np.ndarray) -> dict[str, np.ndarray]:

### 2022-05-17 application/algorithms/keyboard_map.py

@@ -6,8 +6,7 @@ keyDict = {
     "LeftSuper": [45, 0, 0],
     "LeftAlt": [58, 35, 0],
     "Spacebar": [62, 55, 0],
-    "RightAlt": [0, 50, 56],
-    "RightMeta": [0, 28, 44],
+    "RightAlt": [0, 28, 56],
     "RightSuper": [0, 3, 73],
     "RightControl": [24, 0, 76],
     "`": [255, 0, 0],
@@ -23,7 +22,8 @@ keyDict = {
     "0": [0, 102, 255],
     "-": [0, 10, 255],
     "=": [82, 0, 255],
-    "Backspace": [189, 0, 255],
+    "|": [128, 0, 255],
+    "Delete": [189, 0, 255],
     "Tab": [224, 0, 0],
     "Q": [214, 128, 0],
     "W": [225, 203, 0],
@@ -37,7 +37,7 @@ keyDict = {
     "P": [0, 79, 196],
     "[": [0, 8, 211],
     "]": [63, 0, 196],
-    "|": [134, 0, 181],
+    "Backspace": [134, 0, 181],
     "CapsLock": [195, 0, 0],
     "A": [167, 100, 0],
     "S": [178, 161, 0],

### 2022-05-17 application/algorithms/common.py
train: fix keyDict, correct data

@@ -3,7 +3,7 @@ import numpy as np
 from cv2 import cv2 as cv
 
 
-DEBUG = True
+DEBUG = False
 
 
 def findContours(frame: np.ndarray) -> tuple[np.ndarray]:

### 2022-05-17 application/algorithms/keyboard_map.py

@@ -6,7 +6,7 @@ keyDict = {
     "LeftSuper": [45, 0, 0],
     "LeftAlt": [58, 35, 0],
     "Spacebar": [62, 55, 0],
-    "RightAlt": [0, 28, 56],
+    "RightAlt": [0, 28, 44],
     "RightSuper": [0, 3, 73],
     "RightControl": [24, 0, 76],
     "`": [255, 0, 0],

### 2022-05-17 application/algorithms/common.py
train: try uneven scaling

@@ -51,7 +51,7 @@ def sortContourPoints(contour: np.ndarray) -> np.ndarray:
 
 
 def scaleContours(contours: np.ndarray) -> np.ndarray:
-    scaleFactor = 10
+    scaleFactor = 7
 
     upperLeft = [contours[0][0] - scaleFactor, contours[0][1] - scaleFactor]
     upperRight = [contours[1][0] + scaleFactor, contours[1][1] - scaleFactor]
@@ -59,6 +59,18 @@ def scaleContours(contours: np.ndarray) -> np.ndarray:
     lowerLeft = [contours[3][0] - scaleFactor, contours[3][1] + scaleFactor]
 
     return np.array([upperLeft, upperRight, lowerRight, lowerLeft])
+
+
+# def scaleContours(contours: np.ndarray) -> np.ndarray:
+#     upperScaleFactor = 10
+#     lowerScaleFactor = 5
+
+#     upperLeft = [contours[0][0] - upperScaleFactor, contours[0][1] - upperScaleFactor]
+#     upperRight = [contours[1][0] + upperScaleFactor, contours[1][1] - upperScaleFactor]
+#     lowerRight = [contours[2][0] + lowerScaleFactor, contours[2][1] + lowerScaleFactor]
+#     lowerLeft = [contours[3][0] - lowerScaleFactor, contours[3][1] + lowerScaleFactor]
+
+#     return np.array([upperLeft, upperRight, lowerRight, lowerLeft])
 
 
 # def scaleContours(contours: np.ndarray) -> np.ndarray:

### 2022-05-17 application/algorithms/finger_detection.py

@@ -4,7 +4,7 @@ from cv2 import cv2 as cv
 import mediapipe as mp
 import numpy as np
 
-from common import DEBUG, convertResultToStr
+from common import DEBUG, DEBUG_displayContours, convertResultToStr
 
 # Heavily lifted from https://google.github.io/mediapipe/getting_started/python.html.
 # Cited as (Lugaresi et al., n.d.) in the paper
@@ -103,7 +103,10 @@ def classifyPressedFinger(image: np.ndarray, roiPoints: np.ndarray) -> list[str]
                     mp_drawing_styles.get_default_hand_connections_style(),
                 )
 
-                cv.imshow("debug", annotated_image)
+                DEBUG_displayContours(
+                    cv.cvtColor(annotated_image, cv.COLOR_BGR2GRAY),
+                    np.reshape(roiPoints, (4, 1, 2)),
+                )
                 cv.waitKey(0)
 
         # Get all fingertip landmarks

### 2022-05-17 application/algorithms/integration.py

@@ -5,7 +5,7 @@ import numpy as np
 
 # import time
 
-from common import DEBUG
+from common import DEBUG, DEBUG_displayContours
 
 from keyboard_detection import getContourPointsFromImage
 from finger_detection import classifyPressedFinger
@@ -20,7 +20,7 @@ def main(keyboardImage: np.ndarray, handImage: np.ndarray):
     # Step 1
     allContourPoints = getContourPointsFromImage(keyboardImage)
 
-    contourPoints = allContourPoints["Spacebar"]
+    contourPoints = allContourPoints["Y"]
 
     # elapsed = time.time() - start
 
@@ -49,6 +49,6 @@ def main(keyboardImage: np.ndarray, handImage: np.ndarray):
 
 
 if __name__ == "__main__":
-    keyboardImage = cv.imread("images/keyboard.jpg")
-    handImage = cv.imread("images/hand.jpg")
+    keyboardImage = cv.imread("images/medium-06-10.png")
+    handImage = cv.imread("images/medium-06-418.png")
     main(keyboardImage, handImage)

### 2022-05-17 application/algorithms/test-scripts/get-image.py

@@ -1,0 +1,32 @@
+#!/usr/bin/env python3
+
+from cv2 import cv2 as cv
+
+
+def main():
+    file = "medium-06"
+    keyboardFrame = 10
+    fingerFrame = 418
+    fileLocation = f"../dataset/training/{file}"
+
+    cap = cv.VideoCapture(f"{fileLocation}.mp4")
+
+    cap.set(cv.CAP_PROP_POS_FRAMES, keyboardFrame)
+    ret, frame = cap.read()
+
+    if not ret:
+        return
+
+    cv.imwrite(f"../images/{file}-{keyboardFrame}.png", frame)
+
+    cap.set(cv.CAP_PROP_POS_FRAMES, fingerFrame)
+    ret, frame = cap.read()
+
+    if not ret:
+        return
+
+    cv.imwrite(f"../images/{file}-{fingerFrame}.png", frame)
+
+
+if __name__ == "__main__":
+    main()

### 2022-05-18 paper/paper.tex
methodology: add methodology

@@
 
+\usepackage{caption}
+
 %%% LT-SKIP-END
@@
 \printacronyms{}
+\newpage
+
+\listoftables
 \newpage
@@
 
-\chapter{Advanced Review of Related Literature}
+\chapter{Review of Related Literature}
 
@@
 
-The experimental setup and configuration will be composed of three elements: the
-camera, the keyboard, and the environment. Figure~\ref{fig:metho-setup}
-shows the sketch of the complete experimental setup.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.4\textwidth]{setup.png}
-	\caption{Sketch of the experimental setup}
+The experimental setup and configuration was composed of three elements: the
+camera, the keyboard, and the environment. Figure~\ref{fig:metho-setup} is an
+image of the experimental setup
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{full-setup.jpg}
+	\caption{The experimental setup}
 	\label{fig:metho-setup}
@@
 \subsection{Camera}
-The camera setup will use a single monocular camera positioned in a top-down
-view. The camera will capture the entirety of the keyboard and the movement of
-the ten fingers. To do so, it will be mounted on top of the monitor and the
-camera will point down towards the table. Figure~\ref{fig:metho-setup-keeb}
-illustrates what the camera will capture once placed in its correct position.
+The setup used a single monocular camera positioned in a top-down view. The
+camera captured the entirety of the keyboard and the movement of the ten
+fingers. To do so, it was mounted on top of the monitor with the camera pointed
+down towards the table.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{camera-placement.jpg}
+	\caption{The camera angled downwards to capture the keyboard}
+	\label{fig:metho-setup-camera-placement}
+\end{figure}
 
@@
 	\includegraphics[width=0.8\textwidth]{actual-keeb.png}
-	\caption{Keyboard, camera angle, and placement to be used for testing.}
+	\caption{Image captured by the camera in one of the training iterations}
 	\label{fig:metho-setup-keeb}
@@
 
-The specific camera to be used will be a Logitech C920.
-Figure~\ref{fig:metho-setup-cam} is a picture of this camera. It is a 3 mega pixel
-webcam that is capable of capturing color video in 1080p/30fps and 720p/30fps
-with a diagonal field of view of 78\degree. This camera has a universal mounting
-clip that will allow the camera to be correctly positioned within the
-experimental setup \parencite{logitech}.
+The specific camera used was a Logitech C920. It is a 3 mega pixel webcam that
+is capable of capturing color video in 1080p/30fps and 720p/30fps with a
+diagonal field of view of 78\degree. This camera has a universal mounting clip
+that allows the camera to be correctly positioned within the experimental setup
+\parencite{logitech}.
 
@@
 
-The keyboard will be a 60\% keyboard as shown in
-Figure~\ref{fig:metho-setup-keeb}. This will limit the necessary mapping for the
-algorithm to the alphanumeric portion of the keyboard. This keyboard choice will
-also lessen the area that the camera will need to capture, as this keyboard type
-is considerably smaller compared to a full-size keyboard. The keyboard will have
-keycaps and a case in a light color that will contrast the dark surface that it
-will be placed on. This is to improve initial keyboard detection. In addition,
-the keyboard layout will also be ANSI, due to the availability and widespread
-adoption of the layout in the Philippines.
+The keyboard used was a 60\% keyboard as shown in
+Figure~\ref{fig:metho-setup-keeb}. This limited the necessary mapping for the
+algorithm to the alphanumeric portion of the keyboard. This keyboard choice also
+reduced the area that the camera captured, as this keyboard type is considerably
+smaller compared to a full-size keyboard. The keyboard also had its keycaps and
+case in a light color that contrasted the dark surface it was placed on. It also
+had a dark USB-C cable to blend with the dark surface. These color coordination
+steps improved initial keyboard detection. In addition, the keyboard layout was
+ANSI, due to the availability and widespread adoption of the layout in the
+Philippines.
 
 \subsection{Environment}
-The setup would be placed in would be a well lit environment with a light source
-beside the camera. This light source will evenly light the keyboard and the
-fingers used for typing. The light source to be used will be a common LED desk
-lamp rated at 9 Watts with 700 lumens. This light will be white with a color
+The setup was lit with a lamp besides the camera. This light source evenly lit
+the keyboard and the fingers used for typing. The light source used was a common
+LED bulb rated at 9 Watts with 700 lumens. This light was white with a color
 temperature of 6500k.
 
-In addition, the surface where the keyboard is to be placed on will be solid
-black without any variation of color. This is to improve initial keyboard
-detection.
+In addition, the surface where the keyboard was placed on was solid black
+without any variation of color. This also improved initial keyboard detection.
 
@@
 
-\subsection{Computer Vision based Keyboard Detection, and Mapping}
+\subsection{Computer Vision based Keyboard Detection, and Key Mapping}
 \label{section:metho-algo-keyboard}
 
-A computer vision algorithm will be created as a starting point for detecting,
-and mapping the keyboard within a video. The flowchart of the algorithm is shown
-in Figure~\ref{fig:metho-algo-key-flow}.
+A computer vision algorithm was created as a starting point
+mapping the keys of the keyboard within a video. A rough flowchart of the algorithm
+is shown in Figure~\ref{fig:metho-algo-key-flow}.
 
@@
 			inner sep = 3mm
-		] (gray) {Convert to Grayscale};
-
-		\node[draw,
-			below=of gray,
+		] (map) {Get Image Map};
+
+		\node[draw,
+			below=of map,
 			minimum width=3.5cm,
@@
 			inner sep = 3mm
-		] (sobel) {Edge Detection using Sobel Filter};
-
-		\node[draw,
-			below=of sobel,
-			minimum width=3.5cm,
-			minimum height=1cm,
-			inner sep = 3mm
-		] (contours) {Find Contours (Simple)};
-
-		\node[draw,
-			below=of contours,
-			minimum width=3.5cm,
-			minimum height=1cm,
-			inner sep = 3mm
-		] (map) {Overlay Virtual Keyboard Map using Contour Points};
+		] (contours) {Get Key Contour Points};
 
@@
 			trapezium stretches,
-			below=of map,
+			below=of contours,
 			minimum width=3.5cm,
@@
 			inner sep = 3mm
-		] (output) {Output Frame};
+		] (output) {Key-Edge Coordinatess Map};
 
@@
 		\draw[-latex] (start) edge (input)
-		(input) edge (gray)
-		(gray) edge (sobel)
-		(sobel) edge (contours)
-		(contours) edge (map)
-		(map) edge (output)
+		(input) edge (map)
+		(map) edge (contours)
+		(contours) edge (output)
 		(output) edge (end);
@@
 
-\subsubsection{Convert to Grayscale}
-The frame will be converted to a 256 level grayscale image. This step will be
-performed because future steps of the algorithm will not require color values to
-work. In addition, the performance of the following steps will be also improved
-as the number of dimensions to be analyzed is reduced.
-
-\subsubsection{Edge Detection using Sobel Filter}
-A Sobel Filter is an edge detection filter that uses a 3$\times$3 kernel that is
-convolved twice. Once horizontally, and another vertically to produce a
-grayscale image of the outlines within the frame. The kernels used by the Sobel
-Filter \parencite{sobel2014} is shown in Figure~\ref{fig:metho-algo-key-sobel}.
+\newpage
+
+\subsubsection{Get Image Map}
+The first portion of the algorithm creates an image map that is overlaid over
+the detected edges of the keyboard. The flowchart for this function is shown in
+Figure~\ref{fig:metho-algo-key-image-map}.
+
+\begin{figure}[H]
+	\centering
+	\begin{tikzpicture}[font=\small,thick]
+		\node[draw,
+			rounded rectangle,
+			minimum width = 2.5cm,
+			minimum height = 1cm,
+			inner sep = 3mm
+		] (start) {START};
+
+		\node[draw,
+			trapezium,
+			trapezium left angle = 65,
+			trapezium right angle = 115,
+			trapezium stretches,
+			below=of start,
+			minimum width=3.5cm,
+			minimum height=1cm,
+			inner sep = 3mm
+		] (input) {Input Frame};
+
+		\node[draw,
+			below=of input,
+			minimum width=3.5cm,
+			minimum height=1cm,
+			inner sep = 3mm
+		] (gray) {Convert to Grayscale};
+
+		\node[draw,
+			below=of gray,
+			minimum width=3.5cm,
+			minimum height=1cm,
+			inner sep = 3mm
+		] (denoise) {Denoising};
+
+		\node[draw,
+			below=of denoise,
+			minimum width=3.5cm,
+			minimum height=1cm,
+			inner sep = 3mm
+		] (edge) {Edge Detection};
+
+		\node[draw,
+			below=of edge,
+			minimum width=3.5cm,
+			minimum height=1cm,
+			inner sep = 3mm
+		] (thresh) {Thresholding};
+
+		\node[draw,
+			right=of gray,
+			minimum width=3.5cm,
+			minimum height=1cm,
+			inner sep = 3mm
+		] (contours) {Find Contours};
+
+		\node[draw,
+			below=of contours,
+			minimum width=3.5cm,
+			minimum height=1cm,
+			inner sep = 3mm
+		] (sort) {Simplify Contour};
+
+		\node[draw,
+			below=of sort,
+			minimum width=3.5cm,
+			minimum height=1cm,
+			inner sep = 3mm
+		] (transform) {Transform Image Map};
+
+		\node[draw,
+			trapezium,
+			trapezium left angle = 65,
+			trapezium right angle = 115,
+			trapezium stretches,
+			below=of transform,
+			minimum width=3.5cm,
+			minimum height=1cm,
+			inner sep = 3mm
+		] (output) {Virtual Image Map};
+
+		\node[draw,
+			below=of output,
+			rounded rectangle,
+			minimum width=2.5cm,
+			minimum height=1cm,
+			inner sep = 3mm
+		] (end) {END};
+
+		\node[coordinate, right = 0.5cm of thresh] (below) {};
+		\node[coordinate, right = 0.5cm of gray] (top) {};
+
+		\draw[-latex] (start) edge (input)
+		(input) edge (gray)
+		(gray) edge (denoise)
+		(denoise) edge (edge)
+		(edge) edge (thresh)
+		(thresh) -- (below) |- (top) |- (contours)
+		(contours) edge (sort)
+		(sort) edge (transform)
+		(transform) edge (output)
+		(output) edge (end);
+
+	\end{tikzpicture}
+	\caption{Flowchart of Get Image Map}
+	\label{fig:metho-algo-key-image-map}
+\end{figure}
+
+\newpage
+
+\paragraph{Convert to Grayscale}
+The input frame from the camera was converted to a 256 level grayscale image
+using \texttt{cvtColor} of OpenCV \parencite{opencv-cvtColor}. This step was
+performed because future steps of the algorithm did not require color values to
+work. In addition, this step optimized the algorithm as the number of dimensions
+analyzed was reduced.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{grayscale.png}
+	\caption{Camera capture converted to grayscale}
+	\centering
+\end{figure}
+
+\paragraph{Denoising}
+The algorithm denoised the grayscale image using a bilateral filter as
+implemented by OpenCV \parencite{opencv-bilateral-filter}. This filter takes the
+range of the image into account, rather than just the domain. This resulted in
+an image that is smoothed while preserving its edges
+\parencite{bilateral-filter}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{denoised.png}
+	\caption{Smoothed image with intact edges}
+	\centering
+\end{figure}
+
+\paragraph{Edge Detection}
+The algorithm used a Sobel filter for edge detection. This filter uses a
+3$\times$3 kernel that is convolved twice. Once horizontally, and another
+vertically to produce a grayscale image of the outlines within the frame. The
+kernels used by the Sobel filter \parencite{sobel2014} is shown in
+Figure~\ref{fig:metho-algo-key-sobel}.
 
@@
 		\end{bmatrix}$
-	and
 	$\begin{bmatrix}
@@
 
-The expected result of this step is a black and white image that highlights the
-edges of the keyboard in white.
-
-\subsubsection{Find Contours (Simple)}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{sobel.png}
+	\caption{Output of the Sobel filter}
+	\centering
+\end{figure}
+
+\paragraph{Thresholding}
+The output of the Sobel filter is a grayscale image of 255 values, with each
+gray pixel indicating edges within the image. There is a need to reduce the
+range of these values to improve the ability of the algorithm to find the
+contours of these edges. To do so, the algorithm utilized Otsu's algorithm to
+perform automated thresholding. Otsu's algorithm determines a single threshold
+that is most optimal for the image \parencite{otsu}. This outputs a
+black-and-white image, with only two values.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{threshed.png}
+	\caption{Threshed image of the edges}
+	\centering
+\end{figure}
+
+\paragraph{Find Contours}
 \label{section:metho-algo-key-contours}
 Contours are curves joining all continuous points that have the same color or
-intensity \parencite{opencv-contours}. The OpenCV function \texttt{findContours}
-will be used to find the contours of the outlines of the object found using the
-previous step. This function accepts a contour approximation method as one of
-its arguments. Two methods are provided by OpenCV:\@\texttt{CHAIN\_APPROX\_NONE}
-and \texttt{CHAIN\_APPROX\_SIMPLE}. The former returns all contours in the
-shape, while the latter removes redundant points and returns the least amount of
-points that describes the shape. The algorithm will use the latter, as only the
-extreme edges of the keyboard needs to be detected. This OpenCV function
-implements the algorithm of~\cite{contours} in their paper
-\citetitle{contours}.
-
-The expected result of this algorithm is four contour points which are
-positioned on the four edges of the keyboard.
-
-\subsubsection{Overlay Virtual Keyboard Map using Contour Points}
+intensity \parencite{opencv-contours}. The OpenCV function
+\texttt{findContours}, with the \texttt{CHAIN\_APPROX\_SIMPLE} contour
+approximation method, was used to find the contours of the outlines of the
+object found using the previous step. This approximation method removed
+redundant points and returned the least amount of points that describes the
+shape. This OpenCV function implements the algorithm of~\cite{contours} in their
+paper \citetitle{contours}.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{find-contours.png}
+	\caption{All the contours found in the image}
+	\centering
+\end{figure}
+
+\paragraph{Simplify Contour}
+However, the points returned by the function can contain more than the four
+extreme points at the edges of the keyboard. In addition, more than one contour
+may be found within the image. As such, the algorithm sorted the contours by
+area, and selected the largest one.
+
+After sorting, the algorithm performed the Douglas-Peucker algorithm for Line
+Simplification \parencite{douglas-peucker}. This reduced the number of points in
+the largest contour to the minimum amount. In ideal cases, the number of points
+would be four, with each point corresponding to the edges of the keyboard.
+However, there were times when other objects would be within the frame, or they
+intersected with the keyboard. This would result in a contour that would be
+defined by more than four points --- which caused the entire algorithm to fail.
+
+The contour points that described the contour were then sorted clockwise. This
+was a necessity since the algorithm of~\cite{contours} does not guarantee that
+the largest contour's points were returned in a clockwise arrangement, which is
+a requirement for the next step
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{get-largest-contour.png}
+	\caption{Simplified contour}
+	\centering
+\end{figure}
+
+\paragraph{Transform Image Map}
 The virtual keyboard map is a rectangular image that contains individual
@@
 color assigned to it and this color fills the region where this key is located
-at. Figure~\ref{fig:metho-algo-key-map} is the virtual keyboard map. Appendix
-Chapter~\ref{appendix:key-map} is a table that shows the hex color code and key
-mapping present in the image map.
+at. Appendix Chapter~\ref{appendix:key-map} is a table that shows the hex color
+code and key mapping present in the image map.
 
@@
 
-The virtual map will then be stretched over the object, with the four contour
-points as the four edges of the virtual keyboard map. The image generated will
-then be returned as the final output of the graph.
-Figure~\ref{fig:metho-algo-key-overlay} illustrates this image, and
-Figure~\ref{fig:metho-algo-key-half} shows how accurate the image map is when
-overlaid over the image.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1\textwidth]{image-map-final.png}
-	\caption{Expected output of the algorithm}
-	\label{fig:metho-algo-key-overlay}
-	\centering
-\end{figure}
-
-\hfill
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1\textwidth]{image-map-half.png}
-	\caption{Overlaid image map shown at 60\% opacity}
-	\label{fig:metho-algo-key-half}
-	\centering
-\end{figure}
+The virtual map was stretched over the keyboard using OpenCV's
+\texttt{warpPerspective} function. The function requires a perspective transform
+that was calculated using \texttt{getPerspectiveTransform}. This function takes
+in two arrays with four points each. The input points are the edges of the
+virtual map, ordered clockwise. The output points are the four points of the
+contour, ordered clockwise \parencite{opencv-image-transform}. The resulting
+transform was then used by \texttt{warpPerspective} to apply the transform to
+the virtual map.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.6\textwidth]{transformed-image-map.png}
+	\caption{Transformed image map}
+	\centering
+\end{figure}
+
+\subsubsection{Get Key Contour Points}
+The second portion of the algorithm pre-computes the contour points of each key
+in the keyboard based on the generated virtual map and the Key-Color Values map.
+The flowchart of this function is shown in
+Figure~\ref{fig:metho-algo-key-contour-points}
+
+\begin{figure}[H]
+	\centering
+	\begin{tikzpicture}[font=\small,thick]
+		\node[draw,
+			rounded rectangle,
+			minimum width = 2.5cm,
+			minimum height = 1cm,
+			inner sep = 3mm
+		] (start) {START};
+
+		\node[draw,
+			trapezium,
+			trapezium left angle = 65,
+			trapezium right angle = 115,
+			trapezium stretches,
+			below = of start,
+			minimum width = 3.5cm,
+			minimum height = 1cm,
+			align = center,
+			inner sep = 3mm
+		] (input) {Virtual Map \\ Key-Color Values Map};
+
+		\node[draw,
+			diamond,
+			aspect = 2,
+			below = of input,
+			minimum width = 4cm,
+			inner sep = 2mm] (for) {for Key, Color in Key-Color Values Map};
+
+		\node[draw,
+			below=of for,
+			minimum width=3.5cm,
+			minimum height=1cm,
+			inner sep = 3mm
+		] (filter) {Filter Virtual Map};
+
+		\node[draw,
+			below=of filter,
+			minimum width=3.5cm,
+			minimum height=1cm,
+			inner sep = 3mm
+		] (contours) {Find and Simplify Contour};
+
+		\node[draw,
+			below=of contours,
+			minimum width=3.5cm,
+			minimum height=1cm,
+			inner sep = 3mm
+		] (scale) {Scale Contours};
+
+		\node[draw,
+			trapezium,
+			trapezium left angle = 65,
+			trapezium right angle = 115,
+			trapezium stretches,
+			below=of scale,
+			minimum width=3.5cm,
+			minimum height=1cm,
+			inner sep = 3mm
+		] (output) {Key-Edge Coordinatess Map};
+
+		\node[draw,
+			below=of output,
+			rounded rectangle,
+			minimum width=2.5cm,
+			minimum height=1cm,
+			inner sep = 3mm
+		] (end) {END};
+
+		\node[coordinate, right = 1cm of for] (top-right) {};
+		\node[coordinate, right = 1cm of output] (below-right) {};
+		\node[coordinate, left = 1cm of for] (top-left) {};
+		\node[coordinate, left = 2cm of scale] (below-left) {};
+
+		\draw[-latex] (start) edge (input)
+		(input) edge (for)
+		(for) edge (filter)
+		(for) -- (top-right) -- (below-right) |- (output)
+		(filter) edge (contours)
+		(contours) edge (scale)
+		(scale) -- (below-left) -- (top-left) |- (for)
+		(output) edge (end);
+
+	\end{tikzpicture}
+	\caption{Flowchart of Get Key Contour Points}
+	\label{fig:metho-algo-key-contour-points}
+\end{figure}
+
+\paragraph{Filter Virtual Map}
+The color that corresponds to each key in the virtual map is stored in the
+Key-Color Value Map. This color was used to create a mask for the virtual map
+for a specific key. This mask was obtained by using this snippet of code
+\texttt{(virtualMap == key).all(axis=2)}. This mask was then used to black out
+the rest of the virtual map, leaving only the \ac{roi} of the key in the image.
+
+This specific method was used as the other method that was trialed,
+\texttt{np.where(virtualMap != key)}, did not consider all 3 channels when
+comparing each pixel --- i.e.\ a key assigned with the color value of
+\texttt{[100, 0, 0]} will not be blacked out if the color value of the key to be
+isolated is \texttt{[100, 243, 0]} as the blue channel of the pixel,
+\texttt{100} is equal in both. This would result in other keys remaining in the
+virtual map, even if only one key corresponds to that color.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{filtered-key.png}
+	\caption{Filtered key}
+	\centering
+\end{figure}
+
+\paragraph{Find and Simplify Contours}
+The same method in finding and simplifying the contour of the keyboard was used
+to find and simplify the contour of the \ac{roi} of the key.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{key-contour.png}
+	\caption{Contour points of the \ac{roi}}
+	\centering
+\end{figure}
+
+\paragraph{Scale Contours}
+During testing using training data, a lot of the failures in finger-key
+identification was due to the tight fit of the contour to the key. The contour
+on its own did not give enough buffer for the placement of the finger. This
+buffer was needed as, in some instances, the finger used to press the key was
+not exactly on top of the key, but rather to its side. The algorithm accounts
+for this situation by adding a buffer. This was achieved through scaling the
+contour points by seven pixels on each side. This value was obtained after
+testing other possible values. A buffer of five pixels did not have a lot of
+effect in reducing the number of failures. A buffer of ten pixels did reduce the
+number of failures, but it also greatly increased the algorithm's uncertainty in
+determining the finger --- uncertainty is defined as detecting two or more
+fingers within one \ac{roi}. Seven was a good middle ground in decreasing
+failures, without greatly increasing uncertainty. More information about this
+can be found in the Results and Discussion at Section~\ref{section:results}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{scaled-contour.png}
+	\caption{Contour points of the \ac{roi}}
+	\centering
+\end{figure}
+
+\paragraph{Key-Edge Coordinatess Map}
+After all the keys have passed through the loop, a Key-Edge Coordinates Map was
+generated. Each key now has four points that serve as the coordinates of the
+edge of its \ac{roi}.
+
+The coordinate system that the contour points are based on had its origin at the
+top left, starting at 0, 0. Going to the right increased the value of the
+x-axis, and going to the bottom increased the value of the y-axis. This results
+in a coordinate system that only has positive values with each value
+corresponding to a pixel.
 
@@
 \label{section:metho-algo-finger}
-A computer vision algorithm will be chosen as a starting point for detecting,
-and tracking fingers within a video. The specific algorithm used for finger
-detection and tracking will be the solution offered by MediaPipe dubbed
-MediaPipe Hands. This algorithm is composed of two ML models working in
-conjunction to be able to detect the different parts of the hands and track them
-accurately.
+MediaPipe Hands was used as the finger detection and tracking solution. This
+algorithm is composed of two ML models working in conjunction to be able to
+detect the different parts of the hands and track them accurately.
 
@@
 
-The two previously chosen algorithms will be combined to accomplish finger-key
-identification and mapping. Finger-key identification refers to identifying
-which finger is used to press which key. As an example, the key Q was pressed
-with the pinky finger of the left hand.
-
-The integration of the algorithm is a two-step process. The first step is to get
-the mapping of the keyboard using the algorithm shown in
+The two previously chosen algorithms was combined to accomplish finger-key
+identification and mapping.
+
+The integration of the algorithm was a two-step process. The first step was to
+get the Key-Edge Coordinates Map shown in
 Section~\ref{section:metho-algo-keyboard}.
 
-The second step runs whenever a key press has been detected. This step uses the
-mapping of the keyboard in conjunction with the finger tracking algorithm
-selected in Section~\ref{section:metho-algo-finger}. The flowchart of the
-algorithm is shown in Figure~\ref{fig:metho-algo-integration}
+The second step runs whenever a key press has been detected. This step used the
+Key-Edge Coordinates Map in conjunction with the finger tracking algorithm selected
+in Section~\ref{section:metho-algo-finger}. The flowchart of the algorithm is
+shown in Figure~\ref{fig:metho-algo-integration}
 
@@
 			inner sep = 3mm
-		] (map) {Virtual Keyboard Map \\ Finger Tracking Data};
-
-		\node[draw,
-			below = of map,
-			minimum width = 3.5cm,
-			minimum height = 1cm,
-			inner sep = 3mm
-		] (find-roi) {Find \ac{roi} of the pressed key in Virtual Keyboard Map};
-
-		\node[draw,
-			below=of find-roi,
+		] (map) {Key-Edge Coordinates Map \\ Finger Tracking Data};
+
+		\node[draw,
+			below=of map,
 			minimum width = 3.5cm,
@@
 		\draw[-latex] (start) edge (if-packet)
-		(map) edge (find-roi)
-		(find-roi) edge (find-finger)
+		(map) edge (find-finger)
 		(find-finger) edge (return)
@@
 
-\subsubsection{Find \ac{roi} of the pressed key in Virtual Keyboard Map}
-Since there is a known color-key mapping based on the Virtual Keyboard Map, the
-algorithm will find the correct color to search for based on the received key
-press packet. This color is isolated from the Virtual Keyboard Map, and the
-\ac{roi} and its coordinates are obtained. The edges of this singular \ac{roi}
-will then be obtained by finding its contours. The specific implementation of
-this is explained in Section~\ref{section:metho-algo-key-contours}.
-
-The coordinate system will have its origin at the top left, starting at 0, 0.
-Going to the right increases the value of the x-axis, and going to the bottom
-increases the value of the y-axis. This will result in a coordinate system that
-only has positive values with each value corresponding to a pixel.
-
-The expected output of this is four sets of pairs. Each pair corresponds to an
-edge of the single \ac{roi}.
-
 \subsubsection{Find which fingertip is on that \ac{roi}}
-The Finger Tracking Data will contain the pixel positions of each landmark of
-the hand. For this step, the algorithm will find the landmark which is
-positioned within the single \ac{roi} obtained from the previous step.
-This will be done using a series of checks.
-
-Each landmark's coordinates will be compared to the coordinates of the edges of
-the \ac{roi}. A landmark will be determined as positioned within the \ac{roi} if all the
+The Finger Tracking Data contained the pixel positions of each landmark of the
+hand. For this step, the algorithm finds the landmark which is positioned within
+the single \ac{roi} obtained from the previous step. This was done using a
+series of checks.
+
+Each landmark's coordinates was compared to the coordinates of the edges of the
+\ac{roi}. A landmark was determined as positioned within the \ac{roi} if all the
 following conditions are true:
@@
 
-These conditions maximize the total area of the \ac{roi} and is not strict about
-exact accuracy. In essence, these conditions will create a perfectly rectangular
-box that contains the quadrilateral formed by the four points from the previous
-step.
+These conditions maximized the total area of the \ac{roi}, and it is not strict
+about exact accuracy. In essence, these conditions creates a perfectly
+rectangular box that contained the quadrilateral formed by the four points from
+the coordinates.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{detected-finger.png}
+	\caption{Finger tip within an \ac{roi}}
+	\centering
+\end{figure}
 
 \subsubsection{Return which fingertip}
-The landmark will then be used to determine which specific finger corresponds to
-the key press. Figure~\ref{fig:metho-algo-integration-landmarks} shows each
-possible landmark that may be returned from the previous step. This step will
-then return the name of the landmark, up until the first underscore. As an
-example, if the landmark found is \texttt{MIDDLE\_FINGER\_TIP}, this step will
-return \texttt{MIDDLE} denoting that the middle finger is the finger that
-corresponds with the key press. In addition, the specific hand will also be
-returned as one of two strings, \texttt{Left} and \texttt{Right}, since
-this information is also bundled together with the landmarks.
+The landmark was then used to determine which specific finger corresponds to the
+key press. Figure~\ref{fig:metho-algo-integration-landmarks} shows each possible
+landmark that may be returned from the previous step. This step returned the
+name of the landmark, up until the first underscore. As an example, if the
+landmark found is \texttt{MIDDLE\_FINGER\_TIP}, this step will return
+\texttt{MIDDLE} denoting that the middle finger is the finger that corresponds
+with the key press. In addition, the specific hand will also be returned as one
+of two strings, \texttt{LEFT} and \texttt{RIGHT}, since this information is also
+bundled together with the landmarks. The two strings were then concatenated with
+an underscore. An example return value is \texttt{RIGHT\_RING}.
 
@@
 \label{section:metho-algo-implementation}
-The previously discussed algorithms will be implemented using Python. The
-Keyboard Detection, and Mapping Algorithm in
-Section~\ref{section:metho-algo-keyboard} will be implemented as a separate
-submodule in Python using OpenCV.\@The Finger Detection, and Tracking Solution
-in Section~\ref{section:metho-algo-finger} will be consumed using the prebuilt
-Python package offered by the MediaPipe team. Finally, the integration of the
-two will be done as another separate submodule in Python.
-
-\section{Trainer}
+The previously discussed algorithms was implemented using Python. The Keyboard
+Detection, and Mapping Algorithm in Section~\ref{section:metho-algo-keyboard}
+was implemented as a separate submodule in Python using OpenCV. The Finger
+Detection, and Tracking Solution in Section~\ref{section:metho-algo-finger} was
+consumed using the prebuilt Python package offered by the MediaPipe team.
+Finally, the integration of the two will be done as another separate submodule
+in Python.
+
+\section{Training and Testing}
 
 \subsection{Typing Test Sequences}
-Typing test sequences are strings that will be used in testing the user in their
-ability to type. The test sequences that will be used in the trainer will come
-from text found in the public domain obtained from Project Gutenberg and the
-Internet Archive. Sentences will be isolated from these text and used as test
-sequences if they fit the criteria.
-
-The criteria for choosing test sequences is as follows: (1) $\approx80\%$ of the
+Typing test sequences are strings that was used in testing the user in their
+ability to type. The test sequences that was used came from text found in the
+public domain obtained from Project Gutenberg and the Internet Archive.
+Sentences were isolated from these text and used as test sequences if they fit
+the criteria.
+
+The criteria for choosing test sequences were as follows: (1) $\approx80\%$ of the
 characters in the keyboard is present in a test sequence. (2) The number of
@@
 
-There will be a total of 10 test sequences that will be gathered. An example
-test sequence is ``What of it, if some old hunks of a sea-captain orders me to
-get a broom and sweep down the decks?'' from Moby Dick by~\cite{moby-dick}. The
-9 other test sequences can be found in Appendix
+There was a total of 10 test sequences that was gathered. An example test
+sequence is ``What of it, if some old hunks of a sea-captain orders me to get a
+broom and sweep down the decks?'' from Moby Dick by~\cite{moby-dick}. The 9
+other test sequences can be found in Appendix
 Chapter~\ref{appendix:test-sequences}.
 
-\subsection{UX/UI design and Front-end development for the touch typing trainer}
-\subsubsection{Design}
-This design will be based on other type tests such as Monkeytype
-\parencite{bartnik2021}, and Keybr \parencite{keybr}, with additional components
-for showing real-time finger-key identification and mapping.
-Figure~\ref{fig:ui-test} and~\ref{fig:ui-stat} illustrates the two main pages of
-the design. Additional screens, such as historical statistics, in-platform
-tutorial, and a written guide are shown in the appendix.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{ui-test.png}
-	\caption{Test screen of the front-end design}
-	\label{fig:ui-test}
-	\centering
-\end{figure}
-
-Figure~\ref{fig:ui-test} has four main components. From the top, the first
-component is the navigation bar. This directs the user to the different pages
-within the website. The second component in the upper left displays the number
-of remaining words to be typed. This is 16 in the example. Below that is the
-test sequence. This shows the user what characters to type, and highlights
-mistakes in two distinct colors. Red if the character pressed was incorrect and
-purple if the character typed was correct, but the finger used to type it is
-wrong. The last main component in this screen is the keyboard that illustrates
-the correct finger positioning, with each finger corresponding to a different
-color.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{ui-stats.png}
-	\caption{Statistics screen of the front-end design}
-	\label{fig:ui-stat}
-	\centering
-\end{figure}
-
-Figure~\ref{fig:ui-stat} has two main components. The first component at the top
-shows the three major statistics for the typing test: Words per minute,
-Accuracy, and Finger Placement Accuracy. Finger Placement Accuracy will be
-explained in Section~\ref{section:metric}. The component below the statistics is a
-graph plotting the major statistics of the user in time. This includes when the
-user made the mistake and when the user slowed down, or increased their speed.
-
-\subsubsection{Implementation}
-The design will be created in Figma and the front-end will be web based. The
-framework the front-end will be developed in will be Svelte, due to its
-community resources, performance, and the researcher's familiarity with the
-framework. The front-end will serve as a display for the information and the
-interface which the user interacts and inputs information into the trainer. No
-calculation will be done within the front-end.
-
-\subsection{Back-end development of a touch typing trainer using the program created
-	in~\ref{section:metho-algo} and the experimental setup identified in~\ref{section:metho-setup}}
-
-The back-end will handle all the computation for the trainer---mainly
-finger-key identification. The backend will also handle authentication and data
-storage.
-
-The back-end will expose a REST API.\@This is a type of application programming
-interface (API) that conforms to the REST architectural style. One key
-characteristic of this style is its statelessness and cacheable data
-\parencite{rest}. This will be implemented using Django, a Python REST API
-framework.
-
-\subsubsection{Finger-key Identification}
-The submodule that integrates all the algorithms will be called within the
-framework. The framework will then respond to requests for finger-key
-identification using this submodule. This will allow the back-end to expose its
-finger-key identification abilities and allow the front-end to use this data for
-the trainer.
-
-\subsubsection{Authentication}
-The core authentication system will rely on \ac{jwt}. According to~\cite{jwt},
-``JSON Web Token (JWT) is a compact, URL-safe means of representing claims to be
-transferred between two parties.'' This allows for a user to claim that they are
-that specific user securely as the claim is cryptographically signed. For the
-backend of the trainer, the \ac{jwt} will be signed using a secret, and it will
-be encoded in HMAC SHA256.
-
-\subsubsection{Data Storage}
-The trainer will store all the data in a relational database using
-PostgreSQL.\@This specific technology was chosen due to its maturity, community,
-and industry recommendations and the researchers' familiarity with the
-technology.
-
-\subsection{Testing of finger-key identification and mapping accuracy}
-
-There will be 3 groups of 10 videos to be captured by the researcher, for a
+\subsection{Data Gathering}
+\subsubsection{Video Capture}
+There were 3 groups of 10 videos that were captured by the researcher, for a
 total of 30 videos. Each group corresponds to three predetermined speeds of
-typing: slow (15wpm), average (35wpm), and fast (80wpm). These speeds are based
-on test data from~\cite{keybr}. The researcher will type the 10 typing test
-sequences at these predetermined speeds. During typing, errors will not be
-consciously taken into account, and errors will happen naturally as part of the
+typing: slow (15wpm), average (35wpm), and fast (80wpm). These speeds were based
+on test data from~\cite{keybr}. The researcher typed the 10 typing test
+sequences at these predetermined speeds. During typing, errors were not
+consciously taken into account, and errors happened naturally as part of the
 typing process.
 
-The researcher will then manually perform finger-key identification for all key
-presses present in the video. The annotated data will then be divided into
-training and test data at a ratio of 60:40. The finger-key identification
-submodule will then perform the finger-key identification for the same videos.
-The accuracy of the submodule will then be obtained by verifying if its
-identification is as expected based on the manual annotation.
+A Python script was created to facilitate this process. The script captured a
+video of the researcher as the researcher was typing the test sequences.
+Whenever the researcher pressed a key, the corresponding frame count was
+captured, and the key pressed was recorded in conjunction. This was stored in
+the format \texttt{FRAME\_NO:KEY\_PRESSED}, and each keypress was then stored as
+a line in a file.
+
+\subsubsection{Keypress Labelling}
+The researcher then manually perform finger-key identification for all key
+presses present in the video. This was also done through a Python script. The
+script parsed the file associated with a video. The frame number found at each
+line was then shown, and the associated data with the frame was superimposed
+over it. The researcher then pressed keys that corresponded with the finger used
+to press the key. This was then again stored in the format
+\texttt{FRAME\_NO:KEY\_PRESSED:FINGER\_USED}, and each keypress was then stored
+as a line in a file.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.6\textwidth]{labelling.png}
+	\caption{Screen showing the labelling process}
+	\centering
+\end{figure}
+
+\subsubsection{Test-Training Split}
+The annotated data was then divided into training and test data at a ratio of
+60:40.
+
+\subsection{Training}
+A script was created to test the accuracy of the submodule. This script opened
+each video, and the associated labeled file. The script then parsed the labeled
+file, and opened the frame for each keypress. The finger-key identification
+submodule was then run, using the data needed for the submodule: the key
+pressed. The result of the submodule was then compared with the manual
+annotation of the keypress. The resulting data from the test was then stored in
+a \texttt{.csv} file with the following content:
+
+\begin{table}[H]
+	\centering
+	\begin{tabular}{ p{0.3\textwidth} p{0.6\textwidth} }
+		\hline
+		Column Name                           & Definition                                             \\
+		\hline
+		\texttt{file\_name}                   & Name of the file                                       \\
+		\texttt{frame}                        & Frame number                                           \\
+		\texttt{key}                          & Key pressed                                            \\
+		\texttt{finger}                       & Manual annotation of the finger used to press the key  \\
+		\texttt{is\_correct}                  & If the submodule was correct                           \\
+		\texttt{fingertips\_detected}         & The fingertips detected by the submodule               \\
+		\texttt{keyboard\_detection\_time}    & Time spent in creating the Key-Edge Coordinates Map    \\
+		\texttt{finger\_classification\_time} & Time spent in classifying which finger pressed the key \\
+	\end{tabular}
+	\caption{\label{tab:csv-output}CSV File Contents}
+\end{table}
+
+Adjustments were then made to the algorithm after each training pass to improve
+its performance. Some runs had improvements in its statistics without doing any
+adjustments. These were runs that showed errors in the manual annotation and
+were corrected afterwards.
+
+\subsection{Analysis}
+The resulting \texttt{.csv} file was imported into Google Sheets and key
+statistics were obtained based on the data. These statistics can be found on
+Table~\ref{tab:acquired-stats}. These key statistics informed what to adjust in
+the algorithm and its overall performance.
+
+\begin{table}[H]
+	\centering
+	\begin{tabular}{ p{0.3\textwidth} p{0.6\textwidth} }
+		\hline
+		Identifications                                                                                                                       \\
+		\hline
+		Success                                     & Number of successful finger-key identifications                                         \\
+		Failed                                      & Number of failed finger-key identifications                                             \\
+		Success Percentage                          & Percentage of successful finger-key identifications                                     \\
+		Failed Percentage                           & Percentage of failed finger-key identifications                                         \\
+		\\[0.25cm]
+		\hline
+		Failure Types                                                                                                                         \\
+		\hline
+		Mismatched Identification                   & Submodule output did not match manual annotation                                        \\
+		No Identification                           & Submodule did not detect finger tips in \ac{roi}                                        \\
+		\\[0.25cm]
+		\hline
+		Uncertain Successes                                                                                                                   \\
+		\hline
+		Total                                       & Total uncertain successes, defined as detecting two or more fingers within one \ac{roi} \\
+		Spacebar                                    & All uncertain successes that were on the spacebar                                       \\
+		Spacebar Percentage                         & Percentage of spacebar uncertain success over all uncertain successes                   \\
+		Non-spacebar                                & All other uncertain successes                                                           \\
+		Non-spacebar Percentage                     & Percentage of non-spacebar uncertain success over all uncertain successes               \\
+		Non-spacebar over total \newline Percentage & Percentage of non-spacebar uncertain success over all total keypresses                  \\
+		\\[0.25cm]
+		\hline
+		Average Running Time                                                                                                                  \\
+		\hline
+		Keyboard Detection Time                     & Average keyboard detection time of all videos                                           \\
+		Finger Identification Time                  & Average finger identification time of all key presses                                   \\
+	\end{tabular}
+	\caption{\label{tab:acquired-stats}Acquired statistics}
+\end{table}
+
+\subsection{Testing}
+The same script was run through the test data. The same format of output was
+gathered, and the same key statistics were recorded.
+
+%% \section{Trainer}
+
+%% \subsection{UX/UI design and Front-end development for the touch typing trainer}
+%% \subsubsection{Design}
+%% This design will be based on other type tests such as Monkeytype
+%% \parencite{bartnik2021}, and Keybr \parencite{keybr}, with additional components
+%% for showing real-time finger-key identification and mapping.
+%% Figure~\ref{fig:ui-test} and~\ref{fig:ui-stat} illustrates the two main pages of
+%% the design. Additional screens, such as historical statistics, in-platform
+%% tutorial, and a written guide are shown in the appendix.
+
+%% \begin{figure}[H]
+%% 	\centering
+%% 	\includegraphics[width=0.8\textwidth]{ui-test.png}
+%% 	\caption{Test screen of the front-end design}
+%% 	\label{fig:ui-test}
+%% 	\centering
+%% \end{figure}
+
+%% Figure~\ref{fig:ui-test} has four main components. From the top, the first
+%% component is the navigation bar. This directs the user to the different pages
+%% within the website. The second component in the upper left displays the number
+%% of remaining words to be typed. This is 16 in the example. Below that is the
+%% test sequence. This shows the user what characters to type, and highlights
+%% mistakes in two distinct colors. Red if the character pressed was incorrect and
+%% purple if the character typed was correct, but the finger used to type it is
+%% wrong. The last main component in this screen is the keyboard that illustrates
+%% the correct finger positioning, with each finger corresponding to a different
+%% color.
+
+%% \begin{figure}[H]
+%% 	\centering
+%% 	\includegraphics[width=0.8\textwidth]{ui-stats.png}
+%% 	\caption{Statistics screen of the front-end design}
+%% 	\label{fig:ui-stat}
+%% 	\centering
+%% \end{figure}
+
+%% Figure~\ref{fig:ui-stat} has two main components. The first component at the top
+%% shows the three major statistics for the typing test: Words per minute,
+%% Accuracy, and Finger Placement Accuracy. Finger Placement Accuracy will be
+%% explained in Section~\ref{section:metric}. The component below the statistics is a
+%% graph plotting the major statistics of the user in time. This includes when the
+%% user made the mistake and when the user slowed down, or increased their speed.
+
+%% \subsubsection{Implementation}
+%% The design will be created in Figma and the front-end will be web based. The
+%% framework the front-end will be developed in will be Svelte, due to its
+%% community resources, performance, and the researcher's familiarity with the
+%% framework. The front-end will serve as a display for the information and the
+%% interface which the user interacts and inputs information into the trainer. No
+%% calculation will be done within the front-end.
+
+%% \subsection{Back-end development of a touch typing trainer using the program created
+%% 	in~\ref{section:metho-algo} and the experimental setup identified in~\ref{section:metho-setup}}
+
+%% The back-end will handle all the computation for the trainer---mainly
+%% finger-key identification. The backend will also handle authentication and data
+%% storage.
+
+%% The back-end will expose a REST API.\@This is a type of application programming
+%% interface (API) that conforms to the REST architectural style. One key
+%% characteristic of this style is its statelessness and cacheable data
+%% \parencite{rest}. This will be implemented using Django, a Python REST API
+%% framework.
+
+%% \subsubsection{Finger-key Identification}
+%% The submodule that integrates all the algorithms will be called within the
+%% framework. The framework will then respond to requests for finger-key
+%% identification using this submodule. This will allow the back-end to expose its
+%% finger-key identification abilities and allow the front-end to use this data for
+%% the trainer.
+
+%% \subsubsection{Authentication}
+%% The core authentication system will rely on \ac{jwt}. According to~\cite{jwt},
+%% ``JSON Web Token (JWT) is a compact, URL-safe means of representing claims to be
+%% transferred between two parties.'' This allows for a user to claim that they are
+%% that specific user securely as the claim is cryptographically signed. For the
+%% backend of the trainer, the \ac{jwt} will be signed using a secret, and it will
+%% be encoded in HMAC SHA256.
+
+%% \subsubsection{Data Storage}
+%% The trainer will store all the data in a relational database using
+%% PostgreSQL.\@This specific technology was chosen due to its maturity, community,
+%% and industry recommendations and the researchers' familiarity with the
+%% technology.
 
@@
 
-There will be two total metrics obtained pertaining to finger and key mapping.
-The first will be per test sequence, and the second will be per key.
+There was a two total metrics obtained pertaining to finger and key mapping.
+The first was per test sequence, and the second was per key.
 
 \subsection{Per Test Sequence}
-The metric to be calculated per test sequence is \ac{fpacc}. This metric will
-compute the percentage of accurate keys pressed with the correct finger over the
+The metric which calculates per test sequence is \ac{fpacc}. This metric
+computes the percentage of accurate keys pressed with the correct finger over the
 length of the typing test sequence. Inputting the wrong character, even if
-pressed with the correct finger will reduce \ac{fpacc} since \ac{fpacc} measures
-\emph{accurate key presses}, and incorrect characters will be considered as
+pressed with the correct finger, reduces \ac{fpacc} since \ac{fpacc} measures
+\emph{accurate key presses}, and incorrect characters areconsidered as
 inaccurate key presses.
@@
 \subsection{Per Key}
-This metric will compute the \ac{hfpacc} for a certain key. This metric will
-take the times the user pressed the wrong key when asked for a certain key into
-account, the same as in \ac{fpacc}.\@This will allow the user to easily verify
-which keys need attention and retraining if there is a high frequency of error.
+This metric computes the \ac{hfpacc} for a certain key. This metric takes the
+times the user pressed the wrong key when asked for a certain key into account,
+the same as in \ac{fpacc}.\@This allows the user to easily verify which keys
+need attention and retraining if there is a high frequency of error.
 
@@
 for the user.
+
+\chapter{Results and Discussion}
+\chapter{Conclusion}
+\chapter{Future Work}
 

### 2022-05-18 application/algorithms/common.py

@@ -3,7 +3,7 @@ import numpy as np
 from cv2 import cv2 as cv
 
 
-DEBUG = False
+DEBUG = True
 
 
 def findContours(frame: np.ndarray) -> tuple[np.ndarray]:
@@ -101,6 +101,7 @@ def DEBUG_displayContours(frame: np.ndarray, contour: np.ndarray):
                 thickness=cv.FILLED,
             )
         cv.imshow("debug", contourFrame)
+        cv.imwrite("./key-contour.png", contourFrame)
         cv.waitKey(0)
 
 

### 2022-05-18 application/algorithms/integration.py

@@ -20,7 +20,7 @@ def main(keyboardImage: np.ndarray, handImage: np.ndarray):
     # Step 1
     allContourPoints = getContourPointsFromImage(keyboardImage)
 
-    contourPoints = allContourPoints["Y"]
+    contourPoints = allContourPoints["B"]
 
     # elapsed = time.time() - start
 
@@ -49,6 +49,6 @@ def main(keyboardImage: np.ndarray, handImage: np.ndarray):
 
 
 if __name__ == "__main__":
-    keyboardImage = cv.imread("images/medium-06-10.png")
-    handImage = cv.imread("images/medium-06-418.png")
+    keyboardImage = cv.imread("images/fast-02-20.png")
+    handImage = cv.imread("images/fast-02-380.png")
     main(keyboardImage, handImage)

### 2022-05-18 application/algorithms/keyboard_detection.py

@@ -35,8 +35,16 @@ def getKeyboardImage() -> np.ndarray | None:
 def getEdgeFrame(frame: np.ndarray) -> np.ndarray:
     grayscaleFrame = cv.cvtColor(frame, cv.COLOR_BGR2GRAY)
 
+    if DEBUG:
+        cv.imshow("debug", grayscaleFrame)
+        cv.waitKey(0)
+
     # Denoise while maintaining edges
     bilateralFilteredFrame = cv.bilateralFilter(grayscaleFrame, 11, 17, 17)
+
+    if DEBUG:
+        cv.imshow("debug", bilateralFilteredFrame)
+        cv.waitKey(0)
 
     # Perform Sobel Filtering
     sobelX = cv.Sobel(
@@ -61,6 +69,10 @@ def getEdgeFrame(frame: np.ndarray) -> np.ndarray:
     absSobelY = cv.convertScaleAbs(sobelY)
 
     finalSobel = cv.addWeighted(absSobelX, 0.5, absSobelY, 0.5, 0)
+
+    if DEBUG:
+        cv.imshow("debug", finalSobel)
+        cv.waitKey(0)
 
     # Use Otsu's Threshold to get better defined edges while not using arbitrary
     # values
@@ -170,7 +182,10 @@ def getKeyContourPoints(virtualMap: np.ndarray, key: list[int]) -> np.ndarray:
     grayMap = cv.cvtColor(filteredMap, cv.COLOR_BGR2GRAY)
 
     contours = findContours(grayMap)
+
     largestContour = approximateLargestContour(contours)
+
+    DEBUG_displayContours(grayMap, np.reshape(largestContour, (4, 1, 2)))
 
     largestContour = np.reshape(largestContour, (4, 2))
     sortedContour = sortContourPoints(largestContour)

### 2022-05-18 application/algorithms/test-scripts/get-image.py

@@ -4,9 +4,9 @@ from cv2 import cv2 as cv
 
 
 def main():
-    file = "medium-06"
-    keyboardFrame = 10
-    fingerFrame = 418
+    file = "fast-02"
+    keyboardFrame = 20
+    fingerFrame = 380
     fileLocation = f"../dataset/training/{file}"
 
     cap = cv.VideoCapture(f"{fileLocation}.mp4")

### 2022-06-02 paper/paper.tex
rd: add research and discussion

@@
 \graphicspath{ {./images/} }
-
-% Table of Content
-\setcounter{tocdepth}{2}
 
@@
 % Links
-\usepackage{hyperref}
+\usepackage[linktoc=all]{hyperref}
 \hypersetup{
-	pdftitle={Development of a finger-key identification submodule for a touch typing trainer},
+	pdftitle={Development of a Finger-Key Identification Module for a Touch Typing Trainer},
 	pdfstartview={FitH},
@@
 
+% Graphs
+\usepackage{pgfplots}
+\usepgfplotslibrary{external}
+\pgfplotsset{width=1\textwidth,compat=1.9}
+
 % Tables
@@
 
-\usepackage{caption}
+% Center table with controlled width
+\usepackage{array}
+\newcolumntype{P}[1]{>{\centering\arraybackslash}p{#1}}
+\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+
+% List of Equations
+\usepackage[titles]{tocloft}
+\newcommand{\listequationsname}{List of Equations}
+\newlistof{equations}{equ}{\listequationsname}
+\newcommand{\labelequations}[1]{%
+	\addcontentsline{equ}{equations}{\numberline{\theequation}#1}\par}
+\setlength{\cftequationsindent}{1.5em}
+\setlength{\cftequationsnumwidth}{2.3em}
+
+\addto\captionsamerican{% Replace "english" with the language you use
+	\renewcommand{\contentsname}%
+	{Table of Contents}%
+}
 
@@
 
-	\textsc{Development of a finger-key identification submodule\\for a touch typing trainer}
+	\textsc{Development of a Finger-Key Identification Module\\for a Touch Typing Trainer}
 	\par\vspace{0.5cm}
@@
 
-	February 2022
+	June 2022
 	\vfill
@@
 
+\begin{center}
+
+	\hspace{0pt}
+	\vfill
+
+	\includegraphics[width=0.2\textwidth]{upc.png}
+	\includegraphics[width=0.2\textwidth]{dcs.png}
+	\par\vspace{0.25cm}
+
+	\textbf{\uppercase{University of the Philippines Cebu}}\\
+	College of Science\\
+	Department of Computer Science
+
+	\par\vspace{1cm}
+
+	\textbf{\uppercase{Development of a finger-key identification module\\for a touch typing trainer}}
+
+	\par\vspace{1cm}
+
+	\textbf{Permission is given for the following people to have access to this thesis:}
+
+	\par\vspace{0.25cm}
+
+	\begin{tabular}{ | P{0.8\textwidth} | P{0.1\textwidth} | }
+		\hline
+		Available to the general public                                 & Yes \\
+		\hline
+		Available only after consultation with author or thesis adviser & Yes \\
+		\hline
+		Available to those bound by confidentiality agreement           & Yes \\
+		\hline
+	\end{tabular}
+	\par\vspace{1cm}
+
+	\line(1,0){0.6\textwidth}\\
+	Oscar Vian L. Valles\\
+	Student
+	\par\vspace{0.5cm}
+
+	\line(1,0){0.6\textwidth}\\
+	Dhong Fhel K. Gom-os\\
+	Special Problem Adviser
+
+	\vfill
+	\hspace{0pt}
+
+\end{center}
+
+\newpage
+
+\noindent
+This thesis entitled ``\textbf{\uppercase{Development of a finger-key
+		identification \\module for a touch typing trainer}}'' prepared and
+submitted by \\\uppercase{Oscar Vian L. Valles} is approved by:
+
+\par\vspace{2cm}
+
+\begin{center}
+	\begin{minipage}[t]{0.45\textwidth}
+		\centering
+		\line(1,0){0.9\textwidth}\\
+		Prof. Aileen Joan O. Vicente\\
+		Panel Member
+
+		\par\vspace{0.5cm}
+		\line(1,0){0.9\textwidth}\\
+		Date
+	\end{minipage}
+	\begin{minipage}[t]{0.45\textwidth}
+		\centering
+		\line(1,0){0.9\textwidth}\\
+		Prof. Aileen Joan O. Vicente\\
+		Panel Member
+
+		\par\vspace{0.5cm}
+		\line(1,0){0.9\textwidth}\\
+		Date
+	\end{minipage}
+
+	\par\vspace{2cm}
+
+	\line(1,0){0.405\textwidth}\\
+	Prof. Aileen Joan O. Vicente\\
+	Panel Chair
+
+	\par\vspace{0.5cm}
+	\line(1,0){0.405\textwidth}\\
+	Date
+
+	\par\vspace{2cm}
+
+	\begin{minipage}{0.45\textwidth}
+		\centering
+		\line(1,0){0.9\textwidth}\\
+		Prof. Aileen Joan O. Vicente\\
+		Chair, Department of Computer Science
+
+		\par\vspace{0.5cm}
+		\line(1,0){0.9\textwidth}\\
+		Date
+	\end{minipage}
+	\begin{minipage}{0.45\textwidth}
+		\centering
+		\line(1,0){0.9\textwidth}\\
+		Prof. Nelia S. Ereno\\
+		Dean, College of Science
+
+		\par\vspace{0.5cm}
+		\line(1,0){0.9\textwidth}\\
+		Date
+	\end{minipage}
+\end{center}
+
+\newpage
+
+\pagenumbering{roman}
+\addcontentsline{toc}{chapter}{Abstract}
+\begin{center}
+	\LARGE\textbf{Abstract}
+\end{center}
+
+\newpage
+
+\addcontentsline{toc}{chapter}{Acronyms}
 \printacronyms{}
@@
 
+\addcontentsline{toc}{chapter}{List of Tables}
 \listoftables
@@
 
+\addcontentsline{toc}{chapter}{List of Figures}
 \listoffigures
+\newpage
+
+\addcontentsline{toc}{chapter}{List of Equations}
+\listofequations
+
 \newpage
@@
 
-Thus, there is a need for automatically detecting which finger is used during
-typing. One way to do this is through finger and hand tracking. One solution for
-tracking is by using image processing and machine learning. An example of this
-is MediaPipe by~\cite{mediapipe}.
-
-MediaPipe allows for various applications for machine learning in the field of
-image processing. This includes, hand tracking, pose estimation, object
-detection, and others. Another example of a library that allows for hand and
-finger tracking is OpenCV by~\cite{opencv}. This is a tool that simplifies
-computer vision and image processing. Machine learning can also be used with
-OpenCV.\@
+Thus, there is a need for automatically identifying which finger is used to
+press which key during typing --- hereby defined as finger-key identification.
 
@@
 	\item What algorithm or technique using computer vision is capable of
-	      identifying which keys are being pressed by which fingers during keyboard
-	      typing?
+	      finger-key identification.
 	\item What setup and configuration of a single optical camera can be used to
-	      capture images during keyboard typing for real-time finger-key identification
+	      capture images during keyboard typing for time finger-key identification
 	      using the algorithm identified in 1?
-	\item How to design and develop a touch typing trainer using the algorithm in 1
-	      and camera setup and configuration in 2?
-	\item How accurate is the touch typing trainer in identifying which finger is
-	      used to press which key?
-	\item What new keyboard typing metric can be developed in order to consider both
-	      the key and finger used to press as parameters?
+	\item How to design and develop a finger-key identification module using
+	      the algorithm in 1 and camera setup and configuration in 2?
+	\item How accurate is the module developed in 3 in finger-key identification?
+	\item Is the module developed in 3 fast enough to run in real-time?
+	\item What new keyboard typing metric to develop that considers both the key and
+	      finger used to press as parameters?
 \end{enumerate}
@@
 \subsection{General Objectives}
-To create a touch typing trainer that detects poor finger placement and hand
-position to help develop better typing habits and healthier typing ergonomics.
+To create a finger-key identification module that accurately identifies what
+finger was used to press a key at a point in time to help develop better typing
+habits and healthier typing ergonomics.
 
@@
 	\item Develop an algorithm or technique using computer vision that is capable
-	      of identifying which keys are being pressed by which fingers during keyboard
-	      typing --- i.e.\ finger-key identification.
+	      of finger-key identification.
 	\item Identify the setup and configuration of a single optical camera that can
-	      be used to capture images during keyboard typing for real-time finger-key
+	      be used to capture images during keyboard typing for finger-key
 	      identification using the algorithm identified in 1
-	\item Design and develop a touch typing trainer using the algorithm in 1 and
-	      camera setup and configuration in 2
-	\item Determine the accuracy of the touch typing trainer in finger-key
+	\item Design and develop a module using the algorithm in 1 and camera setup
+	      and configuration in 2
+	\item Determine the accuracy of the module developed in 3 in finger-key
 	      identification
+	\item Determine if the module developed in 3 is fast-enough to run in
+	      real-time
 	\item Develop a new keyboard typing metric that uses both the key and
@@
 The setup that will be captured will be simplified and fixed for the development
-and testing of the finger-key identification submodule. This includes the
+and testing of the finger-key identification module. This includes the
 keyboard, surface it is placed on, and lighting of the entire area. Furthermore,
@@
 --- as the focus is on the development and testing of the finger-key
-identification submodule and not on the touch typing trainer and its effects on
+identification module and not on the touch typing trainer and its effects on
 typing.
@@
 	\centering
-	\includegraphics[width=0.7\textwidth]{60.jpg}
-	\caption{A 60\% keyboard in \ac{ansi} layout. Reprinted from \fullcite{60}}
+	\includegraphics[width=0.7\textwidth]{60.png}
+	\caption{A 60\% keyboard in \ac{ansi} layout.}
 	\label{fig:60}
@@
 This research is beneficial for all users of physical keyboards. These include a
-vast majority of the population as there are a lot of professions that heavily
-rely on keyboards. Examples include developers, physicians, educators, and
+majority of the population as there are a lot of professions that heavily rely
+on keyboards. Examples include developers, physicians, educators, and
 accountants. By having better ergonomics while typing --- by touch typing
@@
 This research also helps educators, especially early educators teaching beginner
-typists. By automatically checking for ergonomics, posture, and correct
-technique, the burden of checking each student is lessened, and directed
-interventions for bad habits can be easily created as students with these bad
-habits are easily identified
-
-This research has a direct impact on people that has hand or wrist injuries that
-are caused by poor typing habits. By correcting these poor habits, pain from
-these injuries will be lessened, and even be prevented from occurring in the
+typists. By automatically checking for correct technique, the burden of checking
+each student is reduced, and directed interventions for bad habits can be
+easily created as students with these bad habits are easily identified
+
+This research has a direct impact on people that have hand or wrist injuries
+that are caused by poor typing habits. By correcting these poor habits, pain
+from these injuries will be reduced, and even be prevented from occurring in the
 first place. A specific example of this is by reducing ulnar deviation which
@@
 using a physical keyboard to input characters in a computer system.
-
-% Insert more about keyboard typing here %
 
@@
 	WPM = \frac{|T| - 1}{S} \cdot 60 \cdot \frac{1}{5}
+	\label{equ:wpm}
 \end{equation}
+\labelequations{Words per Minute}
 
@@
 \end{equation}
+\labelequations{Accuracy}
 
@@
 \end{figure}
+
+\subsubsection{Average Typing Speed}
+
+Average typing speed using a keyboard with QWERTY layout, based on test data
+from \cite{keybr}, is $\approx37.5$ \ac{wpm}. This average is taken from all of
+the test results they have gathered from all of their users.
+\cite{arif2009} also performed a survey with 12 participants using
+multiple typing methods. They got an average \ac{wpm} of 75.85 with a standard
+deviation of 15.61 when their participants used a physical QWERTY layout. This
+average \ac{wpm} was also the highest among the other typing methods the
+researchers studied.
 
@@
 
+\subsection{Computer}
+The computer used to run the algorithm is a desktop computer with the following
+specification:
+
+\begin{table}[H]
+	\small
+	\centering
+	\begin{tabular}{ p{0.3\textwidth} p{0.6\textwidth} }
+		\toprule
+		Component & Specification                          \\
+		\midrule
+		CPU       & AMD Ryzen 5 3600                       \\
+		GPU       & AMD Radeon RX 5600 XT                  \\
+		RAM       & G.Skill Trident Z Neo RGB 16GB 3200mhz \\
+		Storage   & Samsung SSD 850 EVO                    \\
+		OS        & Arch Linux                             \\
+		\bottomrule
+	\end{tabular}
+	\caption{\label{tab:table-specs}Computer Specifications}
+\end{table}
+
 \subsection{Environment}
@@
 			inner sep = 3mm
-		] (output) {Key-Edge Coordinatess Map};
+		] (output) {Key-Edge Coordinates Map};
 
@@
 fingers within one \ac{roi}. Seven was a good middle ground in decreasing
-failures, without greatly increasing uncertainty. More information about this
-can be found in the Results and Discussion at Section~\ref{section:results}
+failures, without greatly increasing uncertainty.
 
@@
 	\includegraphics[width=0.8\textwidth]{detected-finger.png}
-	\caption{Finger tip within an \ac{roi}}
+	\caption{Fingertip within an \ac{roi}}
 	\centering
@@
 
-As such, the expected output of the submodule is the hand and the finger used to
+As such, the expected output of the module is the hand and the finger used to
 press the key. This allows for a more flexible utilization of the data as it
@@
 Detection, and Mapping Algorithm in Section~\ref{section:metho-algo-keyboard}
-was implemented as a separate submodule in Python using OpenCV. The Finger
+was implemented as a separate module in Python using OpenCV. The Finger
 Detection, and Tracking Solution in Section~\ref{section:metho-algo-finger} was
 consumed using the prebuilt Python package offered by the MediaPipe team.
-Finally, the integration of the two will be done as another separate submodule
+Finally, the integration of the two will be done as another separate module
 in Python.
@@
 
+One caveat of the camera used was it's autofocus capabilities. This meant that
+there was a set amount of time during the start of the recording where the
+camera captured blurry images of the keyboard. This necessitates around three
+seconds for the autofocus to complete and focus on the keyboard. There was one
+video capture where the whole test sequence was scrapped as the researcher
+started typing even if the camera was not focused.
+
 \subsubsection{Keypress Labelling}
@@
 \subsection{Training}
-A script was created to test the accuracy of the submodule. This script opened
-each video, and the associated labeled file. The script then parsed the labeled
-file, and opened the frame for each keypress. The finger-key identification
-submodule was then run, using the data needed for the submodule: the key
-pressed. The result of the submodule was then compared with the manual
-annotation of the keypress. The resulting data from the test was then stored in
-a \texttt{.csv} file with the following content:
+A script was created to test the accuracy of the module. This script opened each
+video, and the associated labeled file. There were three frames taken for each
+video to perform step one of the module: frames ten, fifteen, and twenty. The
+first frame that returned a Key-Edge Coordinates Map after going through step
+one of the module was used. The rest of the frames were not passed through the
+module and were discarded.
+
+The script then parsed the labeled file, and opened the frame for each keypress.
+The second algorithm of the module was then run, using the data needed for the
+algorithm: the key pressed. The result of the module was then compared with the
+manual labelling of the keypress. The resulting data from the test was then
+stored in a \texttt{.csv} file with the following content:
 
 \begin{table}[H]
+	\small
 	\centering
 	\begin{tabular}{ p{0.3\textwidth} p{0.6\textwidth} }
-		\hline
+		\toprule
 		Column Name                           & Definition                                             \\
-		\hline
+		\midrule
 		\texttt{file\_name}                   & Name of the file                                       \\
@@
 		\texttt{key}                          & Key pressed                                            \\
-		\texttt{finger}                       & Manual annotation of the finger used to press the key  \\
-		\texttt{is\_correct}                  & If the submodule was correct                           \\
-		\texttt{fingertips\_detected}         & The fingertips detected by the submodule               \\
+		\texttt{finger}                       & Manual labelling of the finger used to press the key   \\
+		\texttt{is\_correct}                  & If the module was correct                              \\
+		\texttt{fingertips\_detected}         & The fingertips detected by the module                  \\
 		\texttt{keyboard\_detection\_time}    & Time spent in creating the Key-Edge Coordinates Map    \\
 		\texttt{finger\_classification\_time} & Time spent in classifying which finger pressed the key \\
+		\bottomrule
 	\end{tabular}
@@
 its performance. Some runs had improvements in its statistics without doing any
-adjustments. These were runs that showed errors in the manual annotation and
-were corrected afterwards.
+adjustments. These were runs that showed errors in the manual labelling and
+were corrected afterward.
 
@@
 \begin{table}[H]
+	\small
 	\centering
 	\begin{tabular}{ p{0.3\textwidth} p{0.6\textwidth} }
-		\hline
+		\toprule
 		Identifications                                                                                                                    \\
-		\hline
-		Success                                     & Number of successful finger-key identifications                                         \\
-		Failed                                      & Number of failed finger-key identifications                                             \\
+		\midrule
+		Successes                                & Number of successful finger-key identifications                                         \\
+		Failures                                 & Number of failed finger-key identifications                                             \\
 		Success Percentage                       & Percentage of successful finger-key identifications                                     \\
-		Failed Percentage                           & Percentage of failed finger-key identifications                                         \\
+		Failure Percentage                       & Percentage of failed finger-key identifications
 		\\[0.25cm]
-		\hline
+		\midrule
 		Failure Types                                                                                                                      \\
-		\hline
-		Mismatched Identification                   & Submodule output did not match manual annotation                                        \\
-		No Identification                           & Submodule did not detect finger tips in \ac{roi}                                        \\
+		\midrule
+		Mismatched Identification                & Module output did not match manual labelling                                            \\
+		No Identification                        & Module did not detect fingertips in \ac{roi}
 		\\[0.25cm]
-		\hline
+		\midrule
 		Uncertain Successes                                                                                                                \\
-		\hline
+		\midrule
 		Total                                    & Total uncertain successes, defined as detecting two or more fingers within one \ac{roi} \\
 		Spacebar                                 & All uncertain successes that were on the spacebar                                       \\
-		Spacebar Percentage                         & Percentage of spacebar uncertain success over all uncertain successes                   \\
+		Spacebar Percentage                      & Percentage of spacebar uncertain successes over all uncertain successes                 \\
+		Spacebar Percentage (Total)              & Percentage of spacebar uncertain successes over total keypresses                        \\
 		Non-spacebar                             & All other uncertain successes                                                           \\
-		Non-spacebar Percentage                     & Percentage of non-spacebar uncertain success over all uncertain successes               \\
-		Non-spacebar over total \newline Percentage & Percentage of non-spacebar uncertain success over all total keypresses                  \\
+		Non-spacebar Percentage                  & Percentage of non-spacebar uncertain successes over all uncertain successes             \\
+		Non-spacebar Percentage \newline (Total) & Percentage of non-spacebar uncertain successes over total keypresses
 		\\[0.25cm]
-		\hline
+		\midrule
 		Average Running Time                                                                                                               \\
-		\hline
+		\midrule
 		Keyboard Detection Time                  & Average keyboard detection time of all videos                                           \\
 		Finger Identification Time               & Average finger identification time of all key presses                                   \\
+		\bottomrule
 	\end{tabular}
@@
 %% \subsubsection{Finger-key Identification}
-%% The submodule that integrates all the algorithms will be called within the
+%% The module that integrates all the algorithms will be called within the
 %% framework. The framework will then respond to requests for finger-key
-%% identification using this submodule. This will allow the back-end to expose its
+%% identification using this module. This will allow the back-end to expose its
 %% finger-key identification abilities and allow the front-end to use this data for
@@
 
-\section{Metric}
+\section{Metrics}
 \label{section:metric}
@@
 pressed with the correct finger, reduces \ac{fpacc} since \ac{fpacc} measures
-\emph{accurate key presses}, and incorrect characters areconsidered as
+\emph{accurate key presses}, and incorrect characters are considered as
 inaccurate key presses.
@@
 \end{equation}
+\labelequations{Finger Placement Accuracy}
 
@@
 times the user pressed the wrong key when asked for a certain key into account,
-the same as in \ac{fpacc}.\@This allows the user to easily verify which keys
+the same as in \ac{fpacc}. This allows the user to easily verify which keys
 need attention and retraining if there is a high frequency of error.
@@
 \end{equation}
+\labelequations{Historical Finger Placement Accuracy}
 
@@
 \chapter{Results and Discussion}
+\label{section:rd}
+
+\section{Accuracy}
+Accuracy, in the context of the finger-key identification module calculates the
+percentage of successful finger-key identifications that the module has
+performed. In this case, success is defined as having the expected finger, based
+on the manual labelling, equal the return value of the module. Accuracy is
+calculated as
+
+\begin{equation}
+	ACC = \frac{|S|}{|T|}
+\end{equation}
+\labelequations{Finger-Key Identification Module Accuracy}
+
+Where $|S|$ is the number of successful finger-key identifications, and $|T|$ is
+the total number of keypresses for all typing test sequences. This is also
+referred to as the success percentage.
+
+\subsection{Training Results}
+There were a total of seven training iterations. These training iterations
+served as a way to improve the performance of the module by adjusting its
+parameters.
+
+\subsubsection{Identifications}
+
+\begin{figure}[H]
+	\centering
+	\begin{tikzpicture}
+		\begin{axis}[
+				ybar stacked,
+				bar width=30pt,
+				xlabel={Training Iteration},
+				ylabel={Percentage over total keypresses},
+				xmin=1, xmax=7,
+				ymin=86, ymax=100,
+				xtick={1,2,3,4,5,6,7},
+				ytick={86, 86, 88, 90, 92, 94, 96, 98, 100},
+				enlargelimits=0.15,
+				area legend,
+				tick label style={font=\footnotesize},
+				legend style={font=\footnotesize},
+				label style={font=\footnotesize},
+				every axis legend/.append style={cells={anchor=west}},
+				ymajorgrids=true,
+			]
+
+			\addplot[
+				CornflowerBlue,
+				fill=CornflowerBlue,
+			]
+			coordinates {
+					(1, 88.47) (2, 88.95) (3, 98.51) (4, 99.55) (5, 98.47) (6, 99.17) (7, 99.49)
+				};
+			\addplot[
+				Salmon,
+				fill=Salmon,
+			]
+			coordinates {
+					(1, 11.46) (2, 10.98) (3, 1.42) (4, 0.38) (5, 1.46) (6, 0.76) (7, 0.45)
+				};
+			\legend{Successes, Failures}
+
+		\end{axis}
+	\end{tikzpicture}
+	\caption{Identification results of the training iterations}
+	\label{fig:rd-training-identification}
+\end{figure}
+
+Over the seven training iterations, there was an increase in success percentage.
+Consequently, there was also a decrease in failure percentage. There was a total
+of 1475 total keypresses for iterations one to three, and 1571 total keypresses
+for iteration four to seven.
+
+The increase in total keypresses starting from iteration four was due to adding
+multiple tries to the training routine in obtaining the Key-Edge Coordinates
+Map. Iterations one through three only tried frame 10 as the source frame for
+the module while iterations four through seven tried frame 10, 15, and 20.
+
+There was a need to try different frames since the camera was not consistent in
+its ability to focus on the keyboard for each video. In some videos video, the
+camera was not focused on the keyboard on frame 10, but it was able to focus on
+frame 15. As a result, the total number of available keypresses to test
+increased.
+
+The 0.48\% jump between iteration one and two can be attributed to correcting
+erroneous finger-key identifications that was obtained from manual labelling.
+The 9.56\% jump in success percentage between iteration two and three was due to
+adding buffers around the \ac{roi} by scaling the contours by 10 pixels on each
+side. Iteration four's increase was due to fixing the image map. There were
+certain keys that had incorrect color values assigned in the Key-Color Values
+Map, and some keys were of incorrect shape. Iterations five to seven were
+performed to test different values and shapes for scaling the \ac{roi}.
+
+\subsubsection{Uncertain Successes}
+
+\begin{figure}[H]
+	\centering
+	\begin{tikzpicture}
+		\begin{axis}[
+				ybar stacked,
+				bar width=30pt,
+				xlabel={Training Iteration},
+				ylabel={Percentage over total keypresses},
+				xmin=1, xmax=7,
+				ymin=0, ymax=30,
+				xtick={1,2,3,4,5,6,7},
+				ytick={0,5,10,15,20,25,30},
+				enlargelimits=0.15,
+				area legend,
+				tick label style={font=\footnotesize},
+				legend style={font=\footnotesize},
+				label style={font=\footnotesize},
+				every axis legend/.append style={cells={anchor=west}},
+				ymajorgrids=true,
+			]
+
+			\addplot[
+				CornflowerBlue,
+				fill=CornflowerBlue
+			]
+			coordinates {
+					(1, 5.69) (2, 5.69) (3, 12.61) (4, 12.86) (5, 9.32) (6, 10.76) (7, 12.86)
+				};
+			\addplot[
+				Salmon,
+				fill=Salmon,
+			]
+			coordinates {
+					(1, 0.61) (2, 0.61) (3, 17.15) (4, 17.06) (5, 3.95) (6, 6.49) (7, 9.68)
+				};
+			\legend{Spacebar, Non-spacebar}
+
+		\end{axis}
+	\end{tikzpicture}
+	\caption{Uncertain Successes of the training iterations}
+	\label{fig:rd-training-uncertain-successes}
+\end{figure}
+
+Uncertain Successes are instances where the finger-key identification module was
+able to detect two or more fingertips within the \ac{roi} of the key. As such,
+the module is uncertain which of the multiple fingertips actually pressed the
+key. However, it is still classified as a success since the expected finger from
+the manual labelling exists within the array of fingertips returned by the
+module.
+
+\paragraph{Spacebar}
+In a 60\% keyboard, the spacebar is the longest key on the keyboard. In
+addition, this key is right below the non-dominant hand's thumb default resting
+position when the hand's posture follows the proper touch typing posture. As
+such, it was to be expected that a majority of the uncertain successes were
+because of the spacebar. However, this was not the case for iterations three and
+four. This was because the buffer for the \ac{roi} was too big and this caused
+the module to detect more than one fingertip within that \ac{roi}.
+
+\paragraph{Non-spacebar}
+
+There were cases where the module had uncertain successes that were not over the
+spacebar. In almost practically all cases, these were over the alphanumeric
+keys. This uncertain successes stem from the increased buffer of the \ac{roi}.
+This type of uncertain successes carry more weight compared to spacebar
+uncertain successes since the keys where these uncertain successes come from are
+the smallest keys in the keyboard. This means that the chances that a person
+actually has two or more fingers over the keys is rare, compared to the
+spacebar.
+
+\subsubsection{Uncertain Successes vs Success Percentage}
+\label{section:rd-uncertain vs success}
+
+\begin{figure}[H]
+	\centering
+	\begin{tikzpicture}
+		\begin{axis}[
+				xlabel={Training Iteration},
+				ylabel={Percentage over total keypresses},
+				xmin=1, xmax=7,
+				ymin=0, ymax=100,
+				xtick={1,2,3,4,5,6,7},
+				ytick={0,20,40,60,80,100},
+				tick label style={font=\footnotesize},
+				legend style={font=\footnotesize},
+				label style={font=\footnotesize},
+				enlargelimits=0.15,
+				ymajorgrids=true,
+				xmajorgrids=true,
+				nodes near coords,
+				every node near coord/.append style={font=\footnotesize},
+				every axis legend/.append style={at={(0.98, 0.02)}, anchor=south east, cells={anchor=west}}
+			]
+
+			\addplot[
+				color=CornflowerBlue,
+				mark=square,
+			]
+			coordinates {
+					(1, 88.47) (2, 88.95) (3, 98.51) (4, 99.55) (5, 98.47) (6, 99.17) (7, 99.49)
+				};
+			\addplot[
+				color=Salmon,
+				mark=square,
+			]
+			coordinates {
+					(1, 6.31) (2, 6.31) (3, 29.76) (4, 29.92) (5, 13.18) (6, 17.25) (7, 22.53)
+				};
+
+			\legend{Successes, Uncertain Successes}
+
+		\end{axis}
+	\end{tikzpicture}
+	\caption{Uncertain Successes of the training iterations}
+	\label{fig:rd-training-uncertain-successes}
+\end{figure}
+
+Tuning the algorithm required managing both uncertain successes and the success
+percentage and making compromises between one or the other. The goal of the
+training process was to maximize the success percentage, while limiting the
+number of uncertain successes.
+
+It can be noted that in iterations one and two, the success rate was lower
+compared to the other iterations. However, the percentage of uncertain successes
+was also the lowest. Iteration three was the first iteration that created a
+buffer around each key, however this buffer of 10 pixels was too aggressive,
+with the uncertain successes spiking upwards to nearly 30\%. Iteration five
+reduced the buffer size in half to five pixels. This did not affect success rate
+that much, with only a reduction of 1.07\%. This also greatly decreased the
+uncertain success percentage by 16.74\%. The sixth iteration met halfway between
+iteration four and five, by adding a buffer of seven pixels. This increased the
+success rate back to 99.17\% without greatly increasing the number of uncertain
+successes. In comparison with the fourth iteration, the sixth iteration had its
+success rate decrease by 0.38\%, but its percentage of uncertain successes
+decreased by 12.67\%. A seventh iteration was trialed where the increase in
+buffer was asymmetrical. For each key, the points nearer the user were increased
+by 10 pixels, while the points farther away were increased by five pixels only.
+This slightly increased the success rate by 0.32\%. However, the percentage of
+uncertain successes also increased by 5.28\%.
+
+Based on these results, the final selected parameters for the module was the
+parameters set in iteration six, with the buffer set to seven pixels as it was a
+good middle ground between success rate, and uncertain successes.
+
+\subsubsection{Failure Types}
+
+\begin{figure}[H]
+	\centering
+	\begin{tikzpicture}
+		\begin{axis}[
+				ybar stacked,
+				bar width=30pt,
+				xlabel={Training Iteration},
+				ylabel={Percentage over total keypresses},
+				xmin=1, xmax=7,
+				ymin=0, ymax=15,
+				xtick={1,2,3,4,5,6,7},
+				ytick={0,3,6,9,12,15},
+				enlargelimits=0.15,
+				area legend,
+				tick label style={font=\footnotesize},
+				legend style={font=\footnotesize},
+				label style={font=\footnotesize},
+				every axis legend/.append style={cells={anchor=west}},
+				ymajorgrids=true,
+			]
+
+			\addplot[
+				CornflowerBlue,
+				fill=CornflowerBlue
+			]
+			coordinates {
+					(1, 0.47) (2, 0) (3, 0.14) (4, 0) (5, 0) (6, 0) (7, 0)
+				};
+			\addplot[
+				Salmon,
+				fill=Salmon,
+			]
+			coordinates {
+					(1, 10.98) (2, 10.98) (3, 1.29) (4, 0.38) (5, 1.46) (6, 0.76) (7, 0.45)
+				};
+			\legend{Mismatched Detection, No Detection}
+
+		\end{axis}
+	\end{tikzpicture}
+	\caption{Uncertain Successes of the training iterations}
+	\label{fig:rd-training-uncertain-successes}
+\end{figure}
+
+\paragraph{Mismatched Identification}
+These are failures where the module did not return the expected fingertip based
+on the manual labelling. This type of failure only manifested in iterations one
+and three. The failures in iteration one, after double-checking the video, were
+due to an error during manual labelling. In iteration three, these errors were
+due to an incorrect image map and Key-Color Values map. The information for some
+keys were not consistent between both.
+
+Based on this information and from the gathered data, This failure type is
+non-existent for the module in real-life conditions.
+
+\paragraph{No Identification}
+These are failures where the module did not detect any fingertips within the
+\ac{roi}. These made up all the failures from the test results. These types of
+failures occurred because the fingertips used to press the key were not directly
+above the key, but rather offset to its side. Scaling the \ac{roi} would reduce
+the number of no identification failures, but it would also increase the number
+of uncertain successes. See Section~\ref{section:rd-uncertain vs success},
+Uncertain Successes vs Success Percentage for more information.
+
+\subsubsection{Occlusion}
+The module was also successful in finger-key identification, even if both the
+key and finger was occluded. Figure~\ref{fig:rd-occluded} is one frame where
+this occurred. The module was successful in identifying that the Left Thumb was
+used to press the Left Shift.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{occluded.png}
+	\caption{Occluded finger and key}
+	\label{fig:rd-occluded}
+\end{figure}
+
+\subsection{Test Results}
+\begin{table}[H]
+	\small
+	\centering
+	\begin{tabular}{ p{0.4\textwidth} R{0.2\textwidth} }
+		\toprule
+		Category                        & Value   \\
+		\midrule
+		Total Keypresses                & 942     \\[0.25cm]
+		\midrule
+		Identifications                           \\
+		\midrule
+		Successes                       & 937     \\
+		Failures                        & 4       \\
+		Success Percentage / Accuracy   & 99.47\% \\
+		Failure Percentage              & 0.42\%  \\[0.25cm]
+		\midrule
+		Failure Types                             \\
+		\midrule
+		Mismatched Identification       & 0       \\
+		No Identification               & 4       \\[0.25cm]
+		\midrule
+		Uncertain Successes                       \\
+		\midrule
+		Total                           & 133     \\
+		Spacebar                        & 72      \\
+		Spacebar Percentage             & 54.14\% \\
+		Spacebar Percentage (Total)     & 7.64\%  \\
+		Non-spacebar                    & 61      \\
+		Non-spacebar Percentage         & 45.86\% \\
+		Non-spacebar Percentage (Total) & 6.48\%  \\
+		\bottomrule
+	\end{tabular}
+	\caption{\label{tab:rd-accuracy}Accuracy Results of the Module with Test Data}
+\end{table}
+
+The results were satisfactory after running the module with the parameters from
+the sixth training iteration on the test data. There was another run with the
+same test data and the same parameters, however these results were omitted since
+this run had inconclusive results as there was a single identification that was
+incorrectly labeled during manual finger-key identification.
+
+\section{Speed}
+
+Based on the definition of \ac{wpm}
+from \cite{arif2009} as seen in Equation~\ref{equ:wpm}, it can be extrapolated that calculating Characters per
+Second is:
+
+\begin{equation}
+	CPS = \frac{WPM}{60} \cdot 5
+\end{equation}
+\labelequations{Characters per Second}
+
+The average \ac{wpm} when typing using the QWERTY layout, according to test
+data from \citeauthor{keybr} is $\approx37.5$ \ac{wpm}. As such, we can
+calculate that the average Characters per Second when typing using the QWERTY
+layout is 3.125. This means that each character, on average, requires 0.32
+seconds to type. This sets the maximum time spent for finger-key identification
+for a single character. Any more than this would result in noticeable slowdown
+during typing.
+
+\subsection{Results}
+
+\subsubsection{Training Iterations}
+
+\begin{figure}[H]
+	\centering
+	\begin{tikzpicture}
+		\begin{axis}[
+				xlabel={Training Iteration},
+				ylabel={Seconds},
+				xmin=1, xmax=7,
+				ymin=0, ymax=1,
+				xtick={1,2,3,4,5,6,7},
+				ytick={0, 0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0},
+				tick label style={font=\footnotesize},
+				legend style={font=\footnotesize},
+				label style={font=\footnotesize},
+				enlargelimits=0.15,
+				ymajorgrids=true,
+				xmajorgrids=true,
+				nodes near coords,
+				every node near coord/.append style={font=\footnotesize},
+				every axis legend/.append style={cells={anchor=west}}
+			]
+
+			\addplot[
+				color=CornflowerBlue,
+				mark=square,
+			]
+			coordinates {
+					(1, 0.9123) (2, 0.8225) (3, 0.8144) (4, 0.8211) (5, 0.8258) (6, 0.8366) (7, 0.8118)
+				};
+			\addplot[
+				color=Salmon,
+				mark=square,
+			]
+			coordinates {
+					(1, 0.0895) (2, 0.0903) (3, 0.0823) (4, 0.0835) (5, 0.0845) (6, 0.0828) (7, 0.0797)
+
+				};
+			\legend{Avg. Keyboard Detection Time, Avg. Finger Identification Time}
+
+		\end{axis}
+	\end{tikzpicture}
+	\caption{Uncertain Successes of the training iterations}
+	\label{fig:rd-training-uncertain-successes}
+\end{figure}
+
+Training iterations had an average keyboard detection time of $0.8620\pm0.0502s$
+and average finger identification time of $0.0863\pm0.0040s$. Over the course of
+the training, there was no noticeable difference between running time over the
+same dataset. However, The first training iteration had issues with gathering
+the speed of the module, as the average keyboard detection time included other
+operations that was not part of getting the Key-Edge Coordinates Map. Removing
+this iteration, the average keyboard detection time becomes more consistent,
+with it being $0.8295\pm0.0070s$.
+
+\subsubsection{Test Results}
+
+\begin{table}[H]
+	\small
+	\centering
+	\begin{tabular}{ p{0.4\textwidth} R{0.2\textwidth} }
+		\toprule
+		Category                        & Seconds \\
+		\midrule
+		Avg. Keyboard Detection Time    & 0.838   \\
+		Avg. Finger Identification Time & 0.083   \\
+		\bottomrule
+	\end{tabular}
+	\caption{\label{tab:rd-speed}Speed Results of the Module with Test Data}
+\end{table}
+
+\subsection{Average Keyboard Detection Time}
+This is the time that was measured as the module generated the Key-Edge
+Coordinates Map. This corresponds with the first step of the module. It is to be
+noted that this time does not affect keyboard typing and is not limited by the
+0.32 threshold as the user is not yet typing at this point.
+
+\subsection{Average Finger Identification Time}
+This is the time spent in finger-key identification for a single key. As such,
+this time should be below 0.32 seconds for the module to be capable of
+performing finger-key identification in real-time without noticeable slowdown
+from the perspective of the user.
+
+The average finger identification time that the module got, 0.08389272103s was
+well below 0.32 seconds. This can be attributed to the speed of the MediaPipe
+library, and the pre-calculation of the edge coordinates.
+
 \chapter{Conclusion}
+
 \chapter{Future Work}
@@
 \centering
+\small
 \begin{supertabular}{ll}
@@
 \label{appendix:test-sequences}
-
+\normalsize
 \begin{enumerate}

### 2022-06-02 application/algorithms/integration.py

@@ -20,7 +20,7 @@ def main(keyboardImage: np.ndarray, handImage: np.ndarray):
     # Step 1
     allContourPoints = getContourPointsFromImage(keyboardImage)
 
-    contourPoints = allContourPoints["B"]
+    contourPoints = allContourPoints["LeftShift"]
 
     # elapsed = time.time() - start
 
@@ -49,6 +49,6 @@ def main(keyboardImage: np.ndarray, handImage: np.ndarray):
 
 
 if __name__ == "__main__":
-    keyboardImage = cv.imread("images/fast-02-20.png")
-    handImage = cv.imread("images/fast-02-380.png")
+    keyboardImage = cv.imread("images/medium-04-10.png")
+    handImage = cv.imread("images/medium-04-437.png")
     main(keyboardImage, handImage)

### 2022-06-02 application/algorithms/test-scripts/get-image.py

@@ -4,9 +4,9 @@ from cv2 import cv2 as cv
 
 
 def main():
-    file = "fast-02"
-    keyboardFrame = 20
-    fingerFrame = 380
+    file = "medium-04"
+    keyboardFrame = 10
+    fingerFrame = 437
     fileLocation = f"../dataset/training/{file}"
 
     cap = cv.VideoCapture(f"{fileLocation}.mp4")

### 2022-06-08 paper/paper.tex
trainer: add initial trainer

@@
 
-% Code Formatting
-\usepackage{minted}
-\usemintedstyle{pastie}
-
-\def\code#1{\texttt{#1}}
-
 % Positioning
@@
 	long=Historical Finger Placement Accuracy
+}
+
+\DeclareAcronym{cps}{
+	short=CPS,
+	long=Characters per Second
 }
@@
 
-\addto\captionsamerican{% Replace "english" with the language you use
-	\renewcommand{\contentsname}%
-	{Table of Contents}%
+\addto\captionsamerican{
+	\renewcommand{\contentsname}
+	{Table of Contents}
 }
+
+% Subfigure
+\usepackage{subcaption}
 
@@
 	\par\vspace{0.25cm}
-	A Special Project Proposal Presented to the\\
+	A Special Project Presented to the\\
 	Faculty of the Department of Computer Science,\\
@@
 
-	\textbf{\uppercase{Development of a finger-key identification module\\for a touch typing trainer}}
+	\textbf{\uppercase{Development of a Finger-Key Identification Module\\for a Touch Typing Trainer}}
 
@@
 
+\begin{center}
+	\uppercase{\textbf{APPROVAL SHEET}}
+\end{center}
+
+\par\vspace{0.5cm}
 \noindent
-This thesis entitled ``\textbf{\uppercase{Development of a finger-key
-		identification \\module for a touch typing trainer}}'' prepared and
-submitted by \\\uppercase{Oscar Vian L. Valles} is approved by:
+The faculty of the University of the Philippines Cebu, College of Science,
+Department of Computer Science approves this Special Problem entitled:
+\par\vspace{0.5cm}
+
+\begin{center}
+	\textbf{Development of a Finger-Key Identification Module\\for a Touch Typing Trainer}
+	\par\vspace{0.5cm}
+	by
+	Oscar Vian L. Valles
 
@@
 
-\begin{center}
 	\begin{minipage}[t]{0.45\textwidth}
@@
 		\line(1,0){0.9\textwidth}\\
-		Prof. Aileen Joan O. Vicente\\
-		Panel Member
+		Dhong Fhel K. Gom-os\\
+		Special Problem Adviser
 
@@
 		\line(1,0){0.9\textwidth}\\
-		Date
+		Date Signed
 	\end{minipage}
 	\begin{minipage}[t]{0.45\textwidth}
-		\centering
-		\line(1,0){0.9\textwidth}\\
-		Prof. Aileen Joan O. Vicente\\
-		Panel Member
-
-		\par\vspace{0.5cm}
-		\line(1,0){0.9\textwidth}\\
-		Date
-	\end{minipage}
-
-	\par\vspace{2cm}
-
-	\line(1,0){0.405\textwidth}\\
-	Prof. Aileen Joan O. Vicente\\
-	Panel Chair
-
-	\par\vspace{0.5cm}
-	\line(1,0){0.405\textwidth}\\
-	Date
-
-	\par\vspace{2cm}
-
-	\begin{minipage}{0.45\textwidth}
 		\centering
@@
 		\line(1,0){0.9\textwidth}\\
-		Date
+		Date Signed
 	\end{minipage}
-	\begin{minipage}{0.45\textwidth}
-		\centering
-		\line(1,0){0.9\textwidth}\\
+
+	\par\vspace{2cm}
+
+	\line(1,0){0.405\textwidth}\\
 	Prof. Nelia S. Ereno\\
@@
 	\par\vspace{0.5cm}
-		\line(1,0){0.9\textwidth}\\
-		Date
-	\end{minipage}
+	\line(1,0){0.405\textwidth}\\
+	Date Signed
+
+	\par\vspace{2cm}
 \end{center}
@@
 \end{center}
+
+Typing tests lists out words for the user to repeat. These tests measure metrics
+that quantify the user's ability to type. However, conventional typing metrics
+do not measure correct finger placement. This aspect of typing may affect the
+user's health. It is also a crucial aspect of keyboard typing education. As
+such, a method to identify which finger is used to press which key is
+beneficial. This paper introduces a technique to do so by developing a
+finger-key identification module that utilizes computer vision algorithms to
+detect the keys in a keyboard, and a ready-made machine learning solution to
+track fingers while typing. This module was successful in finger-key
+identification with an accuracy of 99.47\% in a dataset of 937 keypresses. The
+module also had an average finger-key identification time of 0.083 seconds,
+which allows the module to perform real-time finger-key identification for
+typing speeds up to 143.030 Words per Minute which is well beyond the median
+typing speed of the population. Finally, two metrics for correct finger
+placement were defined. Finger Placement Accuracy which computes the number of
+keys pressed with the correct finger in a typing test sequence. Historical
+Finger Placement Accuracy which evaluates the user's ability to press a certain
+key with the correct finger over multiple typing test sequences.
 
@@
 time taken is calculated, and certain metrics are given. These metrics include
-\ac{wpm} and Accuracy \parencite{bartnik2021}.
-
-However, this method of examination leaves out a crucial part of typing —
-ergonomics. Ergonomic typing prevents a lot of health issues in the future like
-repetitive strain injury or carpal tunnel. One important factor that affects
-ergonomics is the typing procedure and posture. This means the proper placement of
-the wrist, hands, and hitting the keys using the right finger that is assigned
-to the key.
+\ac{wpm} and Accuracy \parencite{arif2009}. However, this method of examination
+leaves out a crucial part of typing --— correct finger placement.
+
+Incorrect finger placement may affect hand and finger health. Incorrect
+placement of the fingers may cause these hand and wrist positions: ulnar
+deviation, forearm pronation, and wrist extension \parencite{serina1999}. These
+three are hand and wrist positions that are common in all activities, however,
+prolonged periods in these positions, such as in typing, may cause injuries such
+as \ac{cts} \parencite{toosi2015}.
 
@@
 with each key being associated with a specific finger. For instance, the letter
-Q in a QWERTY layout should be hit using the fifth digit, or the little finger,
-of the left hand, and this is shown by coloring the fifth digit and the key Q
-with the same color or by placing the letters directly on the fingers
-\parencite{dobson2009touch}.
-
-Incorrect finger placement may cause these hand and wrist positions: ulnar
-deviation, forearm pronation, and wrist extension \parencite{serina1999}. These
-three are hand and wrist positions that are common in all activities, however,
-prolonged periods in these positions may cause injuries such as \ac{cts}
-\parencite{toosi2015}
-
-In addition, this type of typing is frequently taught in the beginner level
-\parencite{donica2018}. This means that there is a need to weed out bad habits
-that may develop, like using the index finger for pressing the spacebar or
-backspace. However, it is impractical for an educator to check each student if
-they are not performing these movements as these may only show for a small
-period which may not be caught in time.
+Q in a QWERTY layout should be pressed using the fifth digit, or the little
+finger, of the left hand, and this is shown on instruction manuals by coloring
+the fifth digit and the key Q with the same color or by placing the letters
+directly on the fingers \parencite{dobson2009touch}.
+
+Furthermore, Touch typing using correct finger placement is frequently taught in
+the beginner level \parencite{donica2018}. This means that there is a need to
+weed out bad habits that may develop, like using the index finger for pressing
+the spacebar or backspace. However, it is impractical for an educator to check
+each student if they are not performing these movements as these may only show
+for a small period and may not be caught in time.
 
@@
 	      finger-key identification.
-	\item What setup and configuration of a single optical camera can be used to
-	      capture images during keyboard typing for time finger-key identification
-	      using the algorithm identified in 1?
+	\item What is the setup and configuration of a single optical camera can be
+	      used to capture images during keyboard typing for real-time finger-key
+	      identification using the algorithm identified in 1?
 	\item How to design and develop a finger-key identification module using
@@
 \begin{enumerate}
-	\item Develop an algorithm or technique using computer vision that is capable
-	      of finger-key identification.
+	\item Develop an algorithm or technique using computer vision capable of
+	      finger-key identification.
 	\item Identify the setup and configuration of a single optical camera that can
@@
 	      identification
-	\item Determine if the module developed in 3 is fast-enough to run in
+	\item Determine if the module developed in 3 is fast enough to run in
 	      real-time
@@
 \section{Scope and Limitation}
-This research will focus on typing on a 60\% keyboard. Figure~\ref{fig:60}
-illustrates this type of keyboard. This type of keyboard only has the
-alphanumeric part of the keyboard. This limits the number of keys to be checked
-and the expected movement of the hand. Furthermore, the keycaps will also be of
-a light color, while the surface that the keyboard rests upon will be of a dark
-color.
-
-In addition, the keyboard layout will be \ac{ansi}. This layout is described by
-the American National Standards Institute \parencite{ansi}. This is the most
-common layout in the United States. However, it is also used in numerous
-English-speaking countries such as the Philippines, Malaysia, and India
-\parencite{apple-layout}.
-
-The setup that will be captured will be simplified and fixed for the development
-and testing of the finger-key identification module. This includes the
-keyboard, surface it is placed on, and lighting of the entire area. Furthermore,
-the setup will not be tested by other respondents, but by the researchers only
---- as the focus is on the development and testing of the finger-key
-identification module and not on the touch typing trainer and its effects on
-typing.
-
-The program will expect the that user has all ten digits and has no hand,
-finger, or wrist deformities. In addition, only the placement of the hands,
-fingers, and wrists will be taken into account when determining if the
-ergonomics of the user while typing is healthy. The program will not check
-seating position, angle of elbows, and other metrics for an ergonomic typing
-posture while typing.
-
-Capturing of the video to be analyzed by the program would be limited to a
-single 720p webcam that is capturing in 30 frames per second. The camera will be
-pointed downwards facing the keyboard and the hand.
+The experimental setup was simplified and tightly controlled. It was not
+representative of real-world conditions. The colors of the keyboard, cables, and
+surface were all controlled. So were the lighting and camera used.
+
+The supported keyboard of the module was limited to one specific type of
+keyboard. A 60\% Keyboard in a slightly modified \ac{ansi} layout with
+light-colored keys as shown in Figure~\ref{fig:60}. This type of keyboard only
+has the alphanumeric portion of a full-size keyboard. This limited the number of
+keys to be checked and reduced the overall movement of the hand during typing.
+The keyboard layout was a slightly modified \ac{ansi} layout. The majority of
+the layout still followed the base \ac{ansi} layout, with the alphanumeric
+portion remaining exactly the same.
+
+The module's effect on typing was not tested. This includes the module's effect
+on typing education and typing health, as the focus of the research was on the
+development and testing of the finger-key identification module only.
+
+In addition, the module expected the user to have no deformities in their hands,
+and they were capable of typing on a keyboard with difficulties.
 
@@
 system. In the context of this paper, keyboard typing will refer to the act of
-using a physical keyboard to input characters in a computer system.
+using a physical keyboard to input characters into a computer system.
 
@@
 \citeauthor{ansi} is the standard that first defined the \ac{ansi} layout.
-Figure~\ref{fig:ansi} illustrates what the \ac{ansi} layout looks like. This layout
-is also used by countries other than America. Examples of countries that use
-this layout as its standard is the Philippines, China, and Korea
+Figure~\ref{fig:ansi} illustrates what the \ac{ansi} layout looks like. This
+layout is also used by countries other than America. Examples of countries that
+use this layout as its standard are the Philippines, China, and Korea
 \parencite{apple-layout}. However, these countries also opt to modify the layout
@@
 create other layouts. Layouts created from this standard are colloquially called
-\ac{iso} Layouts. Countries around the world use this framework to create layouts
-that fit the characters in their language. Examples of countries that use this
-framework to create their own layout are France, Greece, Canada, and Sweden
+\ac{iso} Layouts. Countries around the world use this framework to create
+layouts that fit the characters in their language. Examples of countries that
+use this framework to create their layout are France, Greece, Canada, and Sweden
 \parencite{apple-layout}.
@@
 \ac{iso} and \ac{ansi} standards by adding or removing certain keys to fit the
-character set of a language, or for additional keys. Other layouts are also
-exactly the same as \ac{ansi} or \ac{iso}, however, these layouts change the
-arrangement of the alphabet within the keyboard.
+character set of a language, or for additional keys. Other layouts are also the
+same as \ac{ansi} or \ac{iso}, however, these layouts change the arrangement of
+the alphabet within the keyboard.
 
@@
 memorization of the layout which reduces the ease of learning, reduces performance
-in typing by reducing speed, and increases of typing errors
+in typing by reducing speed, and increases typing errors
 \parencite{ciobanu2015}.
@@
 layout. This includes the alphanumeric keys, the function keys, the navigation
-cluster, and the numpad.
+cluster, and the Numpad.
 
@@
 Speed, also called as entry rate by \citeauthor{arif2009}, measures the number
-of characters entered in a specific time frame. The most common metric that
+of characters entered within a specific time frame. The most common metric that
 measures speed is \ac{wpm}. \ac{wpm} as defined by \citeauthor{arif2009} is:
@@
 
-\subsubsection{Average Typing Speed}
-
-Average typing speed using a keyboard with QWERTY layout, based on test data
-from \cite{keybr}, is $\approx37.5$ \ac{wpm}. This average is taken from all of
-the test results they have gathered from all of their users.
+\subsection{Typing Speed Statistics}
+
+Median typing speed using a keyboard with QWERTY layout, based on test data
+from \cite{keybr}, is $\approx37.5$ \ac{wpm}. This median is taken from all the test results they have gathered from all of their users.
+
+\cite{monkey-stats} released a dataset showing the speed of 12678 members of the
+Monkeytype community with the speed role. The dataset shows the fastest time
+that these members have achieved on 60-second tests. The median of the dataset
+lies in the 100--109 \ac{wpm} range, with 52.74\% of the community having top
+speeds slower than 100 \ac{wpm}.
+
 \cite{arif2009} also performed a survey with 12 participants using
@@
 
-These studies start teaching kids in the kindergarten level and the studies try
+These studies start teaching kids at the kindergarten level and the studies try
 to optimize the teaching methods to improve the speed and accuracy of typing of
@@
 previous stage are refined. At this stage, the learners are expected to rely less
-on visuals to locate the keys and more on kinesthesis. By the final stage, the
+on visuals to locate the keys and more on kinesthesis. In the final stage, the
 autonomous stage, the learner does not rely on visuals at all and focuses on
@@
 information gathered is then used to generate new words for the next iteration.
-As an example, if a learner has difficulty in typing the letter q, the next
+As an example, if a learner has difficulty typing the letter q, the next
 iterations will have a lot of words that contain the letter q.
@@
 There have been a lot of studies that show the effect of keyboard typing, and
-its associated movements (or lack thereof), has an effect on the human body.
-These studies have shown that keyboard typing has an effect on our neck, shoulder,
+its associated movements (or lack thereof), affects the human body. These
+studies have shown that keyboard typing affects our neck, shoulder,
 upper limb, wrist, arms, and fingers \parencite{szeto2005, baker2007digit}
@@
 \subsection{Finger and Wrist Kinematics}
-The way people move their hands, wrists, and fingers differ between each person.
-This can be attributed to the different typing styles each person has. One key
-difference between people is the angle of the fifth digit.
+The way people move their hands, wrists, and fingers differs between each
+person. This can be attributed to the different typing styles each person has.
+One key difference between people is the angle of the fifth digit.
 
@@
 For those that reach their keys by stretching their fingers, there is an
-increased probability that the wrists and fingers adapt non-neutral postures.
+increased probability that the wrists and fingers adapt to non-neutral postures.
 These include wrist extension, ulnar deviation, and pronation, which may cause
@@
 using motion capture systems or computer vision. This technique allows computers
-to perform actions and analyzes on the motions and positions of these body
-parts.
+to perform actions and analyzes the motions and positions of these body parts.
 
@@
 better performance compared to previously used algorithms, however, the issue of
-cost and self-occlusion still persists.
+cost and self-occlusion persists.
 
@@
 At its core, Computer Vision aims to perform tasks that the human visual system
-can do \parencite{cern}. This includes object classification, tracking, and
-gesture recognition, and face recognition. At the present, most computer vision
-systems utilize deep learning algorithms, and convolutional networks to gather
+can do \parencite{cern}. This includes object classification, tracking, gesture
+recognition, and face recognition. At the present, most computer vision systems
+utilize deep learning algorithms, and convolutional networks to gather
 information from an image, or a set of images. One such example of a
@@
 This framework allows for performant object detection, hand and finger tracking,
-human pose detection. The framework also allows for combining multiple features,
-by adding them to the graph as calculators. MediaPipe has C++, Python, JS, and
-Coral interfaces. It also supports Android and iOS devices
+and human pose detection. The framework also allows for combining multiple
+features, by adding them to the graph as calculators. MediaPipe has C++, Python,
+JS, and Coral interfaces. It also supports Android and iOS devices
 \parencite{mediapipe}.
@@
 
-A toolbox offered by MATLAB is the Computer Vision Toolbox that contains
+A toolbox offered by MATLAB is the Computer Vision Toolbox which contains
 algorithms, and functions for use in the development of computer vision, 3D
@@
 
-\cite{dorf2001} presents a use case for finger tracking in augmented
+\cite{dorf2001} present a use case for finger tracking in augmented
 environments. In the paper, interaction in a virtual environment through the use
@@
 \section{Summary of the Research Gap}
-While there are a lot of applications and curriculum aimed at teaching touch
+While there are a lot of applications and curricula aimed at teaching touch
 typing, there is no automated system available that detects if a person uses the
@@
 
-The experimental setup and configuration was composed of three elements: the
+The experimental setup and configuration were composed of three elements: the
 camera, the keyboard, and the environment. Figure~\ref{fig:metho-setup} is an
@@
 fingers. To do so, it was mounted on top of the monitor with the camera pointed
-down towards the table.
+down towards the table. Figure~\ref{fig:metho-setup-camera-placement}
+illustrates how the camera was mounted over the monitor and
+Figure~\ref{fig:metho-setup-keeb} is a picture taken with the camera at this
+position.
 
@@
 \parencite{logitech}.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.4\textwidth]{webcam.png}
-	\caption{Logitech C920. Reprinted from \fullcite{logitech}}
-	\label{fig:metho-setup-cam}
-	\centering
-\end{figure}
 
@@
 had a dark USB-C cable to blend with the dark surface. These color coordination
-steps improved initial keyboard detection. In addition, the keyboard layout was
-ANSI, due to the availability and widespread adoption of the layout in the
-Philippines.
+steps improved initial keyboard detection.
+
+In addition, the keyboard layout was a modified \ac{ansi} layout. The majority
+of the keys remained the same, however some keys were changed: The pipe key was
+replaced by the backspace key, the backspace key was replaced by the pipe key
+and a delete key, and the bottom row has a 7u length spacebar. Overall, these
+are minor changes that does not affect the alphanumeric keys that are
+predominantly used in typing.
 
@@
 
-In addition, the surface where the keyboard was placed on was solid black
+In addition, the surface where the keyboard was placed was solid black
 without any variation of color. This also improved initial keyboard detection.
@@
 
-A computer vision algorithm was created as a starting point
-mapping the keys of the keyboard within a video. A rough flowchart of the algorithm
-is shown in Figure~\ref{fig:metho-algo-key-flow}.
+A computer vision algorithm was created as a starting point for mapping the keys
+of the keyboard within a video. A rough flowchart of the algorithm is shown in
+Figure~\ref{fig:metho-algo-key-flow}.
 
@@
 work. In addition, this step optimized the algorithm as the number of dimensions
-analyzed was reduced.
+analyzed was reduced. Figure~\ref{fig:metho-algo-convert} is the output of this
+step.
 
@@
 	\caption{Camera capture converted to grayscale}
+	\label{fig:metho-algo-convert}
 	\centering
@@
 an image that is smoothed while preserving its edges
-\parencite{bilateral-filter}.
+\parencite{bilateral-filter}. Figure~\ref{fig:metho-algo-denoise} is the output
+of this step.
 
@@
 	\centering
+	\label{fig:metho-algo-denoise}
 \end{figure}
@@
 kernels used by the Sobel filter \parencite{sobel2014} is shown in
-Figure~\ref{fig:metho-algo-key-sobel}.
+Figure~\ref{fig:metho-algo-key-sobel} and Figure~\ref{fig:metho-algo-sobel} is
+the output of this step.
 
@@
 	\centering
+	\label{fig:metho-algo-sobel}
 \end{figure}
@@
 that is most optimal for the image \parencite{otsu}. This outputs a
-black-and-white image, with only two values.
+black-and-white image, with only two values. Figure~\ref{fig:metho-algo-thresh}
+is the resulting output of this step.
 
@@
 	\caption{Threshed image of the edges}
+	\label{fig:metho-algo-thresh}
 	\centering
@@
 object found using the previous step. This approximation method removed
-redundant points and returned the least amount of points that describes the
+redundant points and returned the least amount of points that describe the
 shape. This OpenCV function implements the algorithm of~\cite{contours} in their
-paper \citetitle{contours}.
+paper \citetitle{contours}. Figure~\ref{fig:metho-algo-contours} shows the contours
+found in the threshed image.
 
@@
 	\caption{All the contours found in the image}
+	\label{fig:metho-algo-contours}
 	\centering
@@
 the largest contour's points were returned in a clockwise arrangement, which is
-a requirement for the next step
+a requirement for the next step. Figure~\ref{fig:metho-algo-largest-contour}
+shows the four contour points found at the edges of the keyboard within the
+image.
 
@@
 	\caption{Simplified contour}
+	\label{fig:metho-algo-largest-contour}
 	\centering
@@
 color assigned to it and this color fills the region where this key is located
-at. Appendix Chapter~\ref{appendix:key-map} is a table that shows the hex color
-code and key mapping present in the image map.
+at. Figure~\ref{fig:metho-algo-key-map} is the initial image map and Appendix
+Chapter~\ref{appendix:key-map} is a table that shows the hex color code and key
+mapping present in the image map.
 
@@
 transform was then used by \texttt{warpPerspective} to apply the transform to
-the virtual map.
+the virtual map. Figure~\ref{fig:metho-algo-transform} is the image obtained
+after warping the image map.
 
@@
 	\caption{Transformed image map}
+	\label{fig:metho-algo-transform}
 	\centering
@@
 virtual map, even if only one key corresponds to that color.
+Figure~\ref{fig:metho-algo-filtered} shows the image where the Left Control key
+was isolated.
 
@@
 	\caption{Filtered key}
+	\label{fig:metho-algo-filtered}
 	\centering
@@
 to find and simplify the contour of the \ac{roi} of the key.
+Figure~\ref{fig:metho-algo-contour-key} illustrates the four edges of the
+isolated key.
 
@@
 	\caption{Contour points of the \ac{roi}}
+	\label{fig:metho-algo-contour-key}
 	\centering
@@
 During testing using training data, a lot of the failures in finger-key
-identification was due to the tight fit of the contour to the key. The contour
+identification were due to the tight fit of the contour to the key. The contour
 on its own did not give enough buffer for the placement of the finger. This
@@
 not exactly on top of the key, but rather to its side. The algorithm accounts
-for this situation by adding a buffer. This was achieved through scaling the
-contour points by seven pixels on each side. This value was obtained after
+for this situation by adding a buffer. This was achieved by scaling the contour
+points by seven pixels on each side. Figure~\ref{fig:metho-algo-scale-roi} shows
+the increased buffer in relation to the key. This value was obtained after
 testing other possible values. A buffer of five pixels did not have a lot of
@@
 	\caption{Contour points of the \ac{roi}}
-	\centering
-\end{figure}
-
-\paragraph{Key-Edge Coordinatess Map}
+	\label{fig:metho-algo-scale-roi}
+	\centering
+\end{figure}
+
+\paragraph{Key-Edge Coordinates Map}
 After all the keys have passed through the loop, a Key-Edge Coordinates Map was
@@
 to detecting hands with articulated fingers since estimating a bounding box
-around a rigid object, i.e.\ a palm, is much simpler. Furthermore, a palm can be
+around a rigid object, i.e.\ a palm is much simpler. Furthermore, a palm can be
 modeled using only square anchors reducing the number of anchors by a factor of
@@
 
-The two previously chosen algorithms was combined to accomplish finger-key
+The two previously chosen algorithms were combined to accomplish finger-key
 identification and mapping.
@@
 
-Each landmark's coordinates was compared to the coordinates of the edges of the
+Each landmark's coordinates were compared to the coordinates of the edges of the
 \ac{roi}. A landmark was determined as positioned within the \ac{roi} if all the
 following conditions are true:
+
 \begin{enumerate}
-	\item In the X axis, the landmark's coordinates is greater than one or both of
-	      the two coordinates found of the left side of the \ac{roi}
-	\item In the X axis, the landmark's coordinates is less than one or both of
-	      the two coordinates found of the right side of the \ac{roi}
-	\item In the Y axis, the landmark's coordinates is greater than one or both of
-	      the two coordinates found of the top side of the \ac{roi}
-	\item In the Y axis, the landmark's coordinates is less than one or both of
+	\item On the X-axis, the landmark's coordinates are greater than one or both of
+	      the two coordinates found on the left side of the \ac{roi}
+	\item On the X-axis, the landmark's coordinates are less than one or both of
+	      the two coordinates found on the right side of the \ac{roi}
+	\item On the Y-axis, the landmark's coordinates are greater than one or both of
+	      the two coordinates found on the top side of the \ac{roi}
+	\item On the Y-axis, the landmark's coordinates are less than one or both of
 	      the two coordinates found of the bottom side of the \ac{roi}
@@
 These conditions maximized the total area of the \ac{roi}, and it is not strict
-about exact accuracy. In essence, these conditions creates a perfectly
+about exact accuracy. In essence, these conditions create a perfectly
 rectangular box that contained the quadrilateral formed by the four points from
-the coordinates.
+the coordinates. Figure~\ref{fig:metho-algo-sample} is an example of a frame with a
+successful finger-key identification using the conditions previously mentioned.
 
@@
 	\caption{Fingertip within an \ac{roi}}
+	\label{fig:metho-algo-sample}
 	\centering
@@
 \label{section:metho-algo-implementation}
-The previously discussed algorithms was implemented using Python. The Keyboard
-Detection, and Mapping Algorithm in Section~\ref{section:metho-algo-keyboard}
-was implemented as a separate module in Python using OpenCV. The Finger
-Detection, and Tracking Solution in Section~\ref{section:metho-algo-finger} was
-consumed using the prebuilt Python package offered by the MediaPipe team.
-Finally, the integration of the two will be done as another separate module
-in Python.
+The previously discussed algorithms were implemented using Python. The Keyboard
+Detection and Mapping Algorithm in Section~\ref{section:metho-algo-keyboard} was
+implemented as a separate module in Python using OpenCV. The Finger Detection
+and Tracking Solution in Section~\ref{section:metho-algo-finger} was consumed
+using the prebuilt Python package offered by the MediaPipe team. Finally, the
+integration of the two will be done as another separate module in Python.
 
@@
 \subsection{Typing Test Sequences}
-Typing test sequences are strings that was used in testing the user in their
-ability to type. The test sequences that was used came from text found in the
+Typing test sequences are strings that were used in testing the user in their
+ability to type. The test sequences that were used came from text found in the
 public domain obtained from Project Gutenberg and the Internet Archive.
-Sentences were isolated from these text and used as test sequences if they fit
+Sentences were isolated from these texts and used as test sequences if they fit
 the criteria.
 
-The criteria for choosing test sequences were as follows: (1) $\approx80\%$ of the
-characters in the keyboard is present in a test sequence. (2) The number of
-words in a test sequence do not exceed 25. (3) Numbers and punctuations should
+The criteria for choosing test sequences were as follows: (1) $\approx80\%$ of
+the characters on the keyboard is present in a test sequence. (2) The number of
+words in a test sequence does not exceed 25. (3) Numbers and punctuations should
 be present in at least $\approx30\%$ of the total test sequences.
 
-There was a total of 10 test sequences that was gathered. An example test
+There was a total of 10 test sequences that were gathered. An example test
 sequence is ``What of it, if some old hunks of a sea-captain orders me to get a
@@
 
-One caveat of the camera used was it's autofocus capabilities. This meant that
+One caveat of the camera used was its autofocus capabilities. This meant that
 there was a set amount of time during the start of the recording where the
@@
 
-\subsubsection{Keypress Labelling}
+\subsubsection{Keypress Labeling}
 The researcher then manually perform finger-key identification for all key
@@
 line was then shown, and the associated data with the frame was superimposed
-over it. The researcher then pressed keys that corresponded with the finger used
-to press the key. This was then again stored in the format
-\texttt{FRAME\_NO:KEY\_PRESSED:FINGER\_USED}, and each keypress was then stored
-as a line in a file.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.6\textwidth]{labelling.png}
-	\caption{Screen showing the labelling process}
+over it. Figure~\ref{fig:metho-data-process} is an example of a frame with these
+data overlaid over the captured image. The researcher then pressed keys that
+corresponded with the finger used to press the key. This was then again stored
+in the format \texttt{FRAME\_NO:KEY\_PRESSED:FINGER\_USED}, and each keypress
+was then stored as a line in a file.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.6\textwidth]{labeling.png}
+	\caption{Screen showing the labeling process}
+	\label{fig:metho-data-process}
 	\centering
@@
 A script was created to test the accuracy of the module. This script opened each
-video, and the associated labeled file. There were three frames taken for each
+video and the associated labeled file. There were three frames taken for each
 video to perform step one of the module: frames ten, fifteen, and twenty. The
@@
 
-The script then parsed the labeled file, and opened the frame for each keypress.
+The script then parsed the labeled file and opened the frame for each keypress.
 The second algorithm of the module was then run, using the data needed for the
 algorithm: the key pressed. The result of the module was then compared with the
-manual labelling of the keypress. The resulting data from the test was then
+manual labeling of the keypress. The resulting data from the test was then
 stored in a \texttt{.csv} file with the following content:
@@
 		\texttt{key}                          & Key pressed                                            \\
-		\texttt{finger}                       & Manual labelling of the finger used to press the key   \\
+		\texttt{finger}                       & Manual labeling of the finger used to press the key    \\
 		\texttt{is\_correct}                  & If the module was correct                              \\
@@
 its performance. Some runs had improvements in its statistics without doing any
-adjustments. These were runs that showed errors in the manual labelling and
+adjustments. These were runs that showed errors in the manual labeling and
 were corrected afterward.
@@
 		\midrule
-		Mismatched Identification                & Module output did not match manual labelling                                            \\
+		Mismatched Identification                & Module output did not match manual labeling                                             \\
 		No Identification                        & Module did not detect fingertips in \ac{roi}
@@
 
-There was a two total metrics obtained pertaining to finger and key mapping.
-The first was per test sequence, and the second was per key.
+There were two total metrics obtained pertaining to finger and key mapping. The
+first was per test sequence, and the second was per key.
 
@@
 The metric which calculates per test sequence is \ac{fpacc}. This metric
-computes the percentage of accurate keys pressed with the correct finger over the
-length of the typing test sequence. Inputting the wrong character, even if
+computes the percentage of accurate keys pressed with the correct finger over
+the length of the typing test sequence. Inputting the wrong character, even if
 pressed with the correct finger, reduces \ac{fpacc} since \ac{fpacc} measures
-\emph{accurate key presses}, and incorrect characters are considered as
-inaccurate key presses.
+\emph{accurate key presses}, and incorrect characters are considered inaccurate
+key presses.
 
@@
 \subsection{Per Key}
-This metric computes the \ac{hfpacc} for a certain key. This metric takes the
-times the user pressed the wrong key when asked for a certain key into account,
-the same as in \ac{fpacc}. This allows the user to easily verify which keys
-need attention and retraining if there is a high frequency of error.
+This metric, \ac{hfpacc}, computes the accuracy of the user in pressing a
+certain key with the correct finger over all test sequences. The same criteria
+in determining accurate key presses This allows the user to easily verify which
+keys need attention and training if there is a high frequency of error.
 
@@
 \section{Accuracy}
-Accuracy, in the context of the finger-key identification module calculates the
+Accuracy, in the context of the finger-key identification module, calculates the
 percentage of successful finger-key identifications that the module has
 performed. In this case, success is defined as having the expected finger, based
-on the manual labelling, equal the return value of the module. Accuracy is
-calculated as
+on the manual labeling, equal to the return value of the module. Accuracy is
+calculated as:
 
@@
 
-Over the seven training iterations, there was an increase in success percentage.
-Consequently, there was also a decrease in failure percentage. There was a total
-of 1475 total keypresses for iterations one to three, and 1571 total keypresses
-for iteration four to seven.
+Over the seven training iterations, there was an increase in success percentage
+as shown in Figure~\ref{fig:rd-training-identification}. Consequently, there was
+also a decrease in failure percentage. There was a total of 1475 total
+keypresses for iterations one to three, and 1571 total keypresses for iterations
+four to seven.
 
@@
 Map. Iterations one through three only tried frame 10 as the source frame for
-the module while iterations four through seven tried frame 10, 15, and 20.
+the module while iterations four through seven tried frames 10, 15, and 20.
 
@@
 its ability to focus on the keyboard for each video. In some videos video, the
-camera was not focused on the keyboard on frame 10, but it was able to focus on
+camera was not focused on the keyboard in frame 10, but it was able to focus on
 frame 15. As a result, the total number of available keypresses to test
@@
 
-The 0.48\% jump between iteration one and two can be attributed to correcting
-erroneous finger-key identifications that was obtained from manual labelling.
-The 9.56\% jump in success percentage between iteration two and three was due to
-adding buffers around the \ac{roi} by scaling the contours by 10 pixels on each
-side. Iteration four's increase was due to fixing the image map. There were
-certain keys that had incorrect color values assigned in the Key-Color Values
-Map, and some keys were of incorrect shape. Iterations five to seven were
-performed to test different values and shapes for scaling the \ac{roi}.
+The 0.48\% jump between iterations one and two can be attributed to correcting
+erroneous finger-key identifications that were obtained from manual labeling.
+The 9.56\% jump in success percentage between iterations two and three was due
+to adding buffers around the \ac{roi} by scaling the contours by 10 pixels on
+each side. Iteration four's increase was due to fixing the image map. Certain
+keys that had incorrect color values assigned in the Key-Color Values Map, and
+some keys were of incorrect shape. Iterations five to seven were performed to
+test different values and shapes for scaling the \ac{roi}.
 
@@
 able to detect two or more fingertips within the \ac{roi} of the key. As such,
-the module is uncertain which of the multiple fingertips actually pressed the
-key. However, it is still classified as a success since the expected finger from
-the manual labelling exists within the array of fingertips returned by the
-module.
+the module is uncertain which of the multiple fingertips pressed the key.
+However, it is still classified as a success since the expected finger from the
+manual labeling exists within the array of fingertips returned by the module.
+Figure~\ref{fig:rd-training-uncertain-successes} shows the trend of this metric
+throughout the different training iterations.
 
@@
 \paragraph{Non-spacebar}
-
 There were cases where the module had uncertain successes that were not over the
 spacebar. In almost practically all cases, these were over the alphanumeric
-keys. This uncertain successes stem from the increased buffer of the \ac{roi}.
-This type of uncertain successes carry more weight compared to spacebar
+keys. These uncertain successes stem from the increased buffer of the \ac{roi}.
+This type of uncertain success carries more weight compared to spacebar
 uncertain successes since the keys where these uncertain successes come from are
-the smallest keys in the keyboard. This means that the chances that a person
-actually has two or more fingers over the keys is rare, compared to the
-spacebar.
+the smallest keys on the keyboard. This means that the chances that a person has
+two or more fingers over the keys are rare, compared to the spacebar.
 
@@
 
-			\legend{Successes, Uncertain Successes}
+			\legend{Successes, Total Uncertain Successes}
 
@@
 	\caption{Uncertain Successes of the training iterations}
-	\label{fig:rd-training-uncertain-successes}
+	\label{fig:rd-training-uncertain-successes-vs-success}
 \end{figure}
@@
 percentage and making compromises between one or the other. The goal of the
-training process was to maximize the success percentage, while limiting the
+training process was to maximize the success percentage while limiting the
 number of uncertain successes.
 
-It can be noted that in iterations one and two, the success rate was lower
-compared to the other iterations. However, the percentage of uncertain successes
-was also the lowest. Iteration three was the first iteration that created a
-buffer around each key, however this buffer of 10 pixels was too aggressive,
-with the uncertain successes spiking upwards to nearly 30\%. Iteration five
-reduced the buffer size in half to five pixels. This did not affect success rate
-that much, with only a reduction of 1.07\%. This also greatly decreased the
-uncertain success percentage by 16.74\%. The sixth iteration met halfway between
-iteration four and five, by adding a buffer of seven pixels. This increased the
+As seen in Figure~\ref{fig:rd-training-uncertain-successes-vs-success}, it can
+be noted that in iterations one and two, the success rate was lower compared to
+the other iterations. However, the percentage of uncertain successes was also
+the lowest. Iteration three was the first iteration that created a buffer around
+each key, however, this buffer of 10 pixels was too aggressive, with the
+uncertain successes spiking upwards to nearly 30\%. Iteration five reduced the
+buffer size by half to five pixels. This did not affect the success rate that
+much, with only a reduction of 1.07\%. This also greatly decreased the uncertain
+success percentage by 16.74\%. The sixth iteration met halfway between
+iterations four and five, by adding a buffer of seven pixels. This increased the
 success rate back to 99.17\% without greatly increasing the number of uncertain
@@
 
-Based on these results, the final selected parameters for the module was the
+While the percentage of uncertain successes in iterations five through seven was
+in the range of 13\% to 20\%, The majority of these uncertain successes were
+spacebar uncertain successes, as shown in
+Figure~\ref{fig:rd-training-uncertain-successes}.
+
+Based on these results, the final selected parameters for the module were the
 parameters set in iteration six, with the buffer set to seven pixels as it was a
-good middle ground between success rate, and uncertain successes.
+good middle ground between success rate and uncertain successes.
 
@@
 	\caption{Uncertain Successes of the training iterations}
-	\label{fig:rd-training-uncertain-successes}
+	\label{fig:rd-training-failure-types}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.995\linewidth]{failure-3.png}
+\end{subfigure}\\
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.995\linewidth]{failure-1.png}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=.995\linewidth]{failure-2.png}
+\end{subfigure}
+\caption{Example of No Identification Failures}
+\label{fig:rd-training-sample}
 \end{figure}
@@
 These are failures where the module did not return the expected fingertip based
-on the manual labelling. This type of failure only manifested in iterations one
+on the manual labeling. This type of failure only manifested in iterations one
 and three. The failures in iteration one, after double-checking the video, were
-due to an error during manual labelling. In iteration three, these errors were
+due to an error during manual labeling. In iteration three, these errors were
 due to an incorrect image map and Key-Color Values map. The information for some
-keys were not consistent between both.
-
-Based on this information and from the gathered data, This failure type is
-non-existent for the module in real-life conditions.
+keys was not consistent between both.
+
+Based on this information and from the gathered data as shown in
+Figure~\ref{fig:rd-training-failure-types}, This failure type is non-existent
+for the module in real-life conditions.
 
@@
 These are failures where the module did not detect any fingertips within the
-\ac{roi}. These made up all the failures from the test results. These types of
-failures occurred because the fingertips used to press the key were not directly
-above the key, but rather offset to its side. Scaling the \ac{roi} would reduce
-the number of no identification failures, but it would also increase the number
-of uncertain successes. See Section~\ref{section:rd-uncertain vs success},
-Uncertain Successes vs Success Percentage for more information.
+\ac{roi}. Figure~\ref{fig:rd-training-sample} shows examples of frames where
+these errors occurred. These made up all the failures from the test results.
+These types of failures occurred because the fingertips used to press the key
+were not directly above the key, but rather offset to its side. Scaling the
+\ac{roi} would reduce the number of no identification failures, but it would
+also increase the number of uncertain successes. See
+Section~\ref{section:rd-uncertain vs success}, Uncertain Successes vs Success
+Percentage for more information.
 
@@
 
-The results were satisfactory after running the module with the parameters from
-the sixth training iteration on the test data. There was another run with the
-same test data and the same parameters, however these results were omitted since
-this run had inconclusive results as there was a single identification that was
-incorrectly labeled during manual finger-key identification.
+Table~\ref{tab:rd-accuracy} provides the accuracy results of the module with
+test data. These results were satisfactory after running the module with the
+parameters from the sixth training iteration on the test data. There was another
+run with the same test data and the same parameters, however, these results were
+omitted since this run had inconclusive results as there was a single
+identification that was incorrectly labeled during manual finger-key
+identification.
 
@@
 
-Based on the definition of \ac{wpm}
-from \cite{arif2009} as seen in Equation~\ref{equ:wpm}, it can be extrapolated that calculating Characters per
-Second is:
+Based on the definition of \ac{wpm} from \cite{arif2009} as seen in
+Equation~\ref{equ:wpm}, it can be extrapolated that calculating \ac{cps} is:
 
@@
 	CPS = \frac{WPM}{60} \cdot 5
+	\label{eq:cps}
 \end{equation}
@@
 
-The average \ac{wpm} when typing using the QWERTY layout, according to test
-data from \citeauthor{keybr} is $\approx37.5$ \ac{wpm}. As such, we can
-calculate that the average Characters per Second when typing using the QWERTY
-layout is 3.125. This means that each character, on average, requires 0.32
-seconds to type. This sets the maximum time spent for finger-key identification
-for a single character. Any more than this would result in noticeable slowdown
-during typing.
+The average \ac{wpm} when typing using the QWERTY layout, according to test data
+from \citeauthor{keybr} is $\approx37.5$ \ac{wpm}. As such, we can calculate
+that the average Characters per Second when typing using the QWERTY layout is
+3.125. We can calculate the seconds needed per character by this equation:
+
+\begin{equation}
+	s = \frac{1}{CPS}
+	\label{eq:s}
+\end{equation}
+\labelequations{Seconds per Character}
+
+This means that each character, on average, requires 0.32 seconds to type. This
+sets the maximum time spent for finger-key identification for a single
+character. Any more than this would result in a noticeable slowdown during
+typing.
 
@@
 	\end{tikzpicture}
-	\caption{Uncertain Successes of the training iterations}
-	\label{fig:rd-training-uncertain-successes}
-\end{figure}
-
-Training iterations had an average keyboard detection time of $0.8620\pm0.0502s$
-and average finger identification time of $0.0863\pm0.0040s$. Over the course of
-the training, there was no noticeable difference between running time over the
-same dataset. However, The first training iteration had issues with gathering
-the speed of the module, as the average keyboard detection time included other
-operations that was not part of getting the Key-Edge Coordinates Map. Removing
-this iteration, the average keyboard detection time becomes more consistent,
-with it being $0.8295\pm0.0070s$.
+	\caption{Keyboard detection time of the training iterations}
+	\label{fig:rd-training-keyboard-detection-time}
+\end{figure}
+
+The training iterations had an average keyboard detection time of
+$0.8620\pm0.0502s$ and average finger identification time of $0.0863\pm0.0040s$.
+Throughout the training, there was no noticeable difference in running time over
+the same dataset as shown in
+Figure~\ref{fig:rd-training-keyboard-detection-time}. However, The first
+training iteration had issues with gathering the speed of the module, as the
+average keyboard detection time included other operations that were not part of
+getting the Key-Edge Coordinates Map. Removing this iteration, the average
+keyboard detection time becomes more consistent, with it being
+$0.8295\pm0.0070s$.
 
@@
 
+These results, as tabulated in Table~\ref{tab:rd-speed} were consistent with the
+training iterations. This means that the module has repeatable and predictable
+running times.
+
 \subsection{Average Keyboard Detection Time}
@@
 
-The average finger identification time that the module got, 0.08389272103s was
-well below 0.32 seconds. This can be attributed to the speed of the MediaPipe
+The average finger identification time that the module got, 0.083s was
+well below 0.32s. This can be attributed to the speed of the MediaPipe
 library, and the pre-calculation of the edge coordinates.
 
+\subsubsection{Real-Time Finger-Key Identification}
+With each keypress only requiring 0.083s for finger-key identification, it can
+be gathered that the maximum \ac{cps} which the module can still perform
+real-time finger-key identification is 11.9200 \ac{cps}, based on
+Equation~\ref{eq:s}. It follows that the maximum \ac{wpm} that still allows for
+real-time finger-key identification is 143.040 \ac{wpm}, based on
+Equation~\ref{eq:cps}. This is well above the median of 37.5 \ac{wpm} on Keybr.
+The module can also perform finger-key identification for 90.99\% of the
+community members of Monkeytype with the speed role.
+
 \chapter{Conclusion}
 
+The finger key identification module was successfully developed and implemented
+in this study.
+
+The finger-key identification solution was split into two parts, keyboard
+detection and key mapping, and finger detection. Keyboard detection and key
+mapping used the following algorithm and techniques: (1) Edge Detection using
+Sobel filter \parencite{sobel2014}, (2) Thresholding using Otsu's algorithm
+\parencite{otsu}, (3) Finding contours using the algorithm of \cite{contours},
+(4) Line Simplification using the Douglas-Peucker algorithm
+\parencite{douglas-peucker}, and (5) Perspective transform
+\parencite{opencv-image-transform}. Finger detection utilized a ready-made
+solution called MediaPipe Hands by \cite{mediapipe}. The two were combined to
+perform finger-key identification.
+
+The keyboard used was a 60\% keyboard in ANSI. It was light in color, and was
+connected with a black USB-C cable. The surface of the desk was dark, and a LED
+Bulb rated at 9 Watts, 700 lumens, and 6500k color temperature was used to
+uniformly light the capture area. The single optical camera, a Logitech C920,
+was placed above the keyboard, pointing directly downwards. It captured videos
+in 720p/30fps with a diagonal field of view of 78\degree \parencite{logitech}.
+The computer used had an AMD Ryzen 5 3600 CPU, AMD Radeon RX 5600 XT GPU,
+G.Skill Trident Z Neo RGB 16GB 3200mhz RAM, and Samsung SSD 850 EVO SSD.
+
+The development of the module utilized OpenCV as the main image manipulation
+platform that implemented the algorithms mentioned previously. OpenCV was also
+used to capture videos from the optical camera. It was all written in Python.
+
+The module was accurate in 99.47\% of identifications in a data set composed of
+942 keypresses. However, 14.12\% of the identifications were uncertain
+successes. These are successes where more than one fingertip was found in an
+\ac{roi}. Of these, 6.48\% were non-spacebar uncertain successes. These are
+uncertain successes where that was not over the spacebar. This bears more weight
+since these keys were the smallest on the keyboard, and in most cases, only one
+fingertip could fit within the key. These uncertain successes were due to the
+addition of a buffer around each \ac{roi} to increase the success rate.
+
+The module is fast enough for real-time finger-key identification up to typing
+speeds of 143.040 \ac{wpm}. This is above the median typing speed of 37.5
+\ac{wpm} on Keybr \parencite{keybr}. This also handles 90.99\% of the 12678
+members with the speed role on MonkeyType \parencite{monkey-stats}.
+
+Two metrics were developed: (1) Finger Placement Accuracy which computes the
+percentage of keys pressed with the correct figure over the length of the test
+sequence, and (2) Historical Finger Placement Accuracy which computes the number
+of times the user is successful if pressing a key with the correct finger over
+all test sequences.
+
 \chapter{Future Work}
+
+\section{Trainer}
+The paper only presented a module for finger-key identification and did not
+include a trainer that utilized the module. This would allow the module to be
+used by actual users and allow for further testing to be done.
+
+\section{Effects on Touch Typing Education}
+The paper created a method for finger-key identification in hopes of improving
+touch typing education, however, there was no test done to prove that the module
+will affect the learner's skill in touch typing
+
+\section{Effects on Health}
+In connection with the previous section, the module also has no testing done to
+quantify if training using the module will improve the posture and reduce
+problematic hand positions and movements that can cause nerve and muscular
+disorders.
+
+\section{Environmental Setup}
+The environmental setup was tightly controlled, with the keyboard and the desk
+of contrasting color, lighting was nearly perfect, and the camera was of great
+quality. This is not realistic and is not representative of the real world.
+Future work could improve upon this area. An idea would be to have a separate
+process where the user could just select four points rather than relying on an
+automated computer vision process to detect the four edges.
+
+\section{Detecting Keyboard Movement Between Frames}
+The module is limited in its ability to track keyboards that move between
+keypresses. While this is not something that usually happens during keyboard
+typing, having the ability to track movements would be a great addition to
+improving the module's performance.
+
+\section{Different Keyboard Types, Layouts, and Colors}
+The module is limited to detecting keypresses with a 60\% keyboard. This type of
+keyboard is not that commonly used. Other keyboard types, like the 100\%, 75\%,
+and TKLs would be good starting points for adding support. In addition,
+supporting other layouts, like ISO, would also greatly increase the number of
+people that could use the module. Furthermore, supporting other color schemes
+would also widen the number of keyboards that can be used with the module.
 
@@
 
-\tablehead{\toprule Hex Color Code&\multicolumn{1}{l}{Corresponding Key} \\ \midrule}
+\tablehead{\toprule Key &\multicolumn{1}{l}{Color (in BGR)} \\ \midrule}
 \tabletail{%
@@
 \begin{supertabular}{ll}
-	\#FF0000 & $\sim$            \\
-	\#FF9900 & 1                 \\
-	\#FFE600 & 2                 \\
-	\#BDFF00 & 3                 \\
-	\#8FFF00 & 4                 \\
-	\#33FF00 & 5                 \\
-	\#00FF66 & 6                 \\
-	\#00FFC2 & 7                 \\
-	\#00E0FF & 8                 \\
-	\#00A3FF & 9                 \\
-	\#0066FF & 0                 \\
-	\#000AFF & -                 \\
-	\#5200FF & +                 \\
-	\#BD00FF & Backspace         \\
-
-	\#E00000 & Tab               \\
-	\#D68000 & Q								 \\
-	\#E1CB00 & W								 \\
-	\#A0D900 & E								 \\
-	\#7EE000 & R								 \\
-	\#2BD900 & T								 \\
-	\#00D856 & Y								 \\
-	\#00C395 & U								 \\
-	\#00B7D0 & I								 \\
-	\#0088D4 & O								 \\
-	\#004FC5 & P								 \\
-	\#0008D3 & [								 \\
-			\#3F00C5 & ]								 \\
-	\#8600B5 & $\vert$					 \\
-
-	\#C30000 & CapsLock          \\
-	\#A76400 & A								 \\
-	\#B2A100 & S								 \\
-	\#7CA800 & D								 \\
-	\#65B400 & F								 \\
-	\#1F9D00 & G								 \\
-	\#008B38 & H								 \\
-	\#018B6A & J								 \\
-	\#008294 & K								 \\
-	\#005E93 & L								 \\
-	\#003482 & :								 \\
-	\#0007AA & "								 \\
-	\#23006D & Enter						 \\
-
-	\#960000 & LeftShift         \\
-	\#683E00 & Z								 \\
-	\#827500 & X								 \\
-	\#4D6800 & C								 \\
-	\#3F7100 & V								 \\
-	\#125800 & B								 \\
-	\#00421A & N								 \\
-	\#005440 & M								 \\
-	\#004B55 & ,								 \\
-	\#00466D & .								 \\
-	\#002050 & /								 \\
-	\#000580 & RightShift				 \\
-
-	\#550000 & LeftControl       \\
-	\#2D0000 & LeftSuper         \\
-	\#3A2300 & LeftAlt           \\
-	\#3E3700 & Spacebar					 \\
-	\#003238 & RightAlt					 \\
-	\#001C2C & RightMeta 				 \\
-	\#000349 & RightSuper				 \\
-	\#18004D & RightControl 		 \\
+	LeftControl & (85, 0, 0)\\
+	LeftSuper & (45, 0, 0)\\
+	LeftAlt & (58, 35, 0)\\
+	Spacebar & (62, 55, 0)\\
+	RightAlt & (0, 28, 44)\\
+	RightSuper & (0, 3, 73)\\
+	RightControl & (24, 0, 76)\\
+	` & (255, 0, 0)\\
+	1 & (255, 153, 0)\\
+	2 & (255, 229, 0)\\
+	3 & (189, 255, 0)\\
+	4 & (143, 255, 0)\\
+	5 & (51, 255, 0)\\
+	6 & (0, 255, 102)\\
+	7 & (0, 255, 194)\\
+	8 & (0, 224, 255)\\
+	9 & (0, 163, 255)\\
+	0 & (0, 102, 255)\\
+	- & (0, 10, 255)\\
+	= & (82, 0, 255)\\
+	$\vert$ & (128, 0, 255)\\
+	Delete & (189, 0, 255)\\
+	Tab & (224, 0, 0)\\
+	Q & (214, 128, 0)\\
+	W & (225, 203, 0)\\
+	E & (160, 217, 0)\\
+	R & (126, 224, 0)\\
+	T & (43, 217, 0)\\
+	Y & (0, 216, 86)\\
+	U & (0, 195, 149)\\
+	I & (0, 183, 208)\\
+	O & (0, 136, 212)\\
+	P & (0, 79, 196)\\
+	] & (0, 8, 211)\\
+	{[} & (63, 0, 196)\\
+	Backspace & (134, 0, 181)\\
+	CapsLock & (195, 0, 0)\\
+	A & (167, 100, 0)\\
+	S & (178, 161, 0)\\
+	D & (124, 168, 0)\\
+	F & (101, 180, 0)\\
+	G & (31, 157, 0)\\
+	H & (0, 139, 56)\\
+	J & (1, 139, 106)\\
+	K & (0, 130, 148)\\
+	L & (0, 94, 147)\\
+	; & (0, 52, 130)\\
+	' & (0, 7, 170)\\
+	Enter & (35, 0, 109)\\
+	LeftShift & (150, 0, 0)\\
+	Z & (104, 62, 0)\\
+	X & (130, 117, 0)\\
+	C & (77, 104, 0)\\
+	V & (63, 113, 0)\\
+	B & (18, 88, 0)\\
+	N & (0, 66, 26)\\
+	M & (0, 84, 64)\\
+	, & (0, 75, 85)\\
+	. & (0, 70, 109)\\
+	/ & (0, 32, 80)\\
+	RightShift & (0, 5, 128)\\
+
 \end{supertabular}
@@
 
-\chapter{Front-End Screens}
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{frontend-statistics.png}
-	\caption{Page showing historical statistics of the user}
-	\centering
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{frontend-guide-1.png}
-	\caption{Introduction to the platform during first open}
-	\centering
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{frontend-guide-2.png}
-	\caption{Step 1 of the tour of the platform}
-	\centering
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{frontend-guide-3.png}
-	\caption{Step 2 of the tour of the platform}
-	\centering
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{frontend-guide-4.png}
-	\caption{Step 3 of the tour of the platform}
-	\centering
-\end{figure}
-
 \chapter{Gantt Chart}

### 2022-06-08 application/algorithms/common.py

@@ -1,109 +1,0 @@
-#!/usr/bin/env python3
-import numpy as np
-from cv2 import cv2 as cv
-
-
-DEBUG = True
-
-
-def findContours(frame: np.ndarray) -> tuple[np.ndarray]:
-    contours, _ = cv.findContours(frame, cv.RETR_EXTERNAL, cv.CHAIN_APPROX_SIMPLE)
-
-    if DEBUG:
-        contourFrame = cv.cvtColor(frame, cv.COLOR_GRAY2BGR)
-        cv.drawContours(contourFrame, contours, -1, (0, 255, 0), 2)
-        cv.imshow("debug", contourFrame)
-        cv.waitKey(0)
-
-    return contours
-
-
-def approximateLargestContour(contours: tuple[np.ndarray]) -> np.ndarray:
-    # Sort contours by area
-    largestContour = sorted(contours, key=cv.contourArea, reverse=True)[0]
-
-    # Use Douglas-Peucker algorithm to approximate a shape
-    # Set accuracy using the epsilon. 1% accuracy as the simplest shape, a
-    # rectangle, is expected
-    epsilon = 0.01 * cv.arcLength(largestContour, True)
-    approximation = cv.approxPolyDP(largestContour, epsilon, True)
-
-    return approximation
-
-
-def sortContourPoints(contour: np.ndarray) -> np.ndarray:
-    # Sort by x axis
-    indices = contour[:, 0].argsort()
-    contour = contour[indices]
-
-    # Get the two left points and sort by y axis
-    left = contour[:2]
-    leftIndices = left[:, 1].argsort()
-    left = left[leftIndices]
-
-    # Get the two right points and sort by y axis
-    right = contour[2:]
-    rightIndices = right[:, 1].argsort()
-    right = right[rightIndices]
-
-    # arrange clockwise
-    return np.array([left[0], right[0], right[1], left[1]])
-
-
-def scaleContours(contours: np.ndarray) -> np.ndarray:
-    scaleFactor = 7
-
-    upperLeft = [contours[0][0] - scaleFactor, contours[0][1] - scaleFactor]
-    upperRight = [contours[1][0] + scaleFactor, contours[1][1] - scaleFactor]
-    lowerRight = [contours[2][0] + scaleFactor, contours[2][1] + scaleFactor]
-    lowerLeft = [contours[3][0] - scaleFactor, contours[3][1] + scaleFactor]
-
[…]

### 2022-06-08 application/algorithms/finger_detection.py

@@ -1,128 +1,0 @@
-#!/usr/bin/env python3
-
-from cv2 import cv2 as cv
-import mediapipe as mp
-import numpy as np
-
-from common import DEBUG, DEBUG_displayContours, convertResultToStr
-
-# Heavily lifted from https://google.github.io/mediapipe/getting_started/python.html.
-# Cited as (Lugaresi et al., n.d.) in the paper
-def classifyPressedFinger(image: np.ndarray, roiPoints: np.ndarray) -> list[str]:
-
-    mp_drawing = mp.solutions.drawing_utils
-    mp_drawing_styles = mp.solutions.drawing_styles
-    mp_hands = mp.solutions.hands
-
-    image_height, image_width, _ = image.shape
-
-    with mp_hands.Hands(
-        static_image_mode=True, max_num_hands=2, min_detection_confidence=0.5
-    ) as hands:
-        results = hands.process(cv.cvtColor(image, cv.COLOR_BGR2RGB))
-
-        if not results.multi_hand_landmarks:
-            raise Exception("finger_detection: no landmarks found")
-
-        # Get all possible landmarks that fit witihin the ROI. This is because
-        # some landmarks such as INDEX_FINGER_PIP may fall within the ROI, and
-        # it is highly unlikely that someone will click using that part of the
-        # finger. However, if no other matches are to be found, this is taken as
-        # the fallback
-        possibleMatches: list[tuple[str, str]] = []
-
-        for handIndex, hand_landmarks in enumerate(results.multi_hand_landmarks):
-            handedness = results.multi_handedness[handIndex].classification[0].label
-
-            # Flip handedness. Mediapipe expects the images to be flipped
-            # horizontally, much like taking a selfie. However, this wont work
-            # for our case
-            if handedness == "Left":
-                handedness = "RIGHT"
-            else:
-                handedness = "LEFT"
-
-            for landmarkIndex, landmark in enumerate(hand_landmarks.landmark):
-                x = round(landmark.x * image_width)
-                y = round(landmark.y * image_height)
-                landmarkName = mp_hands.HandLandmark(landmarkIndex).name
-
-                if DEBUG:
-                    print(handedness)
-                    print(landmarkName)
-                    print(x, y)
-                    print(landmark)
-
-                # The following if statements check if the landmark is within
-                # the ROI. If it fails, the next landmark is tested
-
-                # In the X axis, the landmark's coordinates is greater than one
-                # or both of the two coordinates found on the left side of the
-                # ROI
[…]

### 2022-06-08 application/algorithms/integration.py

@@ -1,54 +1,0 @@
-#!/usr/bin/env python3
-
-from cv2 import cv2 as cv
-import numpy as np
-
-# import time
-
-from common import DEBUG, DEBUG_displayContours
-
-from keyboard_detection import getContourPointsFromImage
-from finger_detection import classifyPressedFinger
-
-
-def main(keyboardImage: np.ndarray, handImage: np.ndarray):
-
-    # start = time.time()
-    if DEBUG:
-        cv.namedWindow("debug")
-
-    # Step 1
-    allContourPoints = getContourPointsFromImage(keyboardImage)
-
-    contourPoints = allContourPoints["LeftShift"]
-
-    # elapsed = time.time() - start
-
-    # Step 2
-    if DEBUG:
-        cv.imshow("debug", handImage)
-        cv.waitKey(0)
-
-    result = classifyPressedFinger(handImage, contourPoints)
-
-    if result is not None:
-        print(result)
-
-    # total = time.time() - start
-    # classifying = total - elapsed
-
-    # print("Init Time: " + str(elapsed))
-    # print("Classifying Time: " + str(classifying))
-    # print("Total Time: " + str(total))
-
-    if DEBUG:
-        print(result)
-        cv.destroyAllWindows()
-
-    return result
-
-
-if __name__ == "__main__":
-    keyboardImage = cv.imread("images/medium-04-10.png")
-    handImage = cv.imread("images/medium-04-437.png")
-    main(keyboardImage, handImage)

### 2022-06-08 application/algorithms/keyboard_detection.py

@@ -1,216 +1,0 @@
-from cv2 import cv2 as cv
-import numpy as np
-
-from copy import deepcopy
-
-from keyboard_map import keyDict
-from common import (
-    findContours,
-    approximateLargestContour,
-    sortContourPoints,
-    scaleContours,
-    DEBUG_displayContours,
-    DEBUG,
-)
-
-
-def getKeyboardImage() -> np.ndarray | None:
-    cam = cv.VideoCapture(0)
-
-    while True:
-        ret, frame = cam.read()
-
-        if not ret:
-            return None
-
-        cv.imshow("keyboard", frame)
-
-        key = cv.waitKey(1)
-
-        if key % 256 == 155:
-            cv.destroyWindow("keyboard")
-            return frame
-
-
-def getEdgeFrame(frame: np.ndarray) -> np.ndarray:
-    grayscaleFrame = cv.cvtColor(frame, cv.COLOR_BGR2GRAY)
-
-    if DEBUG:
-        cv.imshow("debug", grayscaleFrame)
-        cv.waitKey(0)
-
-    # Denoise while maintaining edges
-    bilateralFilteredFrame = cv.bilateralFilter(grayscaleFrame, 11, 17, 17)
-
-    if DEBUG:
-        cv.imshow("debug", bilateralFilteredFrame)
-        cv.waitKey(0)
-
-    # Perform Sobel Filtering
-    sobelX = cv.Sobel(
-        src=bilateralFilteredFrame,
-        ddepth=cv.CV_64F,
-        dx=1,
-        dy=0,
-        ksize=3,
-        borderType=cv.BORDER_DEFAULT,
-    )
-
-    sobelY = cv.Sobel(
-        src=bilateralFilteredFrame,
-        ddepth=cv.CV_64F,
-        dx=0,
-        dy=1,
-        ksize=3,
-        borderType=cv.BORDER_DEFAULT,
-    )
-
-    absSobelX = cv.convertScaleAbs(sobelX)
-    absSobelY = cv.convertScaleAbs(sobelY)
-
-    finalSobel = cv.addWeighted(absSobelX, 0.5, absSobelY, 0.5, 0)
-
-    if DEBUG:
-        cv.imshow("debug", finalSobel)
-        cv.waitKey(0)
-
-    # Use Otsu's Threshold to get better defined edges while not using arbitrary
-    # values
-    _, threshedFrame = cv.threshold(
[…]

### 2022-06-08 application/algorithms/keyboard_map.py

@@ -1,66 +1,0 @@
-#!/usr/bin/env python3
-
-
-keyDict = {
-    "LeftControl": [85, 0, 0],
-    "LeftSuper": [45, 0, 0],
-    "LeftAlt": [58, 35, 0],
-    "Spacebar": [62, 55, 0],
-    "RightAlt": [0, 28, 44],
-    "RightSuper": [0, 3, 73],
-    "RightControl": [24, 0, 76],
-    "`": [255, 0, 0],
-    "1": [255, 153, 0],
-    "2": [255, 229, 0],
-    "3": [189, 255, 0],
-    "4": [143, 255, 0],
-    "5": [51, 255, 0],
-    "6": [0, 255, 102],
-    "7": [0, 255, 194],
-    "8": [0, 224, 255],
-    "9": [0, 163, 255],
-    "0": [0, 102, 255],
-    "-": [0, 10, 255],
-    "=": [82, 0, 255],
-    "|": [128, 0, 255],
-    "Delete": [189, 0, 255],
-    "Tab": [224, 0, 0],
-    "Q": [214, 128, 0],
-    "W": [225, 203, 0],
-    "E": [160, 217, 0],
-    "R": [126, 224, 0],
-    "T": [43, 217, 0],
-    "Y": [0, 216, 86],
-    "U": [0, 195, 149],
-    "I": [0, 183, 208],
-    "O": [0, 136, 212],
-    "P": [0, 79, 196],
-    "[": [0, 8, 211],
-    "]": [63, 0, 196],
-    "Backspace": [134, 0, 181],
-    "CapsLock": [195, 0, 0],
-    "A": [167, 100, 0],
-    "S": [178, 161, 0],
-    "D": [124, 168, 0],
-    "F": [101, 180, 0],
-    "G": [31, 157, 0],
-    "H": [0, 139, 56],
-    "J": [1, 139, 106],
-    "K": [0, 130, 148],
-    "L": [0, 94, 147],
-    ";": [0, 52, 130],
-    "'": [0, 7, 170],
-    "Enter": [35, 0, 109],
-    "LeftShift": [150, 0, 0],
-    "Z": [104, 62, 0],
-    "X": [130, 117, 0],
-    "C": [77, 104, 0],
-    "V": [63, 113, 0],
-    "B": [18, 88, 0],
-    "N": [0, 66, 26],
-    "M": [0, 84, 64],
-    ",": [0, 75, 85],
[…]

### 2022-06-08 application/algorithms/test-scripts/clean-data.py

@@ -1,18 +1,0 @@
-#!/usr/bin/env python3
-from os import walk
-
-
-def main():
-    (_, _, filenames) = next(walk("../dataset/training"))
-    files = [file for file in filenames if not "mp4" in file]
-
-    for file in files:
-        with open(f"../dataset/training/{file}", "r+") as f:
-            content = f.read().replace(" ", "").replace("//", "/")
-            f.seek(0)
-            f.write(content)
-            f.truncate()
-
-
-if __name__ == "__main__":
-    main()

### 2022-06-08 application/algorithms/test-scripts/collect.py

@@ -1,194 +1,0 @@
-#!/usr/bin/env python3
-
-import sys
-import random
-from cv2 import cv2 as cv
-
-asciiDict = {
-    8: "Backspace",
-    9: "Tab",
-    13: "Enter",
-    27: "LeftControl",
-    32: "Spacebar",
-    33: "1",
-    34: "'",
-    35: "3",
-    36: "4",
-    37: "5",
-    38: "7",
-    39: "'",
-    40: "9",
-    41: "0",
-    42: "8",
-    43: "=",
-    44: ",",
-    45: "-",
-    46: ".",
-    47: "/",
-    48: "0",
-    49: "1",
-    50: "2",
-    51: "3",
-    52: "4",
-    53: "5",
-    54: "6",
-    55: "7",
-    56: "8",
-    57: "9",
-    58: ";",
-    59: ";",
-    60: ",",
-    61: "=",
-    62: ".",
-    63: "//",
-    64: "2",
-    65: "A",
-    66: "B",
-    67: "C",
-    68: "D",
-    69: "E",
-    70: "F",
-    71: "G",
-    72: "H",
-    73: "I",
-    74: "J",
-    75: "K",
-    76: "L",
-    77: "M",
-    78: "N",
-    79: "O",
-    80: "P",
-    81: "Q",
-    82: "R",
-    83: "S",
-    84: "T",
-    85: "U",
-    86: "V",
-    87: "W",
-    88: "X",
-    89: "Y",
-    90: "Z",
-    91: "[",
-    92: "\\",
-    93: "]",
-    94: "6",
-    95: "-",
-    96: "`",
-    97: "A",
-    98: "B",
-    99: "C",
-    100: "D",
-    101: "E",
-    102: "F",
-    103: "G",
-    104: "H",
-    105: "I",
-    106: "J",
-    107: "K",
-    108: "L",
-    109: "M",
-    110: "N",
-    111: "O",
-    112: "P",
-    113: "Q",
-    114: "R",
-    115: "S",
-    116: "T",
-    117: "U",
-    118: "V",
-    119: "W",
-    120: "X",
-    121: "Y",
-    122: "Z",
-    123: "[",
-    124: "\\",
-    125: "]",
-    126: "`",
-    225: "LeftShift",
-    226: "RightShift",
-    233: "LeftAlt",
-    234: "RightAlt",
-    235: "LeftSuper",
-    255: "Delete",
-}
-
-
[…]

### 2022-06-08 application/algorithms/test-scripts/confirm-label.py

@@ -1,106 +1,0 @@
-#!/usr/bin/env python3
-from os import walk, path
-from cv2 import cv2 as cv
-
-
-keypressDict = {
-    97: "LEFT_PINKY",
-    115: "LEFT_RING",
-    100: "LEFT_MIDDLE",
-    102: "LEFT_INDEX",
-    106: "RIGHT_INDEX",
-    107: "RIGHT_MIDDLE",
-    108: "RIGHT_RING",
-    59: "RIGHT_PINKY",
-    233: "LEFT_THUMB",
-    32: "RIGHT_THUMB",
-    27: "ESCAPE",
-    72: "INVALID",
-}
-
-
-def main():
-    (_, _, filenames) = next(walk("../dataset/training"))
-    files = sorted(
-        [file for file in filenames if not "mp4" in file and not "labeled" in file]
-    )
-
-    for index, file in enumerate(files):
-        fileLocation = f"../dataset/training/{file}"
-        cap = cv.VideoCapture(fileLocation + ".mp4")
-        labeledFile = f"{fileLocation}-labeled"
-
-        newFileContent = []
-
-        with open(labeledFile) as fp:
-            fileContent = [line.rstrip() for line in fp]
-
-        for keypressIndex, keypress in enumerate(fileContent):
-            data = keypress.split(":")
-            cap.set(cv.CAP_PROP_POS_FRAMES, int(data[0]))
-            ret, frame = cap.read()
-
-            if not ret:
-                raise Exception(f"label: frame {data[0]} not found for {file}")
-
-            frame = cv.rotate(frame, cv.ROTATE_180)
-            cv.putText(
-                frame,
-                keypress,
-                (200, 150),
-                cv.FONT_HERSHEY_SIMPLEX,
-                0.5,
-                (0, 255, 255),
-                1,
-                cv.LINE_AA,
-            )
-
-            cv.putText(
-                frame,
-                f"kp: {keypressIndex + 1}/{len(fileContent)} | file: {index + 1}/{len(files)}",
-                (10, 465),
-                cv.FONT_HERSHEY_SIMPLEX,
-                0.5,
-                (0, 255, 255),
-                1,
-                cv.LINE_AA,
-            )
-
-            cv.imshow("debug", frame)
-
-            keypressString = ""
-
-            while True:
-                key = cv.waitKey(0)
-
-                if key % 256 == 27:
-                    cv.destroyAllWindows()
-                    return
-
-                if key % 256 == 13:
-                    keypressString = ""
-                    break
-
-                try:
[…]

### 2022-06-08 application/algorithms/test-scripts/get-image.py

@@ -1,32 +1,0 @@
-#!/usr/bin/env python3
-
-from cv2 import cv2 as cv
-
-
-def main():
-    file = "medium-04"
-    keyboardFrame = 10
-    fingerFrame = 437
-    fileLocation = f"../dataset/training/{file}"
-
-    cap = cv.VideoCapture(f"{fileLocation}.mp4")
-
-    cap.set(cv.CAP_PROP_POS_FRAMES, keyboardFrame)
-    ret, frame = cap.read()
-
-    if not ret:
-        return
-
-    cv.imwrite(f"../images/{file}-{keyboardFrame}.png", frame)
-
-    cap.set(cv.CAP_PROP_POS_FRAMES, fingerFrame)
-    ret, frame = cap.read()
-
-    if not ret:
-        return
-
-    cv.imwrite(f"../images/{file}-{fingerFrame}.png", frame)
-
-
-if __name__ == "__main__":
-    main()

### 2022-06-08 application/algorithms/test-scripts/label.py

@@ -1,103 +1,0 @@
-#!/usr/bin/env python3
-from os import walk, path
-from cv2 import cv2 as cv
-
-
-keypressDict = {
-    97: "LEFT_PINKY",
-    115: "LEFT_RING",
-    100: "LEFT_MIDDLE",
-    102: "LEFT_INDEX",
-    106: "RIGHT_INDEX",
-    107: "RIGHT_MIDDLE",
-    108: "RIGHT_RING",
-    59: "RIGHT_PINKY",
-    233: "LEFT_THUMB",
-    32: "RIGHT_THUMB",
-    27: "ESCAPE",
-    72: "INVALID",
-}
-
-
-def main():
-    (_, _, filenames) = next(walk("../dataset/test"))
-    files = sorted(
-        [file for file in filenames if not "mp4" in file and not "labeled" in file]
-    )
-
-    for index, file in enumerate(files):
-        fileLocation = f"../dataset/test/{file}"
-        cap = cv.VideoCapture(fileLocation + ".mp4")
-
-        newFileContent = []
-
-        if path.exists(f"{fileLocation}-labeled"):
-            continue
-
-        with open(fileLocation) as fp:
-            fileContent = [line.rstrip() for line in fp]
-
-        cap.set(cv.CAP_PROP_POS_FRAMES, 10)
-        ret, frame = cap.read()
-        frame = cv.rotate(frame, cv.ROTATE_180)
-
-        cv.imshow("setup", frame)
-
-        for keypressIndex, keypress in enumerate(fileContent):
-            data = keypress.split(":")
-            cap.set(cv.CAP_PROP_POS_FRAMES, int(data[0]))
-            ret, frame = cap.read()
-
-            if not ret:
-                raise Exception(f"label: frame {data[0]} not found for {file}")
-
-            frame = cv.rotate(frame, cv.ROTATE_180)
-            cv.putText(
-                frame,
-                keypress,
-                (250, 100),
-                cv.FONT_HERSHEY_SIMPLEX,
-                0.5,
-                (0, 255, 255),
-                1,
-                cv.LINE_AA,
-            )
-
-            cv.putText(
-                frame,
-                f"kp: {keypressIndex + 1}/{len(fileContent)} | file: {index + 1}/{len(files)}",
-                (10, 465),
-                cv.FONT_HERSHEY_SIMPLEX,
-                0.5,
-                (0, 255, 255),
-                1,
-                cv.LINE_AA,
-            )
-
[…]

### 2022-06-08 application/algorithms/test.py

@@ -1,146 +1,0 @@
-#!/usr/bin/env python3
-
-from os import walk, mkdir
-from cv2 import cv2 as cv
-import time
-import csv
-
-from finger_detection import classifyPressedFinger
-from keyboard_detection import getContourPointsFromImage
-
-
-def main():
-    (_, _, filenames) = next(walk("./dataset/training"))
-
-    # Create output folder
-    uuid = f"training-{time.time()}"
-    folder = f"./results/{uuid}"
-    mkdir(folder)
-
-    files = sorted(
-        [file for file in filenames if not "mp4" in file and not "labeled" in file]
-    )
-
-    # Create log for entire run
-    logFileContent = []
-    resultFileContent = [
-        [
-            "file_name",
-            "frame",
-            "key",
-            "finger",
-            "is_correct",
-            "fingertips_detected",
-            "keyboard_detection_time",
-            "finger_classification_time",
-        ]
-    ]
-
-    for file in files:
-        # Output Files
-        logFileContent.append(f"INFO: Start {file}")
-
-        fileLocation = f"./dataset/training/{file}"
-        labeledFile = f"{fileLocation}-labeled"
-
-        cap = cv.VideoCapture(fileLocation + ".mp4")
-
-        keyboardDetectionTimeStart = time.time()
-
-        allContourPoints = {}
-
-        for i in [10, 15, 20]:
-            cap.set(cv.CAP_PROP_POS_FRAMES, i)
-            ret, frame = cap.read()
-            if not ret:
-                logFileContent.append(f"ERROR: Failed to get frame {i} for {file}")
-                break
-
-            try:
-                keyboardDetectionTimeStart = time.time()
-                allContourPoints = getContourPointsFromImage(frame)
-            except Exception as e:
-                logFileContent.append(
-                    f"ERROR: Failed to get contour points for {file} with frame {i}"
-                )
-                logFileContent.append(f"ERROR: {e}")
-                cv.imwrite(f"{folder}/{file}:{i}.png", frame)
-            else:
-                break
-
-        if len(allContourPoints) == 0:
-            logFileContent.append(f"ERROR: Failed to get contour points for {file}")
-            continue
-
-        keyboardDetectionTimeEnd = time.time() - keyboardDetectionTimeStart
-
-        with open(labeledFile) as fp:
-            fileContent = [line.rstrip() for line in fp]
-
-        for keypress in fileContent:
-            data = keypress.split(":")
[…]

### 2022-06-08 application/backend/algorithm/common.py

@@ -1,0 +1,109 @@
+#!/usr/bin/env python3
+import numpy as np
+from cv2 import cv2 as cv
+
+
+DEBUG = False
+
+
+def findContours(frame: np.ndarray) -> tuple[np.ndarray]:
+    contours, _ = cv.findContours(frame, cv.RETR_EXTERNAL, cv.CHAIN_APPROX_SIMPLE)
+
+    if DEBUG:
+        contourFrame = cv.cvtColor(frame, cv.COLOR_GRAY2BGR)
+        cv.drawContours(contourFrame, contours, -1, (0, 255, 0), 2)
+        cv.imshow("debug", contourFrame)
+        cv.waitKey(0)
+
+    return contours
+
+
+def approximateLargestContour(contours: tuple[np.ndarray]) -> np.ndarray:
+    # Sort contours by area
+    largestContour = sorted(contours, key=cv.contourArea, reverse=True)[0]
+
+    # Use Douglas-Peucker algorithm to approximate a shape
+    # Set accuracy using the epsilon. 1% accuracy as the simplest shape, a
+    # rectangle, is expected
+    epsilon = 0.01 * cv.arcLength(largestContour, True)
+    approximation = cv.approxPolyDP(largestContour, epsilon, True)
+
+    return approximation
+
+
+def sortContourPoints(contour: np.ndarray) -> np.ndarray:
+    # Sort by x axis
+    indices = contour[:, 0].argsort()
+    contour = contour[indices]
+
+    # Get the two left points and sort by y axis
+    left = contour[:2]
+    leftIndices = left[:, 1].argsort()
+    left = left[leftIndices]
+
+    # Get the two right points and sort by y axis
+    right = contour[2:]
+    rightIndices = right[:, 1].argsort()
+    right = right[rightIndices]
+
+    # arrange clockwise
+    return np.array([left[0], right[0], right[1], left[1]])
+
+
+def scaleContours(contours: np.ndarray) -> np.ndarray:
+    scaleFactor = 7
+
+    upperLeft = [contours[0][0] - scaleFactor, contours[0][1] - scaleFactor]
+    upperRight = [contours[1][0] + scaleFactor, contours[1][1] - scaleFactor]
+    lowerRight = [contours[2][0] + scaleFactor, contours[2][1] + scaleFactor]
+    lowerLeft = [contours[3][0] - scaleFactor, contours[3][1] + scaleFactor]
+
[…]

### 2022-06-08 application/backend/algorithm/finger_detection.py

@@ -1,0 +1,128 @@
+#!/usr/bin/env python3
+
+from cv2 import cv2 as cv
+import mediapipe as mp
+import numpy as np
+
+from common import DEBUG, DEBUG_displayContours, convertResultToStr
+
+# Heavily lifted from https://google.github.io/mediapipe/getting_started/python.html.
+# Cited as (Lugaresi et al., n.d.) in the paper
+def classifyPressedFinger(image: np.ndarray, roiPoints: np.ndarray) -> list[str]:
+
+    mp_drawing = mp.solutions.drawing_utils
+    mp_drawing_styles = mp.solutions.drawing_styles
+    mp_hands = mp.solutions.hands
+
+    image_height, image_width, _ = image.shape
+
+    with mp_hands.Hands(
+        static_image_mode=True, max_num_hands=2, min_detection_confidence=0.5
+    ) as hands:
+        results = hands.process(cv.cvtColor(image, cv.COLOR_BGR2RGB))
+
+        if not results.multi_hand_landmarks:
+            raise Exception("finger_detection: no landmarks found")
+
+        # Get all possible landmarks that fit witihin the ROI. This is because
+        # some landmarks such as INDEX_FINGER_PIP may fall within the ROI, and
+        # it is highly unlikely that someone will click using that part of the
+        # finger. However, if no other matches are to be found, this is taken as
+        # the fallback
+        possibleMatches: list[tuple[str, str]] = []
+
+        for handIndex, hand_landmarks in enumerate(results.multi_hand_landmarks):
+            handedness = results.multi_handedness[handIndex].classification[0].label
+
+            # Flip handedness. Mediapipe expects the images to be flipped
+            # horizontally, much like taking a selfie. However, this wont work
+            # for our case
+            if handedness == "Left":
+                handedness = "RIGHT"
+            else:
+                handedness = "LEFT"
+
+            for landmarkIndex, landmark in enumerate(hand_landmarks.landmark):
+                x = round(landmark.x * image_width)
+                y = round(landmark.y * image_height)
+                landmarkName = mp_hands.HandLandmark(landmarkIndex).name
+
+                if DEBUG:
+                    print(handedness)
+                    print(landmarkName)
+                    print(x, y)
+                    print(landmark)
+
+                # The following if statements check if the landmark is within
+                # the ROI. If it fails, the next landmark is tested
+
+                # In the X axis, the landmark's coordinates is greater than one
+                # or both of the two coordinates found on the left side of the
+                # ROI
[…]

### 2022-06-08 application/backend/algorithm/integration.py

@@ -1,0 +1,54 @@
+#!/usr/bin/env python3
+
+from cv2 import cv2 as cv
+import numpy as np
+
+# import time
+
+from common import DEBUG, DEBUG_displayContours
+
+from keyboard_detection import getContourPointsFromImage
+from finger_detection import classifyPressedFinger
+
+
+def main(keyboardImage: np.ndarray, handImage: np.ndarray):
+
+    # start = time.time()
+    if DEBUG:
+        cv.namedWindow("debug")
+
+    # Step 1
+    allContourPoints = getContourPointsFromImage(keyboardImage)
+
+    contourPoints = allContourPoints["R"]
+
+    # elapsed = time.time() - start
+
+    # Step 2
+    if DEBUG:
+        cv.imshow("debug", handImage)
+        cv.waitKey(0)
+
+    result = classifyPressedFinger(handImage, contourPoints)
+
+    if result is not None:
+        print(result)
+
+    # total = time.time() - start
+    # classifying = total - elapsed
+
+    # print("Init Time: " + str(elapsed))
+    # print("Classifying Time: " + str(classifying))
+    # print("Total Time: " + str(total))
+
+    if DEBUG:
+        print(result)
+        cv.destroyAllWindows()
+
+    return result
+
+
+if __name__ == "__main__":
+    keyboardImage = cv.imread("images/fast-05-10.png")
+    handImage = cv.imread("images/fast-05-233.png")
+    main(keyboardImage, handImage)

### 2022-06-08 application/backend/algorithm/keyboard_detection.py

@@ -1,0 +1,209 @@
+from cv2 import cv2 as cv
+import numpy as np
+
+from copy import deepcopy
+
+from algorithm import keyboard_map, common
+
+
+def getKeyboardImage() -> np.ndarray | None:
+    cam = cv.VideoCapture(0)
+
+    while True:
+        ret, frame = cam.read()
+
+        if not ret:
+            return None
+
+        cv.imshow("keyboard", frame)
+
+        key = cv.waitKey(1)
+
+        if key % 256 == 155:
+            cv.destroyWindow("keyboard")
+            return frame
+
+
+def getEdgeFrame(frame: np.ndarray) -> np.ndarray:
+    grayscaleFrame = cv.cvtColor(frame, cv.COLOR_BGR2GRAY)
+
+    if common.DEBUG:
+        cv.imshow("debug", grayscaleFrame)
+        cv.waitKey(0)
+
+    # Denoise while maintaining edges
+    bilateralFilteredFrame = cv.bilateralFilter(grayscaleFrame, 11, 17, 17)
+
+    if common.DEBUG:
+        cv.imshow("debug", bilateralFilteredFrame)
+        cv.waitKey(0)
+
+    # Perform Sobel Filtering
+    sobelX = cv.Sobel(
+        src=bilateralFilteredFrame,
+        ddepth=cv.CV_64F,
+        dx=1,
+        dy=0,
+        ksize=3,
+        borderType=cv.BORDER_DEFAULT,
+    )
+
+    sobelY = cv.Sobel(
+        src=bilateralFilteredFrame,
+        ddepth=cv.CV_64F,
+        dx=0,
+        dy=1,
+        ksize=3,
+        borderType=cv.BORDER_DEFAULT,
+    )
+
+    absSobelX = cv.convertScaleAbs(sobelX)
+    absSobelY = cv.convertScaleAbs(sobelY)
+
+    finalSobel = cv.addWeighted(absSobelX, 0.5, absSobelY, 0.5, 0)
+
+    if common.DEBUG:
+        cv.imshow("debug", finalSobel)
+        cv.waitKey(0)
+
+    # Use Otsu's Threshold to get better defined edges while not using arbitrary
+    # values
+    _, threshedFrame = cv.threshold(
+        finalSobel, 0, 255, cv.THRESH_BINARY + cv.THRESH_OTSU
+    )
+
+    if common.DEBUG:
+        cv.imshow("debug", threshedFrame)
[…]

### 2022-06-08 application/backend/algorithm/keyboard_map.py

@@ -1,0 +1,66 @@
+#!/usr/bin/env python3
+
+
+keyDict = {
+    "LeftControl": [85, 0, 0],
+    "LeftSuper": [45, 0, 0],
+    "LeftAlt": [58, 35, 0],
+    "Spacebar": [62, 55, 0],
+    "RightAlt": [0, 28, 44],
+    "RightSuper": [0, 3, 73],
+    "RightControl": [24, 0, 76],
+    "`": [255, 0, 0],
+    "1": [255, 153, 0],
+    "2": [255, 229, 0],
+    "3": [189, 255, 0],
+    "4": [143, 255, 0],
+    "5": [51, 255, 0],
+    "6": [0, 255, 102],
+    "7": [0, 255, 194],
+    "8": [0, 224, 255],
+    "9": [0, 163, 255],
+    "0": [0, 102, 255],
+    "-": [0, 10, 255],
+    "=": [82, 0, 255],
+    "|": [128, 0, 255],
+    "Delete": [189, 0, 255],
+    "Tab": [224, 0, 0],
+    "Q": [214, 128, 0],
+    "W": [225, 203, 0],
+    "E": [160, 217, 0],
+    "R": [126, 224, 0],
+    "T": [43, 217, 0],
+    "Y": [0, 216, 86],
+    "U": [0, 195, 149],
+    "I": [0, 183, 208],
+    "O": [0, 136, 212],
+    "P": [0, 79, 196],
+    "[": [0, 8, 211],
+    "]": [63, 0, 196],
+    "Backspace": [134, 0, 181],
+    "CapsLock": [195, 0, 0],
+    "A": [167, 100, 0],
+    "S": [178, 161, 0],
+    "D": [124, 168, 0],
+    "F": [101, 180, 0],
+    "G": [31, 157, 0],
+    "H": [0, 139, 56],
+    "J": [1, 139, 106],
+    "K": [0, 130, 148],
+    "L": [0, 94, 147],
+    ";": [0, 52, 130],
+    "'": [0, 7, 170],
+    "Enter": [35, 0, 109],
+    "LeftShift": [150, 0, 0],
+    "Z": [104, 62, 0],
+    "X": [130, 117, 0],
+    "C": [77, 104, 0],
+    "V": [63, 113, 0],
+    "B": [18, 88, 0],
+    "N": [0, 66, 26],
+    "M": [0, 84, 64],
+    ",": [0, 75, 85],
[…]

### 2022-06-08 application/backend/algorithm/test-scripts/clean-data.py

@@ -1,0 +1,18 @@
+#!/usr/bin/env python3
+from os import walk
+
+
+def main():
+    (_, _, filenames) = next(walk("../dataset/training"))
+    files = [file for file in filenames if not "mp4" in file]
+
+    for file in files:
+        with open(f"../dataset/training/{file}", "r+") as f:
+            content = f.read().replace(" ", "").replace("//", "/")
+            f.seek(0)
+            f.write(content)
+            f.truncate()
+
+
+if __name__ == "__main__":
+    main()

### 2022-06-08 application/backend/algorithm/test-scripts/collect.py

@@ -1,0 +1,194 @@
+#!/usr/bin/env python3
+
+import sys
+import random
+from cv2 import cv2 as cv
+
+asciiDict = {
+    8: "Backspace",
+    9: "Tab",
+    13: "Enter",
+    27: "LeftControl",
+    32: "Spacebar",
+    33: "1",
+    34: "'",
+    35: "3",
+    36: "4",
+    37: "5",
+    38: "7",
+    39: "'",
+    40: "9",
+    41: "0",
+    42: "8",
+    43: "=",
+    44: ",",
+    45: "-",
+    46: ".",
+    47: "/",
+    48: "0",
+    49: "1",
+    50: "2",
+    51: "3",
+    52: "4",
+    53: "5",
+    54: "6",
+    55: "7",
+    56: "8",
+    57: "9",
+    58: ";",
+    59: ";",
+    60: ",",
+    61: "=",
+    62: ".",
+    63: "//",
+    64: "2",
+    65: "A",
+    66: "B",
+    67: "C",
+    68: "D",
+    69: "E",
+    70: "F",
+    71: "G",
+    72: "H",
+    73: "I",
+    74: "J",
+    75: "K",
+    76: "L",
+    77: "M",
+    78: "N",
+    79: "O",
+    80: "P",
+    81: "Q",
+    82: "R",
+    83: "S",
+    84: "T",
+    85: "U",
+    86: "V",
+    87: "W",
+    88: "X",
+    89: "Y",
+    90: "Z",
+    91: "[",
+    92: "\\",
+    93: "]",
+    94: "6",
+    95: "-",
+    96: "`",
+    97: "A",
+    98: "B",
+    99: "C",
+    100: "D",
+    101: "E",
+    102: "F",
+    103: "G",
+    104: "H",
+    105: "I",
+    106: "J",
+    107: "K",
+    108: "L",
+    109: "M",
+    110: "N",
+    111: "O",
+    112: "P",
+    113: "Q",
+    114: "R",
+    115: "S",
+    116: "T",
+    117: "U",
+    118: "V",
+    119: "W",
+    120: "X",
+    121: "Y",
+    122: "Z",
+    123: "[",
+    124: "\\",
+    125: "]",
+    126: "`",
+    225: "LeftShift",
+    226: "RightShift",
+    233: "LeftAlt",
+    234: "RightAlt",
+    235: "LeftSuper",
+    255: "Delete",
+}
+
+
[…]

### 2022-06-08 application/backend/algorithm/test-scripts/confirm-label.py

@@ -1,0 +1,106 @@
+#!/usr/bin/env python3
+from os import walk, path
+from cv2 import cv2 as cv
+
+
+keypressDict = {
+    97: "LEFT_PINKY",
+    115: "LEFT_RING",
+    100: "LEFT_MIDDLE",
+    102: "LEFT_INDEX",
+    106: "RIGHT_INDEX",
+    107: "RIGHT_MIDDLE",
+    108: "RIGHT_RING",
+    59: "RIGHT_PINKY",
+    233: "LEFT_THUMB",
+    32: "RIGHT_THUMB",
+    27: "ESCAPE",
+    72: "INVALID",
+}
+
+
+def main():
+    (_, _, filenames) = next(walk("../dataset/training"))
+    files = sorted(
+        [file for file in filenames if not "mp4" in file and not "labeled" in file]
+    )
+
+    for index, file in enumerate(files):
+        fileLocation = f"../dataset/training/{file}"
+        cap = cv.VideoCapture(fileLocation + ".mp4")
+        labeledFile = f"{fileLocation}-labeled"
+
+        newFileContent = []
+
+        with open(labeledFile) as fp:
+            fileContent = [line.rstrip() for line in fp]
+
+        for keypressIndex, keypress in enumerate(fileContent):
+            data = keypress.split(":")
+            cap.set(cv.CAP_PROP_POS_FRAMES, int(data[0]))
+            ret, frame = cap.read()
+
+            if not ret:
+                raise Exception(f"label: frame {data[0]} not found for {file}")
+
+            frame = cv.rotate(frame, cv.ROTATE_180)
+            cv.putText(
+                frame,
+                keypress,
+                (200, 150),
+                cv.FONT_HERSHEY_SIMPLEX,
+                0.5,
+                (0, 255, 255),
+                1,
+                cv.LINE_AA,
+            )
+
+            cv.putText(
+                frame,
+                f"kp: {keypressIndex + 1}/{len(fileContent)} | file: {index + 1}/{len(files)}",
+                (10, 465),
+                cv.FONT_HERSHEY_SIMPLEX,
+                0.5,
+                (0, 255, 255),
+                1,
+                cv.LINE_AA,
+            )
+
+            cv.imshow("debug", frame)
+
+            keypressString = ""
+
+            while True:
+                key = cv.waitKey(0)
+
+                if key % 256 == 27:
+                    cv.destroyAllWindows()
+                    return
+
+                if key % 256 == 13:
+                    keypressString = ""
+                    break
+
+                try:
[…]

### 2022-06-08 application/backend/algorithm/test-scripts/get-image.py

@@ -1,0 +1,32 @@
+#!/usr/bin/env python3
+
+from cv2 import cv2 as cv
+
+
+def main():
+    file = "fast-05"
+    keyboardFrame = 10
+    fingerFrame = 233
+    fileLocation = f"../dataset/training/{file}"
+
+    cap = cv.VideoCapture(f"{fileLocation}.mp4")
+
+    cap.set(cv.CAP_PROP_POS_FRAMES, keyboardFrame)
+    ret, frame = cap.read()
+
+    if not ret:
+        return
+
+    cv.imwrite(f"../images/{file}-{keyboardFrame}.png", frame)
+
+    cap.set(cv.CAP_PROP_POS_FRAMES, fingerFrame)
+    ret, frame = cap.read()
+
+    if not ret:
+        return
+
+    cv.imwrite(f"../images/{file}-{fingerFrame}.png", frame)
+
+
+if __name__ == "__main__":
+    main()

### 2022-06-08 application/backend/algorithm/test-scripts/label.py

@@ -1,0 +1,103 @@
+#!/usr/bin/env python3
+from os import walk, path
+from cv2 import cv2 as cv
+
+
+keypressDict = {
+    97: "LEFT_PINKY",
+    115: "LEFT_RING",
+    100: "LEFT_MIDDLE",
+    102: "LEFT_INDEX",
+    106: "RIGHT_INDEX",
+    107: "RIGHT_MIDDLE",
+    108: "RIGHT_RING",
+    59: "RIGHT_PINKY",
+    233: "LEFT_THUMB",
+    32: "RIGHT_THUMB",
+    27: "ESCAPE",
+    72: "INVALID",
+}
+
+
+def main():
+    (_, _, filenames) = next(walk("../dataset/test"))
+    files = sorted(
+        [file for file in filenames if not "mp4" in file and not "labeled" in file]
+    )
+
+    for index, file in enumerate(files):
+        fileLocation = f"../dataset/test/{file}"
+        cap = cv.VideoCapture(fileLocation + ".mp4")
+
+        newFileContent = []
+
+        if path.exists(f"{fileLocation}-labeled"):
+            continue
+
+        with open(fileLocation) as fp:
+            fileContent = [line.rstrip() for line in fp]
+
+        cap.set(cv.CAP_PROP_POS_FRAMES, 10)
+        ret, frame = cap.read()
+        frame = cv.rotate(frame, cv.ROTATE_180)
+
+        cv.imshow("setup", frame)
+
+        for keypressIndex, keypress in enumerate(fileContent):
+            data = keypress.split(":")
+            cap.set(cv.CAP_PROP_POS_FRAMES, int(data[0]))
+            ret, frame = cap.read()
+
+            if not ret:
+                raise Exception(f"label: frame {data[0]} not found for {file}")
+
+            frame = cv.rotate(frame, cv.ROTATE_180)
+            cv.putText(
+                frame,
+                keypress,
+                (250, 100),
+                cv.FONT_HERSHEY_SIMPLEX,
+                0.5,
+                (0, 255, 255),
+                1,
+                cv.LINE_AA,
+            )
+
+            cv.putText(
+                frame,
+                f"kp: {keypressIndex + 1}/{len(fileContent)} | file: {index + 1}/{len(files)}",
+                (10, 465),
+                cv.FONT_HERSHEY_SIMPLEX,
+                0.5,
+                (0, 255, 255),
+                1,
+                cv.LINE_AA,
+            )
+
[…]

### 2022-06-08 application/backend/algorithm/test.py

@@ -1,0 +1,146 @@
+#!/usr/bin/env python3
+
+from os import walk, mkdir
+from cv2 import cv2 as cv
+import time
+import csv
+
+from finger_detection import classifyPressedFinger
+from keyboard_detection import getContourPointsFromImage
+
+
+def main():
+    (_, _, filenames) = next(walk("./dataset/training"))
+
+    # Create output folder
+    uuid = f"training-{time.time()}"
+    folder = f"./results/{uuid}"
+    mkdir(folder)
+
+    files = sorted(
+        [file for file in filenames if not "mp4" in file and not "labeled" in file]
+    )
+
+    # Create log for entire run
+    logFileContent = []
+    resultFileContent = [
+        [
+            "file_name",
+            "frame",
+            "key",
+            "finger",
+            "is_correct",
+            "fingertips_detected",
+            "keyboard_detection_time",
+            "finger_classification_time",
+        ]
+    ]
+
+    for file in files:
+        # Output Files
+        logFileContent.append(f"INFO: Start {file}")
+
+        fileLocation = f"./dataset/training/{file}"
+        labeledFile = f"{fileLocation}-labeled"
+
+        cap = cv.VideoCapture(fileLocation + ".mp4")
+
+        keyboardDetectionTimeStart = time.time()
+
+        allContourPoints = {}
+
+        for i in [10, 15, 20]:
+            cap.set(cv.CAP_PROP_POS_FRAMES, i)
+            ret, frame = cap.read()
+            if not ret:
+                logFileContent.append(f"ERROR: Failed to get frame {i} for {file}")
+                break
+
+            try:
+                keyboardDetectionTimeStart = time.time()
+                allContourPoints = getContourPointsFromImage(frame)
+            except Exception as e:
+                logFileContent.append(
+                    f"ERROR: Failed to get contour points for {file} with frame {i}"
+                )
+                logFileContent.append(f"ERROR: {e}")
+                cv.imwrite(f"{folder}/{file}:{i}.png", frame)
+            else:
+                break
+
+        if len(allContourPoints) == 0:
+            logFileContent.append(f"ERROR: Failed to get contour points for {file}")
+            continue
+
+        keyboardDetectionTimeEnd = time.time() - keyboardDetectionTimeStart
+
+        with open(labeledFile) as fp:
+            fileContent = [line.rstrip() for line in fp]
+
+        for keypress in fileContent:
+            data = keypress.split(":")
[…]

### 2022-06-08 application/backend/app.py

@@ -1,0 +1,41 @@
+from flask import Flask, request
+from flask_cors import CORS
+
+from cv2 import cv2 as cv
+import numpy as np
+import base64
+
+from algorithm import keyboard_detection
+
+app = Flask(__name__)
+CORS(app)
+
+
+@app.route("/")
+def root():
+    return "<p>hello, world</p>"
+
+
+@app.route("/edge-coordinates", methods=["POST"])
+def getEdgeCoordinates():
+    jsonRequest = request.json
+
+    if jsonRequest is None:
+        return {"error", "No data received"}, 400
+
+    imageURI: str = jsonRequest["image"]
+
+    if imageURI is None:
+        return {"error", "No image data received"}, 400
+
+    imageBase64 = imageURI.split(",")[1]
+
+    imageArray = np.frombuffer(base64.b64decode(imageBase64), np.uint8)
+    keyboardImage = cv.imdecode(imageArray, cv.IMREAD_COLOR)
+
+    try:
+        edgeCoordinates = keyboard_detection.getContourPointsFromImage(keyboardImage)
+    except:
+        return {"error": "Failure in getting edge coordinates"}, 500
+
+    return {"coordinates": edgeCoordinates}

### 2022-06-08 application/frontend/src/app.d.ts

@@ -1,0 +1,13 @@
+/// <reference types="@sveltejs/kit" />
+
+// See https://kit.svelte.dev/docs/types#app
+// for information about these interfaces
+declare namespace TeeCubed {
+	type Coordinate = [number, number];
+	type Key = [Coordinate, Coordinate, Coordinate, Coordinate];
+	type Map = Record<string, Key>;
+	// interface Locals {}
+	// interface Platform {}
+	// interface Session {}
+	// interface Stuff {}
+}

### 2022-06-08 application/frontend/src/constants/keymap.ts

@@ -1,0 +1,60 @@
+export const keyMap = {
+	' ': 'Spacebar',
+	'?': '/',
+	'>': '.',
+	'<': ',',
+	':': ';',
+	'"': "'",
+	'{': '[',
+	'}': ']',
+	_: '-',
+	'+': '=',
+	'|': '\\',
+	'~': '`',
+	'!': '1',
+	'@': '2',
+	'#': '3',
+	$: '4',
+	'%': '5',
+	'^': '6',
+	'&': '7',
+	'*': '8',
+	'(': '9',
+	')': '0'
+};
+
+export const expectedKeys = {
+	Q: 'LEFT_PINKY',
+	A: 'LEFT_PINKY',
+	Z: 'LEFT_PINKY',
+	W: 'LEFT_RING',
+	S: 'LEFT_RING',
+	X: 'LEFT_RING',
+	E: 'LEFT_MIDDLE',
+	D: 'LEFT_MIDDLE',
+	C: 'LEFT_MIDDLE',
+	R: 'LEFT_INDEX',
+	F: 'LEFT_INDEX',
+	V: 'LEFT_INDEX',
+	T: 'LEFT_INDEX',
+	G: 'LEFT_INDEX',
+	B: 'LEFT_INDEX',
+	Y: 'RIGHT_INDEX',
+	H: 'RIGHT_INDEX',
+	N: 'RIGHT_INDEX',
+	U: 'RIGHT_INDEX',
+	J: 'RIGHT_INDEX',
+	M: 'RIGHT_INDEX',
+	I: 'RIGHT_MIDDLE',
+	K: 'RIGHT_MIDDLE',
+	',': 'RIGHT_MIDDLE',
+	O: 'RIGHT_RING',
+	L: 'RIGHT_RING',
+	'.': 'RIGHT_RING',
+	P: 'RIGHT_PINKY',
+	';': 'RIGHT_PINKY',
+	'/': 'RIGHT_PINKY',
+	"'": 'RIGHT_PINKY',
+	'[': 'RIGHT_PINKY',
+	']': 'RIGHT_PINKY'
+};

### 2022-06-08 application/frontend/src/constants/test-sequences.ts

@@ -1,0 +1,12 @@
+export const testSequences = [
+	'What of it, if some old hunks of a sea-captain orders me to get a broom and sweep down the decks?',
+	'I sat down on an old wooden settle, carved all over like a bench on the Battery.',
+	'I lay there dismally calculating that sixteen entire hours must elapse before I could hope for a resurrection.',
+	'How slowly the time passes here, encompassed as I am by frost and snow!',
+	'I listened to my father in silence and remained for some time incapable of offering any reply.',
+	"Think you're escaping and run into yourself. Longest way round is the shortest way home.",
+	'History, Stephen said, is a nightmare from which I am trying to awake',
+	'A man of genius makes no mistakes. His errors are volitional and are the portals of discovery.',
+	'Never trust to general impressions, my boy, but concentrate yourself upon details',
+	'I have no data yet. It is a capital mistake to theorise before one has data.'
+];

### 2022-06-08 application/frontend/src/store/edge-coordinates.ts

@@ -1,0 +1,4 @@
+import { writable } from 'svelte/store';
+
+export const photo = writable<string | null>(null);
+export const edgeCoordinates = writable<TeeCubed.Map | null>(null);

### 2022-06-08 application/frontend/src/store/mediapipe.ts

@@ -1,0 +1,7 @@
+import { writable } from 'svelte/store';
+
+import type { Hands } from '@mediapipe/hands';
+
+export const hands = writable<Hands | null>(null);
+export const video = writable<HTMLVideoElement | null>(null);
+export const fingersDetected = writable<string[]>([]);

### 2022-06-08 application/frontend/src/store/test-sequence.ts

@@ -1,0 +1,9 @@
+import { writable } from 'svelte/store';
+
+export const testSequence = writable<number>(Math.floor(Math.random() * 9));
+export const timeElapsed = writable<number>(0);
+export const totalCharactersTyped = writable<number>(0);
+export const inputData = writable<{ coordinates: TeeCubed.Key | null; key: string }>({
+	coordinates: null,
+	key: 'z'
+});

### 2022-06-08 application/frontend/svelte.config.js

@@ -1,0 +1,30 @@
+import path from 'path';
+
+import adapter from '@sveltejs/adapter-auto';
+import preprocess from 'svelte-preprocess';
+import autoprefixer from 'autoprefixer';
+
+/** @type {import('@sveltejs/kit').Config} */
+const config = {
+	// Consult https://github.com/sveltejs/svelte-preprocess
+	// for more information about preprocessors
+	preprocess: [
+		preprocess({
+			postcss: true,
+			scss: { includePaths: ['src', 'node_modules'] }
+		})
+	],
+
+	kit: {
+		adapter: adapter(),
+		vite: {
+			resolve: {
+				alias: {
+					src: path.resolve('./src')
+				}
+			}
+		}
+	}
+};
+
+export default config;

### 2022-06-08 application/frontend/src/store/mediapipe.ts
trainer: working finger-key identification tester

@@ -4,4 +4,12 @@ import type { Hands } from '@mediapipe/hands';
 
 export const hands = writable<Hands | null>(null);
 export const video = writable<HTMLVideoElement | null>(null);
-export const fingersDetected = writable<string[]>([]);
+export const result = writable<{
+	fingersDetected: string[];
+	key: string;
+	currentLetter: number;
+}>({
+	fingersDetected: ['LEFT_PINKY'],
+	key: 'Z',
+	currentLetter: -1
+});

### 2022-06-08 application/frontend/src/store/test-sequence.ts

@@ -3,7 +3,12 @@ import { writable } from 'svelte/store';
 export const testSequence = writable<number>(Math.floor(Math.random() * 9));
 export const timeElapsed = writable<number>(0);
 export const totalCharactersTyped = writable<number>(0);
-export const inputData = writable<{ coordinates: TeeCubed.Key | null; key: string }>({
+export const inputData = writable<{
+	coordinates: TeeCubed.Key | null;
+	key: string;
+	currentLetter: number;
+}>({
 	coordinates: null,
-	key: 'z'
+	key: 'Z',
+	currentLetter: 0
 });

### 2022-06-12 paper/paper.tex
paper: final manuscript

@@
 \usepackage[american]{babel}
-\usepackage[style=apa, citestyle=apa, backend=biber]{biblatex}
+\usepackage[style=apa, citestyle=apa, sortcites=true, backend=biber]{biblatex}
 \DeclareLanguageMapping{american}{american-apa}
@@
 \usepackage{pgfplots}
+\usepackage{pgfplotstable}
 \usepgfplotslibrary{external}
@@
 \begin{center}
-	\textbf{Development of a Finger-Key Identification Module\\for a Touch Typing Trainer}
+	\textbf{Development of a Finger-Key Identification Module\\for a Touch Typing Trainer}\\
+	by\\
+	Oscar Vian L. Valles
+
 	\par\vspace{0.5cm}
-	by
-	Oscar Vian L. Valles
-
-	\par\vspace{2cm}
 
@@
 
-Typing tests lists out words for the user to repeat. These tests measure metrics
+Typing tests list out words for the user to repeat. These tests measure metrics
 that quantify the user's ability to type. However, conventional typing metrics
@@
 such, a method to identify which finger is used to press which key is
-beneficial. This paper introduces a technique to do so by developing a
+beneficial. This paper introduces a technique to achieve this by developing a
 finger-key identification module that utilizes computer vision algorithms to
@@
 track fingers while typing. This module was successful in finger-key
-identification with an accuracy of 99.47\% in a dataset of 937 keypresses. The
+identification with an accuracy of 99.58\% in a dataset of 938 keypresses. The
 module also had an average finger-key identification time of 0.083 seconds,
@@
 typing speeds up to 143.030 Words per Minute which is well beyond the median
-typing speed of the population. Finally, two metrics for correct finger
-placement were defined. Finger Placement Accuracy which computes the number of
-keys pressed with the correct finger in a typing test sequence. Historical
-Finger Placement Accuracy which evaluates the user's ability to press a certain
-key with the correct finger over multiple typing test sequences.
+typing speed of the population. Furthermore, two metrics for correct finger
+placement were defined and were implemented in a proof-of-concept trainer: (1)
+Finger Placement Accuracy which computes the number of keys pressed using the
+correct finger in a typing test sequence, and (2) Historical Finger Placement
+Accuracy which evaluates the user's ability to press a certain key with the
+correct finger over multiple typing test sequences.
 
@@
 
-Furthermore, Touch typing using correct finger placement is frequently taught in
+Furthermore, touch typing using correct finger placement is frequently taught in
 the beginner level \parencite{donica2018}. This means that there is a need to
@@
 keyboard. A 60\% Keyboard in a slightly modified \ac{ansi} layout with
-light-colored keys as shown in Figure~\ref{fig:60}. This type of keyboard only
-has the alphanumeric portion of a full-size keyboard. This limited the number of
-keys to be checked and reduced the overall movement of the hand during typing.
-The keyboard layout was a slightly modified \ac{ansi} layout. The majority of
-the layout still followed the base \ac{ansi} layout, with the alphanumeric
-portion remaining exactly the same.
+light-colored keys was used as shown in Figure~\ref{fig:60}. This type of
+keyboard only has the alphanumeric portion of a full-size keyboard. This limited
+the number of keys to be checked and reduced the overall movement of the hand
+during typing. The keyboard layout was a slightly modified \ac{ansi} layout. The
+majority of the layout still followed the base \ac{ansi} layout, with the
+alphanumeric portion remaining exactly the same.
 
 The module's effect on typing was not tested. This includes the module's effect
-on typing education and typing health, as the focus of the research was on the
-development and testing of the finger-key identification module only.
-
-In addition, the module expected the user to have no deformities in their hands,
-and they were capable of typing on a keyboard with difficulties.
+on typing in education and typing in health, as the focus of the research was on
+the development and testing of the finger-key identification module only.
 
@@
 accountants. By having better ergonomics while typing --- by touch typing
-correctly --- wrist injuries can be prevented, and typing speed may be increased
+correctly --- wrist injuries can be prevented, and typing speed may be
+increased.
 
@@
 typists. By automatically checking for correct technique, the burden of checking
-each student is reduced, and directed interventions for bad habits can be
-easily created as students with these bad habits are easily identified
+each student is reduced, and directed interventions for bad habits can be easily
+created as students with these bad habits are easily identified.
 
@@
 
-Median typing speed using a keyboard with QWERTY layout, based on test data
-from \cite{keybr}, is $\approx37.5$ \ac{wpm}. This median is taken from all the test results they have gathered from all of their users.
-
-\cite{monkey-stats} released a dataset showing the speed of 12678 members of the
+Median typing speed using a keyboard with QWERTY layout, based on test data from
+\textcite{keybr}, is $\approx37.5$ \ac{wpm}. This median is taken from all the test
+results they have gathered from all of their users.
+
+\textcite{monkey-stats} released a dataset showing the speed of 12678 members of the
 Monkeytype community with the speed role. The dataset shows the fastest time
@@
 
-\cite{arif2009} also performed a survey with 12 participants using
+\textcite{arif2009} also performed a survey with 12 participants using
 multiple typing methods. They got an average \ac{wpm} of 75.85 with a standard
@@
 These include wrist extension, ulnar deviation, and pronation, which may cause
-musculoskeletal disorders of the upper extremity (\cite{marklin1999} as cited in
-~\cite{baker2007})
+musculoskeletal disorders of the upper extremity (\textcite{marklin1999} as cited in
+~\textcite{baker2007})
 
@@
 
-OpenCV is an open-source computer vision and machine learning software library
-that houses ${\approx2500}$ optimized algorithms. This library is widely used by
-companies, researchers, and open source communities that utilize computer vision
-and machine learning in their projects. Examples of companies that use OpenCV
-include Google, Sony, and Honda.
+OpenCV is an open-source library that provides ${\approx2500}$ optimized
+algorithms for computer vision and machine learning. This library is widely used
+by companies, researchers, and open source communities that utilize computer
+vision and machine learning in their projects. Examples of companies that use
+OpenCV include Google, Sony, and Honda.
 
@@
 
+Examples of projects that utilized MediaPipe and its solutions include a
+real-time hand gesture recognition system by \textcite{mediapipe-sample-1} which
+utilized MediaPipe Hands. Another example is by \textcite{mediapipe-sample-2}
+which utilized MediaPipe Hands to track movements during pen spinning
+competitions. \textcite{mediapipe-sample-3} developed a real-time sign language
+recognition program using MediaPipe Hands.
+
 \subsubsection{MATLAB}
@@
 
-\cite{dorf2001} present a use case for finger tracking in augmented
+\textcite{dorf2001} present a use case for finger tracking in augmented
 environments. In the paper, interaction in a virtual environment through the use
@@
 
-\cite{chiang2014} used a Kinect, a 3D sensing device by Microsoft that
+\textcite{chiang2014} used a Kinect, a 3D sensing device by Microsoft that
 uses depth data, to track fingers to play virtual instruments. Virtual Pianos
@@
 
-\cite{yousaf2014} created a virtual keyboard that operates using finger
+\textcite{yousaf2014} created a virtual keyboard that operates using finger
 tracking. The tracking uses the movement of the finger joints as the basis for
-selecting which key to press. A camera captures the movement, and the resulting
-video stream is used for hand region detection and finger joint localization.
-Using probabilistic regional density-based kernel tracking, finger joint
-trajectories are gathered. Feature vectors are then interpreted from the
-trajectories. These feature vectors are used in logic-based techniques and
-Dynamic Bayesian Network for classification, detection, and recognition of
-keystrokes.
+selecting which key to press. The tracking first detects where the hands are.
+The result of the detection is then piped to a different algorithm for finger
+joint localization. These are then used to generate finger joint trajectories
+using kernel tracking. A Dynamic Bayesian Network then uses feature vectors from
+these trajectories to classify, detect, and recognize keystrokes.
 
@@
 
+This section provides an overview of the development and testing of a real-time
+finger-key identification module. First, the experimental setup used for data
+collection will be presented. Following that, the algorithm used to implement
+real-time finger-key identification will be discussed. The process of training
+and testing of the developed algorithm comes next. Afterwards, the development
+and design of the proof of concept trainer will be described. Finally, this
+section introduces two metrics that were developed to quantify correct finger
+placement during typing tests.
+
 \section{Setup}
@@
 The experimental setup and configuration were composed of three elements: the
-camera, the keyboard, and the environment. Figure~\ref{fig:metho-setup} is an
-image of the experimental setup
+camera, the keyboard, and the environment. This setup was used to collect data
+points to measure the algorithm's ability to perform finger-key identification
+in a controlled environment. Figure~\ref{fig:metho-setup} is an image of the
+experimental setup.
 
@@
 The setup used a single monocular camera positioned in a top-down view. The
-camera captured the entirety of the keyboard and the movement of the ten
+camera captures the entirety of the keyboard and the movement of the ten
 fingers. To do so, it was mounted on top of the monitor with the camera pointed
@@
 is capable of capturing color video in 1080p/30fps and 720p/30fps with a
-diagonal field of view of 78\degree. This camera has a universal mounting clip
-that allows the camera to be correctly positioned within the experimental setup
+diagonal field of view of 78\degree. For the setup, the camera captured the
+video in 480p/30fps. This camera has a universal mounting clip that allows the
+camera to be correctly positioned within the experimental setup
 \parencite{logitech}.
@@
 In addition, the keyboard layout was a modified \ac{ansi} layout. The majority
-of the keys remained the same, however some keys were changed: The pipe key was
+of the keys remained the same, however some keys were changed: the pipe key was
 replaced by the backspace key, the backspace key was replaced by the pipe key
-and a delete key, and the bottom row has a 7u length spacebar. Overall, these
-are minor changes that does not affect the alphanumeric keys that are
-predominantly used in typing.
+and a delete key, and the bottom row has a longer spacebar. Overall, these are
+minor changes that does not affect the alphanumeric keys that are predominantly
+used in typing.
 
 \subsection{Computer}
-The computer used to run the algorithm is a desktop computer with the following
-specification:
+The computer used to run the algorithm is a desktop computer. The specifications
+of this computer is shown in Table~\ref{tab:table-specs}.
 
@@
 	\centering
+	\caption{\label{tab:table-specs}Computer Specifications}
 	\begin{tabular}{ p{0.3\textwidth} p{0.6\textwidth} }
@@
 	\end{tabular}
-	\caption{\label{tab:table-specs}Computer Specifications}
 \end{table}
@@
 the keyboard and the fingers used for typing. The light source used was a common
-LED bulb rated at 9 Watts with 700 lumens. This light was white with a color
+LED bulb rated at 9 watts with 700 lumens. This light is white with a color
 temperature of 6500k.
@@
 The algorithm used a Sobel filter for edge detection. This filter uses a
-3$\times$3 kernel that is convolved twice. Once horizontally, and another
-vertically to produce a grayscale image of the outlines within the frame. The
-kernels used by the Sobel filter \parencite{sobel2014} is shown in
-Figure~\ref{fig:metho-algo-key-sobel} and Figure~\ref{fig:metho-algo-sobel} is
-the output of this step.
+3$\times$3 kernel that is convolved twice --- once horizontally, and another
+vertically --- to produce a grayscale image of the outlines within the frame.
+The kernels used by the Sobel filter \parencite{sobel2014} is shown in
+Figure~\ref{fig:metho-algo-key-sobel} and the output of this step is shown in
+Figure~\ref{fig:metho-algo-sobel}.
 
@@
 \paragraph{Thresholding}
-The output of the Sobel filter is a grayscale image of 255 values, with each
+The output of the Sobel filter is a grayscale image of 256 values, with each
 gray pixel indicating edges within the image. There is a need to reduce the
@@
 redundant points and returned the least amount of points that describe the
-shape. This OpenCV function implements the algorithm of~\cite{contours} in their
+shape. This OpenCV function implements the algorithm of~\textcite{contours} in their
 paper \citetitle{contours}. Figure~\ref{fig:metho-algo-contours} shows the contours
@@
 The contour points that described the contour were then sorted clockwise. This
-was a necessity since the algorithm of~\cite{contours} does not guarantee that
-the largest contour's points were returned in a clockwise arrangement, which is
-a requirement for the next step. Figure~\ref{fig:metho-algo-largest-contour}
-shows the four contour points found at the edges of the keyboard within the
-image.
+was a necessity since the algorithm of~\textcite{contours} does not guarantee
+that the largest contour's points were returned in a clockwise arrangement.
+There is a need to have consistent ordering of these points since the
+transformation that will be performed in the next step pairs the contour points
+to a hard-coded array that indicates the edges of the image map. This hard-coded
+array was sorted clockwise. As such, if the contour points were not sorted
+clockwise, there will be no guarantee that the pairing would result in a
+transformation that creates a transformed image map that is oriented correctly,
+since the pairing may not pair the correct points.
+
+Figure~\ref{fig:metho-algo-largest-contour} shows the four contour points found
+at the edges of the keyboard within the image.
 
@@
 color assigned to it and this color fills the region where this key is located
-at. Figure~\ref{fig:metho-algo-key-map} is the initial image map and Appendix
-Chapter~\ref{appendix:key-map} is a table that shows the hex color code and key
-mapping present in the image map.
+at. Figure~\ref{fig:metho-algo-key-map} is the initial image map and
+Appendix~\ref{appendix:key-map} is a table that lays out the mapping of the
+color code in BGR format and its associated character in the image map.
 
@@
 buffer was needed as, in some instances, the finger used to press the key was
-not exactly on top of the key, but rather to its side. The algorithm accounts
-for this situation by adding a buffer. This was achieved by scaling the contour
-points by seven pixels on each side. Figure~\ref{fig:metho-algo-scale-roi} shows
-the increased buffer in relation to the key. This value was obtained after
-testing other possible values. A buffer of five pixels did not have a lot of
-effect in reducing the number of failures. A buffer of ten pixels did reduce the
-number of failures, but it also greatly increased the algorithm's uncertainty in
-determining the finger --- uncertainty is defined as detecting two or more
-fingers within one \ac{roi}. Seven was a good middle ground in decreasing
-failures, without greatly increasing uncertainty.
+not exactly on top of the key, but rather to its side. In addition, it can also
+be observed that the finger tracking does not annotate the very edge of the
+fingertip, but rather, it labels somewhere in the middle of the fingernail.
+
+The algorithm accounts for this situation by adding a buffer. This was achieved
+by scaling the contour points by seven pixels on each side.
+Figure~\ref{fig:metho-algo-scale-roi} shows the increased buffer in relation to
+the key. This value was obtained after testing other possible values. A buffer
+of five pixels did not have a lot of effect in reducing the number of failures.
+A buffer of ten pixels did reduce the number of failures, but it also greatly
+increased the algorithm's uncertainty in determining the finger --- uncertainty
+is defined as detecting two or more fingers within one \ac{roi}. Seven was a
+good middle ground in decreasing failures, without greatly increasing
+uncertainty.
 
@@
 	\includegraphics[width=0.8\textwidth]{scaled-contour.png}
-	\caption{Contour points of the \ac{roi}}
+	\caption{Scaled contour points of the \ac{roi} with the buffer of seven pixels}
 	\label{fig:metho-algo-scale-roi}
@@
 The first model, the Palm Detection Model detects the initial hand locations
-using a single shot detector model based on the paper by~\cite{ssd}. This model
+using a single shot detector model based on the paper by~\textcite{ssd}. This model
 achieves an average precision of 95.7\% in palm detection
@@
 implemented as a separate module in Python using OpenCV. The Finger Detection
-and Tracking Solution in Section~\ref{section:metho-algo-finger} was consumed
+and Tracking Solution in Section~\ref{section:metho-algo-finger} was implemented
 using the prebuilt Python package offered by the MediaPipe team. Finally, the
-integration of the two will be done as another separate module in Python.
+integration of the two was done as another separate module in Python.
 
@@
 sequence is ``What of it, if some old hunks of a sea-captain orders me to get a
-broom and sweep down the decks?'' from Moby Dick by~\cite{moby-dick}. The 9
+broom and sweep down the decks?'' from Moby Dick by~\textcite{moby-dick}. The 9
 other test sequences can be found in Appendix
@@
 \subsubsection{Video Capture}
-There were 3 groups of 10 videos that were captured by the researcher, for a
-total of 30 videos. Each group corresponds to three predetermined speeds of
-typing: slow (15wpm), average (35wpm), and fast (80wpm). These speeds were based
-on test data from~\cite{keybr}. The researcher typed the 10 typing test
-sequences at these predetermined speeds. During typing, errors were not
-consciously taken into account, and errors happened naturally as part of the
-typing process.
+There were 3 sets of 10 videos that were captured by the researcher, for a total
+of 30 videos. Each set corresponds to three predetermined speeds of typing: slow
+(15wpm), average (35wpm), and fast (80wpm). These speeds were based on test data
+from~\textcite{keybr}. The researcher typed the 10 typing test sequences at
+these predetermined speeds. Best effort was made to stick to the predetermined
+speeds. During typing, errors were not consciously taken into account, and
+errors happened naturally as part of the typing process --- with more errors
+happening more frequently as the typing speed increased.
 
@@
 \subsubsection{Keypress Labeling}
-The researcher then manually perform finger-key identification for all key
+The researcher then manually performed finger-key identification for all key
 presses present in the video. This was also done through a Python script. The
@@
 
+\subsubsection{Data Inventory}
+
+\begin{figure}[H]
+	\centering
+	\begin{tikzpicture}
+		\begin{axis}[
+				xbar, xmin=0,
+				bar width=4pt,
+				width=0.8\textwidth,
+				ylabel=Key,
+				xlabel=Count,
+				symbolic y coords={
+						J,
+						Q,
+						Dash,
+						ForwardSlash,
+						Z,
+						1,
+						LeftAlt,
+						X,
+						Comma,
+						K,
+						V,
+						G,
+						B,
+						Backspace,
+						W,
+						Period,
+						P,
+						U,
+						C,
+						LeftShift,
+						F,
+						Y,
+						M,
+						H,
+						D,
+						L,
+						I,
+						R,
+						N,
+						T,
+						S,
+						A,
+						O,
+						E,
+						Spacebar,
+					},
+				ytick=data,
+				tick label style={font=\footnotesize},
+				label style={font=\footnotesize},
+				xmajorgrids=true
+			]
+			\addplot table[x=Value,y=Key]{training-data.txt};
+		\end{axis}
+	\end{tikzpicture}
+	\caption{Training Dataset Inventory}
+	\label{fig:metho-inventory-training}
+\end{figure}
+
+Individual counts of the characters in the training iteration data set is shown
+in Figure~\ref{fig:metho-inventory-training}. The letters were not represented
+equally in the data set. Of note, J and Q were not in the dataset.
+
+\begin{figure}[H]
+	\centering
+	\begin{tikzpicture}
+		\begin{axis}[
+				xbar, xmin=0,
+				bar width=4pt,
+				width=0.8\textwidth,
+				ylabel=Key,
+				xlabel=Count,
+				symbolic y coords={
+						J,
+						Q,
+						Z,
+						Dash,
+						ForwardSlash,
+						X,
+						V,
+						Period,
+						K,
+						B,
+						Comma,
+						G,
+						F,
+						Backspace,
+						P,
+						W,
+						C,
+						U,
+						M,
+						Y,
+						D,
+						LeftShift,
+						L,
+						H,
+						N,
+						R,
+						S,
+						I,
+						O,
+						A,
+						T,
+						E,
+						Spacebar,
+					},
+				ytick=data,
+				tick label style={font=\footnotesize},
+				label style={font=\footnotesize},
+				xmajorgrids=true
+			]
+			\addplot table[x=Value,y=Key]{test-data.txt};
+		\end{axis}
+	\end{tikzpicture}
+	\caption{Test Dataset Inventory}
+	\label{fig:metho-inventory-test}
+\end{figure}
+
+The letters were also not represented equally in the test dataset as shown in
+Figure~\ref{fig:metho-inventory-test}. Three characters in the alphabet were not
+represented in the dataset: J, Q, and Z. The inequality of representation in
+both of these datasets can be attributed to the source materials of the typing
+test sequences --- novels and texts. While there were criteria to select which
+text was sourced, there was no conscious decision to equally represent all
+characters in the data set.
+
 \subsection{Training}
@@
 	\centering
+	\caption{\label{tab:csv-output}CSV File Contents}
 	\begin{tabular}{ p{0.3\textwidth} p{0.6\textwidth} }
@@
 	\end{tabular}
-	\caption{\label{tab:csv-output}CSV File Contents}
 \end{table}
@@
 Adjustments were then made to the algorithm after each training pass to improve
-its performance. Some runs had improvements in its statistics without doing any
-adjustments. These were runs that showed errors in the manual labeling and
+its performance. Some runs had improvements in their statistics without doing
+any adjustments. These were runs that showed errors in the manual labeling and
 were corrected afterward.
@@
 The resulting \texttt{.csv} file was imported into Google Sheets and key
-statistics were obtained based on the data. These statistics can be found on
-Table~\ref{tab:acquired-stats}. These key statistics informed what to adjust in
-the algorithm and its overall performance.
+statistics were obtained based on the data. Appendix~\ref{appendix:sheets} shows
+the screenshots where this was performed. The statistics obtained can be found
+in Table~\ref{tab:acquired-stats}. These key statistics informed what to adjust
+in the algorithm and its overall performance.
 
@@
 	\centering
+	\caption{\label{tab:acquired-stats}Acquired statistics}
 	\begin{tabular}{ p{0.3\textwidth} p{0.6\textwidth} }
@@
 	\end{tabular}
-	\caption{\label{tab:acquired-stats}Acquired statistics}
 \end{table}
@@
 
-%% \section{Trainer}
-
-%% \subsection{UX/UI design and Front-end development for the touch typing trainer}
-%% \subsubsection{Design}
-%% This design will be based on other type tests such as Monkeytype
-%% \parencite{bartnik2021}, and Keybr \parencite{keybr}, with additional components
-%% for showing real-time finger-key identification and mapping.
-%% Figure~\ref{fig:ui-test} and~\ref{fig:ui-stat} illustrates the two main pages of
-%% the design. Additional screens, such as historical statistics, in-platform
-%% tutorial, and a written guide are shown in the appendix.
-
-%% \begin{figure}[H]
-%% 	\centering
-%% 	\includegraphics[width=0.8\textwidth]{ui-test.png}
-%% 	\caption{Test screen of the front-end design}
-%% 	\label{fig:ui-test}
-%% 	\centering
-%% \end{figure}
-
-%% Figure~\ref{fig:ui-test} has four main components. From the top, the first
-%% component is the navigation bar. This directs the user to the different pages
-%% within the website. The second component in the upper left displays the number
-%% of remaining words to be typed. This is 16 in the example. Below that is the
-%% test sequence. This shows the user what characters to type, and highlights
-%% mistakes in two distinct colors. Red if the character pressed was incorrect and
-%% purple if the character typed was correct, but the finger used to type it is
-%% wrong. The last main component in this screen is the keyboard that illustrates
-%% the correct finger positioning, with each finger corresponding to a different
-%% color.
-
-%% \begin{figure}[H]
-%% 	\centering
-%% 	\includegraphics[width=0.8\textwidth]{ui-stats.png}
-%% 	\caption{Statistics screen of the front-end design}
-%% 	\label{fig:ui-stat}
-%% 	\centering
-%% \end{figure}
-
-%% Figure~\ref{fig:ui-stat} has two main components. The first component at the top
-%% shows the three major statistics for the typing test: Words per minute,
-%% Accuracy, and Finger Placement Accuracy. Finger Placement Accuracy will be
-%% explained in Section~\ref{section:metric}. The component below the statistics is a
-%% graph plotting the major statistics of the user in time. This includes when the
-%% user made the mistake and when the user slowed down, or increased their speed.
-
-%% \subsubsection{Implementation}
-%% The design will be created in Figma and the front-end will be web based. The
-%% framework the front-end will be developed in will be Svelte, due to its
-%% community resources, performance, and the researcher's familiarity with the
-%% framework. The front-end will serve as a display for the information and the
-%% interface which the user interacts and inputs information into the trainer. No
-%% calculation will be done within the front-end.
-
-%% \subsection{Back-end development of a touch typing trainer using the program created
-%% 	in~\ref{section:metho-algo} and the experimental setup identified in~\ref{section:metho-setup}}
-
-%% The back-end will handle all the computation for the trainer---mainly
-%% finger-key identification. The backend will also handle authentication and data
-%% storage.
-
-%% The back-end will expose a REST API.\@This is a type of application programming
-%% interface (API) that conforms to the REST architectural style. One key
-%% characteristic of this style is its statelessness and cacheable data
-%% \parencite{rest}. This will be implemented using Django, a Python REST API
-%% framework.
-
-%% \subsubsection{Finger-key Identification}
-%% The module that integrates all the algorithms will be called within the
-%% framework. The framework will then respond to requests for finger-key
-%% identification using this module. This will allow the back-end to expose its
-%% finger-key identification abilities and allow the front-end to use this data for
-%% the trainer.
-
-%% \subsubsection{Authentication}
-%% The core authentication system will rely on \ac{jwt}. According to~\cite{jwt},
-%% ``JSON Web Token (JWT) is a compact, URL-safe means of representing claims to be
-%% transferred between two parties.'' This allows for a user to claim that they are
-%% that specific user securely as the claim is cryptographically signed. For the
-%% backend of the trainer, the \ac{jwt} will be signed using a secret, and it will
-%% be encoded in HMAC SHA256.
-
-%% \subsubsection{Data Storage}
-%% The trainer will store all the data in a relational database using
-%% PostgreSQL.\@This specific technology was chosen due to its maturity, community,
-%% and industry recommendations and the researchers' familiarity with the
-%% technology.
+\section{Trainer}
+A trainer was developed as a proof of concept of a real-world implementation of
+the finger-key identification module.
+
+\subsection{Design}
+The design was heavily based on Monkeytype by \textcite{bartnik2021}. Additional
+components were added to enable real-time finger-key identification. There are
+three main screens:
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1\textwidth]{frontend-edge-coords.png}
+	\caption{Frontend Screen for obtaining Key-Edge Coordinates Map}
+	\label{fig:metho-front-coords}
+	\centering
+\end{figure}
+
+Figure~\ref{fig:metho-front-coords} shows the first screen the user sees. This page
+captured the keyboard and passed it to the backend to obtain the Key-Edge
+Coordinates Map.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1\textwidth]{frontend-test.png}
+	\caption{Testing page of the trainer}
+	\label{fig:metho-front-test}
+	\centering
+\end{figure}
+
+Figure~\ref{fig:metho-front-test} is the screen where the user performs the
+typing test. On this screen, the typing test sequence was placed above. If a
+letter was typed correctly with the correct finger, the character was colored
+white. If the wrong character was typed, the character was colored red. If the
+correct character was typed, but the finger used was incorrect, the character
+was colored purple.
+
+At the bottom, the video captured by the webcam was shown on the left. Right
+beside it, important coordinates of the finger-key identification module are
+overlaid over the specific frame that was used as the input.
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=1\textwidth]{frontend-metrics.png}
+	\caption{Page presenting the metrics obtained from the trainer}
+	\label{fig:metho-front-metrics}
+	\centering
+\end{figure}
+
+The metrics gathered from the test sequences are shown on a page as illustrated
+in Figure~\ref{fig:metho-front-metrics}. This page presents metrics that will be
+discussed in Section~\ref{section:metric}. Of note, the metrics shown on this
+page that use multiple test sequences as an input uses tests that were obtained
+only during the time the page was opened. This proof of concept does not store
+the data of a user's previous test sequences that were performed from a
+different session.
+
+\subsubsection{Implementation}
+The frontend was developed in Svelte, and the backend was developed with Flask.
+The frontend handled capturing the video, presenting the layout, and capturing
+input events. Finger Tracking was done on the frontend through the use of
+MediaPipe Hand's JavaScript library. The frontend also handled checking what
+fingertip was inside the \ac{roi} of the key that was entered. The backend
+handled the generation of the Key-Edge Coordinates Map using the image passed
+from the frontend.
+
+Two of the key steps in the module: finger tracking and integration were
+offloaded in the frontend as the wall of latency that exists between the
+frontend and the backend would cause performance issues. While this issue also
+exists in obtaining the Key-Edge Coordinates Map, this wouldn't affect real-time
+finger-key identification since this is a preprocessing step and is not ran
+during the time the user performs the typing test.
 
@@
 certain key with the correct finger over all test sequences. The same criteria
-in determining accurate key presses This allows the user to easily verify which
+in determining accurate key presses. This allows the user to easily verify which
 keys need attention and training if there is a high frequency of error.
@@
 \begin{equation}
-	HFP ACC = \frac{|F_{char}|}{|C_{char}|} \cdot 100\%
+	HFP ACC_{char} = \frac{|F_{char}|}{|C_{char}|} \cdot 100\%
 \end{equation}
@@
 			coordinates {
-					(1, 88.47) (2, 88.95) (3, 98.51) (4, 99.55) (5, 98.47) (6, 99.17) (7, 99.49)
+					(1, 88.54) (2, 89.02) (3, 98.58) (4, 99.62) (5, 98.54) (6, 99.24) (7, 99.55)
 				};
@@
 each side. Iteration four's increase was due to fixing the image map. Certain
-keys that had incorrect color values assigned in the Key-Color Values Map, and
-some keys were of incorrect shape. Iterations five to seven were performed to
-test different values and shapes for scaling the \ac{roi}.
+keys had incorrect color values assigned in the Key-Color Values Map, and some
+keys were of incorrect shape. Iterations five to seven were performed to test
+different values and shapes for scaling the \ac{roi}.
 
@@
 			coordinates {
-					(1, 88.47) (2, 88.95) (3, 98.51) (4, 99.55) (5, 98.47) (6, 99.17) (7, 99.49)
+					(1, 88.54) (2, 89.02) (3, 99.58) (4, 99.62) (5, 98.54) (6, 99.24) (7, 99.55)
 				};
@@
 buffer size by half to five pixels. This did not affect the success rate that
-much, with only a reduction of 1.07\%. This also greatly decreased the uncertain
+much, with only a reduction of 1.08\%. This also greatly decreased the uncertain
 success percentage by 16.74\%. The sixth iteration met halfway between
 iterations four and five, by adding a buffer of seven pixels. This increased the
-success rate back to 99.17\% without greatly increasing the number of uncertain
+success rate back to 99.24\% without greatly increasing the number of uncertain
 successes. In comparison with the fourth iteration, the sixth iteration had its
@@
 by 10 pixels, while the points farther away were increased by five pixels only.
-This slightly increased the success rate by 0.32\%. However, the percentage of
+This slightly increased the success rate by 0.31\%. However, the percentage of
 uncertain successes also increased by 5.28\%.
@@
 Based on this information and from the gathered data as shown in
-Figure~\ref{fig:rd-training-failure-types}, This failure type is non-existent
+Figure~\ref{fig:rd-training-failure-types}, this failure type is non-existent
 for the module in real-life conditions.
@@
 The module was also successful in finger-key identification, even if both the
-key and finger was occluded. Figure~\ref{fig:rd-occluded} is one frame where
+key and finger were occluded. Figure~\ref{fig:rd-occluded} is one frame where
 this occurred. The module was successful in identifying that the Left Thumb was
@@
 	\centering
+	\caption{\label{tab:rd-accuracy}Accuracy Results of the Module with Test Data}
 	\begin{tabular}{ p{0.4\textwidth} R{0.2\textwidth} }
@@
 		\midrule
-		Successes                       & 937     \\
+		Successes                       & 938     \\
 		Failures                        & 4       \\
-		Success Percentage / Accuracy   & 99.47\% \\
+		Success Percentage / Accuracy   & 99.58\% \\
 		Failure Percentage              & 0.42\%  \\[0.25cm]
@@
 	\end{tabular}
-	\caption{\label{tab:rd-accuracy}Accuracy Results of the Module with Test Data}
 \end{table}
@@
 
-Based on the definition of \ac{wpm} from \cite{arif2009} as seen in
+Based on the definition of \ac{wpm} from \textcite{arif2009} as seen in
 Equation~\ref{equ:wpm}, it can be extrapolated that calculating \ac{cps} is:
@@
 the same dataset as shown in
-Figure~\ref{fig:rd-training-keyboard-detection-time}. However, The first
+Figure~\ref{fig:rd-training-keyboard-detection-time}. However, the first
 training iteration had issues with gathering the speed of the module, as the
@@
 	\centering
+	\caption{\label{tab:rd-speed}Speed Results of the Module with Test Data}
 	\begin{tabular}{ p{0.4\textwidth} R{0.2\textwidth} }
@@
 	\end{tabular}
-	\caption{\label{tab:rd-speed}Speed Results of the Module with Test Data}
 \end{table}
@@
 
-The average finger identification time that the module got, 0.083s was
-well below 0.32s. This can be attributed to the speed of the MediaPipe
-library, and the pre-calculation of the edge coordinates.
+The average finger identification time that the module got was 0.083s which is
+well below 0.32s. This can be attributed to the speed of the MediaPipe library,
+and the pre-calculation of the edge coordinates.
 
@@
 Sobel filter \parencite{sobel2014}, (2) Thresholding using Otsu's algorithm
-\parencite{otsu}, (3) Finding contours using the algorithm of \cite{contours},
-(4) Line Simplification using the Douglas-Peucker algorithm
+\parencite{otsu}, (3) Finding contours using the algorithm of
+\textcite{contours}, (4) Line Simplification using the Douglas-Peucker algorithm
 \parencite{douglas-peucker}, and (5) Perspective transform
 \parencite{opencv-image-transform}. Finger detection utilized a ready-made
-solution called MediaPipe Hands by \cite{mediapipe}. The two were combined to
-perform finger-key identification.
-
-The keyboard used was a 60\% keyboard in ANSI. It was light in color, and was
+solution called MediaPipe Hands by \textcite{mediapipe}. The two were combined
+to perform finger-key identification.
+
+The keyboard used was a 60\% keyboard in ANSI. It was light in color and was
 connected with a black USB-C cable. The surface of the desk was dark, and a LED
@@
 
-The module was accurate in 99.47\% of identifications in a data set composed of
+The module was accurate in 99.58\% of identifications in a data set composed of
 942 keypresses. However, 14.12\% of the identifications were uncertain
@@
 \ac{roi}. Of these, 6.48\% were non-spacebar uncertain successes. These are
-uncertain successes where that was not over the spacebar. This bears more weight
-since these keys were the smallest on the keyboard, and in most cases, only one
-fingertip could fit within the key. These uncertain successes were due to the
-addition of a buffer around each \ac{roi} to increase the success rate.
+uncertain successes over keys that are not the spacebar. This bears more weight
+since most of these keys are the smallest on the keyboard, and in most cases,
+only one fingertip could fit within the key. These uncertain successes were due
+to the addition of a buffer around each \ac{roi} to increase the success rate.
 
@@
 
-Two metrics were developed: (1) Finger Placement Accuracy which computes the
-percentage of keys pressed with the correct figure over the length of the test
-sequence, and (2) Historical Finger Placement Accuracy which computes the number
-of times the user is successful if pressing a key with the correct finger over
-all test sequences.
+Two metrics were developed and implemented in a proof of concept trainer: (1)
+Finger Placement Accuracy which computes the percentage of keys pressed with the
+correct figure over the length of the test sequence, and (2) Historical Finger
+Placement Accuracy which computes the number of times the user is successful if
+pressing a key with the correct finger over all test sequences.
 
@@
 \section{Trainer}
-The paper only presented a module for finger-key identification and did not
-include a trainer that utilized the module. This would allow the module to be
-used by actual users and allow for further testing to be done.
+A proof of concept trainer with real-time finger-key identification was
+developed, however this trainer is lacking in functionality to be an actual
+usable trainer that will be used for educational purposes.
+
+In addition, the wall of latency that exists during th Key-Edge Coordinates Map
+can be removed by doing this process on the browser. This can be done by redoing
+the process in C, and compiling that to WebAssembly. The MediaPipe team did the
+same thing for MediaPipe Hands, and performance, for this use case, was more or
+less equal.
+
+\section{Standardized Finger-Key Mapping}
+While there are mappings for the default ANSI and ISO layouts, these layouts
+were not official and were defacto standards. While the trainer uses this
+defacto standard, it wouldn't be difficult to adapt the trainer to use another
+standard for finger-key mapping.
+
+\section{Character Representation}
+The test and training data did not have equal representation of the keys within
+the keyboard. For the module to undergo more rigorous testing, it would be
+better to have a dataset that has equal representation of all of the characters
+in the keyboard.
 
@@
 automated computer vision process to detect the four edges.
+
+\section{ROI Buffer}
+The buffer was selected after trail and error in the training phase of the
+algorithm. However, this buffer may not be the best and adaptive techniques or a
+machine learning algorithm may select better values for the buffer.
 
@@
 
+\chapter{Screenshots of Data Analysis}
+\label{appendix:sheets}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{sheets-raw-data.png}
+	\caption{Raw data as seen on Google Sheets}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{sheets-uncertain.png}
+	\caption{Filtered raw data to show uncertain successes}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{sheets-failures.png}
+	\caption{Filtered raw data to show failures}
+\end{figure}
+
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\textwidth]{sheets-analysis.png}
+	\caption{Metrics calculated from the raw data}
+\end{figure}
+
 \chapter{Gantt Chart}

### 2022-06-12 application/frontend/src/store/test-sequence.ts

@@ -1,6 +1,6 @@
 import { writable } from 'svelte/store';
 
-export const testSequence = writable<number>(Math.floor(Math.random() * 9));
+export const testSequence = writable<number>(0);
 export const timeElapsed = writable<number>(0);
 export const totalCharactersTyped = writable<number>(0);
 export const inputData = writable<{

### 2022-06-14 paper/paper.tex
paper: final manuscript

@@
 \pagenumbering{roman}
+
+\addcontentsline{toc}{chapter}{Abstract}
+\begin{center}
+	\LARGE\textbf{Acknowledgment}
+\end{center}
+
+I would like to express my deepest appreciation for the guidance that my Special
+Project Adviser, Dhong Fhel K. Gom-os, has given me throughout the duration of
+the project. His technical knowledge has been a key factor to the success of the
+finger-key identification module. The paper would not have been possible without
+his unwavering support and commitment.
+
+I also greatly appreciate the advice that Prof. Aileen Joan O. Vicente and Ms.
+Sram J Isle L. Santillan have given me. The constructive criticism that I
+received was instrumental in giving me a clear direction on where I want to take
+this research.
+
+I am particularly grateful to Jack Bartnik, the developer of Monkeytype. The
+application that they developed was the main source of inspiration for this
+project. I would also like to thank Jack for their generosity in sharing
+statistics concerning the typing speed of the users of Monkeytype.
+
+I would like to acknowledge that this research was partially supported by the
+Department of Science and Technology --- Science Education Institute.
+
+I give my heartfelt thanks to Ana Janelle H. Ca\~{n}ete for her love and comfort
+in these trying times.
+
+To friends, family, and other significant people in my life, thank you for
+supporting me thus far.
+
+\newpage
 \addcontentsline{toc}{chapter}{Abstract}
@@
 track fingers while typing. This module was successful in finger-key
-identification with an accuracy of 99.58\% in a dataset of 938 keypresses. The
+identification with an accuracy of 99.58\% in a dataset of 942 keypresses. The
 module also had an average finger-key identification time of 0.083 seconds,
@@
 	\includegraphics[width=0.7\textwidth]{60.png}
-	\caption{A 60\% keyboard in \ac{ansi} layout.}
+	\caption{A 60\% keyboard in modified \ac{ansi} layout.}
 	\label{fig:60}
@@
 This research is beneficial for all users of physical keyboards. These include a
-majority of the population as there are a lot of professions that heavily rely
+majority of the population as there are a lot of professions that rely heavily
 on keyboards. Examples include developers, physicians, educators, and
@@
 utilized MediaPipe Hands. Another example is by \textcite{mediapipe-sample-2}
-which utilized MediaPipe Hands to track movements during pen spinning
+who utilized MediaPipe Hands to track movements during pen spinning
 competitions. \textcite{mediapipe-sample-3} developed a real-time sign language
@@
 vision, and video processing systems. By using the available algorithms in the
-toolbox, such as YOLOv2, and ACF, hand detection and gesture recognition is made
-possible in the platform \parencite{matlab}.
+toolbox, such as YOLOv2, and ACF, hand detection and gesture recognition are
+made possible in the platform \parencite{matlab}.
 
@@
 real-time finger-key identification will be discussed. The process of training
-and testing of the developed algorithm comes next. Afterwards, the development
-and design of the proof of concept trainer will be described. Finally, this
-section introduces two metrics that were developed to quantify correct finger
-placement during typing tests.
+and testing the developed algorithm comes next. Afterward, the development and
+design of the proof of concept trainer will be described. Finally, this section
+introduces two metrics that were developed to quantify correct finger placement
+during typing tests.
 
@@
 
-The specific camera used was a Logitech C920. It is a 3 mega pixel webcam that
+The specific camera used was a Logitech C920. It is a 3 megapixel webcam that
 is capable of capturing color video in 1080p/30fps and 720p/30fps with a
@@
 Equation~\ref{eq:cps}. This is well above the median of 37.5 \ac{wpm} on Keybr.
-The module can also perform finger-key identification for 90.99\% of the
-community members of Monkeytype with the speed role.
+The module can also perform real-time finger-key identification for 90.99\% of
+the community members of Monkeytype with the speed role.
 
@@
 \ac{wpm} on Keybr \parencite{keybr}. This also handles 90.99\% of the 12678
-members with the speed role on MonkeyType \parencite{monkey-stats}.
+members with the speed role on Monkeytype \parencite{monkey-stats}.
 

### 2022-06-14 application/frontend/src/store/metrics.ts

@@ -1,0 +1,44 @@
+import { writable } from 'svelte/store';
+
+const baseHistory = {
+	Q: 0,
+	A: 0,
+	Z: 0,
+	W: 0,
+	S: 0,
+	X: 0,
+	E: 0,
+	D: 0,
+	C: 0,
+	R: 0,
+	F: 0,
+	V: 0,
+	T: 0,
+	G: 0,
+	B: 0,
+	Y: 0,
+	H: 0,
+	N: 0,
+	U: 0,
+	J: 0,
+	M: 0,
+	I: 0,
+	K: 0,
+	',': 0,
+	O: 0,
+	L: 0,
+	'.': 0,
+	P: 0,
+	';': 0,
+	'/': 0,
+	"'": 0,
+	'[': 0,
+	']': 0,
+	Spacebar: 0
+};
+
+export const testSequenceEnteredCharacters = writable<number>(0);
+export const testSequenceWrongCharacters = writable<number>(0);
+export const testSequenceCharacters = writable<number>(0);
+export const keyPressesHistory = writable<Record<string, number>>({ ...baseHistory });
+export const wrongKeyPressesHistory = writable<Record<string, number>>({ ...baseHistory });

### 2022-06-23 journal-article/journal-article.tex
journal-article: finish journal article

--- 2022-06-12 paper/paper.tex
@@
-\documentclass{report}
-%%% LT-SKIP-BEGIN
-
-% Page Size
-\usepackage[letterpaper, portrait, margin=1in, left=1.5in]{geometry}
-
-% Spacing
-\usepackage{setspace}
-\setlength{\parskip}{\baselineskip}
-\doublespacing{}
-
-% Bibliography
-\usepackage[american]{babel}
-\usepackage[style=apa, citestyle=apa, sortcites=true, backend=biber]{biblatex}
-\DeclareLanguageMapping{american}{american-apa}
-\addbibresource{references.bib}
-\usepackage[babel,threshold=2]{csquotes}
-
-% Graphics
+\documentclass[preprint,12pt,authoryear]{elsarticle}
+
+\usepackage[utf8]{inputenc}
+
+% Bibliograpy
+\usepackage{natbib}
+\usepackage{url}
+
+% Acronyms
+\usepackage{acro}
+
+% Symbols
+\usepackage{gensymb}
+\usepackage{amsmath}
+
+% Positioning
+\usepackage{float}
+
+% Colors
+\usepackage[dvipsnames]{xcolor}
+
+% Flowchart
+\usepackage{tikz}
+\usetikzlibrary{shapes,positioning,calc}
+
+% Graphs
+\usepackage{pgfplots}
+\usepackage{pgfplotstable}
+\usepgfplotslibrary{external}
+\pgfplotsset{width=1\textwidth,height=0.7\textwidth,compat=1.9}
+
 \usepackage{graphicx}
@@
 
-% Positioning
-\usepackage{float}
-
-% Acronyms
-\usepackage{acro}
-
-\DeclareAcronym{iso}{
-	short=ISO,
-	long=International Organization for Standardization
-}
-
-\DeclareAcronym{ansi}{
-	short=ANSI,
-	long=American National Standards Institute
-}
-
-\DeclareAcronym{jis}{
-	short=JIS,
-	long=Japanese International Standards
-}
-
-\DeclareAcronym{wpm}{
-	short=WPM,
-	long=Words per Minute
-}
-
-\DeclareAcronym{neck}{
-	short=WRNULD,
-	long=Work-related neck and upper limb disorders
-}
-
-\DeclareAcronym{cv}{
-	short=CV,
-	long=Computer Vision
-}
-
-\DeclareAcronym{roi}{
-	short=ROI,
-	long=Region of Interest
-}
-
-\DeclareAcronym{cts}{
-	short=CTS,
-	long=Carpal Tunnel Syndrome
-}
-
-\DeclareAcronym{jwt}{
-	short=JWT,
-	long=JSON Web Tokens
-}
-
-\DeclareAcronym{fpacc}{
-	short=FP ACC,
-	long=Finger Placement Accuracy
-}
-
-\DeclareAcronym{hfpacc}{
-	short=HFP ACC,
-	long=Historical Finger Placement Accuracy
-}
-
-\DeclareAcronym{cps}{
-	short=CPS,
-	long=Characters per Second
-}
-
-% Chapter Formatting
-\usepackage{titlesec}
-\titleformat{\chapter}[display]
-{\centering\LARGE\normalfont\bfseries}
-{Chapter \thechapter}
-{0em}
-{\vspace{-1.5ex}}
-
-% Colors
-\usepackage[dvipsnames]{xcolor}
+\usepackage{subcaption}
 
@@
 
-% Misc
-\usepackage{gensymb}
-\usepackage{amsmath}
-
-% Flowchart
-\usepackage{tikz}
-\usetikzlibrary{shapes,positioning,calc}
-
-% Graphs
-\usepackage{pgfplots}
-\usepackage{pgfplotstable}
-\usepgfplotslibrary{external}
-\pgfplotsset{width=1\textwidth,compat=1.9}
-
 % Tables
 \usepackage{booktabs}
-\usepackage{supertabular}
-
-% Adjust Figure Spacing
-\setlength\intextsep{1cm}
-
-% Add Gantt Chart
-\usepackage{pdfpages}
-\usepackage{pdflscape}
-
-% Center table with controlled width
 \usepackage{array}
@@
 
-% List of Equations
-\usepackage[titles]{tocloft}
-\newcommand{\listequationsname}{List of Equations}
-\newlistof{equations}{equ}{\listequationsname}
-\newcommand{\labelequations}[1]{%
-	\addcontentsline{equ}{equations}{\numberline{\theequation}#1}\par}
-\setlength{\cftequationsindent}{1.5em}
-\setlength{\cftequationsnumwidth}{2.3em}
-
-\addto\captionsamerican{
-	\renewcommand{\contentsname}
-	{Table of Contents}
-}
-
-% Subfigure
-\usepackage{subcaption}
-
-%%% LT-SKIP-END
+\DeclareAcronym{iso}{
+	short=ISO,
+	long=International Organization for Standardization
+}
+
+\DeclareAcronym{ansi}{
+	short=ANSI,
+	long=American National Standards Institute
+}
+
+\DeclareAcronym{jis}{
+	short=JIS,
+	long=Japanese International Standards
+}
+
+\DeclareAcronym{wpm}{
+	short=WPM,
+	long=Words per Minute
+}
+
+\DeclareAcronym{neck}{
+	short=WRNULD,
+	long=Work-related neck and upper limb disorders
+}
+
+\DeclareAcronym{cv}{
+	short=CV,
+	long=Computer Vision
+}
+
+\DeclareAcronym{roi}{
+	short=ROI,
+	long=Region of Interest,
+	long-plural-form=Regions of Interest
+}
+
+\DeclareAcronym{cts}{
+	short=CTS,
+	long=Carpal Tunnel Syndrome
+}
+
+\DeclareAcronym{jwt}{
+	short=JWT,
+	long=JSON Web Tokens
+}
+
+\DeclareAcronym{fpacc}{
+	short=FP ACC,
+	long=Finger Placement Accuracy
+}
+
+\DeclareAcronym{hfpacc}{
+	short=HFP ACC,
+	long=Historical Finger Placement Accuracy
+}
+
+\DeclareAcronym{cps}{
+	short=CPS,
+	long=Characters per Second
+}
+
+\journal{International Journal of Human-Computer Studies}
 
@@
 
-\pagenumbering{gobble}
-\begin{titlepage}
-	\centering
-
-	\hspace{0pt}
-	\vfill
-
-	\includegraphics[width=0.2\textwidth]{upc.png}
-	\includegraphics[width=0.2\textwidth]{dcs.png}
-	\par\vspace{1cm}
-
-	\textsc{Development of a Finger-Key Identification Module\\for a Touch Typing Trainer}
-	\par\vspace{0.5cm}
-
-	\hrulefill{}
-	\par\vspace{0.25cm}
-	A Special Project Presented to the\\
-	Faculty of the Department of Computer Science,\\
-	University of the Philippines Cebu
-
-	\par\vspace{0.25cm}
-	In Partial Fulfillment\\
-	Of the Requirements for the Degree\\
-	Bachelor of Science in Computer Science\\
-	\par\vspace{0.25cm}
-	\hrulefill{}
-	\par\vspace{0.5cm}
-
-	Oscar Vian L. Valles\\
-	BS Computer Science
-	\par\vspace{0.5cm}
-
-	Dhong Fhel K. Gom-os\\
-	Adviser
-	\par\vspace{0.5cm}
-
-	June 2022
-	\vfill
-	\hspace{0pt}
-\end{titlepage}
-
-\newpage
-
-\begin{center}
-
-	\hspace{0pt}
-	\vfill
-
-	\includegraphics[width=0.2\textwidth]{upc.png}
-	\includegraphics[width=0.2\textwidth]{dcs.png}
-	\par\vspace{0.25cm}
-
-	\textbf{\uppercase{University of the Philippines Cebu}}\\
-	College of Science\\
-	Department of Computer Science
-
-	\par\vspace{1cm}
-
-	\textbf{\uppercase{Development of a Finger-Key Identification Module\\for a Touch Typing Trainer}}
-
-	\par\vspace{1cm}
-
-	\textbf{Permission is given for the following people to have access to this thesis:}
-
-	\par\vspace{0.25cm}
-
-	\begin{tabular}{ | P{0.8\textwidth} | P{0.1\textwidth} | }
-		\hline
-		Available to the general public                                 & Yes \\
-		\hline
-		Available only after consultation with author or thesis adviser & Yes \\
-		\hline
-		Available to those bound by confidentiality agreement           & Yes \\
-		\hline
-	\end{tabular}
-	\par\vspace{1cm}
-
-	\line(1,0){0.6\textwidth}\\
-	Oscar Vian L. Valles\\
-	Student
-	\par\vspace{0.5cm}
-
-	\line(1,0){0.6\textwidth}\\
-	Dhong Fhel K. Gom-os\\
-	Special Problem Adviser
-
-	\vfill
-	\hspace{0pt}
-
-\end{center}
-
-\newpage
-
-\begin{center}
-	\uppercase{\textbf{APPROVAL SHEET}}
-\end{center}
-
-\par\vspace{0.5cm}
-\noindent
-The faculty of the University of the Philippines Cebu, College of Science,
-Department of Computer Science approves this Special Problem entitled:
-\par\vspace{0.5cm}
-
-\begin{center}
-	\textbf{Development of a Finger-Key Identification Module\\for a Touch Typing Trainer}\\
-	by\\
-	Oscar Vian L. Valles
-
-	\par\vspace{0.5cm}
-
-	\begin{minipage}[t]{0.45\textwidth}
-		\centering
-		\line(1,0){0.9\textwidth}\\
-		Dhong Fhel K. Gom-os\\
-		Special Problem Adviser
-
-		\par\vspace{0.5cm}
-		\line(1,0){0.9\textwidth}\\
-		Date Signed
-	\end{minipage}
-	\begin{minipage}[t]{0.45\textwidth}
-		\centering
-		\line(1,0){0.9\textwidth}\\
-		Prof. Aileen Joan O. Vicente\\
-		Chair, Department of Computer Science
-
-		\par\vspace{0.5cm}
-		\line(1,0){0.9\textwidth}\\
-		Date Signed
-	\end{minipage}
-
-	\par\vspace{2cm}
-
-	\line(1,0){0.405\textwidth}\\
-	Prof. Nelia S. Ereno\\
-	Dean, College of Science
-
-	\par\vspace{0.5cm}
-	\line(1,0){0.405\textwidth}\\
-	Date Signed
-
-	\par\vspace{2cm}
-\end{center}
-
-\newpage
-
-\pagenumbering{roman}
-\addcontentsline{toc}{chapter}{Abstract}
-\begin{center}
-	\LARGE\textbf{Abstract}
-\end{center}
-
+\begin{frontmatter}
+	\title{Development of a Finger-Key Identification Module for a Touch Typing Trainer}
+	\author[inst1]{Oscar Vian L. Valles}
+	\ead{oscarvianvalles@gmail.com}
+	\author[inst1]{Dhong Fhel K. Gom-os\corref{cor1}}
+	\ead{dkgomos@up.edu.ph}
+	\cortext[cor1]{Corresponding author at: Cebu City, Cebu, Philippines}
+	\affiliation[inst1]{organization={Center for Research in Intelligent Systems, Department of Computer Science, University of the Philippines Cebu},
+		country={Philippines}}
+
+	\begin{abstract}
 		Typing tests list out words for the user to repeat. These tests measure metrics
@@
 		track fingers while typing. This module was successful in finger-key
-identification with an accuracy of 99.58\% in a dataset of 938 keypresses. The
+		identification with an accuracy of 99.58\% in a dataset of 942 keypresses. The
 		module also had an average finger-key identification time of 0.083 seconds,
 		which allows the module to perform real-time finger-key identification for
-typing speeds up to 143.030 Words per Minute which is well beyond the median
+		typing speeds up to 143.040 Words per Minute which is well beyond the median
 		typing speed of the population. Furthermore, two metrics for correct finger
@@
 		correct finger over multiple typing test sequences.
-
-\newpage
-
-\addcontentsline{toc}{chapter}{Acronyms}
-\printacronyms{}
-\newpage
-
-\addcontentsline{toc}{chapter}{List of Tables}
-\listoftables
-\newpage
-
-\addcontentsline{toc}{chapter}{List of Figures}
-\listoffigures
-\newpage
-
-\addcontentsline{toc}{chapter}{List of Equations}
-\listofequations
-
-\newpage
-
-\tableofcontents
-\newpage
-
-\pagenumbering{arabic}
-
-\chapter{Introduction}
-
-\section{Background of the Study}
+	\end{abstract}
+
+	\begin{highlights}
+		\item Automated identification of what finger was used to press a key with
+		an accuracy of 99.58\% in a dataset of 942 keypresses.
+		\item Finger-key identifications take an average of 0.083 seconds. This
+			allows for finger-key identification for typing speeds up to 143.040 Words
+			per Minute
+		\item Two metrics for correct finger placement were created: Finger
+		Placement Accuracy and Historical Finger Placement Accuracy
+	\end{highlights}
+
+	\begin{keyword}
+		touch typing \sep keyboard typing metrics \sep touch typing trainer \sep
+		computer vision \sep finger tracking
+	\end{keyword}
+
+\end{frontmatter}
+
+\section{Introduction}
 There are a lot of educational typing tests available that help people learn
@@
 list out words that are then typed out. The entered keys are then compared to
-check if the user has typed the expected letter. At the end of the test, the
-time taken is calculated, and certain metrics are given. These metrics include
-\ac{wpm} and Accuracy \parencite{arif2009}. However, this method of examination
-leaves out a crucial part of typing --— correct finger placement.
-
-Incorrect finger placement may affect hand and finger health. Incorrect
-placement of the fingers may cause these hand and wrist positions: ulnar
-deviation, forearm pronation, and wrist extension \parencite{serina1999}. These
+check if the user has typed the expected letter \citep{bartnik2021}. At the end
+of the test, the time taken is calculated, and certain metrics are given.
+Examples of metrics include \ac{wpm} and Accuracy \citep{arif2009}. However,
+this method of examination leaves out a crucial part of typing --— correct
+finger placement.
+
+There have been studies that show the effect of keyboard typing on the human
+body. These studies have shown that keyboard typing affects our neck, shoulder,
+upper limb, wrist, arms, and fingers \citep{szeto2005, baker2007digit}. In
+addition, it has been shown that 22\% of computer users sustain musculoskeletal
+disorders of the upper extremity. This includes the neck, shoulder, hands, and
+wrists. \citep{gerr2002}
+
+Incorrect placement of the fingers exacerbates these issues. These erroneous
+finger placements may cause the following hand and wrist positions: ulnar
+deviation, forearm pronation, and wrist extension \citep{serina1999}. These
 three are hand and wrist positions that are common in all activities, however,
 prolonged periods in these positions, such as in typing, may cause injuries such
-as \ac{cts} \parencite{toosi2015}.
-
-Correct finger placement is usually taught at the beginning using a diagram,
-with each key being associated with a specific finger. For instance, the letter
-Q in a QWERTY layout should be pressed using the fifth digit, or the little
-finger, of the left hand, and this is shown on instruction manuals by coloring
-the fifth digit and the key Q with the same color or by placing the letters
-directly on the fingers \parencite{dobson2009touch}.
-
-Furthermore, touch typing using correct finger placement is frequently taught in
-the beginner level \parencite{donica2018}. This means that there is a need to
-weed out bad habits that may develop, like using the index finger for pressing
-the spacebar or backspace. However, it is impractical for an educator to check
-each student if they are not performing these movements as these may only show
-for a small period and may not be caught in time.
-
-Thus, there is a need for automatically identifying which finger is used to
-press which key during typing --- hereby defined as finger-key identification.
-
-\section{Research Questions}
-\begin{enumerate}
-	\item What algorithm or technique using computer vision is capable of
-	      finger-key identification.
-	\item What is the setup and configuration of a single optical camera can be
-	      used to capture images during keyboard typing for real-time finger-key
-	      identification using the algorithm identified in 1?
-	\item How to design and develop a finger-key identification module using
-	      the algorithm in 1 and camera setup and configuration in 2?
-	\item How accurate is the module developed in 3 in finger-key identification?
-	\item Is the module developed in 3 fast enough to run in real-time?
-	\item What new keyboard typing metric to develop that considers both the key and
-	      finger used to press as parameters?
-\end{enumerate}
-
-\section{Research Objectives}
-
-\subsection{General Objectives}
-To create a finger-key identification module that accurately identifies what
-finger was used to press a key at a point in time to help develop better typing
-habits and healthier typing ergonomics.
-
-\subsection{Specific Objectives}
-\begin{enumerate}
-	\item Develop an algorithm or technique using computer vision capable of
-	      finger-key identification.
-	\item Identify the setup and configuration of a single optical camera that can
-	      be used to capture images during keyboard typing for finger-key
-	      identification using the algorithm identified in 1
-	\item Design and develop a module using the algorithm in 1 and camera setup
-	      and configuration in 2
-	\item Determine the accuracy of the module developed in 3 in finger-key
-	      identification
-	\item Determine if the module developed in 3 is fast enough to run in
-	      real-time
-	\item Develop a new keyboard typing metric that uses both the key and
-	      finger used to press as parameters
-\end{enumerate}
-
-\section{Scope and Limitation}
-The experimental setup was simplified and tightly controlled. It was not
-representative of real-world conditions. The colors of the keyboard, cables, and
-surface were all controlled. So were the lighting and camera used.
-
-The supported keyboard of the module was limited to one specific type of
-keyboard. A 60\% Keyboard in a slightly modified \ac{ansi} layout with
-light-colored keys was used as shown in Figure~\ref{fig:60}. This type of
-keyboard only has the alphanumeric portion of a full-size keyboard. This limited
-the number of keys to be checked and reduced the overall movement of the hand
-during typing. The keyboard layout was a slightly modified \ac{ansi} layout. The
-majority of the layout still followed the base \ac{ansi} layout, with the
-alphanumeric portion remaining exactly the same.
-
-The module's effect on typing was not tested. This includes the module's effect
-on typing in education and typing in health, as the focus of the research was on
-the development and testing of the finger-key identification module only.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.7\textwidth]{60.png}
-	\caption{A 60\% keyboard in \ac{ansi} layout.}
-	\label{fig:60}
-	\centering
-\end{figure}
-
-\section{Significance of the Research}
-This research is beneficial for all users of physical keyboards. These include a
-majority of the population as there are a lot of professions that heavily rely
-on keyboards. Examples include developers, physicians, educators, and
-accountants. By having better ergonomics while typing --- by touch typing
-correctly --- wrist injuries can be prevented, and typing speed may be
-increased.
-
-This research also helps educators, especially early educators teaching beginner
-typists. By automatically checking for correct technique, the burden of checking
-each student is reduced, and directed interventions for bad habits can be easily
-created as students with these bad habits are easily identified.
-
-This research has a direct impact on people that have hand or wrist injuries
-that are caused by poor typing habits. By correcting these poor habits, pain
-from these injuries will be reduced, and even be prevented from occurring in the
-first place. A specific example of this is by reducing ulnar deviation which
-affects the nerve that is indicative of \ac{cts} \parencite{toosi2015}.
-
-\chapter{Review of Related Literature}
-
-\section{Keyboard Typing}
-
-Keyboard typing is the process of using a keyboard to input characters in a
-system. In the context of this paper, keyboard typing will refer to the act of
-using a physical keyboard to input characters into a computer system.
-
-\subsection{Keyboard Layouts and Form Factors}
-
-One key characteristic of a keyboard is its physical attributes. Keyboards come
-in a lot of layouts and form factors. Keyboard layouts are the shapes, size, and
-positions of a key on a keyboard while the form factor of a keyboard refers to
-its shape and dimensions. The form factor also refers to the number of keys included
-in the keyboard \parencite{parkkinen2018}. By combining different layouts and
-form factors, different permutations of a keyboard can be created.
-
-Different keyboard layouts and form factors also produce different effects for
-the user. This is due to how vastly different some keyboard layouts and form
-factors are from one another. Some layouts focus on ergonomics, while others
-focus on typing speed. Some form factors were designed for aesthetics, while
-others focus on comfort and health. As such, different layouts may affect typing
-performance, ergonomics, and long-term health effects \parencite{ciobanu2015}.
-
-\subsubsection{ANSI and ISO Layout}
-
-There are two common keyboard layouts around the world --- \ac{iso} and \ac{ansi}.
-
-\citeauthor{ansi} is the standard that first defined the \ac{ansi} layout.
-Figure~\ref{fig:ansi} illustrates what the \ac{ansi} layout looks like. This
-layout is also used by countries other than America. Examples of countries that
-use this layout as its standard are the Philippines, China, and Korea
-\parencite{apple-layout}. However, these countries also opt to modify the layout
-by adding extra layers to accommodate other character sets.
-
-\citeauthor{iso} is the standard series that defines a framework that is used to
-create other layouts. Layouts created from this standard are colloquially called
-\ac{iso} Layouts. Countries around the world use this framework to create
-layouts that fit the characters in their language. Examples of countries that
-use this framework to create their layout are France, Greece, Canada, and Sweden
-\parencite{apple-layout}.
-
-Both of these layouts usually utilize the same key ordering. This ordering is
-commonly called QWERTY, based on the first five characters of the first row
-of this specific layout.
-
-There are other layouts available, however, they are not as common as the two
-previously mentioned layouts. Examples of this include \ac{jis}. Other esoteric
-layouts, like Tsangan or split-backspace, also exist. These layouts modify the
-\ac{iso} and \ac{ansi} standards by adding or removing certain keys to fit the
-character set of a language, or for additional keys. Other layouts are also the
-same as \ac{ansi} or \ac{iso}, however, these layouts change the arrangement of
-the alphabet within the keyboard.
-
-Despite the ubiquity of these common layouts, studies have shown that these
-layouts are not ergonomic. The main issue with these layouts is the random
-configurations of the letters. The randomness of the layout necessitates
-memorization of the layout which reduces the ease of learning, reduces performance
-in typing by reducing speed, and increases typing errors
-\parencite{ciobanu2015}.
-
-\subsubsection{Keyboard Form Factors}
-
-There is only one common keyboard form factor used worldwide: the full-size
-keyboard. This keyboard contains all the keys specified in the keyboard
-layout. This includes the alphanumeric keys, the function keys, the navigation
-cluster, and the Numpad.
-
-Other common keyboard form factors are based on the full-size keyboard. The name
-of these layouts, 60\%, 75\%, and 80\% reference the remaining number of keys
-after cutting a portion off from the full-size keyboard. The 60\% keyboards only
-contain the alphanumeric cluster while the 80\% and 75\% layouts retain the
-navigation cluster and the function keys \parencite{parkkinen2018}. The main
-draw for using keyboards with reduced sizes is for aesthetics, space
-constraints, and ergonomics.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{ansi.png}
-	\caption{\ac{ansi} Keyboard layout with form factors. Reprinted from \fullcite{figure-ansi}}
-	\label{fig:ansi}
-	\centering
-\end{figure}
-
-\subsubsection{Ergonomic Keyboards Layouts and Form Factors}
-
-There have been other keyboard layouts and form factors created to mitigate
-common issues associated with QWERTY layouts. These include Colemak, Dvorak, and
-Alphabetical layouts. However, studies have shown that the layout itself does
-not matter as beginners do not necessarily see the keyboard as a structured set,
-but rather as a random collection of characters, even if it is alphabetized
-\parencite{norman1982},
-
-A different form factor has a great effect on ergonomics. One such example of a
-form factor is an ergonomic keyboard developed by Microsoft called Microsoft
-Natural MultiMedia Keyboard. \citeauthor{ripat2010} used this keyboard in
-determining that ergonomic keyboards can help in reducing symptoms of \ac{neck}.
-Figure~\ref{fig:mn} shows the layout of the Microsoft Natural MultiMedia
-Keyboard.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{mn.png}
-	\caption{Microsoft Natural MultiMedia Keyboard. Reprinted from \fullcite{mn}}
-	\label{fig:mn}
-	\centering
-\end{figure}
-
-There are other form factors other than the full-size keyboard and variations
-thereof that focus on ergonomics. One such example is a split keyboard layout
-where the keyboard is split in half, one for the left hand and one for the
-right. One such benefit, according to \citeauthor{ergodox}, is a more relaxed
-position due to typing at shoulder width.
-
-\subsection{Keyboard Typing Metrics}
-
-There are numerous metrics used to quantify keyboard typing performance. Two
-common metrics used in the majority of typing tests include Accuracy and Speed.
-
-\subsubsection{Standardized Keyboard Typing Assessments}
-
-To be able to measure these metrics, a keyboard typing assessment needs to be
-done. However, there are no standardized keyboard typing assessments
-\parencite{donica2018}. As such, teaching methods and assessments, like
-Keyboarding without Tears, Monkeytype, and Keybr, may produce different metrics
-for the same typist due to their difference in conducting the assessment.
-
-\subsubsection{Speed}
-Speed, also called as entry rate by \citeauthor{arif2009}, measures the number
-of characters entered within a specific time frame. The most common metric that
-measures speed is \ac{wpm}. \ac{wpm} as defined by \citeauthor{arif2009} is:
-
-\begin{equation}
-	WPM = \frac{|T| - 1}{S} \cdot 60 \cdot \frac{1}{5}
-	\label{equ:wpm}
-\end{equation}
-\labelequations{Words per Minute}
-
-Where, $|T|$ is the length of the text, $S$ is the time in seconds spent writing
-the text. This time starts directly after the first character has been pressed,
-and ends when the last letter has been entered. As such, $1$ is subtracted from
-$|T|$, as the time spent to find and press the first character cannot be
-accurately determined. However, some typing assessments do not subtract $1$ from
-$|T|$. $60$ refers to the number of seconds in a minute and $\frac{1}{5}$
-normalizes the metric for the average length of words.
-
-Other metrics also measure speed, but they aren't as commonly used as \ac{wpm}.
-These include Characters per Minute, Gestures per Second, Adjusted Words per
-Minute, and Keystrokes per Second
-
-\subsubsection{Accuracy}
-Accuracy measures the number of correctly pressed characters in an input string.
-Accuracy, as defined by \citeauthor{bartnik2021}, is:
-
-\begin{equation}
-	ACC = \frac{|C|}{|T|} \cdot 100\%
-\end{equation}
-\labelequations{Accuracy}
-
-Where $|C|$ is the number of correct characters and $|T|$ is the length of the
-text.
-
-The inverse of accuracy is error rate, where the number of incorrectly pressed
-characters is measured instead. \citeauthor{arif2009} describe 5 common error
-rate metrics: Error Rate, Minimum String Distance Error Rate, Keystroke per
-Character, Erroneous Keystroke Error Rate, and Total Error Rate.
-
-\subsubsection{Limitations of the Metrics}
-These metrics are all based on the inputted characters by the user. These
-metrics do not take into account other aspects of keyboard typing such as
-posture, hand and wrist positions, and finger placement. Consequently, these
-metrics do not give a full picture of the performance of the person typing, and
-they only provide a cursory view of how a person types.
-
-\subsection{Keyboard Typing Methodology}
-Keyboard typing can be accomplished in numerous ways. The main difference
-between the different methodologies is the number of fingers used when typing
-and how the typist navigates the keyboard to find the keys. The methodology
-ranges from Hunt and Peck to Touch Typing, with variations of the two in
-between.
-
-Hunt and Peck uses one finger on one hand to press a key. This method is aided
-by using vision to locate the specific key to press \parencite{hoot1986}. On the
-other hand, Touch typing uses standard QWERTY mapping to type without using
-visual cues. \parencite{dobson2009touch} This mapping involves assigning certain
-fingers to certain keys. Figure~\ref{fig:touch-type} is the standard QWERTY
-mapping used for an \ac{ansi} layout. Kinesthesis and proprioception are used in
-locating the keys \parencite{logan2016}.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{touch-type.png}
-	\caption{Standard QWERTY mapping for \ac{ansi}. Reprinted from \fullcite{logan2016}}
-	\label{fig:touch-type}
-	\centering
-\end{figure}
-
-\subsection{Typing Speed Statistics}
-
-Median typing speed using a keyboard with QWERTY layout, based on test data from
-\textcite{keybr}, is $\approx37.5$ \ac{wpm}. This median is taken from all the test
-results they have gathered from all of their users.
-
-\textcite{monkey-stats} released a dataset showing the speed of 12678 members of the
-Monkeytype community with the speed role. The dataset shows the fastest time
-that these members have achieved on 60-second tests. The median of the dataset
-lies in the 100--109 \ac{wpm} range, with 52.74\% of the community having top
-speeds slower than 100 \ac{wpm}.
-
-\textcite{arif2009} also performed a survey with 12 participants using
-multiple typing methods. They got an average \ac{wpm} of 75.85 with a standard
-deviation of 15.61 when their participants used a physical QWERTY layout. This
-average \ac{wpm} was also the highest among the other typing methods the
-researchers studied.
-
-\section{Keyboard Typing in Education}
-
-Today, students are expected to type essays, articles, and other submissions
-using word processors \parencite{poole2016}. Testing is also commonly done using
-computerized assessments which require the need for keyboards
-\parencite{moodle}. As such, there is a need for students to be well versed in
-keyboard typing and for keyboard typing to be part of the curriculum.
-
-Keyboard typing has been a part of this curriculum for a long time, with studies
-about effective methods to teach keyboard typing reaching as far back as 1986
-\parencite{hoot1986}. Studies have continued to this day to continue to optimize
-and improve methods of teaching keyboard typing to students.
-
-These studies start teaching kids at the kindergarten level and the studies try
-to optimize the teaching methods to improve the speed and accuracy of typing of
-the learners. By starting to teach touch typing to students early, these
-students will develop the potential for higher-level keyboard typing
-\parencite{donica2018}.
-
-\subsection{Expectations of Keyboard Proficiency}
-
-In the United States, keyboard typing is an expected learning outcome for third
-grade in the Common Core State Standards \parencite{ccs}. At this grade level,
-only basic keyboard typing skills are required. By fourth grade, students are
-expected to have enough proficiency to type one page in one sitting. This is
-increased to two pages by fifth grade.
-
-In the Philippine context, the \citeauthor{deped} expects learners with a mental
-age of 4--6.9 years old to use correct posture and locate characters, learners
-with a mental age of 7--11.9 are introduced to home row finger placement, and
-learners with a mental age of 12 and above are expected to ``use proper typing
-technique with efficiency and accuracy without looking at the keyboard''
-\parencite{deped}.
-
-\subsection{Current Teaching Methods}
-
-Current teaching methods involve replicating a given text. Learners then copy
-the text into a given text field that records the typed characters. Correct and
-incorrect characters are then identified, and suitable errors are presented.
-Afterward, metrics, such as \ac{wpm}, and accuracy are given
-\parencite{bartnik2021, typeracer}.
-
-Through this process, the learner goes through the three stages of Motor
-Learning Theory. The student undergoes the cognitive stage where they try to
-understand and create strategies to accomplish the given task. Then the
-associative stage follows where the strategies and skills learned from the
-previous stage are refined. At this stage, the learners are expected to rely less
-on visuals to locate the keys and more on kinesthesis. In the final stage, the
-autonomous stage, the learner does not rely on visuals at all and focuses on
-using kinesthetic feedback to find the keys. By this point, the learner has
-progressed from using Hunt and Peck, to becoming proficient in touch typing.
-\parencite{donica2018}
-
-\subsubsection{Keyboarding without Tears}
-
-Keyboarding without Tears is a web-based application and curriculum that teaches
-students touch typing. However, one key differentiator of this curriculum is the
-usage of a row-based standard mapping, rather than a column-based standard
-mapping that is common in other teaching guides. Figure~\ref{fig:kwt} shows the
-standard mapping used in this curriculum.
-
-This curriculum is self-directed and learners can learn at their own pace. At
-its core, the curriculum is designed to be 36-week long with 5--10 minutes of
-lessons per day. The lessons in the curriculum follow the three stages of
-Motor Learning Theory \parencite{kwt}.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{kwt.png}
-	\caption{Row based standard mapping. Reprinted from \fullcite{kwt}}
-	\label{fig:kwt}
-	\centering
-\end{figure}
-
-\subsubsection{Monkeytype, Typeracer}
-
-These two keyboard typing tests are similar. They follow a common experience
-where users type a predetermined phrase, quote, or random words, and metrics are
-given after the test. Afterward, the learners may try the test again, or choose
-another set of words to type. These typing tests do not have a structured
-curriculum for learning how to touch type. It is left to the learner to practice
-and learn on their own \parencite{bartnik2021, typeracer}.
-
-\subsubsection{Keybr}
-
-Keybr is similar to Monkeytype and Typerace, in that they also have the users
-type a predetermined phrase, quote, or random words. However, this application
-has more guidance compared to the two. Keybr uses statistics to create typing
-lessons that are appropriate to the current typing proficiency of the learner.
-The words selected are random at first, and the skill level of the learner is
-determined by the performance of the user with these words and characters. The
-information gathered is then used to generate new words for the next iteration.
-As an example, if a learner has difficulty typing the letter q, the next
-iterations will have a lot of words that contain the letter q.
-
-Statistics from their website show that this learning method is successful, with
-some learners improving their typing speed by 20--40 \ac{wpm} \parencite{keybr}.
-
-\section{Keyboard Typing in Health}
-
-There have been a lot of studies that show the effect of keyboard typing, and
-its associated movements (or lack thereof), affects the human body. These
-studies have shown that keyboard typing affects our neck, shoulder,
-upper limb, wrist, arms, and fingers \parencite{szeto2005, baker2007digit}
-
-\subsection{Health Issues arising from Keyboard Typing}
-
-\ac{neck} are a common issue that is associated with an elongated length of time
-maintaining a static posture. When using the computer, the posture commonly
-adapted by users has the neck and shoulder regions in a static hold for a long
-time. This results in forward neck flexion and increased muscle tension
-\parencite{szeto2005}.
-
-In addition, it has been shown that 22\% of computer users sustain
-musculoskeletal disorders of the upper extremity. This includes the neck,
-shoulder, hands, and wrists. \parencite{gerr2002}
-
-Carpal tunnel syndrome is also a common issue in the general population. This is
-caused by the chronic compression of the median nerve. There is a common belief
-that typing is one main cause for the disorder \parencite{carpal-myth}. There
-are no definite conclusions if this myth is true, however, a study by
-\citeauthor{toosi2015} found that typing causes ulnar deviation, especially if
-done without proper form. This ulnar deviation contributes to the swelling of
-the median nerve during and after typing. However, the authors noted that it is
-unclear if this swelling leads to long-term nerve injury.
-
-\subsection{Finger and Wrist Kinematics}
-The way people move their hands, wrists, and fingers differs between each
-person. This can be attributed to the different typing styles each person has.
-One key difference between people is the angle of the fifth digit.
-
-However, there are some common movements and positions regardless of typing
-style: flexion, or the curving of the fingers, across the fingers, is decreasing
-across the hand, with the 2nd digit having the least flexion. This may be due to
-the instinct to reduce pronation of the hand, which in turn increases the
-distance of the 2nd digit to the keyboard. In addition, some people isolate or
-extend one of their thumbs, usually the one not used for pressing a key. This is
-also true for some people that do not use their fifth digit during typing
-\parencite{baker2007}.
-
-The movement and angle of the wrists also depend on the typing style of the
-typists. Some people do not reposition their hands, while others do. This
-difference comes from the way these people reach for certain far-away keys. Some
-stretch their fingers to reach far-away keys, while others move their entire
-hand to reach these keys.
-
-For those that reach their keys by stretching their fingers, there is an
-increased probability that the wrists and fingers adapt to non-neutral postures.
-These include wrist extension, ulnar deviation, and pronation, which may cause
-musculoskeletal disorders of the upper extremity (\textcite{marklin1999} as cited in
-~\textcite{baker2007})
-
-\section{Finger and Hand Tracking}
-Finger and Hand tracking is a method of tracking fingers and hands in 3D space
-using motion capture systems or computer vision. This technique allows computers
-to perform actions and analyzes the motions and positions of these body parts.
-
-\subsection{Types of Tracking}
-
-\subsubsection{Hardware Aided Solutions}
-
-Motion Capture Systems allow for capturing detailed skeletal motion in humans.
-These systems usually capture full-body motion, focusing on large parts of the
-human body, such as the torso, limbs, and head.
-
-However, motion capture systems have difficulty in tracking more articulated
-body parts --- with the fingers being one of them. The industry standard for
-capturing finger movements is through the use of an optical marker-based motion
-capture system. This is due to its ability to capture natural motion accurately.
-
-This method uses cameras to triangulate the 3D location of markers attached to
-the limbs of a person. For finger tracking, 13--20 markers are placed on the
-fingers, and cameras are brought closer to track the small movements of the
-finger \parencite{wheatland2015}.
-
-But this method is cost-prohibitive, and cannot handle occlusions well.
-\citeauthor{alexanderson2016} present a method for an optical marker-based
-motion capture system that can predictably recover from self-occlusion and has a
-better performance compared to previously used algorithms, however, the issue of
-cost and self-occlusion persists.
-
-Bend-sensor gloves are also an option for finger tracking. These gloves have
-sensors within them that track joint angles in the hand and fingers. One key
-differentiator of this solution compared to the others is the removal of
-self-occlusion in the data. As such, this is commonly used in sign language, and
-gesture recognition due to its accuracy.
-
-However, these gloves need a lot of time to calibrate as cross-coupling of the
-sensors proves a problem. Cross-coupling is prevalent because the movement of
-one finger also moves other parts of the hand. These movements may cause a
-sensor aimed to track a specific movement of a different part of the hand to
-inadvertently detect a movement when there should be none
-\parencite{wheatland2015}.
-
-\subsubsection{Computer Vision}
-
-At its core, Computer Vision aims to perform tasks that the human visual system
-can do \parencite{cern}. This includes object classification, tracking, gesture
-recognition, and face recognition. At the present, most computer vision systems
-utilize deep learning algorithms, and convolutional networks to gather
-information from an image, or a set of images. One such example of a
-convolutional network used in computer vision is Inception by
-\citeauthor{szegedy2015} which proposes a convolutional neural network
-architecture for object classification and detection.
-
-\subsection{Available CV Solutions for Tracking}
-
-\subsubsection{OpenCV}
-
-OpenCV is an open-source library that provides ${\approx2500}$ optimized
-algorithms for computer vision and machine learning. This library is widely used
-by companies, researchers, and open source communities that utilize computer
-vision and machine learning in their projects. Examples of companies that use
-OpenCV include Google, Sony, and Honda.
-
-The library has C++, Python, Java, and Matlab interfaces. The library also supports
-Windows, Linux, Android, and macOS, allowing for great developer experience, and
-wide deployment capabilities \parencite{opencv}.
-
-\subsubsection{MediaPipe}
-\label{section:rrl-mediapipe}
-
-MediaPipe is an open-source computer vision framework that allows developers to
-create a perception pipeline. This perception pipeline is a directed graph of
-calculators. Data passes through the graph as packets and a group of packets
-constitutes a data stream. As the data passes through the pipeline, the
-calculators, produce the desired output.
-
-This framework allows for performant object detection, hand and finger tracking,
-and human pose detection. The framework also allows for combining multiple
-features, by adding them to the graph as calculators. MediaPipe has C++, Python,
-JS, and Coral interfaces. It also supports Android and iOS devices
-\parencite{mediapipe}.
-
-Examples of projects that utilized MediaPipe and its solutions include a
-real-time hand gesture recognition system by \textcite{mediapipe-sample-1} which
-utilized MediaPipe Hands. Another example is by \textcite{mediapipe-sample-2}
-which utilized MediaPipe Hands to track movements during pen spinning
-competitions. \textcite{mediapipe-sample-3} developed a real-time sign language
-recognition program using MediaPipe Hands.
-
-\subsubsection{MATLAB}
-MATLAB is a programming platform for the analysis and designing of systems.
-MATLAB is commonly used by engineers and scientists for computational
-mathematics \parencite{what-matlab}.
-
-A toolbox offered by MATLAB is the Computer Vision Toolbox which contains
-algorithms, and functions for use in the development of computer vision, 3D
-vision, and video processing systems. By using the available algorithms in the
-toolbox, such as YOLOv2, and ACF, hand detection and gesture recognition is made
-possible in the platform \parencite{matlab}.
-
-\subsection{Applications}
-
-There have been multiple applications and products that utilize hand and finger
-tracking as their main component.
-
-\textcite{dorf2001} present a use case for finger tracking in augmented
-environments. In the paper, interaction in a virtual environment through the use
-of gestures. The tracking system uses an optical marker-based motion capture
-system where the user wears a glove with retroreflective markers.
-
-\textcite{chiang2014} used a Kinect, a 3D sensing device by Microsoft that
-uses depth data, to track fingers to play virtual instruments. Virtual Pianos
-and Guitars were created and played with reliable and stable tracking.
-
-\textcite{yousaf2014} created a virtual keyboard that operates using finger
-tracking. The tracking uses the movement of the finger joints as the basis for
-selecting which key to press. The tracking first detects where the hands are.
-The result of the detection is then piped to a different algorithm for finger
-joint localization. These are then used to generate finger joint trajectories
-using kernel tracking. A Dynamic Bayesian Network then uses feature vectors from
-these trajectories to classify, detect, and recognize keystrokes.
-
-\section{Summary of the Research Gap}
-While there are a lot of applications and curricula aimed at teaching touch
-typing, there is no automated system available that detects if a person uses the
-correct finger to press a key.
-
-By having this system, educators can accurately determine if and when a student
-is having a hard time typing and if these students will need an intervention to
-correct mistakes.
-
-This is also important because certain movements and hand positions will cause
-nerve and muscular disorders that will impact the user. By correcting these
-problematic movements and hand positions, these disorders can be prevented.
-
-\chapter{Methodology}
-
-This section provides an overview of the development and testing of a real-time
-finger-key identification module. First, the experimental setup used for data
-collection will be presented. Following that, the algorithm used to implement
-real-time finger-key identification will be discussed. The process of training
-and testing of the developed algorithm comes next. Afterwards, the development
-and design of the proof of concept trainer will be described. Finally, this
-section introduces two metrics that were developed to quantify correct finger
-placement during typing tests.
-
-\section{Setup}
+as Carpal Tunnel Syndrom \citep{toosi2015} and musculoskeletal disorders of the
+upper extremity (\citet{marklin1999} as cited in~\citet{baker2007}).
+
+These issues can be prevented through keyboard typing education. Keyboard typing
+has been a part of the curriculum for a long time \citep{hoot1986} and studies
+have continued to this day to continue to optimize and improve methods of
+teaching keyboard typing to students. By starting to teach touch typing to
+students early, these students will develop the potential for higher-level
+keyboard typing \citep{donica2018}.
+
+There have been tools created to facilitate the teaching of touch typing.
+Keyboarding without Tears is a web-based application and 36-week curriculum that
+teaches students touch typing by following the three stages of Motor Learning
+Theory \citep{kwt}. Monkeytype and TypeRacer are two keyboard typing test
+websites where users type a predetermined phrase, quote, or random words, and
+metrics are given after the test \citep{bartnik2021, typeracer}. Keybr is
+similar to Monkeytype and Typerace, in that they also have the users type a
+predetermined phrase, quote, or random words. However, this application guides
+the user by using statistics to create typing lessons that are appropriate to
+the current typing proficiency of the learner \citep{keybr}.
+
+However, these typing tests only measure the ability to type the correct
+character and do not check correct finger placement. While educators may be able
+to weed out bad habits manually, it is impractical for them to check each
+student individually.
+
+While there have been other studies that utilize hand and finger tracking as
+input methods. Examples include studies by \citet{dorf2001} which used finger
+tracking as an input method in augmented environments, \citet{chiang2014} used a
+Kinect to track fingers to play virtual instruments, and \citet{yousaf2014}
+created a virtual keyboard that operates using finger tracking. However, these
+studies used hand and finger tracking as an alternative input method, rather
+than supplementing physical keyboard typing.
+
+There is currently no solution to check for correct finger placement while
+typing on a physical keyboard. Thus, there is a need for automatically
+identifying which finger is used to press which key during typing --- hereby
+defined as finger-key identification.
+
+The paper aims to create a finger-key identification module that accurately
+identifies what finger was used to press a key at a point in time to help
+develop better typing habits and healthier typing ergonomics through the use of
+computer vision algorithms. The paper also aims to create metrics to quantify
+finger placement during the duration of keyboard typing.
+
+\section{Algorithm}
+
+\subsection{Setup}
 \label{section:metho-setup}
 
-The experimental setup and configuration were composed of three elements: the
-camera, the keyboard, and the environment. This setup was used to collect data
-points to measure the algorithm's ability to perform finger-key identification
-in a controlled environment. Figure~\ref{fig:metho-setup} is an image of the
-experimental setup.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{full-setup.jpg}
-	\caption{The experimental setup}
-	\label{fig:metho-setup}
-	\centering
-\end{figure}
-
-\subsection{Camera}
-The setup used a single monocular camera positioned in a top-down view. The
-camera captures the entirety of the keyboard and the movement of the ten
-fingers. To do so, it was mounted on top of the monitor with the camera pointed
-down towards the table. Figure~\ref{fig:metho-setup-camera-placement}
-illustrates how the camera was mounted over the monitor and
-Figure~\ref{fig:metho-setup-keeb} is a picture taken with the camera at this
-position.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{camera-placement.jpg}
-	\caption{The camera angled downwards to capture the keyboard}
-	\label{fig:metho-setup-camera-placement}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{actual-keeb.png}
-	\caption{Image captured by the camera in one of the training iterations}
-	\label{fig:metho-setup-keeb}
-\end{figure}
-
-The specific camera used was a Logitech C920. It is a 3 mega pixel webcam that
-is capable of capturing color video in 1080p/30fps and 720p/30fps with a
-diagonal field of view of 78\degree. For the setup, the camera captured the
-video in 480p/30fps. This camera has a universal mounting clip that allows the
-camera to be correctly positioned within the experimental setup
-\parencite{logitech}.
-
-\subsection{Keyboard}
-\label{section:metho-keeb}
-
-The keyboard used was a 60\% keyboard as shown in
-Figure~\ref{fig:metho-setup-keeb}. This limited the necessary mapping for the
-algorithm to the alphanumeric portion of the keyboard. This keyboard choice also
-reduced the area that the camera captured, as this keyboard type is considerably
-smaller compared to a full-size keyboard. The keyboard also had its keycaps and
-case in a light color that contrasted the dark surface it was placed on. It also
-had a dark USB-C cable to blend with the dark surface. These color coordination
-steps improved initial keyboard detection.
-
-In addition, the keyboard layout was a modified \ac{ansi} layout. The majority
-of the keys remained the same, however some keys were changed: the pipe key was
-replaced by the backspace key, the backspace key was replaced by the pipe key
-and a delete key, and the bottom row has a longer spacebar. Overall, these are
-minor changes that does not affect the alphanumeric keys that are predominantly
-used in typing.
-
-\subsection{Computer}
-The computer used to run the algorithm is a desktop computer. The specifications
-of this computer is shown in Table~\ref{tab:table-specs}.
-
-\begin{table}[H]
-	\small
-	\centering
-	\caption{\label{tab:table-specs}Computer Specifications}
-	\begin{tabular}{ p{0.3\textwidth} p{0.6\textwidth} }
-		\toprule
-		Component & Specification                          \\
-		\midrule
-		CPU       & AMD Ryzen 5 3600                       \\
-		GPU       & AMD Radeon RX 5600 XT                  \\
-		RAM       & G.Skill Trident Z Neo RGB 16GB 3200mhz \\
-		Storage   & Samsung SSD 850 EVO                    \\
-		OS        & Arch Linux                             \\
-		\bottomrule
-	\end{tabular}
-\end{table}
-
-\subsection{Environment}
-The setup was lit with a lamp besides the camera. This light source evenly lit
-the keyboard and the fingers used for typing. The light source used was a common
-LED bulb rated at 9 watts with 700 lumens. This light is white with a color
-temperature of 6500k.
-
-In addition, the surface where the keyboard was placed was solid black
-without any variation of color. This also improved initial keyboard detection.
-
-\section{Algorithm}
-\label{section:metho-algo}
-
-\subsection{Computer Vision based Keyboard Detection, and Key Mapping}
+The experimental setup and configuration controlled three key elements: the
+camera, the keyboard, and the environment. The camera used was a Logitech C920.
+It was placed directly above the keyboard and captured in 720p/30fps. The
+keyboard used was a light-colored 60\% keyboard in a modified \ac{ansi} layout.
+This type of keyboard only has the alphanumeric portion of a full-size keyboard.
+The color of the surface the keyboard rested on was dark. The contrasting color
+of the surface and the keyboard was chosen to improve the performance of the
+algorithm in keyboard detection. The entire environment was evenly lit with a
+6500k LED bulb rated at 9 watts outputting 700 lumens.
+
+This setup was used to collect data points to measure the algorithm's ability to
+perform finger-key identification in a controlled environment.
+
+\subsection{Keyboard Detection and Key Mapping}
 \label{section:metho-algo-keyboard}
@@
 A computer vision algorithm was created as a starting point for mapping the keys
-of the keyboard within a video. A rough flowchart of the algorithm is shown in
-Figure~\ref{fig:metho-algo-key-flow}.
-
-\begin{figure}[H]
-	\centering
-	\begin{tikzpicture}[font=\small,thick]
-		\node[draw,
-			rounded rectangle,
-			minimum width = 2.5cm,
-			minimum height = 1cm,
-			inner sep = 3mm
-		] (start) {START};
-
-		\node[draw,
-			trapezium,
-			trapezium left angle = 65,
-			trapezium right angle = 115,
-			trapezium stretches,
-			below=of start,
-			minimum width=3.5cm,
-			minimum height=1cm,
-			inner sep = 3mm
-		] (input) {Input Frame};
-
-		\node[draw,
-			below=of input,
-			minimum width=3.5cm,
-			minimum height=1cm,
-			inner sep = 3mm
-		] (map) {Get Image Map};
-
-		\node[draw,
-			below=of map,
-			minimum width=3.5cm,
-			minimum height=1cm,
-			inner sep = 3mm
-		] (contours) {Get Key Contour Points};
-
-		\node[draw,
-			trapezium,
-			trapezium left angle = 65,
-			trapezium right angle = 115,
-			trapezium stretches,
-			below=of contours,
-			minimum width=3.5cm,
-			minimum height=1cm,
-			inner sep = 3mm
-		] (output) {Key-Edge Coordinates Map};
-
-		\node[draw,
-			below=of output,
-			rounded rectangle,
-			minimum width=2.5cm,
-			minimum height=1cm,
-			inner sep = 3mm
-		] (end) {END};
-
-		\draw[-latex] (start) edge (input)
-		(input) edge (map)
-		(map) edge (contours)
-		(contours) edge (output)
-		(output) edge (end);
-
-	\end{tikzpicture}
-	\caption{Flowchart of Keyboard Detection and Mapping Algorithm}
-	\label{fig:metho-algo-key-flow}
-\end{figure}
-
-\newpage
+of the keyboard within a video. OpenCV \citep{opencv} was used as the image
+processing library for this step.
 
@@
 The first portion of the algorithm creates an image map that is overlaid over
-the detected edges of the keyboard. The flowchart for this function is shown in
-Figure~\ref{fig:metho-algo-key-image-map}.
-
-\begin{figure}[H]
-	\centering
-	\begin{tikzpicture}[font=\small,thick]
-		\node[draw,
-			rounded rectangle,
-			minimum width = 2.5cm,
-			minimum height = 1cm,
-			inner sep = 3mm
-		] (start) {START};
-
-		\node[draw,
-			trapezium,
-			trapezium left angle = 65,
-			trapezium right angle = 115,
-			trapezium stretches,
-			below=of start,
-			minimum width=3.5cm,
-			minimum height=1cm,
-			inner sep = 3mm
-		] (input) {Input Frame};
-
-		\node[draw,
-			below=of input,
-			minimum width=3.5cm,
-			minimum height=1cm,
-			inner sep = 3mm
-		] (gray) {Convert to Grayscale};
-
-		\node[draw,
-			below=of gray,
-			minimum width=3.5cm,
-			minimum height=1cm,
-			inner sep = 3mm
-		] (denoise) {Denoising};
-
-		\node[draw,
-			below=of denoise,
-			minimum width=3.5cm,
-			minimum height=1cm,
-			inner sep = 3mm
-		] (edge) {Edge Detection};
-
-		\node[draw,
-			below=of edge,
-			minimum width=3.5cm,
-			minimum height=1cm,
-			inner sep = 3mm
-		] (thresh) {Thresholding};
-
-		\node[draw,
-			right=of gray,
-			minimum width=3.5cm,
-			minimum height=1cm,
-			inner sep = 3mm
-		] (contours) {Find Contours};
-
-		\node[draw,
-			below=of contours,
-			minimum width=3.5cm,
-			minimum height=1cm,
-			inner sep = 3mm
-		] (sort) {Simplify Contour};
-
-		\node[draw,
-			below=of sort,
-			minimum width=3.5cm,
-			minimum height=1cm,
-			inner sep = 3mm
-		] (transform) {Transform Image Map};
-
-		\node[draw,
-			trapezium,
-			trapezium left angle = 65,
-			trapezium right angle = 115,
-			trapezium stretches,
-			below=of transform,
-			minimum width=3.5cm,
-			minimum height=1cm,
-			inner sep = 3mm
-		] (output) {Virtual Image Map};
-
-		\node[draw,
-			below=of output,
-			rounded rectangle,
-			minimum width=2.5cm,
-			minimum height=1cm,
-			inner sep = 3mm
-		] (end) {END};
-
-		\node[coordinate, right = 0.5cm of thresh] (below) {};
-		\node[coordinate, right = 0.5cm of gray] (top) {};
-
-		\draw[-latex] (start) edge (input)
-		(input) edge (gray)
-		(gray) edge (denoise)
-		(denoise) edge (edge)
-		(edge) edge (thresh)
-		(thresh) -- (below) |- (top) |- (contours)
-		(contours) edge (sort)
-		(sort) edge (transform)
-		(transform) edge (output)
-		(output) edge (end);
-
-	\end{tikzpicture}
-	\caption{Flowchart of Get Image Map}
-	\label{fig:metho-algo-key-image-map}
-\end{figure}
-
-\newpage
-
-\paragraph{Convert to Grayscale}
-The input frame from the camera was converted to a 256 level grayscale image
-using \texttt{cvtColor} of OpenCV \parencite{opencv-cvtColor}. This step was
-performed because future steps of the algorithm did not require color values to
-work. In addition, this step optimized the algorithm as the number of dimensions
-analyzed was reduced. Figure~\ref{fig:metho-algo-convert} is the output of this
-step.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{grayscale.png}
-	\caption{Camera capture converted to grayscale}
-	\label{fig:metho-algo-convert}
-	\centering
-\end{figure}
-
-\paragraph{Denoising}
-The algorithm denoised the grayscale image using a bilateral filter as
-implemented by OpenCV \parencite{opencv-bilateral-filter}. This filter takes the
-range of the image into account, rather than just the domain. This resulted in
-an image that is smoothed while preserving its edges
-\parencite{bilateral-filter}. Figure~\ref{fig:metho-algo-denoise} is the output
-of this step.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{denoised.png}
-	\caption{Smoothed image with intact edges}
-	\centering
-	\label{fig:metho-algo-denoise}
-\end{figure}
-
-\paragraph{Edge Detection}
-The algorithm used a Sobel filter for edge detection. This filter uses a
-3$\times$3 kernel that is convolved twice --- once horizontally, and another
-vertically --- to produce a grayscale image of the outlines within the frame.
-The kernels used by the Sobel filter \parencite{sobel2014} is shown in
-Figure~\ref{fig:metho-algo-key-sobel} and the output of this step is shown in
-Figure~\ref{fig:metho-algo-sobel}.
-
-\begin{figure}[H]
-	\centering
-	$\begin{bmatrix}
-			+1 & 0 & -1 \\
-			+2 & 0 & -2 \\
-			+1 & 0 & -1
-		\end{bmatrix}$
-	$\begin{bmatrix}
-			+1 & +2 & +1 \\
-			0  & 0  & 0  \\
-			-1 & -2 & -1
-		\end{bmatrix}$
-	\caption{Sobel Operator Kernels. Reproduced from \fullcite{sobel2014}}
-	\label{fig:metho-algo-key-sobel}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{sobel.png}
-	\caption{Output of the Sobel filter}
-	\centering
-	\label{fig:metho-algo-sobel}
-\end{figure}
-
-\paragraph{Thresholding}
-The output of the Sobel filter is a grayscale image of 256 values, with each
-gray pixel indicating edges within the image. There is a need to reduce the
-range of these values to improve the ability of the algorithm to find the
-contours of these edges. To do so, the algorithm utilized Otsu's algorithm to
-perform automated thresholding. Otsu's algorithm determines a single threshold
-that is most optimal for the image \parencite{otsu}. This outputs a
-black-and-white image, with only two values. Figure~\ref{fig:metho-algo-thresh}
-is the resulting output of this step.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{threshed.png}
-	\caption{Threshed image of the edges}
-	\label{fig:metho-algo-thresh}
-	\centering
-\end{figure}
-
-\paragraph{Find Contours}
-\label{section:metho-algo-key-contours}
-Contours are curves joining all continuous points that have the same color or
-intensity \parencite{opencv-contours}. The OpenCV function
-\texttt{findContours}, with the \texttt{CHAIN\_APPROX\_SIMPLE} contour
-approximation method, was used to find the contours of the outlines of the
-object found using the previous step. This approximation method removed
-redundant points and returned the least amount of points that describe the
-shape. This OpenCV function implements the algorithm of~\textcite{contours} in their
-paper \citetitle{contours}. Figure~\ref{fig:metho-algo-contours} shows the contours
-found in the threshed image.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{find-contours.png}
-	\caption{All the contours found in the image}
-	\label{fig:metho-algo-contours}
-	\centering
-\end{figure}
-
-\paragraph{Simplify Contour}
-However, the points returned by the function can contain more than the four
-extreme points at the edges of the keyboard. In addition, more than one contour
-may be found within the image. As such, the algorithm sorted the contours by
-area, and selected the largest one.
-
-After sorting, the algorithm performed the Douglas-Peucker algorithm for Line
-Simplification \parencite{douglas-peucker}. This reduced the number of points in
-the largest contour to the minimum amount. In ideal cases, the number of points
-would be four, with each point corresponding to the edges of the keyboard.
-However, there were times when other objects would be within the frame, or they
-intersected with the keyboard. This would result in a contour that would be
-defined by more than four points --- which caused the entire algorithm to fail.
-
-The contour points that described the contour were then sorted clockwise. This
-was a necessity since the algorithm of~\textcite{contours} does not guarantee
-that the largest contour's points were returned in a clockwise arrangement.
-There is a need to have consistent ordering of these points since the
-transformation that will be performed in the next step pairs the contour points
-to a hard-coded array that indicates the edges of the image map. This hard-coded
-array was sorted clockwise. As such, if the contour points were not sorted
-clockwise, there will be no guarantee that the pairing would result in a
-transformation that creates a transformed image map that is oriented correctly,
-since the pairing may not pair the correct points.
-
-Figure~\ref{fig:metho-algo-largest-contour} shows the four contour points found
-at the edges of the keyboard within the image.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{get-largest-contour.png}
-	\caption{Simplified contour}
-	\label{fig:metho-algo-largest-contour}
-	\centering
-\end{figure}
-
-\paragraph{Transform Image Map}
-The virtual keyboard map is a rectangular image that contains individual
-\ac{roi} for each key in a 60\% \ac{ansi} keyboard. Each key has a corresponding
-color assigned to it and this color fills the region where this key is located
-at. Figure~\ref{fig:metho-algo-key-map} is the initial image map and
-Appendix~\ref{appendix:key-map} is a table that lays out the mapping of the
-color code in BGR format and its associated character in the image map.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.6\textwidth]{image-map.png}
-	\caption{Initial Image Map}
-	\label{fig:metho-algo-key-map}
-	\centering
-\end{figure}
-
-The virtual map was stretched over the keyboard using OpenCV's
-\texttt{warpPerspective} function. The function requires a perspective transform
-that was calculated using \texttt{getPerspectiveTransform}. This function takes
-in two arrays with four points each. The input points are the edges of the
-virtual map, ordered clockwise. The output points are the four points of the
-contour, ordered clockwise \parencite{opencv-image-transform}. The resulting
-transform was then used by \texttt{warpPerspective} to apply the transform to
-the virtual map. Figure~\ref{fig:metho-algo-transform} is the image obtained
-after warping the image map.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.6\textwidth]{transformed-image-map.png}
+the detected edges of the keyboard. This image map subdivides the keyboard into
+multiple \acp{roi} where each \ac{roi} represents a key.
+
+To do so, the input frame from the camera was converted to an 8-bit grayscale
+image. The frame was then denoised by using a bilateral filter. This filter
+removes noise while maintaining the edges. A Sobel filter was then used for edge
+detection \citep{sobel2014}. Thresholding was then performed using Otsu's
+algorithm \citep{otsu} to reduce the bit-depth of the image from 8 to 1. This
+reduction of bit depth improves the performance in finding contours. Contours
+were found using the algorithm of \citet{contours}.
+
+The detected contours were then simplified as the output of the algorithm
+results in multiple points. To find the four points of the edges of the
+keyboard, the contours were sorted and the largest one was selected. The
+Douglas-Peucker algorithm for line simplification \citet{douglas-peucker} was
+then performed to reduce the contour points to the minimum amount. The contour
+points were then sorted clockwise, as the algorithm of \citet{contours} does not
+guarantee the arrangement of the contour points. There is a need for the points
+to be ordered clockwise since each point will be paired to the edges of the
+image map, which were sorted in clockwise order.
+
+The image map was then stretched over the keyboard by using a perspective warp,
+where the four points of the edges of the keyboard were the end position of the
+four points on the edges of the image map. Figure~\ref{fig:metho-image-map}
+presents a grayscale frame and the resulting image map from that frame.
+
+In ideal cases, the number of points would be four, with each point
+corresponding to the edges of the keyboard. However, there were times when other
+objects would be within the frame, or they intersected with the keyboard. This
+would result in a contour that would be defined by more than four points ---
+which caused the entire algorithm to fail.
+
+\begin{figure}[h]
+	\centering
+	\begin{subfigure}{.50\textwidth}
+		\centering
+		\includegraphics[width=.995\linewidth]{grayscale.png}
+		\caption{Input image that was converted to grayscale}
+	\end{subfigure}%
+	\begin{subfigure}{.50\textwidth}
+		\centering
+		\includegraphics[width=.995\linewidth]{transformed-image-map.png}
 		\caption{Transformed image map}
-	\label{fig:metho-algo-transform}
-	\centering
+	\end{subfigure}%
+	\caption{First and last step of getting the image map}
+	\label{fig:metho-image-map}
 \end{figure}
@@
 in the keyboard based on the generated virtual map and the Key-Color Values map.
-The flowchart of this function is shown in
-Figure~\ref{fig:metho-algo-key-contour-points}
-
-\begin{figure}[H]
-	\centering
-	\begin{tikzpicture}[font=\small,thick]
-		\node[draw,
-			rounded rectangle,
-			minimum width = 2.5cm,
-			minimum height = 1cm,
-			inner sep = 3mm
-		] (start) {START};
-
-		\node[draw,
-			trapezium,
-			trapezium left angle = 65,
-			trapezium right angle = 115,
-			trapezium stretches,
-			below = of start,
-			minimum width = 3.5cm,
-			minimum height = 1cm,
-			align = center,
-			inner sep = 3mm
-		] (input) {Virtual Map \\ Key-Color Values Map};
-
-		\node[draw,
-			diamond,
-			aspect = 2,
-			below = of input,
-			minimum width = 4cm,
-			inner sep = 2mm] (for) {for Key, Color in Key-Color Values Map};
-
-		\node[draw,
-			below=of for,
-			minimum width=3.5cm,
-			minimum height=1cm,
-			inner sep = 3mm
-		] (filter) {Filter Virtual Map};
-
-		\node[draw,
-			below=of filter,
-			minimum width=3.5cm,
-			minimum height=1cm,
-			inner sep = 3mm
-		] (contours) {Find and Simplify Contour};
-
-		\node[draw,
-			below=of contours,
-			minimum width=3.5cm,
-			minimum height=1cm,
-			inner sep = 3mm
-		] (scale) {Scale Contours};
-
-		\node[draw,
-			trapezium,
-			trapezium left angle = 65,
-			trapezium right angle = 115,
-			trapezium stretches,
-			below=of scale,
-			minimum width=3.5cm,
-			minimum height=1cm,
-			inner sep = 3mm
-		] (output) {Key-Edge Coordinates Map};
-
-		\node[draw,
-			below=of output,
-			rounded rectangle,
-			minimum width=2.5cm,
-			minimum height=1cm,
-			inner sep = 3mm
-		] (end) {END};
-
-		\node[coordinate, right = 1cm of for] (top-right) {};
-		\node[coordinate, right = 1cm of output] (below-right) {};
-		\node[coordinate, left = 1cm of for] (top-left) {};
-		\node[coordinate, left = 2cm of scale] (below-left) {};
-
-		\draw[-latex] (start) edge (input)
-		(input) edge (for)
-		(for) edge (filter)
-		(for) -- (top-right) -- (below-right) |- (output)
-		(filter) edge (contours)
-		(contours) edge (scale)
-		(scale) -- (below-left) -- (top-left) |- (for)
-		(output) edge (end);
-
-	\end{tikzpicture}
-	\caption{Flowchart of Get Key Contour Points}
-	\label{fig:metho-algo-key-contour-points}
-\end{figure}
-
-\paragraph{Filter Virtual Map}
+
 The color that corresponds to each key in the virtual map is stored in the
-Key-Color Value Map. This color was used to create a mask for the virtual map
-for a specific key. This mask was obtained by using this snippet of code
-\texttt{(virtualMap == key).all(axis=2)}. This mask was then used to black out
-the rest of the virtual map, leaving only the \ac{roi} of the key in the image.
-
-This specific method was used as the other method that was trialed,
-\texttt{np.where(virtualMap != key)}, did not consider all 3 channels when
-comparing each pixel --- i.e.\ a key assigned with the color value of
-\texttt{[100, 0, 0]} will not be blacked out if the color value of the key to be
-isolated is \texttt{[100, 243, 0]} as the blue channel of the pixel,
-\texttt{100} is equal in both. This would result in other keys remaining in the
-virtual map, even if only one key corresponds to that color.
-Figure~\ref{fig:metho-algo-filtered} shows the image where the Left Control key
-was isolated.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{filtered-key.png}
-	\caption{Filtered key}
-	\label{fig:metho-algo-filtered}
-	\centering
-\end{figure}
-
-\paragraph{Find and Simplify Contours}
-The same method in finding and simplifying the contour of the keyboard was used
-to find and simplify the contour of the \ac{roi} of the key.
-Figure~\ref{fig:metho-algo-contour-key} illustrates the four edges of the
-isolated key.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{key-contour.png}
-	\caption{Contour points of the \ac{roi}}
-	\label{fig:metho-algo-contour-key}
-	\centering
-\end{figure}
-
-\paragraph{Scale Contours}
+Key-Color Value map. Each individual \ac{roi} was isolated using the color
+values stored in the map. The simplified contours of each \ac{roi} were then
+obtained using the same method in finding and simplifying the contour of the
+keyboard.
+
 During testing using training data, a lot of the failures in finger-key
@@
 buffer was needed as, in some instances, the finger used to press the key was
-not exactly on top of the key, but rather to its side. In addition, it can also
-be observed that the finger tracking does not annotate the very edge of the
-fingertip, but rather, it labels somewhere in the middle of the fingernail.
-
-The algorithm accounts for this situation by adding a buffer. This was achieved
-by scaling the contour points by seven pixels on each side.
-Figure~\ref{fig:metho-algo-scale-roi} shows the increased buffer in relation to
-the key. This value was obtained after testing other possible values. A buffer
-of five pixels did not have a lot of effect in reducing the number of failures.
-A buffer of ten pixels did reduce the number of failures, but it also greatly
-increased the algorithm's uncertainty in determining the finger --- uncertainty
-is defined as detecting two or more fingers within one \ac{roi}. Seven was a
-good middle ground in decreasing failures, without greatly increasing
-uncertainty.
+not exactly on top of the key, but rather to its side. In addition, it was also
+observed that the finger tracking does not annotate the very edge of the
+fingertip, but rather, it labels somewhere in the middle of the fingernail. The
+algorithm accounts for this situation by adding a buffer. This was achieved by
+scaling the contour points by seven pixels on each side.
+
+After the contours of all \acp{roi} were obtained, a Key-Edge Coordinates map
+was generated.
+
+\subsection{Finger Detection and Tracking}
+\label{section:metho-algo-finger}
+MediaPipe Hands \citep{mediapipe} was used as the finger detection and tracking
+solution. This algorithm is composed of two ML models working in conjunction to
+be able to detect the different parts of the hands and track them accurately.
+This algorithm outputs the positions of 21 hand and finger landmarks for each
+hand within the image.
+
+\subsection{Integration}
+
+The two algorithms were combined to accomplish finger-key identification The
+integration of the algorithm was a two-step process. The first step was to get
+the Key-Edge Coordinates map shown in Section~\ref{section:metho-algo-keyboard}.
+The second step runs whenever a key press has been detected. This step used the
+Key-Edge Coordinates map in conjunction with the finger tracking algorithm
+selected in Section~\ref{section:metho-algo-finger}.
+
+The finger tracking data contained the pixel positions of each landmark of the
+hands within the frame. The edge coordinates of the \ac{roi} of the key pressed
+were obtained using the Key-Edge Coordinates map. Any fingertip landmarks that
+fit within the \ac{roi} were then returned.
+
+\section{Data Gathering}
+Ten sentences were gathered from novels and texts in the public domain. These
+sentences were used as typing test sequences. 3 sets of 10 videos were then
+captured with each video corresponding to one typing test sequence. Each set
+corresponds to three predetermined speeds of typing: slow (15wpm), average
+(35wpm), and fast (80wpm). These speeds were based on test data
+from~\citet{keybr}. Best effort was made to stick to the predetermined typing
+speeds, however, there were discrepancies and the exact speed was not maintained
+throughout the entire typing test sequence. During typing, errors were not
+consciously taken into account, and errors happened naturally as part of the
+typing process --- with more errors happening more frequently as the typing
+speed increased.
+
+Manual finger-key identification was performed for all key presses present in
+the video. The annotated data was then divided into training and test data at a
+ratio of 60:40, respectively.
+
+\subsection{Data Inventory}
 
 \begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{scaled-contour.png}
-	\caption{Scaled contour points of the \ac{roi} with the buffer of seven pixels}
-	\label{fig:metho-algo-scale-roi}
-	\centering
-\end{figure}
-
-\paragraph{Key-Edge Coordinates Map}
-After all the keys have passed through the loop, a Key-Edge Coordinates Map was
-generated. Each key now has four points that serve as the coordinates of the
-edge of its \ac{roi}.
-
-The coordinate system that the contour points are based on had its origin at the
-top left, starting at 0, 0. Going to the right increased the value of the
-x-axis, and going to the bottom increased the value of the y-axis. This results
-in a coordinate system that only has positive values with each value
-corresponding to a pixel.
-
-\subsection{Computer Vision based Finger Detection, and Tracking}
-\label{section:metho-algo-finger}
-MediaPipe Hands was used as the finger detection and tracking solution. This
-algorithm is composed of two ML models working in conjunction to be able to
-detect the different parts of the hands and track them accurately.
-
-\subsubsection{Palm Detection Model}
-The first model, the Palm Detection Model detects the initial hand locations
-using a single shot detector model based on the paper by~\textcite{ssd}. This model
-achieves an average precision of 95.7\% in palm detection
-\parencite{mediapipe-hands}.
-
-MediaPipe Hands detects the palms first, instead of whole hands with one model
-because hands lack high contrast patterns. This reduces the model's ability to
-detect hands with accuracy. In addition, detecting a palm is simpler compared
-to detecting hands with articulated fingers since estimating a bounding box
-around a rigid object, i.e.\ a palm is much simpler. Furthermore, a palm can be
-modeled using only square anchors reducing the number of anchors by a factor of
-3--5 \parencite{mediapipe-hands}.
-
-\subsubsection{Hand Landmark Model}
-After the palms have been detected and an appropriate anchor has been
-established, the Hand Landmark Model pinpoints 21 3D hand-knuckle coordinates
-inside the detected hand. This is done using direct coordinate prediction.
-
-This model was trained using 30,000 manually annotated, real-world images with
-21 3D hand-knuckle coordinates. Using this information, the model can also
-accurately add landmarks to partially visible hands and hands with
-self-occlusion. This is also made possible by the model's consistent internal
-hand pose representation \parencite{mediapipe-hands}.
-
-\subsection{Integration for finger-key identification and mapping}
-
-The two previously chosen algorithms were combined to accomplish finger-key
-identification and mapping.
-
-The integration of the algorithm was a two-step process. The first step was to
-get the Key-Edge Coordinates Map shown in
-Section~\ref{section:metho-algo-keyboard}.
-
-The second step runs whenever a key press has been detected. This step used the
-Key-Edge Coordinates Map in conjunction with the finger tracking algorithm selected
-in Section~\ref{section:metho-algo-finger}. The flowchart of the algorithm is
-shown in Figure~\ref{fig:metho-algo-integration}
-
-\begin{figure}[H]
-	\centering
-	\tikzset{
-		rectangle connector/.style={
-				-latex,
-				to path={(\tikztostart) -- ++(6cm,0pt) \tikztonodes |- (\tikztotarget) },
-				pos=0.5
-			},
-	}
-	\begin{tikzpicture}[font=\small,thick]
-		\node[draw,
-			rounded rectangle,
-			minimum width = 2.5cm,
-			minimum height = 1cm,
-			inner sep = 3mm
-		] (start) {START};
-
-		\node[draw,
-			diamond,
-			aspect = 2,
-			below = of start,
-			minimum width = 4cm,
-			inner sep = 3mm] (if-packet) {Keypress Detected};
-
-		\node[draw,
-			trapezium,
-			trapezium left angle = 65,
-			trapezium right angle = 115,
-			trapezium stretches,
-			below = 2cm of if-packet,
-			minimum width = 3.5cm,
-			minimum height = 1cm,
-			align = center,
-			inner sep = 3mm
-		] (map) {Key-Edge Coordinates Map \\ Finger Tracking Data};
-
-		\node[draw,
-			below=of map,
-			minimum width = 3.5cm,
-			minimum height = 1cm,
-			inner sep = 3mm
-		] (find-finger) {Find which fingertip is on that \ac{roi}};
-
-		\node[draw,
-			trapezium,
-			trapezium left angle = 65,
-			trapezium right angle = 115,
-			trapezium stretches,
-			below = of find-finger,
-			minimum width = 3.5cm,
-			minimum height = 1cm,
-			inner sep = 3mm
-		] (return) {Return which fingertip};
-
-		\node[draw,
-			below = of return,
-			rounded rectangle,
-			minimum width = 2.5cm,
-			minimum height = 1cm,
-			inner sep = 3mm
-		] (end) {END};
-
-		\draw[-latex] (start) edge (if-packet)
-		(map) edge (find-finger)
-		(find-finger) edge (return)
-		(return) edge (end);
-
-		\draw[-latex] (if-packet) edge node[pos=0.5, fill=white, inner sep=2] {true} (map);
-		\draw[rectangle connector] (if-packet) to node[fill=white, inner sep=2] {false} (end);
-
-	\end{tikzpicture}
-	\caption{Flowchart of the overall flow}
-	\label{fig:metho-algo-integration}
-\end{figure}
-
-\subsubsection{Find which fingertip is on that \ac{roi}}
-The Finger Tracking Data contained the pixel positions of each landmark of the
-hand. For this step, the algorithm finds the landmark which is positioned within
-the single \ac{roi} obtained from the previous step. This was done using a
-series of checks.
-
-Each landmark's coordinates were compared to the coordinates of the edges of the
-\ac{roi}. A landmark was determined as positioned within the \ac{roi} if all the
-following conditions are true:
-
-\begin{enumerate}
-	\item On the X-axis, the landmark's coordinates are greater than one or both of
-	      the two coordinates found on the left side of the \ac{roi}
-	\item On the X-axis, the landmark's coordinates are less than one or both of
-	      the two coordinates found on the right side of the \ac{roi}
-	\item On the Y-axis, the landmark's coordinates are greater than one or both of
-	      the two coordinates found on the top side of the \ac{roi}
-	\item On the Y-axis, the landmark's coordinates are less than one or both of
-	      the two coordinates found of the bottom side of the \ac{roi}
-\end{enumerate}
-
-These conditions maximized the total area of the \ac{roi}, and it is not strict
-about exact accuracy. In essence, these conditions create a perfectly
-rectangular box that contained the quadrilateral formed by the four points from
-the coordinates. Figure~\ref{fig:metho-algo-sample} is an example of a frame with a
-successful finger-key identification using the conditions previously mentioned.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{detected-finger.png}
-	\caption{Fingertip within an \ac{roi}}
-	\label{fig:metho-algo-sample}
-	\centering
-\end{figure}
-
-\subsubsection{Return which fingertip}
-The landmark was then used to determine which specific finger corresponds to the
-key press. Figure~\ref{fig:metho-algo-integration-landmarks} shows each possible
-landmark that may be returned from the previous step. This step returned the
-name of the landmark, up until the first underscore. As an example, if the
-landmark found is \texttt{MIDDLE\_FINGER\_TIP}, this step will return
-\texttt{MIDDLE} denoting that the middle finger is the finger that corresponds
-with the key press. In addition, the specific hand will also be returned as one
-of two strings, \texttt{LEFT} and \texttt{RIGHT}, since this information is also
-bundled together with the landmarks. The two strings were then concatenated with
-an underscore. An example return value is \texttt{RIGHT\_RING}.
-
-As such, the expected output of the module is the hand and the finger used to
-press the key. This allows for a more flexible utilization of the data as it
-carries greater context, compared to just returning if the finger used was
-correct or not.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.6\textwidth]{hand-landmarks.png}
-	\caption{Hand landmarks that may be returned by the algorithm. Reproduced from \fullcite{mediapipe-hands}}.
-	\label{fig:metho-algo-integration-landmarks}
-	\centering
-\end{figure}
-
-\subsection{Implementation}
-\label{section:metho-algo-implementation}
-The previously discussed algorithms were implemented using Python. The Keyboard
-Detection and Mapping Algorithm in Section~\ref{section:metho-algo-keyboard} was
-implemented as a separate module in Python using OpenCV. The Finger Detection
-and Tracking Solution in Section~\ref{section:metho-algo-finger} was implemented
-using the prebuilt Python package offered by the MediaPipe team. Finally, the
-integration of the two was done as another separate module in Python.
-
-\section{Training and Testing}
-
-\subsection{Typing Test Sequences}
-Typing test sequences are strings that were used in testing the user in their
-ability to type. The test sequences that were used came from text found in the
-public domain obtained from Project Gutenberg and the Internet Archive.
-Sentences were isolated from these texts and used as test sequences if they fit
-the criteria.
-
-The criteria for choosing test sequences were as follows: (1) $\approx80\%$ of
-the characters on the keyboard is present in a test sequence. (2) The number of
-words in a test sequence does not exceed 25. (3) Numbers and punctuations should
-be present in at least $\approx30\%$ of the total test sequences.
-
-There was a total of 10 test sequences that were gathered. An example test
-sequence is ``What of it, if some old hunks of a sea-captain orders me to get a
-broom and sweep down the decks?'' from Moby Dick by~\textcite{moby-dick}. The 9
-other test sequences can be found in Appendix
-Chapter~\ref{appendix:test-sequences}.
-
-\subsection{Data Gathering}
-\subsubsection{Video Capture}
-There were 3 sets of 10 videos that were captured by the researcher, for a total
-of 30 videos. Each set corresponds to three predetermined speeds of typing: slow
-(15wpm), average (35wpm), and fast (80wpm). These speeds were based on test data
-from~\textcite{keybr}. The researcher typed the 10 typing test sequences at
-these predetermined speeds. Best effort was made to stick to the predetermined
-speeds. During typing, errors were not consciously taken into account, and
-errors happened naturally as part of the typing process --- with more errors
-happening more frequently as the typing speed increased.
-
-A Python script was created to facilitate this process. The script captured a
-video of the researcher as the researcher was typing the test sequences.
-Whenever the researcher pressed a key, the corresponding frame count was
-captured, and the key pressed was recorded in conjunction. This was stored in
-the format \texttt{FRAME\_NO:KEY\_PRESSED}, and each keypress was then stored as
-a line in a file.
-
-One caveat of the camera used was its autofocus capabilities. This meant that
-there was a set amount of time during the start of the recording where the
-camera captured blurry images of the keyboard. This necessitates around three
-seconds for the autofocus to complete and focus on the keyboard. There was one
-video capture where the whole test sequence was scrapped as the researcher
-started typing even if the camera was not focused.
-
-\subsubsection{Keypress Labeling}
-The researcher then manually performed finger-key identification for all key
-presses present in the video. This was also done through a Python script. The
-script parsed the file associated with a video. The frame number found at each
-line was then shown, and the associated data with the frame was superimposed
-over it. Figure~\ref{fig:metho-data-process} is an example of a frame with these
-data overlaid over the captured image. The researcher then pressed keys that
-corresponded with the finger used to press the key. This was then again stored
-in the format \texttt{FRAME\_NO:KEY\_PRESSED:FINGER\_USED}, and each keypress
-was then stored as a line in a file.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.6\textwidth]{labeling.png}
-	\caption{Screen showing the labeling process}
-	\label{fig:metho-data-process}
-	\centering
-\end{figure}
-
-\subsubsection{Test-Training Split}
-The annotated data was then divided into training and test data at a ratio of
-60:40.
-
-\subsubsection{Data Inventory}
-
-\begin{figure}[H]
-	\centering
 	\begin{tikzpicture}
@@
 				xbar, xmin=0,
-				bar width=4pt,
-				width=0.8\textwidth,
-				ylabel=Key,
+				bar width=3pt,
+				width=0.9\textwidth,
+				height=1\textwidth,
+				enlarge y limits=0.02,
 				xlabel=Count,
 				symbolic y coords={
-						J,
-						Q,
+						Spacebar,
+						Period,
+						LeftShift,
+						LeftAlt,
+						ForwardSlash,
 						Dash,
-						ForwardSlash,
+						Comma,
+						Backspace,
 						Z,
-						1,
-						LeftAlt,
+						Y,
 						X,
-						Comma,
-						K,
+						W,
 						V,
-						G,
-						B,
-						Backspace,
-						W,
-						Period,
-						P,
 						U,
-						C,
-						LeftShift,
-						F,
-						Y,
-						M,
-						H,
-						D,
-						L,
-						I,
-						R,
-						N,
 						T,
 						S,
+						R,
+						Q,
+						P,
+						O,
+						N,
+						M,
+						L,
+						K,
+						J,
+						I,
+						H,
+						G,
+						F,
+						E,
+						D,
+						C,
+						B,
 						A,
-						O,
-						E,
-						Spacebar,
+						1,
 					},
@@
 				tick label style={font=\footnotesize},
+				legend style={font=\footnotesize},
 				label style={font=\footnotesize},
-				xmajorgrids=true
-			]
-			\addplot table[x=Value,y=Key]{training-data.txt};
+				xmajorgrids=true,
+				area legend,
+			]
+
+			\addplot[
+				CornflowerBlue,
+				fill=CornflowerBlue
+			] table[x=Value,y=Key]{test-data.txt};
+
+			\addplot[
+				Salmon,
+				fill=Salmon,
+			] table[x=Value,y=Key]{training-data.txt};
+			\legend{Test Dataset, Training Dataset}
 		\end{axis}
 	\end{tikzpicture}
-	\caption{Training Dataset Inventory}
+	\caption{Data Inventory}
 	\label{fig:metho-inventory-training}
@@
 
-Individual counts of the characters in the training iteration data set is shown
-in Figure~\ref{fig:metho-inventory-training}. The letters were not represented
-equally in the data set. Of note, J and Q were not in the dataset.
-
-\begin{figure}[H]
-	\centering
-	\begin{tikzpicture}
-		\begin{axis}[
-				xbar, xmin=0,
-				bar width=4pt,
-				width=0.8\textwidth,
-				ylabel=Key,
-				xlabel=Count,
-				symbolic y coords={
-						J,
-						Q,
-						Z,
-						Dash,
-						ForwardSlash,
-						X,
-						V,
-						Period,
-						K,
-						B,
-						Comma,
-						G,
-						F,
-						Backspace,
-						P,
-						W,
-						C,
-						U,
-						M,
-						Y,
-						D,
-						LeftShift,
-						L,
-						H,
-						N,
-						R,
-						S,
-						I,
-						O,
-						A,
-						T,
-						E,
-						Spacebar,
-					},
-				ytick=data,
-				tick label style={font=\footnotesize},
-				label style={font=\footnotesize},
-				xmajorgrids=true
-			]
-			\addplot table[x=Value,y=Key]{test-data.txt};
-		\end{axis}
-	\end{tikzpicture}
-	\caption{Test Dataset Inventory}
-	\label{fig:metho-inventory-test}
-\end{figure}
-
-The letters were also not represented equally in the test dataset as shown in
-Figure~\ref{fig:metho-inventory-test}. Three characters in the alphabet were not
-represented in the dataset: J, Q, and Z. The inequality of representation in
-both of these datasets can be attributed to the source materials of the typing
-test sequences --- novels and texts. While there were criteria to select which
-text was sourced, there was no conscious decision to equally represent all
-characters in the data set.
-
-\subsection{Training}
+Individual counts of the characters in the training and testing data set are
+shown in Figure~\ref{fig:metho-inventory-training}. The letters were not
+represented equally in the data set. Of note, J and Q were not in the training
+dataset, and J, Q, and Z were not in the testing dataset. The inequality of
+representation in both of these datasets can be attributed to the source
+materials of the typing test sequences --- novels and texts. There was no
+conscious decision to equally represent all characters in the data set.
+
+\section{Training and Testing}
 A script was created to test the accuracy of the module. This script opened each
@@
 video to perform step one of the module: frames ten, fifteen, and twenty. The
-first frame that returned a Key-Edge Coordinates Map after going through step
-one of the module was used. The rest of the frames were not passed through the
-module and were discarded.
-
-The script then parsed the labeled file and opened the frame for each keypress.
-The second algorithm of the module was then run, using the data needed for the
-algorithm: the key pressed. The result of the module was then compared with the
-manual labeling of the keypress. The resulting data from the test was then
-stored in a \texttt{.csv} file with the following content:
-
-\begin{table}[H]
-	\small
-	\centering
-	\caption{\label{tab:csv-output}CSV File Contents}
-	\begin{tabular}{ p{0.3\textwidth} p{0.6\textwidth} }
-		\toprule
-		Column Name                           & Definition                                             \\
-		\midrule
-		\texttt{file\_name}                   & Name of the file                                       \\
-		\texttt{frame}                        & Frame number                                           \\
-		\texttt{key}                          & Key pressed                                            \\
-		\texttt{finger}                       & Manual labeling of the finger used to press the key    \\
-		\texttt{is\_correct}                  & If the module was correct                              \\
-		\texttt{fingertips\_detected}         & The fingertips detected by the module                  \\
-		\texttt{keyboard\_detection\_time}    & Time spent in creating the Key-Edge Coordinates Map    \\
-		\texttt{finger\_classification\_time} & Time spent in classifying which finger pressed the key \\
-		\bottomrule
-	\end{tabular}
-\end{table}
-
-Adjustments were then made to the algorithm after each training pass to improve
-its performance. Some runs had improvements in their statistics without doing
-any adjustments. These were runs that showed errors in the manual labeling and
-were corrected afterward.
-
-\subsection{Analysis}
-The resulting \texttt{.csv} file was imported into Google Sheets and key
-statistics were obtained based on the data. Appendix~\ref{appendix:sheets} shows
-the screenshots where this was performed. The statistics obtained can be found
-in Table~\ref{tab:acquired-stats}. These key statistics informed what to adjust
-in the algorithm and its overall performance.
-
-\begin{table}[H]
-	\small
-	\centering
-	\caption{\label{tab:acquired-stats}Acquired statistics}
-	\begin{tabular}{ p{0.3\textwidth} p{0.6\textwidth} }
-		\toprule
-		Identifications                                                                                                                    \\
-		\midrule
-		Successes                                & Number of successful finger-key identifications                                         \\
-		Failures                                 & Number of failed finger-key identifications                                             \\
-		Success Percentage                       & Percentage of successful finger-key identifications                                     \\
-		Failure Percentage                       & Percentage of failed finger-key identifications
-		\\[0.25cm]
-		\midrule
-		Failure Types                                                                                                                      \\
-		\midrule
-		Mismatched Identification                & Module output did not match manual labeling                                             \\
-		No Identification                        & Module did not detect fingertips in \ac{roi}
-		\\[0.25cm]
-		\midrule
-		Uncertain Successes                                                                                                                \\
-		\midrule
-		Total                                    & Total uncertain successes, defined as detecting two or more fingers within one \ac{roi} \\
-		Spacebar                                 & All uncertain successes that were on the spacebar                                       \\
-		Spacebar Percentage                      & Percentage of spacebar uncertain successes over all uncertain successes                 \\
-		Spacebar Percentage (Total)              & Percentage of spacebar uncertain successes over total keypresses                        \\
-		Non-spacebar                             & All other uncertain successes                                                           \\
-		Non-spacebar Percentage                  & Percentage of non-spacebar uncertain successes over all uncertain successes             \\
-		Non-spacebar Percentage \newline (Total) & Percentage of non-spacebar uncertain successes over total keypresses
-		\\[0.25cm]
-		\midrule
-		Average Running Time                                                                                                               \\
-		\midrule
-		Keyboard Detection Time                  & Average keyboard detection time of all videos                                           \\
-		Finger Identification Time               & Average finger identification time of all key presses                                   \\
-		\bottomrule
-	\end{tabular}
-\end{table}
-
-\subsection{Testing}
-The same script was run through the test data. The same format of output was
-gathered, and the same key statistics were recorded.
-
-\section{Trainer}
-A trainer was developed as a proof of concept of a real-world implementation of
-the finger-key identification module.
-
-\subsection{Design}
-The design was heavily based on Monkeytype by \textcite{bartnik2021}. Additional
-components were added to enable real-time finger-key identification. There are
-three main screens:
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1\textwidth]{frontend-edge-coords.png}
-	\caption{Frontend Screen for obtaining Key-Edge Coordinates Map}
-	\label{fig:metho-front-coords}
-	\centering
-\end{figure}
-
-Figure~\ref{fig:metho-front-coords} shows the first screen the user sees. This page
-captured the keyboard and passed it to the backend to obtain the Key-Edge
-Coordinates Map.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1\textwidth]{frontend-test.png}
-	\caption{Testing page of the trainer}
-	\label{fig:metho-front-test}
-	\centering
-\end{figure}
-
-Figure~\ref{fig:metho-front-test} is the screen where the user performs the
-typing test. On this screen, the typing test sequence was placed above. If a
-letter was typed correctly with the correct finger, the character was colored
-white. If the wrong character was typed, the character was colored red. If the
-correct character was typed, but the finger used was incorrect, the character
-was colored purple.
-
-At the bottom, the video captured by the webcam was shown on the left. Right
-beside it, important coordinates of the finger-key identification module are
-overlaid over the specific frame that was used as the input.
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=1\textwidth]{frontend-metrics.png}
-	\caption{Page presenting the metrics obtained from the trainer}
-	\label{fig:metho-front-metrics}
-	\centering
-\end{figure}
-
-The metrics gathered from the test sequences are shown on a page as illustrated
-in Figure~\ref{fig:metho-front-metrics}. This page presents metrics that will be
-discussed in Section~\ref{section:metric}. Of note, the metrics shown on this
-page that use multiple test sequences as an input uses tests that were obtained
-only during the time the page was opened. This proof of concept does not store
-the data of a user's previous test sequences that were performed from a
-different session.
-
-\subsubsection{Implementation}
-The frontend was developed in Svelte, and the backend was developed with Flask.
-The frontend handled capturing the video, presenting the layout, and capturing
-input events. Finger Tracking was done on the frontend through the use of
-MediaPipe Hand's JavaScript library. The frontend also handled checking what
-fingertip was inside the \ac{roi} of the key that was entered. The backend
-handled the generation of the Key-Edge Coordinates Map using the image passed
-from the frontend.
-
-Two of the key steps in the module: finger tracking and integration were
-offloaded in the frontend as the wall of latency that exists between the
-frontend and the backend would cause performance issues. While this issue also
-exists in obtaining the Key-Edge Coordinates Map, this wouldn't affect real-time
-finger-key identification since this is a preprocessing step and is not ran
-during the time the user performs the typing test.
+first frame that returned a Key-Edge Coordinates map was used. The script then
+parsed the labeled file and opened the frame for each keypress. The second
+algorithm of the module was then run. The result of the module was then compared
+with the manual labeling of the keypress. Adjustments were then made to the
+algorithm after each training pass to improve its performance.
+
+The same script was run through the test data.
 
@@
 \end{equation}
-\labelequations{Finger Placement Accuracy}
 
@@
 \end{equation}
-\labelequations{Historical Finger Placement Accuracy}
 
@@
 
-\chapter{Results and Discussion}
-\label{section:rd}
-
-\section{Accuracy}
+\section{Proof of Concept Trainer}
+A trainer was developed as a proof of concept of a real-world implementation of
+the finger-key identification module. It consisted of three screens. The first
+screen is where the keyboard was captured to perform get the Key-Edge
+Coordinates map. The second screen is where the typing test is performed and
+where real-time finger-key identification was done on each keypress. The third
+screen shows the computed metrics as defined in Section~\ref{section:metric}. Of
+note, the metrics shown on this page that uses multiple test sequences as an
+input use tests that were obtained only during the current training session.
+This proof of concept does not store the data of a user's previous test
+sequences that were performed from a different session.
+
+\section{Finger-Key Identification Accuracy}
 Accuracy, in the context of the finger-key identification module, calculates the
@@
 \end{equation}
-\labelequations{Finger-Key Identification Module Accuracy}
 
@@
 
-\subsection{Training Results}
+\subsection{Training Iterations}
 There were a total of seven training iterations. These training iterations
@@
 				label style={font=\footnotesize},
-				every axis legend/.append style={cells={anchor=west}},
+				every axis legend/.append style={at={(0.98, 0.02)}, anchor=south east, cells={anchor=west}},
 				ymajorgrids=true,
@@
 multiple tries to the training routine in obtaining the Key-Edge Coordinates
-Map. Iterations one through three only tried frame 10 as the source frame for
+map. Iterations one through three only tried frame 10 as the source frame for
 the module while iterations four through seven tried frames 10, 15, and 20.
@@
 each side. Iteration four's increase was due to fixing the image map. Certain
-keys had incorrect color values assigned in the Key-Color Values Map, and some
+keys had incorrect color values assigned in the Key-Color Values map, and some
 keys were of incorrect shape. Iterations five to seven were performed to test
@@
 In a 60\% keyboard, the spacebar is the longest key on the keyboard. In
-addition, this key is right below the non-dominant hand's thumb default resting
-position when the hand's posture follows the proper touch typing posture. As
-such, it was to be expected that a majority of the uncertain successes were
-because of the spacebar. However, this was not the case for iterations three and
-four. This was because the buffer for the \ac{roi} was too big and this caused
-the module to detect more than one fingertip within that \ac{roi}.
+addition, this key is right below the non-dominant hand's thumb's default
+resting position when the hand's posture follows the proper touch typing
+posture. As such, it was to be expected that a majority of the uncertain
+successes were because of the spacebar. However, this was not the case for
+iterations three and four. This was because the buffer for the \ac{roi} was too
+big and this caused the module to detect more than one fingertip within that
+\ac{roi}.
 
@@
 
-\begin{figure}[H]
-	\centering
-	\begin{subfigure}{.5\textwidth}
+\begin{figure}[h]
+	\centering
+	\begin{subfigure}{.33\textwidth}
 		\centering
 		\includegraphics[width=.995\linewidth]{failure-3.png}
-	\end{subfigure}\\
-	\begin{subfigure}{.5\textwidth}
+	\end{subfigure}%
+	\begin{subfigure}{.33\textwidth}
 		\centering
@@
 	\end{subfigure}%
-	\begin{subfigure}{.5\textwidth}
+	\begin{subfigure}{.33\textwidth}
 		\centering
@@
 also increase the number of uncertain successes. See
-Section~\ref{section:rd-uncertain vs success}, Uncertain Successes vs Success
-Percentage for more information.
+Section~\ref{section:rd-uncertain vs success} for more information.
 
@@
 
-\section{Speed}
-
-Based on the definition of \ac{wpm} from \textcite{arif2009} as seen in
-Equation~\ref{equ:wpm}, it can be extrapolated that calculating \ac{cps} is:
-
-\begin{equation}
-	CPS = \frac{WPM}{60} \cdot 5
-	\label{eq:cps}
-\end{equation}
-\labelequations{Characters per Second}
-
-The average \ac{wpm} when typing using the QWERTY layout, according to test data
-from \citeauthor{keybr} is $\approx37.5$ \ac{wpm}. As such, we can calculate
-that the average Characters per Second when typing using the QWERTY layout is
-3.125. We can calculate the seconds needed per character by this equation:
-
-\begin{equation}
-	s = \frac{1}{CPS}
-	\label{eq:s}
-\end{equation}
-\labelequations{Seconds per Character}
-
-This means that each character, on average, requires 0.32 seconds to type. This
-sets the maximum time spent for finger-key identification for a single
-character. Any more than this would result in a noticeable slowdown during
-typing.
-
-\subsection{Results}
-
-\subsubsection{Training Iterations}
+\section{Finger-Key Identification Speed}
+
+\subsection{Training Iterations}
 
@@
 The training iterations had an average keyboard detection time of
-$0.8620\pm0.0502s$ and average finger identification time of $0.0863\pm0.0040s$.
-Throughout the training, there was no noticeable difference in running time over
-the same dataset as shown in
+$0.8620\pm0.0502s$ and an average finger identification time of
+$0.0863\pm0.0040s$. Throughout the training, there was no noticeable difference
+in running time over the same dataset as shown in
 Figure~\ref{fig:rd-training-keyboard-detection-time}. However, the first
@@
 average keyboard detection time included other operations that were not part of
-getting the Key-Edge Coordinates Map. Removing this iteration, the average
+getting the Key-Edge Coordinates map. Removing this iteration, the average
 keyboard detection time becomes more consistent, with it being
@@
 
-\subsubsection{Test Results}
-
-\begin{table}[H]
-	\small
-	\centering
-	\caption{\label{tab:rd-speed}Speed Results of the Module with Test Data}
-	\begin{tabular}{ p{0.4\textwidth} R{0.2\textwidth} }
-		\toprule
-		Category                        & Seconds \\
-		\midrule
-		Avg. Keyboard Detection Time    & 0.838   \\
-		Avg. Finger Identification Time & 0.083   \\
-		\bottomrule
-	\end{tabular}
-\end{table}
-
-These results, as tabulated in Table~\ref{tab:rd-speed} were consistent with the
-training iterations. This means that the module has repeatable and predictable
-running times.
-
 \subsection{Average Keyboard Detection Time}
 This is the time that was measured as the module generated the Key-Edge
-Coordinates Map. This corresponds with the first step of the module. It is to be
+Coordinates map. This corresponds with the first step of the module. It is to be
 noted that this time does not affect keyboard typing and is not limited by the
-0.32 threshold as the user is not yet typing at this point.
+0.32 threshold as the user is not yet typing at this point. The module got an
+average keyboard detection time of 0.083s when ran on the test data.
 
@@
 
-The average finger identification time that the module got was 0.083s which is
-well below 0.32s. This can be attributed to the speed of the MediaPipe library,
-and the pre-calculation of the edge coordinates.
+The average finger identification time that the module got when ran on the test
+data was 0.083s which is well below 0.32s. This can be attributed to the speed
+of the MediaPipe library, and the pre-calculation of the edge coordinates.
 
 \subsubsection{Real-Time Finger-Key Identification}
+Based on the definition of \ac{wpm} from \citet{arif2009}, it can be
+extrapolated that calculating \ac{cps} is:
+
+\begin{equation}
+	CPS = \frac{WPM}{60} \cdot 5
+	\label{eq:cps}
+\end{equation}
+
+The average \ac{wpm} when typing using the QWERTY layout, according to test data
+from \citet{keybr} is $\approx37.5$ \ac{wpm}. As such, we can calculate
+that the average Characters per Second when typing using the QWERTY layout is
+3.125. We can calculate the seconds needed per character by this equation:
+
+\begin{equation}
+	s = \frac{1}{CPS}
+	\label{eq:s}
+\end{equation}
+
+This means that each character, on average, requires 0.32 seconds to type. This
+sets the maximum time spent for finger-key identification for a single
+character. Any more than this would result in a noticeable slowdown during
+typing.
+
 With each keypress only requiring 0.083s for finger-key identification, it can
@@
 Equation~\ref{eq:cps}. This is well above the median of 37.5 \ac{wpm} on Keybr.
-The module can also perform finger-key identification for 90.99\% of the
+The module can also perform real-time finger-key identification for 90.99\% of the
 community members of Monkeytype with the speed role.
 
-\chapter{Conclusion}
-
+\section{Conclusion}
 The finger key identification module was successfully developed and implemented
@@
 mapping used the following algorithm and techniques: (1) Edge Detection using
-Sobel filter \parencite{sobel2014}, (2) Thresholding using Otsu's algorithm
-\parencite{otsu}, (3) Finding contours using the algorithm of
-\textcite{contours}, (4) Line Simplification using the Douglas-Peucker algorithm
-\parencite{douglas-peucker}, and (5) Perspective transform
-\parencite{opencv-image-transform}. Finger detection utilized a ready-made
-solution called MediaPipe Hands by \textcite{mediapipe}. The two were combined
-to perform finger-key identification.
+Sobel filter \citep{sobel2014}, (2) Thresholding using Otsu's algorithm
+\citep{otsu}, (3) Finding contours using the algorithm of \citet{contours}, (4)
+line simplification using the Douglas-Peucker algorithm \citep{douglas-peucker},
+and (5) Perspective transform \citep{opencv-docs}. Finger detection utilized a
+ready-made solution called MediaPipe Hands by \citet{mediapipe}. The two were
+combined to perform finger-key identification.
 
@@
 was placed above the keyboard, pointing directly downwards. It captured videos
-in 720p/30fps with a diagonal field of view of 78\degree \parencite{logitech}.
-The computer used had an AMD Ryzen 5 3600 CPU, AMD Radeon RX 5600 XT GPU,
-G.Skill Trident Z Neo RGB 16GB 3200mhz RAM, and Samsung SSD 850 EVO SSD.
+in 720p/30fps with a diagonal field of view of 78\degree. The
+computer used had an AMD Ryzen 5 3600 CPU, AMD Radeon RX 5600 XT GPU, G.Skill
+Trident Z Neo RGB 16GB 3200mhz RAM, and Samsung SSD 850 EVO SSD.
 
@@
 speeds of 143.040 \ac{wpm}. This is above the median typing speed of 37.5
-\ac{wpm} on Keybr \parencite{keybr}. This also handles 90.99\% of the 12678
-members with the speed role on MonkeyType \parencite{monkey-stats}.
+\ac{wpm} on Keybr \citep{keybr}. This also handles 90.99\% of the 12678 members
+with the speed role on Monkeytype.
 
@@
 
-\chapter{Future Work}
-
-\section{Trainer}
-A proof of concept trainer with real-time finger-key identification was
-developed, however this trainer is lacking in functionality to be an actual
-usable trainer that will be used for educational purposes.
-
-In addition, the wall of latency that exists during th Key-Edge Coordinates Map
-can be removed by doing this process on the browser. This can be done by redoing
-the process in C, and compiling that to WebAssembly. The MediaPipe team did the
-same thing for MediaPipe Hands, and performance, for this use case, was more or
-less equal.
-
-\section{Standardized Finger-Key Mapping}
-While there are mappings for the default ANSI and ISO layouts, these layouts
-were not official and were defacto standards. While the trainer uses this
-defacto standard, it wouldn't be difficult to adapt the trainer to use another
-standard for finger-key mapping.
-
-\section{Character Representation}
-The test and training data did not have equal representation of the keys within
-the keyboard. For the module to undergo more rigorous testing, it would be
-better to have a dataset that has equal representation of all of the characters
-in the keyboard.
-
-\section{Effects on Touch Typing Education}
-The paper created a method for finger-key identification in hopes of improving
-touch typing education, however, there was no test done to prove that the module
-will affect the learner's skill in touch typing
-
-\section{Effects on Health}
-In connection with the previous section, the module also has no testing done to
-quantify if training using the module will improve the posture and reduce
-problematic hand positions and movements that can cause nerve and muscular
-disorders.
-
-\section{Environmental Setup}
-The environmental setup was tightly controlled, with the keyboard and the desk
-of contrasting color, lighting was nearly perfect, and the camera was of great
-quality. This is not realistic and is not representative of the real world.
-Future work could improve upon this area. An idea would be to have a separate
-process where the user could just select four points rather than relying on an
-automated computer vision process to detect the four edges.
-
-\section{ROI Buffer}
-The buffer was selected after trail and error in the training phase of the
-algorithm. However, this buffer may not be the best and adaptive techniques or a
-machine learning algorithm may select better values for the buffer.
-
-\section{Detecting Keyboard Movement Between Frames}
-The module is limited in its ability to track keyboards that move between
-keypresses. While this is not something that usually happens during keyboard
-typing, having the ability to track movements would be a great addition to
-improving the module's performance.
-
-\section{Different Keyboard Types, Layouts, and Colors}
-The module is limited to detecting keypresses with a 60\% keyboard. This type of
-keyboard is not that commonly used. Other keyboard types, like the 100\%, 75\%,
-and TKLs would be good starting points for adding support. In addition,
-supporting other layouts, like ISO, would also greatly increase the number of
-people that could use the module. Furthermore, supporting other color schemes
-would also widen the number of keyboards that can be used with the module.
+\section{Future Work}
+There are a couple of ways to improve upon the module and its application. (1)
+While a trainer was developed, this trainer is lacking in functionality to be a
+usable trainer for teaching touch typing. (2) There were no tests done to prove
+the effectiveness of the module on a learner's skill in touch typing, and there
+was also no test done to prove that training using the module will improve
+posture and reduce problematic hand positions and movements. (3) The
+environmental setup was tightly controlled. This is not representative of
+real-world conditions. (4) The keyboard used was specific. Supporting other
+keyboard types, layouts, and color schemes would greatly increase the number of
+users that could use the module. (5) The buffer of the \ac{roi} was selected
+through trial and error. Adaptive techniques or machine learning algorithms
+could select better values for the buffer. (6) The module is limited in its
+ability to track keyboards that move between keypresses. (7) The dataset used
+did not have an equal representation of all characters within the keyboard and
+could be improved.
+
+\section{CRediT authorship contribution statement}
+\textbf{Oscar Vian L. Valles}: Conceptualization, Methodology, Software,
+Investigation, Resources, Data Curation, Writing --- original draft, Writing ---
+review \& editing, Visualization, Funding acquisition. \textbf{Dhong Fhel K.
+	Gom-os}: Conceptualization, Methodology, Writing --- original draft, Writing
+--- review \& editing, Supervision, Project administration
+
+\section{Declaration of Competing Interest}
+The authors declare that they have no known competing financial interests or
+personal relationships that could have appeared to influence the work reported
+in this paper.
+
+\section{Acknowledgment}
+We would like to recognize that this research was partially funded by the
+Department of Science and Technology --- Science Education Institute
+
+\bibliographystyle{elsarticle-harv}
+\bibliography{references}
 
 \newpage
-\appendix
-
-\twocolumn
-\chapter{Key-Color Values Map}
-\label{appendix:key-map}
-
-\tablehead{\toprule Key &\multicolumn{1}{l}{Color (in BGR)} \\ \midrule}
-\tabletail{%
-	\midrule \multicolumn{2}{r}{{Continued on next column}} \\ \midrule}
-\tablelasttail{%
-	\\\midrule
-	\multicolumn{2}{r}{{Concluded}} \\ \bottomrule}
-\centering
-\small
-\begin{supertabular}{ll}
-	LeftControl & (85, 0, 0)\\
-	LeftSuper & (45, 0, 0)\\
-	LeftAlt & (58, 35, 0)\\
-	Spacebar & (62, 55, 0)\\
-	RightAlt & (0, 28, 44)\\
-	RightSuper & (0, 3, 73)\\
-	RightControl & (24, 0, 76)\\
-	` & (255, 0, 0)\\
-	1 & (255, 153, 0)\\
-	2 & (255, 229, 0)\\
-	3 & (189, 255, 0)\\
-	4 & (143, 255, 0)\\
-	5 & (51, 255, 0)\\
-	6 & (0, 255, 102)\\
-	7 & (0, 255, 194)\\
-	8 & (0, 224, 255)\\
-	9 & (0, 163, 255)\\
-	0 & (0, 102, 255)\\
-	- & (0, 10, 255)\\
-	= & (82, 0, 255)\\
-	$\vert$ & (128, 0, 255)\\
-	Delete & (189, 0, 255)\\
-	Tab & (224, 0, 0)\\
-	Q & (214, 128, 0)\\
-	W & (225, 203, 0)\\
-	E & (160, 217, 0)\\
-	R & (126, 224, 0)\\
-	T & (43, 217, 0)\\
-	Y & (0, 216, 86)\\
-	U & (0, 195, 149)\\
-	I & (0, 183, 208)\\
-	O & (0, 136, 212)\\
-	P & (0, 79, 196)\\
-	] & (0, 8, 211)\\
-	{[} & (63, 0, 196)\\
-	Backspace & (134, 0, 181)\\
-	CapsLock & (195, 0, 0)\\
-	A & (167, 100, 0)\\
-	S & (178, 161, 0)\\
-	D & (124, 168, 0)\\
-	F & (101, 180, 0)\\
-	G & (31, 157, 0)\\
-	H & (0, 139, 56)\\
-	J & (1, 139, 106)\\
-	K & (0, 130, 148)\\
-	L & (0, 94, 147)\\
-	; & (0, 52, 130)\\
-	' & (0, 7, 170)\\
-	Enter & (35, 0, 109)\\
-	LeftShift & (150, 0, 0)\\
-	Z & (104, 62, 0)\\
-	X & (130, 117, 0)\\
-	C & (77, 104, 0)\\
-	V & (63, 113, 0)\\
-	B & (18, 88, 0)\\
-	N & (0, 66, 26)\\
-	M & (0, 84, 64)\\
-	, & (0, 75, 85)\\
-	. & (0, 70, 109)\\
-	/ & (0, 32, 80)\\
-	RightShift & (0, 5, 128)\\
-
-\end{supertabular}
-
-\onecolumn
-\chapter{Typing Test Sequences}
-\label{appendix:test-sequences}
-\normalsize
-\begin{enumerate}
-	\item ``What of it, if some old hunks of a sea-captain orders me to get a broom and sweep down the decks?'' \parencite{moby-dick}
-	\item ``I sat down on an old wooden settle, carved all over like a bench on the Battery.'' \parencite{moby-dick}
-	\item ``I lay there dismally calculating that sixteen entire hours must elapse before I could hope for a resurrection.'' \parencite{moby-dick}
-	\item ``How slowly the time passes here, encompassed as I am by frost and snow!'' \parencite{frankenstein}
-	\item ``I listened to my father in silence and remained for some time incapable of offering any reply.'' \parencite{frankenstein}
-	\item ``Think you’re escaping and run into yourself. Longest way round is the shortest way home.'' \parencite{ulysses}
-	\item ``History, Stephen said, is a nightmare from which I am trying to awake'' \parencite{ulysses}
-	\item ``A man of genius makes no mistakes. His errors are volitional and are the portals of discovery.'' \parencite{ulysses}
-	\item ``Never trust to general impressions, my boy, but concentrate yourself upon details'' \parencite{sherlock}
-	\item ``I have no data yet. It is a capital mistake to theorise before one has data.'' \parencite{sherlock}
-\end{enumerate}
-
-\chapter{Screenshots of Data Analysis}
-\label{appendix:sheets}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{sheets-raw-data.png}
-	\caption{Raw data as seen on Google Sheets}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{sheets-uncertain.png}
-	\caption{Filtered raw data to show uncertain successes}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{sheets-failures.png}
-	\caption{Filtered raw data to show failures}
-\end{figure}
-
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.8\textwidth]{sheets-analysis.png}
-	\caption{Metrics calculated from the raw data}
-\end{figure}
-
-\chapter{Gantt Chart}
-\begin{landscape}
-	\includepdf[pages=-,landscape,scale=0.8,pagecommand={}]{images/gantt-chart.pdf}
-\end{landscape}
-
-\newpage
-\printbibliography[heading=bibintoc,title={References}]{}
 
